### 2024-02-05 apples_and_oranges.qmd
several changes

---
title: Fundamental Problems with the Evidence Base of Adolescent Depression Treatments.
subtitle: Insights from the relative efficacy of Control Conditions.
author:
  - name: Argyris Stringaris
    email: a.stringaris@ucl.ac.uk
    affiliations: 
        - id: ucl_1
          name: University College London
          department: Divisions of Psychiatry and Psychology and Language Science
          address: 1-19 Torrington Place
          city: London, UK
          state: 
          postal-code: WC1E 7HB
          country: United Kingdom
    attributes:
        corresponding: true
    note: 
  - name: Charlotte Burman
    email: charlotte.burman@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
  - name: Dayna Bhudia
    email: dayna.bhudia@ucl.ac.uk
    affiliations:
        - ref: ucl_1
  - name: Lucy Foulkes
    email: lucy.foulkes@psych.ox.ac.uk
    affiliations:
        - id: OU
          name: University of Oxford
          department: Experimental Psychology
          address: Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Road
          city: Oxford
          postal-code: OX2 6GG
          country: United Kingdom
    note:
  - name: Carmen Moreno
    email: cmoreno@hggm.es
    affiliations:
        - id: HGM
          name: Hospital Gregorio Marañón
          department: Department of Psychiatry
          address: 46 C. del Dr. Esquerdo
          city: Madrid
          state:
          postal-code: 28007
          country: Spain
  - name: Samuele Cortese
    email: samuele.cortese@soton.ac.uk
    affiliations:
        - id: soton
          name: University of Southampton
          department: Centre for Innovation in Mental Health (CIMH), School of Psychology
          address: Highfield Campus, Building 44
          city: Southampton
          state:
          postal-code: SO17 1BJ
          country: United Kingdom
  - name: Georgina Krebs
    email: g.krebs@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note: 
abstract: |
  To follow
keywords: 
  - depression
  - adolescents
  - antidepressants
  - psychotherapy
date: last-modified
bibliography: emotion_concepts_paper.bib
format: 
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
---

# Introduction

Depression is the leading cause of disability in adolescents and influential guidelines recommend that it be treated with either psychotherapy or anti-depressant medication or their combination. The evidence base for this recommendation typically derives from the appraisal of randomised controlled trials (RCTs) in which one modality (psychotherapy or anti-depressants) are compared against a control condition. In this paper we: a) test the hypothesis that medication control arms are more efficacious than psychotherapy control arms and that this undermines the validity of treatment comparisons and therefore of existing guidelines about depression treatment; b) explore how variation in treatment control arms can itself be exploited to advance our thinking concerning the mechanisms of depression and its treatment.

Clinicians across medicine and other health professions are faced with choices between different treatment modalities for their patients, be it between surgical or medical (drug-based) treatments or, in the case of adolescent depression, between psychotherapy and antidepressant medication. To make informed choices, clinicians rely on the extant evidence, frequently summarised in the form of national guidelines or other treatment recommendations. Such a comparison is inherent in the internationally influential UK National Institute of Clinical Excellence (NICE) guidelines for adolescent depression which recommend that youth be first offered psychological therapy over medication for most presentations of depression. Such comparisons between the two treatment modalities are inherent in other recommendations, e.g. the practice parameter issued by the American Academy of Child and Adolescent Psychiatry or the Royal Australian and New Zealand College of Psychiatrists.

These recommendations are not based on head to head trials between modalities; only one head-to-head trial exists in adolescent depression. Instead, they rely on the comparison of the results of randomised controlled trials (RCTs) of each modality against RCTs of the other. In other words, it rests on a comparison of comparisons.

In the case of antidepressant studies, the control is the placebo pill (standard also in the rest of medicine), whilst controls for psychotherapy studies can vary substantially from waiting list to treatment as usual to an attention control, such as psychoeducation. Stated in simple formal terms, the comparison relies on testing this equality,

$$Effect_{Med} = Effect_{Psy}$$ {#eq-1}

where *Effect* denotes the effect on depression that either medication (*Med* ) or psychotherapy (*Psy*) have. Yet for this indirect comparison to be valid, some basic assumptions need to hold. The most important of these is the *all-else-being-equal,* or more technically *ceteris paribus,* assumption with regards to trial design. By necessity, therefore, the control conditions for each antidepressant and psychotherapy trials must be equal for inferences to be valid. We call this the *equality of controls* assumption.

The importance of this assumption can be easily intuited by expanding equation 1 to include the comparisons on which each side of that equation rests:

$$Effect_{MedActive} - Effect_{MedControls} = Effect_{PsyActive} - Effect_{PsyControls}$$ {#eq-2}

where MedActive and PsyActive are the active arms of antidepressant and psychotherapy trials, respectively, and MedControls and PsyControls the respective control arms. The assumption here has to be that the controls are equal, that is, $$ Effect_{MedControls} = Effect_{PsyControls}$$ A simple example from medicine helps illustrate why this assumption is important Let the two modalities of treatment for appendicitis be antibiotic medication and surgery, and let there be a set of trials for each modality. In the medication trials, on average 40% of participants respond to the antibiotics and only 0% to placebo; in the surgery trial, 80% of people respond to surgery and 40% to sham. If one relied on Equation 1 to compare the treatments they would arrive at the absurd conclusion that the two treatments are equal and recommend to a patient that it does not really matter which one they choose. Playing around with numbers in which no equality of the respective control arms is required, leads to other risible conclusions. Yet, in the dicussion about depression treatments, a focus on the equality of controls is surprisingly absent.

The importance of testing the equality of controls assumption for patients and the public can be vast because treatment guidelines, clinicians' advice to patients, and patients' informed choices are all affected by the validity of treatment comparisons. Similarly, the rational allocation of public funds towards treatment modalities rests on valid comparisons between them.

Beyond its immediate clinical importance, understanding the nature of controls in clinical trials of depression is key for reasons of causal inference and mechanistic understanding of treatment. Adolescent depression is notable for a relatively high placebo effect: the proportion of adolescents responding to placebo is higher in depression than in anxiety, and it is higher for adolescent vs adult depression. The reasons for these differences remain unclear and are typically treated as nuisance effects to be minimised by using trial designs such as a placebo lead in period. However, an alternative view is that placebo effects, or more generally, variability in the response to control conditions, may be informative about mechanisms of improvement. If, for example, the antidepressant effect of placebo is stronger than that of equivalent control arms in psychotherapy, then we ought to ask about the possible explanations of such differences and consider whether they can be exploited for improving treatment delivery, as previously suggested.

In this paper, we have two aims: first, test whether the equality of controls assumption of adolescent depression treatment holds, and therefore scrutinise the validity of current recommendations concerning choice between treatment modalities; second,we explore the variabiliity in trial control arms across and within the two treatment modalities and examine how they relate to anti-depressant effects.

Our main hypothesis is that the within group effect of psychotherapy controls is substantially weaker than that of the placebo pill and we will test this using random effects meta-regression. We will also use series of sensitivity analyses. First we compared CBT trial arms with those of trials involving fluoxetine and escitalopram. CBT is the most widely studied treatment and, similar to fluoxetine, it has metanalytic evidence supporting its use. Both fluoxetine and escitalopram are approved by the FDA as treatments for adolescent depression. We also conduct analyses excluding waitlist control arms as it could be argued that waitlist control arms represent outliers amongst psychotherapy control studies that reduce the estimate of the overall efficacy. Finally, we also conduct sensitivity analyses in the form of simulations, to ensure that results are robust to estimations of the standard error of the change scores.

We then go on to examine the control arms of psychological treatments in both a qualitative/descriptive way (e.g. the content of attention control) and by quantifying characteristics such as session frequency, duration of delivery. We then explore how these characteristics relate to the magnitude of within group change.

@stringarisDevelopmentalPathwaysChildhood2014

# Method

## Studies included

For this review, we primarily drew upon studies included in two recent meta-analyses. Please refer to these original meta-analyses for a detailed description of study selection criteria. Psychotherapy studies were drawn from Cuijpers' (2023) systematic review and meta-analysis of randomised trials comparing psychotherapy for youth depression against control conditions. Cuijpers made available a full dataset of psychotherapy trials (via https://www.metapsy.org/), which we used for the current study. Whilst Cuijpers (2023) excluded those studies for which the primary outcome variable could not be calculated due to missing data, we included these studies and performed the imputations outlined below; hence we have more psychotherapy studies included in this review compared to Cuijpers' (2023) meta-analysis. Whilst the online database is regularly updated, we chose to exclude studies published after the final date of Cuipers' (2023) literature search.

Medication studies were drawn from Cipriani's (2016) network meta-analysis examining the efficacy and tolerability of a range of antidepressants and placebo for major depressive disorder in young people. The study dataset was made available online though did not include means or standard deviations at baseline or post-test. We emailed the study authors requesting a full dataset with this data, though did not receive a reply, and hence conducted data extraction for medication studies. We excluded three studies because they did not include a control arm. We were unable to locate and therefore complete extraction for two papers. Many studies did not report complete data, and so we emailed all corresponding authors to request missing data, though did not receive any responses. **Yet to do** We conducted a systematic search for medication studies published after the final search date of Cipriani's (2016) review up to the final search date of Cuijpers' (2023) review to ensure we analysed an equivalently up-to-date database of medication trials. We used the same search terms outlined in Cipriani (2016).

NOTE: **Charlotte** to please check code lines 986 to 1121. **CB**: Done. NOTE: **Charlotte** please check that the studies I have in the *df_long_for_metan* dataframe makes sense and correspond to what you would expect. In the beginning of the results below, I explain in detail what is missing where. **CB**: I've checked and there are fewer psy active conditions that I expected - it seems some of the trials with multiple active arms have been filtered out. I have now fixed this and the numbers look correct. The numbers look right for med studies. One thing - for psy the no of studies with missing Cohens d for controls should probably be considered 6 (not 7) as it is absent for March/TADS because we removed it (didn't want placebo to be considered as a psy control) rather than there being missing data. <!--# Charlotte, that's all great. What we now need here is an account of exactly how  we searched and how we used the other datasets. Follow previous examples, though in our case, it will be a lot about how we used the accounts of others. Please make sure to say where that we asked about missing information. **CB** Completed above. -->

## Statistical Analysis

### Measures of Effect

As the measure of effect of each individual study, we used the within-group Standardised Mean Difference (SMD), which we defined following NOTE: Charlotte cite Lakens 2013 and in there Cummings 2012, as:

$$SMD_{change} = \frac{Mean_{t_{2}} - Mean_{t_{1}}}{\frac{SD_{t{2}} + SD_{t{1}}}2}\ $$ {#eq-4}

where, $Mean_{t_{2}}$ and $Mean_{t_{1}}$ refer to the means of the main outcome score at the end and beginning of the intervention respectively and $SD_{t_{2}}$ and $SD_{t_{1}}$ to the respective standard deviations. NOTE **Charlotte**, please add note here with what we did to impute SDs where missing, e.g. using the pre to impute the post etc. In general state here the rules we used citing the Cochrane book. **CB** Where individual studies did not report all data required to calculate the SMD, we imputed missing data according to the methods summarised by X (citation), in the following order. If a study reported the standard error of the mean, the SD was obtained simply by multiplying the SE by the square root of the sample size. For conditions where the SD was missing at one time point, the baseline SD was substituted by the post-test SD, and vice versa. If the SD was not available at either time point, missing values were replaced by the mean of the SDs available for comparable cases (defined as same trial type (psy or med), same instrument, same timepoint (pre or post), and same arm (control or active)). Where there were missing means at either baseline or post-test, missing values were calculated using mean change scores, preferring the change scores reported in the paper itself, though where this was unavailable, using the change scores reported in the dataset from Cipriani's meta-analysis (citation).

For the purposes of metanalysis, it is necessary to estimate a standard error of the change score, that is of the SMD. This is calculated according to:

$$ SE_{change} = \sqrt{\frac{r_{t_{1}t_{2}}}{n} + \frac{SMD^2}{2n}}$$ {#eq-5}

where $n$ refers to the study sample size and $r_{t_{1}t_{2}}$ refers to the correlation between the outcome score obtained at baseline and at the end point. This correlation is typically not reported in studies and is often imputed using previously reported correlations for the instruments used. However, this practice has given rise to concerns about misestimation. Whilst such misestimation is possible, there is no reason to expect that it would be systematic, i.e. bias estimation of the effects for the control group of medication compared to those of psychotherapy. Still, to alleviate such concerns we have used a simulations.

In particular, we simulated one thousand truncated distribution of standard errors with the following general characteristics: <!--# Georgina, I have used the Truncated Normal, we can obviously do sensitivity analyses with other ones -->

$$r_{t_{1}t_{2}} \sim \mathcal{TN}(\mu, \sigma, a, b)$$ {#eq-6}

for which we chose the mean to be $\mu = 0.65$, the standard deviation to be $\ sigma = 0.2$, and the upper and lower bounds to be $a = 0.45$ and $b = 0.9$, respectively. We then used these simulated datasets in the subsequent metanalyses.

<!--# Charlotte, I am particularly keen for you to code independently the SE calculation, maybe use all the other parts as they are, but check the SEs. Once you have re-assured yourself about it, could you please read the code for the other steps below, but break it down by taking it out of the function and probably out of the loop and verifying that you would get the same results for one dataset, e.g. the first one in the list of 1000. It would be amazing to verify the results this way. Finally, as discussed, we can vary the parammeters of the distribution if we have better a prioris-->

```{r, warning=FALSE, message = FALSE, echo=FALSE}
###########Libraries needed
library(tidyverse)
library(truncnorm)
library(metafor)
```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#############Some custom made functions#############

###### 1. Simulation Function for SEs
simulate_dataframes_for_st_errors <- function(df, num_repetitions, seed, n, a, b, mean, sd) { # n refers to the number
                                                                                              # of unique ids to which a correlation coef
                                                                                              # is allocated.
  set.seed(seed)

  # Empty list to store simulated dfs
  list_df_simulated <- list()

  for (i in 1:num_repetitions) {
    # Simulate the vector
    simulated_correlations_vector <- truncnorm::rtruncnorm(n, a, b, mean, sd)  #using the truncnorm to create correlations

    # Create a copy of the original dataframe
    df_simulated_copy <- df

    # Add/update the sims column
    df_simulated_copy$correlation_sim_values <- simulated_correlations_vector[match(df_simulated_copy$new_study_id, unique(df_simulated_copy$new_study_id))]

    # Remove duplicated columns
    df_simulated_copy <- df_simulated_copy[, !duplicated(colnames(df_simulated_copy))]

    # Calculate the ses
    df_simulated_copy <- df_simulated_copy %>% 
      group_by(new_study_id, baseline_n) %>% 
      mutate(simulated_se = sqrt(((2*(1-correlation_sim_values))/baseline_n) + 
                                   (cohens_d^2/(2*baseline_n))))

    # Add the simulated dataframe to the list
    list_df_simulated[[i]] <- df_simulated_copy
  }

  return(list_df_simulated)
}

###############2. Simulation function for metaregression
run_metaregression <- function(list_of_datasets, model_formula) {
  list_model_meta_reg <- list()

  for (i in seq_along(list_of_datasets)) {
    list_model_meta_reg[[i]] <- metafor::rma(yi = cohens_d,
                                             sei = simulated_se,
                                             data = list_of_datasets[[i]],
                                             method = "ML",
                                             mods = model_formula,
                                             test = "knha")
  }

  return(list_model_meta_reg)
}

#########2. Function to count studies with missingness etc.

count_studies <- function(df) {
  df_count_studies <- df %>%
    group_by(psy_or_med, arm_effect_size) %>%
    count(is.na(cohens_d)) %>%
    rename(is_missing = `is.na(cohens_d)`) %>%
    mutate(
      condition = case_when(
        psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
        psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
        psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
        psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control"
      )
    )

  df_count_studies$condition <- factor(df_count_studies$condition)  # turn to factor

  # relevel so that medication control becomes the reference category for the regression
  df_count_studies$condition <- relevel(df_count_studies$condition, ref = "medication_control")

  return(df_count_studies)
}

##########3. Function to extract mean coefficients and to extract model characteristics

aggregate_model_results <- function(list_model_1_meta_reg, condition) {
  df_coefs <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  df_se <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  df_z_value <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  df_lower_ci <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  df_upper_ci <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  tau_sq <- rep(0, length(list_model_1_meta_reg))
  i_sq <- rep(0, length(list_model_1_meta_reg))
  k <- rep(0, length(list_model_1_meta_reg))
  r_sq <- rep(0, length(list_model_1_meta_reg))

  for (i in 1:length(list_model_1_meta_reg)) {
    for (j in 1:length(condition)) {
      df_coefs[i, j] <- list_model_1_meta_reg[[i]]$beta[[j]]
      df_se[i, j] <- list_model_1_meta_reg[[i]]$se[[j]]
      df_z_value[i, j] <- list_model_1_meta_reg[[i]]$zval[[j]]
      df_lower_ci[i, j] <- list_model_1_meta_reg[[i]]$ci.lb[[j]]
      df_upper_ci[i, j] <- list_model_1_meta_reg[[i]]$ci.ub[[j]]
      tau_sq[i] <- list_model_1_meta_reg[[i]]$tau2
      i_sq[i] <- list_model_1_meta_reg[[i]]$I2
      k[i] <- list_model_1_meta_reg[[i]]$k
      r_sq[i] <- list_model_1_meta_reg[[i]]$R2
    }
  }

  df_coefficients_model <- data.frame(
    coefficients = colMeans(df_coefs),
    se = colMeans(df_se),
    z_value = colMeans(df_z_value),
    lower_ci = colMeans(df_lower_ci),
    upper_ci = colMeans(df_upper_ci),
    tau_sq = mean(tau_sq),
    i_sq = mean(i_sq),
    k = mean(k),
    r_sq = mean(r_sq)
  )

  return(df_coefficients_model)
}

###########4. Function to extract the SMDs  (SMD, ses, CI) for each level of the dummy #variable. The output is used below in 5.
# it is a slightly awkward one, because I couldnt' come up with a better way to add the intercept to #each coefficient
# given the named output of the built in coef function. 
# here I exctract the coeffcients
extract_coefficients_func <- function(df_with_coefs){
  list_coefs <- list()
  coefs <- 0
  coefficient_output <- 0

  for(i in 1: length(coefficients(df_with_coefs))){

    temp_vec <- c(0, rep(coefficients(df_with_coefs)[[1]], 
                         length(coefficients(df_with_coefs)) - 1 ))
    coefs[i] <- coef(df_with_coefs)[[i]]

    coefficient_output <- temp_vec + coefs

    st_error_output <- df_with_coefs$se

    df_coefficients <- data.frame(cbind(coefficients = coefficient_output, se = st_error_output ))

    df_coefficients$lower_bound <- df_coefficients$coefficients - 1.96*df_coefficients$se
    df_coefficients$upper_bound <- df_coefficients$coefficients + 1.96*df_coefficients$se
  }
  return (df_coefficients )
}

############5. Function to get the means out of the SMDs (this is similar to Func 3 and I should at some point create one more general one)
# here the argument list_of_dfs is the list of extracted coefficients that you get out of the #extract_coefficients_func
# which itself you get from the run_metaregression function. The argument conditions refers to #the levels of the variable that describes each study arm

calculate_mean_coefs_ses <- function(list_of_dfs, conditions) {
  df_coefs <- matrix(NA, nrow = length(list_of_dfs), ncol = length(conditions))
  df_se <- matrix(NA, nrow = length(list_of_dfs), ncol = length(conditions))

  for (i in 1:length(list_of_dfs)) {
    for (j in 1:length(conditions)) {
      df_coefs[i, j] <- list_of_dfs[[i]]$coefficients[j]
      df_se[i, j] <- list_of_dfs[[i]]$se[j]
    }
  }

  df_mean_coefs_ses <- data.frame(
    coef_means = colMeans(df_coefs),
    se_means = colMeans(df_se)
  )

  df_mean_coefs_ses$lower_ci <- df_mean_coefs_ses$coef_means - 1.96 * df_mean_coefs_ses$se_means
  df_mean_coefs_ses$upper_ci <- df_mean_coefs_ses$coef_means + 1.96 * df_mean_coefs_ses$se_means

  return(df_mean_coefs_ses)
}

###############6. Function for plotting the SMDs
plot_means_function <- function(dataframe, subtitle) {
  dataframe$text_label <- 0

  for (i in 1:nrow(dataframe)) {
    dataframe$text_label[i] <- paste(
      dataframe$condition[i], "\n",
      round(dataframe$coef_means[i], 2),
      "[" ,
      round(dataframe$lower_ci[i], 2), 
      round(dataframe$upper_ci[i], 2),
      "]"
    ) 
  }

  # Recode condition for ease of plotting
  dataframe$condition <- str_replace_all(dataframe$condition, "_", " ")

  # Encode colours
  redish_palette <- c("medication control" = "deeppink", "medication active" = "deeppink3")
  blueish_palette <- c("psychotherapy active" = "steelblue1", "psychotherapy control" = "steelblue3")

  # Plotting
  plot_means <- ggplot(dataframe, aes(x = coef_means, y = 1:nrow(dataframe),
                                      colour = condition, label = text_label)) +
    geom_point(label=dataframe$k) +
    geom_errorbar(aes(xmin = lower_ci, xmax = upper_ci), width = 0.2, position = position_dodge(0.5)) +
    geom_text(vjust = +1.5, size = 4) +
    scale_size_continuous(guide = "none") +
    guides(colour = FALSE) + 
    scale_color_manual(values = c(redish_palette, blueish_palette)) +
    theme_minimal() +
    labs(x = "TE-random", y = NULL, title = "Adolescent Depression Trial Efficacy by Treatment Arm",
         subtitle = subtitle) +
    xlab("Standardized Mean Difference (SMD) with 95% CIs") +
    ylab("") +
    ylim(0, nrow(dataframe) + 1) +
    xlim(-3.0, 0.5) +
    geom_vline(xintercept = 0, linetype = "dashed", size = 1.5, colour = "grey") +
    geom_segment(x = -1.25, xend = -1.75, y = 4.7, yend = 4.7, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
    geom_text(x = -1.55, y = 4.9, label = "More Effective", color = "grey", vjust = 0.5, hjust = 1) +
    geom_segment(x = -1.35, xend = -0.85, y = 4.9, yend = 4.9, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
    geom_text(x = -1.05, y = 5.05, label = "Less Effective", color = "grey", vjust = 0.5, hjust = 0)+
    theme(axis.title.x = element_text(size = 12), 
          axis.text.x  = element_text(size = 12),
          axis.text.y = (element_blank()),
          plot.title = element_text(size = 12),
          plot.subtitle = element_text(size = 12)) +
    geom_curve(aes(x = 0.18, y = 4.73, xend = 0.02, yend = 4.5), color = "grey", curvature = -0.2, arrow = arrow(length = unit(0.25, "cm"), type = "closed")) +
    geom_text(x = 0.2, y = 4.75, label = "No Effect", color = "grey", vjust = 0.5, hjust = 0)

  return(plot_means)
}

###########7. Function for baseline SMD results
library(broom)
run_metareg_and_extract <- function(dataframe, model_1, condition) {
  # Run metareg
  inst_baseline_metareg <- rma(yi = baseline_mean,
                               sei = baseline_stand_error,
                               data = dataframe,
                               method = "ML",
                               mods = model_1,
                               test = "knha")

  # Extract coefficients from the list
  baseline_metareg_results <- tidy(inst_baseline_metareg)
  baseline_metareg_results$condition <- condition
  coefs_for_inst_baseline <- tidy(inst_baseline_metareg)$estimate

  # Calculate SMDs
  smds_for_inst_baseline <- c(coefs_for_inst_baseline[1], coefs_for_inst_baseline[-1] + coefs_for_inst_baseline[1])

  # Calculate upper and lower CIs
  upper_cis_for_inst_baseline <- smds_for_inst_baseline + 1.96 * tidy(inst_baseline_metareg)$std.error
  lower_cis_for_inst_baseline <- smds_for_inst_baseline - 1.96 * tidy(inst_baseline_metareg)$std.error

  # Create dataframe with results
  df_baseline_means_with_cis <- data.frame(
    baseline_smds = smds_for_inst_baseline,
    upper_cis_smds = upper_cis_for_inst_baseline,
    lower_cis_smds = lower_cis_for_inst_baseline,
    condition = condition
  )

  # Return the two dataframes
  return(list(baseline_metareg_results = baseline_metareg_results, df_baseline_means_with_cis = df_baseline_means_with_cis))
}

```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
library(tidyverse)
# Start here------------------------------------------------------
# import master dataset

df_appl_v_orange <-readxl:: read_excel("df_appl_v_orange.xlsx")

# Charlotte to add in instrument end point values (mins / maxs)
# We will use these to transform the scale into % as follows: (score - min of the scale)/(max-min)
# We have done this for both means and standard deviations

df_scale_ranges <- read.csv("scale_ranges.csv")
df_appl_v_orange <- full_join(df_appl_v_orange, df_scale_ranges, by = "instrument_name")

df_appl_v_orange <- df_appl_v_orange %>% 
  mutate(perc_baseline_mean_active = (baseline_mean_active - scale_min) / (scale_max - scale_min)) %>% 
  mutate(perc_baseline_sd_active = baseline_sd_active / (scale_max - scale_min)) %>%  # this #is the best way to estimate the SD of the linear transform of the percentage

  mutate(perc_baseline_mean_control = (baseline_mean_control - scale_min) / (scale_max - scale_min)) %>% 
  mutate(perc_baseline_sd_control = baseline_sd_control  / (scale_max - scale_min)) %>% 
  mutate(perc_post_mean_active = (post_mean_active - scale_min) / (scale_max - scale_min)) %>% 
  mutate(perc_post_sd_active = post_sd_active  / (scale_max - scale_min)) %>% 
  mutate(perc_post_mean_control = (post_mean_control - scale_min) / (scale_max - scale_min)) %>% 
  mutate(perc_post_sd_control = post_sd_control  / (scale_max - scale_min))

#Need to use SMDs, ie our Cohen's d and then use Standard error of SMD, to achieve this
# I need reliabilities.

# note re: CDRS reliability from here https://www.liebertpub.com/doi/epdf/10.1089/104454601317261546 
# Using a   2-week interval, and different psychiatrists from the first to the second assessment, 
# Poz-nanski et    al. (1984) demonstrated high reliability (r=   0.86) 
# for the CDRS-R total score in 53 clinic-referred 6- to 12-year-olds.

### A few more tidying things from Argyris before doing metanalyses
# # discovered an error in the percentage women of the Fristad study. I have checked in the
# # cuijpers dataset and the correct percentage is 43.1, but could not verify with the paper as it is not in 
# # our folder and after a quick search I could not find it online either. Messaged Charlotte on Discord to
# # check again. # CB checked and all good.
df_appl_v_orange[df_appl_v_orange$study_ID=="Fristad, 2019_cbt + placebo_placebo",]$percent_women <-43.1

# Create SEs proportions for percentage women --------------------------------------------

# # We also need to calculate SE for proportion women for the baseline calculations
# # for proportions, this is calculated as sqrt(p(1-p)/n), which I implement stepwise below

product_perc_women <-  (df_appl_v_orange$percent_women/100)*
  (1-(df_appl_v_orange$percent_women/100) )

total_n <- df_appl_v_orange$baseline_n_active + 
  df_appl_v_orange$baseline_n_control

df_appl_v_orange$percent_women_std_error <- sqrt(product_perc_women/total_n )

# Calculate SE for baseline severity --------------------------------------

df_appl_v_orange$baseline_st_error_active <- 
  df_appl_v_orange$baseline_sd_active/sqrt(df_appl_v_orange$baseline_n_active)

df_appl_v_orange$baseline_st_error_control <-
  df_appl_v_orange$baseline_sd_control/sqrt(df_appl_v_orange$baseline_n_control)

# Turn into a long database with unique rows ------------------------------

### Important: create a dataset that will have unique control studies (see problem that we identified with Charlotte, 
#namely common control conditions)
# create new id with Charlotte to help with better identification and work with duplicates (see below) 
df_appl_v_orange  <- df_appl_v_orange  %>%
  mutate(new_study_id = case_when(psy_or_med == 0 ~ paste(study,year, sep = ", "),
                                  .default = study ))

# We are going to use the descr_active and descr_control columns rather than active_type and control_type columns, as these have unique values for each study. However the descr columns are empty for med studies, so I'm going to paste values across to fix this error.
df_appl_v_orange <- df_appl_v_orange %>% 
  mutate(descr_active = if_else(is.na(descr_active ), active_type , descr_active ))

df_appl_v_orange <- df_appl_v_orange %>% 
  mutate(descr_control = if_else(is.na(descr_control), control_type, descr_control ))

# Step 1: keep only the rows with the top instrument in our hierarchy
df_with_distinct_instruments <-  df_appl_v_orange %>%          
  group_by(new_study_id) %>% 
  filter(instrument_value == min(instrument_value)) # coded for the smallest number to be best.

write.csv(df_with_distinct_instruments, "df_with_distinct_instruments.csv")

# Step 2: turn dataframes to long
# A: turn long the rows with active
active_df <- df_with_distinct_instruments %>% 
  dplyr:: select(new_study_id, active_type, descr_active, psy_or_med, baseline_n_active, cohens_d_active,
                 baseline_mean_active, baseline_sd_active, post_mean_active, post_sd_active,
                 perc_baseline_mean_active, perc_baseline_sd_active, perc_post_mean_active, perc_post_sd_active) %>% 
  rename(treatment = descr_active) %>% 
  rename(type = active_type) %>% 
  rename_with(~ str_remove(., "_active")) %>% 
  mutate(arm_effect_size = "cohens_d_active") %>% # chose this clunky name because I had used this variable name in previous code and don't want to bread the rest
  group_by(new_study_id) %>% #go through each study id and remove duplicates
  distinct(treatment, .keep_all = TRUE)

control_df <- df_with_distinct_instruments %>% 
  dplyr:: select(new_study_id, control_type, descr_control, psy_or_med, baseline_n_control, cohens_d_control,
                 baseline_mean_control, baseline_sd_control,post_mean_control, post_sd_control,
                 perc_baseline_mean_control, perc_baseline_sd_control, perc_post_mean_control, perc_post_sd_control) %>% 
  rename(treatment = descr_control) %>% 
  rename(type = control_type) %>% 
  rename_with(~ str_remove(., "_control")) %>% 
  mutate(arm_effect_size = "cohens_d_control") %>% #chose this clunky name because I had used this variable name in previous code and don't want to bread the rest
  group_by(new_study_id) %>% #go through each study id and remove duplicates
  distinct(treatment, .keep_all = TRUE)

df_long_for_metan <- rbind(active_df, control_df)

# I'm going to remove the March / TADS psy control row. It is currently filled with NAs which implies missing data, however it is actually the case that this condition doesn't exist.

df_long_for_metan <- df_long_for_metan %>%
  filter(!(new_study_id == "March, 2004" & is.na(treatment)))

#dim(df_long_for_metan ) #check dimension
#make sure no study lost
#length(unique(df_long_for_metan$new_study_id ) ) == # #length(unique(df_appl_v_orange$new_study_id ) )

# Now check again the studies with muliple arms
# Now checking if this worked with the studies that I used to illustrate the problem above. 
# df_long_for_metan[df_long_for_metan$new_study_id == "Stallard, 2012",]
# df_long_for_metan[df_long_for_metan$new_study_id == "Atkinson, 2014",]
# # also check one which is single to make sure it is kept
# df_long_for_metan[df_long_for_metan$new_study_id == "Ackerson, 1998",]

# save this as the master dataframe  --------------------------------------
# use this dataframe to perform all operations below in quarto
# ask Charlotte to check steps above (lines 986 to 1121)
write.csv(df_long_for_metan, "df_long_for_metan.csv", row.names = F)

########### simulate so that each study has a value from a distribution of 
######## of correlations.

# I will generate random numbers per study id from a distribution with these parameters. 
# it is reasonable to generate one random correlation value per study as there is no reason why the correlation should
# systematically vary within studies

# Use the function I have created for this

df_long_for_metan <- read.csv("df_long_for_metan.csv")

list_df_simulated <-  simulate_dataframes_for_st_errors (df = df_long_for_metan, 
                                              num_repetitions = 1000, 
                                              seed  = 1974,
                                              n = length(unique(df_long_for_metan$new_study_id)), 
                                              a = 0.45, 
                                              b = .9, 
                                              mean = 0.65, 
                                              sd = 0.2)

#check this worked

#list_df_simulated[[10]][,c("new_study_id", "correlation_sim_values", "simulated_se")] # looks right
#summary(list_df_simulated[[1000]]$correlation_sim_values, na.rm = T) # as expected # the mean is around the parameter I gave

```

### Random Effects Metaregression

We estimated the pooled standardised mean difference for each arm by using a random effects metanalysis implemented in R's metafor package. The main underlying assumption of random effects metanalysis is that each study's true effect size $\theta_{k}$ is affected not only by sampling error $\epsilon_{k}$, but also by $\zeta$ which represents heterogeneity between studies, allowing each study's estimate to vary along a distribution of effects, and the distribution of true effect sizes termed $\tau^2$. Therefore, we can estimate a two stage model with:

$$Y_i \sim \mathcal{N}(\theta_i, \sigma_i^2)
$$ {#eq-7}

$$\theta_i \sim \mathcal{N}(x_i\beta, \tau_i^2)
$$ {#eq-8}

where $Y_i$ is the estimated effect size for study i, has a normal distribution with $\theta_i$ as its true mean effect and sampling error $\sigma^2$. Whereas $\theta_i$ is a study-specific instantiation of the distribution of effect sizes, with $\tau^2$ representing heterogeneity.

This then gives rise to:

$$Y_i = x_i\beta_i + u_i + \epsilon_i$$ {#eq-9}

where,

$$u_i \sim N(0, \tau^2) $$ {#eq-10}

describes the deviation of each study from the mean of the distribution, and,

$$\epsilon_i \sim N(0, \sigma^2)$$ {#eq-11}

describes the sampling error.

We can then specify the following model to obtain the means of each arm of the trials as follows:

$$ \begin{aligned}
    Υ_i &=
    \begin{cases}
        0 & MedControl:  b_0 + u_i + \epsilon_i\\
        1 & MedActive:   b_0 + b_{1_{i}} +  u_i + \epsilon_i \\
        2 & PsyActive:   b_0 + b_{2_{i}}  +  u_i + \epsilon_i \\
        3 & PsyControl:  b_0 + b_{3_{i}}  +  u_i + \epsilon_i \\
    \end{cases}
\end{aligned}$$ {#eq-12}

where to obtain the mean of each level is the sum of $b_0$, the intercept for the reference category of medication control, with the coefficient of each level, e.g. for level 3, $b_{3_{i}}$ the psychotherapy controls. The confidence intervals of the means are constructed in the standard way using the standard errors of the mean. Similarly, each coefficient represents the contrast between the reference category and each level, for an example and of main interest to us $b_{3_{i}}$ represents the contrast between psychotherapy and medication control arms. Inference on the contrasts is done as follows:

$$
z = \frac{\hat{\beta}}{\text{SE}(\hat{\beta})}
$$ {#eq-13}

We used maximum likelihood (ML) to estimate model and applied Hartung-Knapp adjustment to reduce the chance of false positives (NOTE: **Charlotte**cite Ioannides on this).

We present our main results as means of the estimates across simulated datasets, for example, the SMDs of each level of the dummy variable above are means across the simulations.

### Sensitivity Analyses

We conducted a number of additional analyses to test for the robustness of our results.

First, we tested whether the simulated values for the standard error had a substantial influence on the estimation of the differences between the medication vs psychotherapy control condition. To inspect whether this is the case, we plotted the z-value of the difference between the two coefficients against the number of simulations. We make inference on the stability of the difference, by counting the proportion of times that the z-value is above the

```{r, warning=FALSE, message = FALSE, echo = FALSE}
# my alpha for the z 
target_probability <- 1 - 0.05

#start here 
x <- 1

# two decimals should be fine
tolerance <- 1e-3

#find the z value using the pnorm function
while (abs(pnorm(x)) < target_probability ) {
  x <- x + tolerance  
}

# Adjust the step size based on your problem  
cat("critical value of z =", x, "corresponding to an alpha = 0.05." , "\n")

```

Second, we restricted the metaregression to studies using CBT and the FDA-approved antidepressants fluoxetine and escitalopram.

Third, we restricted the metaregression to studies that *did not* use waitlist as their control.

Fourth, we tested whether our results could be accounted for by differential regression to the mean. In an initial step, we tested in metaregression as above for differences in baseline means. For this we used a min-max normalisation to turn each study arm mean score into a percentage using the following formalism:

$$
\text{outcome}_{percent} = \frac{\bar{X} - \text{scale}_{min}}{\text{scale}_{max} - \text{scale}_{min}}
$$ {#eq-14}

where, $$\bar{X}$$ is the mean score of each study arm on the primary questionnaire, and ${scale}_{min}$ and ${scale}_{max}$ are the minimum and maximum possible value of the scale in question, and for which the the standard deviation is calculated thus:

$$
\text{SD}_{Xpercent} = \frac{\text{SD}_{X}}{\text{scale}_{max} - \text{scale}_{min}}
$$ {#eq-15}

where $\text{SD}_{X}$ is the original standard deviation of the mean for X. Applying equations (12) and (13) on the baseline and outcome means and standard deviations of each study allowed us to then calculate SMDs and standard errors using equations (4) and (5)

# Results

Studies included in the metaregression.

The data for the studies included in this metanalysis are summarised in Supplementary Table 1 and are also available as a csv dataframe on \[<https://github.com/transatlantic-comppsych/apples_oranges>\]. <!--# Charlotte, this should essentially be our df_long dataset, right? Perhaps, for the puproses of the table we can get rid of some of the columns. It would be good to have it in the way Pim had it. CB - I'm not sure I entirely understand. df_long_for_metan is already quite succinct (6 columns). I wonder if you are thinking of df_appl_v_orange which is more complete and includes means and sds etc. I'd say that's probably more appropriate to make available. I agree we should simplify it down a bit. Happy to do that once you confirm I'm on the right track here. If we decide to format it to be consistent with Cuijpers' data, we would have one row per comparison, rather than the long format we have in df_long_for_metan. -->

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# Get Ns, run metareg, prepare graphing ------------------------------------------------------
library(tidyverse)
library(metafor)

# add a new multilevel variable to the simulated data for the regression

for(i in 1: length(list_df_simulated)){

list_df_simulated[[i]]$arm_effect_size <- factor(list_df_simulated[[i]]$arm_effect_size)
  list_df_simulated[[i]] <- list_df_simulated[[i]] %>% 
    mutate(four_level_var = case_when(psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
                                      psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control")
    )

  list_df_simulated[[i]]$four_level_var <- factor(list_df_simulated[[i]]$four_level_var) # turn to factor

  # relevel so that medication control becomes the reference category for the regression
#  list_df_simulated[[i]]$four_level_var <- relevel(list_df_simulated[[1]]$four_level_var, ref = "medication_control")
list_df_simulated[[i]]$four_level_var <-   fct_relevel(list_df_simulated[[i]]$four_level_var,
                                                       "medication_control",
                                                       "medication_active", 
                                                       "psychotherapy_control",
                                                       "psychotherapy_active")
}

# check it worked
# list_df_simulated[[2]] %>% 
# dplyr:: select(psy_or_med, arm_effect_size, four_level_var)

######## NOW RUN METAREGRESSIONS AND GET Coefficients for 
# A) the overall sample
# B) the CBT vs Fluox and Escitalopram sample
# C) the no WL sample
# D) the no WL or CAU sample
# E) the CDRS only sample
# F) percentage variables

# A. Estimate Means and CIs for the overall ----------------------------------

# count studies
# count the number of studies
n_unique_studies <- length(unique(df_long_for_metan$new_study_id)) # CHARLOTTE please check
df_count_studies <- count_studies(df_long_for_metan ) # CHARLOTTE please check
# CB checked both, both correct.

# specify model
model_1 <- as.formula(~ four_level_var)

# run metareg function 
# here you can use the simulated results directly
list_model_1_meta_reg <- run_metaregression(list_df_simulated, model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_overall <-  aggregate_model_results(list_model_1_meta_reg, condition)
aggregate_results_overall <- cbind(aggregate_results_overall, condition)

# extract SMDs from the  list
list_dummy_var_means <-  lapply(list_model_1_meta_reg, extract_coefficients_func)

# calculate mean SMDs and ses
df_mean_coefs_from_sim <- calculate_mean_coefs_ses(list_dummy_var_means, condition)

df_mean_coefs_from_sim <- data.frame(cbind(condition,df_mean_coefs_from_sim))

# add the ns
df_mean_coefs_from_sim <- data.frame(cbind(df_mean_coefs_from_sim, n = df_count_studies[df_count_studies$is_missing == FALSE, ]$n))

# B. Estimate Means and CIs for CBT Fluox Esc --------------------------------
#  grab ids for those studies
cbt_fluox_esc_study_ids <-
  df_appl_v_orange[df_appl_v_orange$active_type == "cbt" | df_appl_v_orange$active_type == "Fluoxetine" | df_appl_v_orange$active_type == "Escitalopram",]$new_study_id

# Use those ids to subset all dataframes in the list
# By searching for IDs we have mistakenly included multiple active arms. E.g. for Atkinson which has dulox and fluox. 
# So we need to further subset the simulated dfs. For the active arms, I am filtering out conditions that do not meet our criteria. 
# This leaves all the control conditions from trials where the active is cbt, fluox or esc. 
# We may have multiple controls per study id for this reason, though this isn't a problem 
# conceptually.

list_cbt_fluox_esc_study <- lapply(list_df_simulated, function(df) {
  df %>%
    filter(new_study_id %in% cbt_fluox_esc_study_ids) %>%
    filter(!(arm_effect_size == "cohens_d_active" & !(type %in% c("Fluoxetine", "Escitalopram", "cbt"))))
})

# count the number of studies
cbt_fluox_esc_study <- df_long_for_metan %>%
  filter(new_study_id %in% cbt_fluox_esc_study_ids,
         !(arm_effect_size == "cohens_d_active" & !(type %in% c("Fluoxetine", "Escitalopram", "cbt")))) 
study_count_cbt_fluox_esc_study <- count_studies(cbt_fluox_esc_study)
# CB checked, looks good now.

# apply the metareg function to the list of cbt, fluox, esc
model_1 <- as.formula(~ four_level_var)
list_model_1_cbt_fluox_esc_study <- run_metaregression(list_cbt_fluox_esc_study , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_cbt_fluox_esc_study <-  aggregate_model_results(list_model_1_cbt_fluox_esc_study, condition)
aggregate_results_cbt_fluox_esc_study <- cbind(aggregate_results_cbt_fluox_esc_study, condition)

# extract SMDs from the list for cbt fluox escit
means_cbt_fluox_esc_study <-  lapply(list_model_1_cbt_fluox_esc_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_cbt_fluox_esc_study <- calculate_mean_coefs_ses(means_cbt_fluox_esc_study, condition)
 coef_and_se_means_cbt_fluox_esc_study <- data.frame(cbind(condition, coef_and_se_means_cbt_fluox_esc_study))

# add the ns to the dataframe
 coef_and_se_means_cbt_fluox_esc_study$n  <- study_count_cbt_fluox_esc_study[study_count_cbt_fluox_esc_study$is_missing ==F,]$n

# C. Estimate Means and CIs without WL ---------------------------------------

 # grab the ids for no waitlist
no_wl_ids <-df_appl_v_orange[df_appl_v_orange$control_type != "wl",]$new_study_id

# use them to loop over the simulated list to subset all the dataframes.

list_no_wl <- lapply(list_df_simulated, function(df) {
  df %>%
    filter(new_study_id %in% no_wl_ids) %>%
    filter(!(arm_effect_size == "cohens_d_control" & type == "wl")) 
})

# count studies 
no_wl_studies <- df_long_for_metan %>% 
  filter(new_study_id %in% no_wl_ids) %>%
  filter(!(arm_effect_size == "cohens_d_control" & type == "wl"))
study_count_no_wl_study <- count_studies(no_wl_studies)
# CB checked, looks good!

# use this model as above
model_1 <- as.formula(~ four_level_var)

# apply the metareg function to the list of no wl
list_model_1_no_wl <- run_metaregression(list_no_wl , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_no_wl <-  aggregate_model_results(list_model_1_no_wl, condition)
aggregate_results_no_wl <- cbind(aggregate_results_no_wl, condition)

# now use the function to extract the coefficients
means_no_wl <-  lapply(list_model_1_no_wl, extract_coefficients_func)

# run the mean function for no wl
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_no_wl <- calculate_mean_coefs_ses(means_no_wl, condition)
coef_and_se_means_no_wl <- data.frame(cbind(condition,coef_and_se_means_no_wl))

coef_and_se_means_no_wl$n <- study_count_no_wl_study[study_count_no_wl_study$is_missing 
                                                     ==FALSE,]$n

# D. Estimate Means and CIs without WL and CAU --------------------------------------- 
 # grab the ids for no waitlist and cau
no_wl_cau_ids <- df_appl_v_orange %>%
  filter(!(control_type %in% c("wl", "cau"))) %>%
  pull(new_study_id)

# use them to loop over the simulated list to subset all the dataframes.

list_no_wl_cau <- lapply(list_df_simulated, function(df) {
  df %>%
    filter(new_study_id %in% no_wl_cau_ids) %>%
    filter(!(arm_effect_size == "cohens_d_control" & type %in% c("wl", "cau")))
})

# count studies 
no_wl_cau_studies <- df_long_for_metan %>% 
  filter(new_study_id %in% no_wl_cau_ids) %>%
  filter(!(arm_effect_size == "cohens_d_control" & type %in% c("wl", "cau")))
study_count_no_wl_cau_study <- count_studies(no_wl_cau_studies)
# CB checked, looks good!

# use this model as above
model_1 <- as.formula(~ four_level_var)

# apply the metareg function to the list of no wl or cau
list_model_1_no_wl_cau <- run_metaregression(list_no_wl_cau , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_no_wl_cau <- aggregate_model_results(list_model_1_no_wl_cau, condition)
aggregate_results_no_wl_cau <- cbind(aggregate_results_no_wl_cau, condition)

# now use the function to extract the coefficients
means_no_wl_cau <-  lapply(list_model_1_no_wl_cau, extract_coefficients_func)

# run the mean function for no wl
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the condition argument
coef_and_se_means_no_wl_cau <- calculate_mean_coefs_ses(means_no_wl_cau, condition)
coef_and_se_means_no_wl_cau <- data.frame(cbind(condition,coef_and_se_means_no_wl_cau))

coef_and_se_means_no_wl_cau$n <- study_count_no_wl_cau_study[study_count_no_wl_cau_study$is_missing 
                                                     ==FALSE,]$n

# here are all three of them OVERALL, CBT and WL together. We may want to display them together. 
# df_all_coefficients <- cbind(df_mean_coefs_from_sim, coef_and_se_means_cbt_fluox_esc_study,coef_and_se_means_no_wl)

# E. CDRS Sensitivity Analysis ---------------------------------------

cdrs_study_ids <-
  df_appl_v_orange[df_appl_v_orange$instrument_name == "cdrs",]$new_study_id

# use those ids to subset all dataframes in the list
list_cdrs_study <- list()

for(i in 1: length(list_df_simulated)){

  list_cdrs_study[[i]] <- list_df_simulated[[i]][list_df_simulated[[i]]$new_study_id
                                                          %in% cdrs_study_ids,]

}

# count the number of studies
study_count_cdrs_study <- count_studies(df_long_for_metan[df_long_for_metan$new_study_id %in% cdrs_study_ids,])
# CB checked, correct.

# apply the metareg function to the list of cdrs studies
model_1 <- as.formula(~ four_level_var)
list_model_1_cdrs_study <- run_metaregression(list_cdrs_study , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_cdrs_study <-  aggregate_model_results(list_model_1_cdrs_study, condition)

# extract SMDs from the list for cdrs
means_cdrs_study <-  lapply(list_model_1_cdrs_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_cdrs_study <- calculate_mean_coefs_ses(means_cdrs_study, condition)
coef_and_se_means_cdrs_study <- data.frame(cbind(condition, coef_and_se_means_cdrs_study))

# add the ns to the dataframe
coef_and_se_means_cdrs_study$n  <- study_count_cdrs_study[study_count_cdrs_study$is_missing ==F,]$n

# F. Estimate Means and CIs for the overall sample using baseline percentage variable ----------------------------------

# the first thing to do is to create a new dataset with the percentage cohen's d

list_df_simulated_percentage <- list_df_simulated

for(i in 1: length(list_df_simulated_percentage)){

  list_df_simulated_percentage[[i]]$cohens_d <- 
            (list_df_simulated_percentage[[i]]$perc_post_mean - list_df_simulated_percentage[[i]]$perc_baseline_mean)/
    ((list_df_simulated_percentage[[i]]$perc_post_sd + list_df_simulated_percentage[[i]]$perc_baseline_sd)/2)

}

# specify model
model_1 <- as.formula(~ four_level_var )

# run metareg function 
# here you can use the simulated results directly
list_model_1_meta_reg_perc <- run_metaregression(list_df_simulated_percentage, model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_perc <-  aggregate_model_results(list_model_1_meta_reg_perc, condition)
aggregate_results_perc <- cbind(aggregate_results_perc, condition)

# extract SMDs from the  list
means_perc <-  lapply(list_model_1_meta_reg_perc, extract_coefficients_func)

# calculate mean SMDs and ses
df_mean_coefs_perc <- calculate_mean_coefs_ses(means_perc, condition)
df_mean_coefs_perc <- data.frame(cbind(condition,df_mean_coefs_perc))

# add the ns
df_mean_coefs_perc <- data.frame(cbind(df_mean_coefs_perc, n = df_count_studies[df_count_studies$is_missing == FALSE, ]$n))

##### now for the baseline controlled variable
# specify model
model_2 <- as.formula(~ four_level_var + perc_baseline_mean )

# run metareg function 
# here you can use the simulated results directly
list_model_2_meta_reg_perc <- run_metaregression(list_df_simulated_percentage, model_2)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_perc_mod_2 <-  aggregate_model_results(list_model_2_meta_reg_perc, condition)
aggregate_results_perc_mod_2 <- cbind(aggregate_results_perc_mod_2, condition)

# extract SMDs from the  list
means_perc_mod_2 <-  lapply(list_model_2_meta_reg_perc, extract_coefficients_func)

# calculate mean SMDs and ses
df_mean_coefs_perc_mod_2 <- calculate_mean_coefs_ses(means_perc_mod_2, condition)
df_mean_coefs_perc_mod_2 <- data.frame(cbind(condition,df_mean_coefs_perc_mod_2))

# add the ns
df_mean_coefs_perc_mod_2 <- data.frame(cbind(df_mean_coefs_perc_mod_2, n = df_count_studies[df_count_studies$is_missing == FALSE, ]$n))

# Plots  -----------------------
# 
# #define groups
# df_long_for_metan <- df_long_for_metan %>%
#   mutate(
#     group = case_when(
#       psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "Medication Active",
#       psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "Medication Control",
#       psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "Psychotherapy Active",
#       psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "Psychotherapy Control",
#       TRUE ~ NA_character_  
#     )
#   )

# group_colors <- c("Medication Active" = "deeppink1", "Medication Control" = "deeppink4", "Psychotherapy Active" = "steelblue1", "Psychotherapy Control" = "steelblue3")
# 
# # Plot cohens d per arm
# 
# ggplot(df_long_for_metan, aes(x = group, y = cohens_d, color = group)) +
#   geom_point(position = position_jitter(width = 0.2), size = 3) +
#   geom_segment(x = 4.5, xend = 4.5, y = -3.5, yend = -3, 
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -3, label = "Less Effective", color = "grey", vjust = 0, hjust = .2) +
#   geom_segment(x = 4.5, xend = 4.5, y = -4, yend = -4.5, 
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -4.5, label = "More Effective", color = "grey", vjust = 0, hjust = .2) +
#   scale_color_manual(values = group_colors) +
#   labs(title = "Scatter plot of Cohen's d by group", x = NULL, y = "Cohen's d") +
#   theme_minimal() +
#   theme(legend.position = "none")
# 
# # Now look at change scores of percentage transformed variables
# 
# df_long_for_metan <- df_long_for_metan %>% 
#   mutate(perc_change = perc_post_mean - perc_baseline_mean)
# 
# ggplot(df_long_for_metan, aes(x = group, y = perc_change, color = group)) +
#   geom_point(position = position_jitter(width = 0.2), size = 3) +
#   geom_segment(x = 4.5, xend = 4.5, y = -0.3, yend = -0.25, 
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -0.3, label = "Less Effective", color = "grey", vjust = 0, hjust = .3) +
#   geom_segment(x = 4.5, xend = 4.5, y = -0.4, yend = -0.45, 
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -0.4, label = "More Effective", color = "grey", vjust = 0, hjust = .3) +
#   scale_color_manual(values = group_colors) +
#   labs(title = "Scatter plot of change scores (percentage transformed) by group", x = NULL, y = "Change") +
#   theme_minimal() +
#   theme(legend.position = "none")
# 
# # # Plot percentage  baseline means 
# # 
# # ggplot(df_long_for_metan, aes(x = group, y = perc_baseline_mean, color = group)) +
# #   geom_point(position = position_jitter(width = 0.2), size = 3) +
# #   scale_color_manual(values = group_colors) +
# #   labs(title = "Scatter plot of % baseline mean by group", x = NULL, y = "Percentage Baseline Mean") +
# #   theme_minimal() +
# #   theme(legend.position = "none")
# 
# # Box plots
# 
# ggplot(df_long_for_metan, aes(x = group, y = cohens_d, fill = group)) +
#   geom_boxplot(outlier.shape = NA, position = position_dodge(0.8), color = "black", alpha = 0.5) +
#   scale_y_continuous(limits = c(-3.5, 1), breaks = seq(-4, 1, by = 1)) +
#   geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
#   geom_segment(x = 4.5, xend = 4.5, y = -2.5, yend = -2, 
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -2.2, label = "Less Effective", color = "grey", vjust = 0, hjust = .2) +
#   geom_segment(x = 4.5, xend = 4.5, y = -3, yend = -3.5, 
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -3.2, label = "More Effective", color = "grey", vjust = 0, hjust = .2) +
#   scale_fill_manual(values = group_colors) +
#   labs(title = "Box plot of cohens_d by group", x = "Group", y = "Cohen's d") +
#   theme_minimal() +
#   theme(legend.position = "none")

```

### Analysis of all studies

#### Number and type of of studies included

In total, there were `r length(unique(df_long_for_metan$new_study_id))` studies which included `r df_count_studies[df_count_studies$condition == "medication_active"& df_count_studies$is_missing == FALSE,]$n` active arms and `r df_count_studies[df_count_studies$condition == "medication_control"& df_count_studies$is_missing == FALSE,]$n` control arms of antidepressant trials; and `r df_count_studies[df_count_studies$condition == "psychotherapy_active"& df_count_studies$is_missing == FALSE,]$n` active arms and `r df_count_studies[df_count_studies$condition == "psychotherapy_control"& df_count_studies$is_missing == FALSE,]$n` control arms from psychotherapy trials. Note that the number of active and control arms does not exactly match because some studies feature more than one control or active arm. There were also missing data for `r df_count_studies[df_count_studies$condition == "medication_active"& df_count_studies$is_missing == TRUE,]$n`, `r df_count_studies[df_count_studies$condition == "medication_control"& df_count_studies$is_missing == TRUE,]$n`, `r df_count_studies[df_count_studies$condition == "psychotherapy_active"& df_count_studies$is_missing == TRUE,]$n`, and `r df_count_studies[df_count_studies$condition == "psychotherapy_control"& df_count_studies$is_missing == TRUE,]$n` trial arms for medication active, medication control, psychotherapy active, and psychotherapy control conditions respectively, as the data needed to calculate the SMD was missing and could not be imputed by any of the methods outlined above.

<!--# Charlotte, for the paragraph below, please provide the numbers for each of the numbers wherever I have put an X. CB - done. I removed psychodynamic because there weren't any active conditions of this type. -->

Placebo was the control condition for all medication trials; the active arm ranged from serotonin reuptake inhibitors, such as 6 fluoxetine and 2 escitalopram, to tricylics, such as 2 nortriptyline. In psychological trials, the control arm included 14 WL controls, 25 care as usual and several other conditions such as 4 attention control conditions; the active arm included 43 CBT and 8 IPT amongst others. All included trials and the types of treatment controls can be found in Suppelmenateray Table 1.

#### Metaregression analysis.

We applied metaregression to obtain the SMDs and confidence intervals of each of the four arms.

As can be seen in Figure 1, there were substantial differences between the four arms of the metanalysis with striking difference being the medication and the psychotherapy control arms. In particular, placebo had an SMD = `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$coef_means, 2)`, 95% CI: `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$lower_ci, 2)` to `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$upper_ci, 2)`, whereas psychotherapy controls had an SMD = `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$coef_means, 2)` , 95% CI: `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$lower_ci, 2)` to `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$upper_ci, 2)`.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
subtitle_text_all <- "Metanalytic estimates of within-group changes: all studies"
plot_means_all <- plot_means_function(df_mean_coefs_from_sim, subtitle_text_all)
#print(plot_means_all)
pdf(file = "plot_means_all.pdf")
plot_means_all
```

![Figure 1.](plot_means_all.pdf)

Table 1 summarises the statistics of the estimated SMDs. <!--# Charlotte, we need to turn these into nice tables with proper columns, rounding etc. -->

```{r, warning=FALSE, message = FALSE, echo=FALSE}

knitr:: kable(df_mean_coefs_from_sim, caption = "Table 1")
```

In Table 2, we present the regression that tests our hypothesis about differences between medication and psychotherapy control. In particular, in this metaregression model, medication control is the reference category (termed intercept) to which all other categories of the dummy variable, including psychotherapy control, are compared. The strongest difference betwen arms, as judged by the z-value, is between the psychotherapy and medication control with a z-value `r round(aggregate_results_overall[aggregate_results_overall$condition == "psychotherapy_control", ]$z_value, 2)` , which yields a very low p-value ( `r sprintf("%.3e",(1-(pnorm(8.949921))))` ).

```{r, warning=FALSE, message = FALSE, echo=FALSE}
knitr:: kable(aggregate_results_overall, caption = "Table 2", digits = 2
              )
```

### Sensitivity Analyses

We then conducted a series of sensitivity analyses of our results.

#### Effect of standard errors of the SMDs .

It could be argued that the choice of standard errors of the changes for the calculation of the confidence intervals could have affect the results in one or the other direction. To address such concerns we have simulated 1000 different datasets with SMDs coming from a broad distribution. If standard error distributions were influential, this should show up as substantial variability across simulations. We test this idea in the Figure 1 which displays across the 1000 simulations the z-value of the contrast between medication and psychotherapy control arms (the mean of which we presented in Table 2). As can be seen, the variability in the z-score is minimal and consistently far away from the threshold for significance, i.e. the value of z = 1.645.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# examine the stability of the simulations

z_psy_con_vs_med_con <- 0
for(i in 1: length(list_model_1_meta_reg)){
z_psy_con_vs_med_con[i] <- list_model_1_meta_reg[[i]]$zval[condition=="psychotherapy_control"]

}

stability_sims <- data.frame(n_sims = 1:1000, z_value = z_psy_con_vs_med_con)

stab_sims <- stability_sims %>%
  ggplot(aes(x = n_sims, y = z_value)) +
  geom_point() +
  geom_hline(yintercept = 1.65, linetype = "dotted", color = "red") +
  geom_segment(aes(x = 300, xend = 300, y = 2, yend = 1.65),
               arrow = arrow(length = unit(0.3, "cm")),
               color = "black") +
  annotate("text", x = 300, y = 2.0, label = "p < 0.05 threshold (values above line significant)", hjust = -0.1, vjust = 0) +
  geom_segment(aes(x = 300, xend = 300, y = 7.5, yend = min(stability_sims$z_value-0.3)),
               arrow = arrow(length = unit(0.3, "cm")),
               color = "black") +
  annotate("text", x = 300, y = 7.5, label = "simulated z-values",
           hjust = -0.1, vjust = 0) +
  theme_minimal() +
  labs(
    x = "Number of Simulations",
    y = "z-value",
    title = "Stability of the Statistic of the Difference between \nMedication and Psychotherapy Control Arms",
    subtitle = "Results from 1000 simulations with within-group standard errors"
  )

pdf(file = "stab_sims.pdf")
stab_sims
dev.off()
```

![](stab_sims.pdf)

#### Comparison of CBT with fluoxetine and escitalopram

We next compared the control and active arms of CBT studies to those of fluoxetine and escitalopram studies. As can be seen, the pattern of results is very similar to that of the ovarall analyses.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
subtitle_text_cbt <- "Metanalytic estimates of within-group changes: cbt, fluox, escit"
plot_means_cbt <- plot_means_function(coef_and_se_means_cbt_fluox_esc_study , subtitle_text_cbt)
pdf(file = "plot_means_cbt.pdf")
plot_means_cbt
dev.off()
```

![](plot_means_cbt.pdf)

#### Excluding studies with waitlist control arms.

We next analysed the data after excluding waitlist control studies. As can be seen, the pattern of results is very similar to that of the overall analyses.

```{r, warning=FALSE, message = FALSE, echo=FALSE}

subtitle_text_no_wl <- "Metanalytic estimates of within-group changes: wl excluded"
plot_means_wl <- plot_means_function(coef_and_se_means_no_wl, subtitle_text_no_wl)
#print(plot_means_wl)

pdf(file = "plot_means_wl.pdf")
plot_means_wl
dev.off()
```

![](plot_means_wl.pdf)

#### Addressing Regression to the Mean

We first conducted a meta-analysis to test for differences in means at baseline in the two instruments, CDRS-R and HAM-D on which there was a sufficient number of studies to metanalyse.

```{r, warning=FALSE, message = FALSE, echo=FALSE}

### do ALL studies using percentage at baseline
all_studies_baseline_metareg <- list_df_simulated_percentage[[1]]

#turn sds to sec
all_studies_baseline_metareg$baseline_stand_error <- all_studies_baseline_metareg$perc_baseline_sd/
                                        sqrt(all_studies_baseline_metareg$baseline_n)

model_1 <- as.formula(~ four_level_var)
condition <-  levels(list_df_simulated[[1]]$four_level_var) # this is the four level var 
                                                              #throughout
# also change this for the function to work
all_studies_baseline_metareg$baseline_mean <- all_studies_baseline_metareg$perc_baseline_mean
results_all_studies_metareg_baseline <- run_metareg_and_extract(all_studies_baseline_metareg,  
                                                         model_1, condition)

### get CDRS studies
cdrs_studies_for_baseline_metareg <- list_cdrs_study[[1]] # created this above, you can take any of these, as the means and sds are repeated.

#turn sds to sec
cdrs_studies_for_baseline_metareg$baseline_stand_error <- cdrs_studies_for_baseline_metareg$baseline_sd/
                                        sqrt(cdrs_studies_for_baseline_metareg$baseline_n)

model_1 <- as.formula(~ four_level_var)
condition <-  levels(list_df_simulated[[1]]$four_level_var) # this is the four level var 
                                                              #throughout

results_cdrs_metareg_baseline <- run_metareg_and_extract(cdrs_studies_for_baseline_metareg,  
                                                         model_1, condition)

### get HAM-D studies
hamd_study_ids <-
  df_appl_v_orange[df_appl_v_orange$instrument_name == "hamd",]$new_study_id

list_hamd_study <- list_df_simulated[[1]][list_df_simulated[[1]]$new_study_id
                                                          %in% hamd_study_ids,]

hamd_studies_for_baseline_metareg <- list_hamd_study #

#turn sds to sec
hamd_studies_for_baseline_metareg$baseline_stand_error <- hamd_studies_for_baseline_metareg$baseline_sd/
                                        sqrt(hamd_studies_for_baseline_metareg$baseline_n)

model_1 <- as.formula(~ four_level_var)
condition <-  levels(list_df_simulated[[1]]$four_level_var) # this is the four level var 
                                                              #throughout
results_hamd_metareg_baseline <- run_metareg_and_extract(hamd_studies_for_baseline_metareg, 
                                                         model_1, condition)

results_hamd_metareg_baseline$condition <-  levels(list_df_simulated[[1]]$four_level_var)

```

As can be seen in Table 3a and 3b, there baseline scores in medication studies were on average substantially higher for the CDRS (10.9 points for the control arms) and the HAM-D (7.3 points for the control arms), respectively. As can be seen in Supplementary Tables XX <!--# Charlotte, please create these out of results_cdrs_metareg_baseline[[1]]  and the same for HAM-D--> the differences were significant at p\<0.05 for CDRS, thought they did not reach statistical significance for the HAM-D (p = 0.0573). <!--# Charlotte, please add the n for each row. Easiest to modify the function so that it appends the baseline n together with the condition. -->

```{r, warning=FALSE, message = FALSE, echo=FALSE}
knitr:: kable(results_cdrs_metareg_baseline[[2]], caption="Table 3a CDRS baselilne scores")
```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
knitr:: kable(results_hamd_metareg_baseline[[2]], caption="Table 3b HAM-D baselilne scores")
```

We next addressed the e

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#### Let's start from the beginning and regress cohens on the baseline dataframe.

# # create a new list with dfs for the sims
# df_long_for_metan_adj_cohens_d <- df_long_for_metan
# # rename the cohens_d value to make space for the new cohen's d
# df_long_for_metan_adj_cohens_d <-
#   df_long_for_metan_adj_cohens_d %>% 
# rename(unadjusted_cohens_d = cohens_d,
# ) 
# # estimate the cohen's d from the residuals
# model_adjust_cohens_d <-  lm(df_long_for_metan_adj_cohens_d$unadjusted_cohens_d ~ df_long_for_metan_adj_cohens_d$baseline_mean, na.action=na.exclude)
# 
# # there is a great effect of the baseline means, as shown here
# results_adjusted_cohens_d <- broom:: tidy(model_adjust_cohens_d )
# 
# # extract the residuals which are the adjusted cohen's d, to get the simulated standard errors
# df_long_for_metan_adj_cohens_d$cohens_d <- residuals(model_adjust_cohens_d )
# 
# # apply function to simulate standard errors
# list_df_simulated_adj_cohens_d <- simulate_dataframes_for_st_errors(df = df_long_for_metan_adj_cohens_d, 
#                                    num_repetitions = 1000, 
#                                    seed  = 1974,
#                                    n = length(unique(df_long_for_metan_adj_cohens_d$new_study_id)), 
#                                    a = 0.45, 
#                                    b = .9, 
#                                    mean = 0.65, 
#                                    sd = 0.2)
# 
# 
# # add the four level variable and re-level
# for(i in 1: length(list_df_simulated_adj_cohens_d)){
#   
#   list_df_simulated_adj_cohens_d[[i]]$arm_effect_size <- factor(list_df_simulated_adj_cohens_d[[i]]$arm_effect_size)
#   list_df_simulated_adj_cohens_d[[i]] <- list_df_simulated_adj_cohens_d[[i]] %>% 
#     mutate(four_level_var = case_when(psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
#                                       psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
#                                       psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
#                                       psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control")
#     )
#   
#   list_df_simulated_adj_cohens_d[[i]]$four_level_var <- factor(list_df_simulated_adj_cohens_d[[i]]$four_level_var) # turn to factor
#   
#   # relevel so that medication control becomes the reference category for the regression
#   #  list_df_simulated[[i]]$four_level_var <- relevel(list_df_simulated[[1]]$four_level_var, ref = "medication_control")
#   list_df_simulated_adj_cohens_d[[i]]$four_level_var <-   fct_relevel(list_df_simulated_adj_cohens_d[[i]]$four_level_var,
#                                                          "medication_control",
#                                                          "medication_active", 
#                                                          "psychotherapy_control",
#                                                          "psychotherapy_active")
# }
# 
# 
# 
# 
# 
# # Run metaregression model (assuming that the four level variable is present given that I create this dataframe from the list_)
# # specify model
# model_1 <- as.formula(~ four_level_var)
# 
# # run metareg function 
# # here you can use the simulated results directly
# list_model_1_meta_reg_adj_cohens_d <- run_metaregression(list_df_simulated_adj_cohens_d, model_1)
# 
# # extract coefficients and model characteristics.
# condition <- levels(list_df_simulated_adj_cohens_d[[1]]$four_level_var)
# aggregate_results_overall_adjusted_cohens_d <-  aggregate_model_results(list_model_1_meta_reg_adj_cohens_d, condition)
# aggregate_results_overall_adjusted_cohens_d <- cbind(aggregate_results_overall_adjusted_cohens_d, condition)
# 
# # extract SMDs from the  list
# list_dummy_var_means_adjusted_cohens_d <-  lapply(list_model_1_meta_reg_adj_cohens_d, extract_coefficients_func)
# 
# # calculate mean SMDs and ses
# df_mean_coefs_from_sim_adjusted_cohens_d <- calculate_mean_coefs_ses(list_dummy_var_means_adjusted_cohens_d, condition)
# 
# df_mean_coefs_from_sim_adjusted_cohens_d <- data.frame(cbind(condition, df_mean_coefs_from_sim_adjusted_cohens_d))
# 
# 
# 
#

```

As can be seen in Table 4, the baseline mean has a very strong association with the SMD.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# knitr:: kable(results_adjusted_cohens_d 
#               
#               , caption = "Table 4. Effect of baseline values on SMDs")
```

Table 5 shows the adjusted SMDs with their confidence intervals that we estimated in metaregression. As can be inferred from the non-overlapping confidence intervals, the difference between the two control conditions was significant (details of the regression model can be found in Supplementary Table XX <!--# Charlotte, please add there the aggregate_results_overall_adjusted_cohens_d -->). Differences were significant between active medication and psychotherapy control, but not between any of the other groups.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
knitr:: kable (df_mean_coefs_from_sim_adjusted_cohens_d, caption = "Table 5. SMDs Adjusted by baseline score" )
```

Finally, we tested whether the simulated correlation values we used for simulation of the standard errors had an appreciable effect on the values of the adjusted SMDs. Supplemental Figure X shows that the variability of the estimated z-value was minimal<!--# Charlotte, the figure below should be supplemental-->.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# 
# # plot the mean z for the difference between controls
# z_psy_con_vs_med_con <- 0
# for(i in 1: length(list_model_1_meta_reg_adj_cohens_d )){
#   z_psy_con_vs_med_con[i] <- list_model_1_meta_reg_adj_cohens_d [[i]]$zval[condition=="psychotherapy_control"]
#   
# }
# 
# 
# 
# stability_sims_adjusted_cohens_d <- data.frame(n_sims = 1:1000, z_value = z_psy_con_vs_med_con)
# 
# 
# stab_sims_adjusted_cohens_d <- stability_sims_adjusted_cohens_d %>%
#   ggplot(aes(x = n_sims, y = z_value)) +
#   geom_point() +
#   geom_hline(yintercept = 1.65, linetype = "dotted", color = "red") +
#   geom_segment(aes(x = 300, xend = 300, y = 2, yend = 1.65),
#                arrow = arrow(length = unit(0.3, "cm")),
#                color = "black") +
#   annotate("text", x = 300, y = 2.0, label = "p < 0.05 threshold (values above line significant)", hjust = -0.1, vjust = 0) +
#   annotate("text", x = 300, y = stability_sims_adjusted_cohens_d$z_value[1] - 0.1, label = "simulated z-values",
#            hjust = -0.1, vjust = 0) +
#   theme_minimal() +
#   labs(
#     x = "Number of Simulations",
#     y = "z-value",
#     title = "Stability of the Statistic of the Difference between \nMedication and Psychotherapy Control Arms",
#     subtitle = "Results from 1000 simulations with within-group standard errors"
#   )
# stab_sims_adjusted_cohens_d
# pdf(file = "stab_sims_adjusted_cohens_d.pdf")
# stab_sims_adjusted_cohens_d
# dev.off()
#

```

We have also run models to take into accountregression to the mean within CDRS and HAMD.

```{# {r, warning=FALSE, message = FALSE, echo=FALSE}
# ####code to extract baseline-adjusted post means for CDRS and HAMD 
# 
#  
# 
# de_meaned_baselilne_mean_cdrs <- (list_cdrs_study[[1]]$baseline_mean - mean(list_cdrs_study[[1]]$baseline_mean, na.rm = TRUE)) 
# 
#                                     
# 
#  
# 
#                                  
# 
# lm_for_cdrs_smds <- lm(list_cdrs_study[[1]]$post_mean ~ de_meaned_baselilne_mean_cdrs +list_cdrs_study[[1]]$four_level_var,
# 
#  
# 
#   na.action=na.exclude,
# 
#    )
# 
#  
# 
#  
# 
# #summary(lm_for_cdrs_smds)
# 
#  
# 
# coefs_lm_for_cdrs_smds  <- broom:: tidy(lm_for_cdrs_smds)
# 
#  
# 
# condition <- coefs_lm_for_cdrs_smds$term
# 
# cdrs_smds <-c(coefs_lm_for_cdrs_smds$estimate[1], coefs_lm_for_cdrs_smds$estimate[-1] + coefs_lm_for_cdrs_smds$estimate[1])
# 
# cdrs_upper_ci <- cdrs_smds + coefs_lm_for_cdrs_smds$std.error*1.96
# 
# cdrs_lower_ci <- cdrs_smds - coefs_lm_for_cdrs_smds$std.error*1.96
# 
#  
# 
# cdrs_smd_with_cis <- data.frame (
# 
#   condition = levels(list_hamd_study$four_level_var),
# 
#   cdrs_smds = cdrs_smds[-2],
# 
#   cdrs_upper_ci = cdrs_upper_ci[-2],
# 
#   cdrs_lower_ci = cdrs_lower_ci[-2]
# 
# )
# 
#  
# 
#  
# 
# # do the same for hamd
# 
# de_meaned_baselilne_mean_hamd <- (list_hamd_study$baseline_mean - mean(list_hamd_study$baseline_mean, na.rm = TRUE)) 
# 
#  
# 
#  
# 
#  
# 
# lm_for_hamd_smds <- lm(list_hamd_study$post_mean ~ de_meaned_baselilne_mean_hamd +list_hamd_study$four_level_var,
# 
#                       
# 
#                        na.action=na.exclude,
# 
# )
# 
#  
# 
#  
# 
# #summary(lm_for_hamd_smds)
# 
#  
# 
# coefs_lm_for_hamd_smds  <- broom:: tidy(lm_for_hamd_smds)
# 
#  
# 
#  
# 
# condition <- coefs_lm_for_hamd_smds$term
# 
# hamd_smds <-c(coefs_lm_for_hamd_smds$estimate[1], coefs_lm_for_hamd_smds$estimate[-1] + coefs_lm_for_hamd_smds$estimate[1])
# 
# hamd_upper_ci <- hamd_smds + coefs_lm_for_hamd_smds$std.error*1.96
# 
# hamd_lower_ci <- hamd_smds - coefs_lm_for_hamd_smds$std.error*1.96
# 
#  
# 
# hamd_smd_with_cis <- data.frame (
# 
# condition = levels(list_hamd_study$four_level_var),
# 
# hamd_smds = hamd_smds[-2],
# 
# hamd_upper_ci = hamd_upper_ci[-2],
# 
# hamd_lower_ci = hamd_lower_ci[-2]
# 
# )
# 
#  
# 
#

```

Tables 6a and 6b show the baeline-mean adjusted SMDs for the CDRS at outcome when the baseline measurement has been taken into account. <!--# Here Intercept refers to teh psychotherapy control. Sorry about the labelling of the rows, will improve. -->

```{r}
knitr:: kable(cdrs_smd_with_cis, caption = "Table 6a. CDRS baseline mean adjsuted SMDs")
```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
knitr:: kable(hamd_smd_with_cis, caption = "Table 6. HAM-D baseline-mean adjusted SMDs")
```