### 2024-03-14 apples_and_oranges.qmd
More tabulating, running t tests for control vs active comparison in psy trials

---
title: Fundamental Problems with the Evidence Base of Adolescent Depression Treatments.
subtitle: Insights from the relative efficacy of Control Conditions.
author:
  - name: Argyris Stringaris
    email: a.stringaris@ucl.ac.uk
    affiliations: 
        - id: ucl_1
          name: University College London
          department: Divisions of Psychiatry and Psychology and Language Science
          address: 1-19 Torrington Place
          city: London, UK
          state: 
          postal-code: WC1E 7HB
          country: United Kingdom
    attributes:
        corresponding: true
    note: 
  - name: Charlotte Burman
    email: charlotte.burman@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
  - name: Dayna Bhudia
    email: dayna.bhudia@ucl.ac.uk
    affiliations:
        - ref: ucl_1
  - name: Lucy Foulkes
    email: lucy.foulkes@psych.ox.ac.uk
    affiliations:
        - id: OU
          name: University of Oxford
          department: Experimental Psychology
          address: Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Road
          city: Oxford
          postal-code: OX2 6GG
          country: United Kingdom
    note:
  - name: Carmen Moreno
    email: cmoreno@hggm.es
    affiliations:
        - id: HGM
          name: Hospital Gregorio Marañón
          department: Department of Psychiatry
          address: 46 C. del Dr. Esquerdo
          city: Madrid
          state:
          postal-code: 28007
          country: Spain
  - name: Samuele Cortese
    email: samuele.cortese@soton.ac.uk
    affiliations:
        - id: soton
          name: University of Southampton
          department: Centre for Innovation in Mental Health (CIMH), School of Psychology
          address: Highfield Campus, Building 44
          city: Southampton
          state:
          postal-code: SO17 1BJ
          country: United Kingdom
  - name: Georgina Krebs
    email: g.krebs@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note: 
abstract: |
  To follow
keywords: 
  - depression
  - adolescents
  - antidepressants
  - psychotherapy
date: last-modified
bibliography: emotion_concepts_paper.bib
format: 
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
---

# Introduction

There are two principal treatment modalities for adolescent depression: psychotherapy, medication, or both. Where does one turn to find the evidence that will inform their treatment options? This question is relevant for patients and their carers, for clinicians and for policy makers when they plan services. However, the question is particularly difficult to answer for adolescent depression, where there are limited data from head-to-head trials of medication and psychotherapy, and where recommendations must therefore be derived from indirect comparisons of treatment efficacy.

Guidelines, such as the ones that the internationally influential UK National Institute of Clinical Excellence (NICE) have produced for adolescent depression, are a principal source of such information. NICE recommend that youth be first offered psychological therapy (specifically cognitive behaviour therapy and interpersonal therapy) over medication for most presentations of depression. This conclusion is in keeping with two sources of evidence. On the one hand, medication meta-analyses that cast doubt on the efficacy of most antidepressants, with the exception of fluoxetine (ref); and on the other hand, psychotherapy meta-analyses that conclude psychotherapy to be effective for adolescent depression (ref). However, such conclusions seem at odds with those of a recent network meta-analysis (ref), an established method of comparing treatments with each other using both direct (head-to-head) and indirect (indirectly treatment A with treatment C, via studies that directly compare treatments A with B and B with C) evidence. Indeed, the network meta-analysis concluded that only fluoxetine alone and fluoxetine administered together with CBT were significantly more effective than pill placebo or psychological controls. Given this confusing evidence base, how should patients, carers, clinicians and policy makers make decisions on treatment options for adolescent depression?

In this paper, we examine whether the existing evidence base for adolescent depression treatment can offer valid answers to such questions. Below, we provide a conceptual framework for answering such questions and state our hypotheses that we will test using data from existing trials. Two points are relevant to indirect comparisons of treatment modalities with each other.

First, whether the participants of trials in one modality are comparable to those in another modality. Second, whether key conditions of the trial, such as the effects of control conditions or the number of sites involved in a trial, are comparable.

Starting with the first point, to be able to compare between different trials one must assume that these trials sample from the same population. If they do not, then the validity of any comparisons, including those conducted through network metanalysis (which rests on the principal of transitivity) are questionable.

Indeed, comparing outcomes (Y) between psychotherapy and medication trials requires us to contrast what is called the sample average treatment effect $\tau$ of each, defined in the following way:

$$
\tau_{psy} =  E(Y_{1} - Y_{0} \,|\, S, psy)
$$
$$
\tau_{med} =  E(Y_{1} - Y_{0} \,|\, S, med)\ 
$$ {#eq-1}

where the operator E denotes the expectation over the differences in outcomes between those who received the intervention (T = 1) and those who received the control condition (T = 0), for each sample, $S$, where psy and med stand for psychotherapy and medication respectively.

Obviously, this comparison rests on the assumption that trials in both modalities sample from the same population, $P$, of patients. Formally, this can be expressed as follows:

$$E(Y_{1} \,|\, \text{S, psy}) = E(Y_{1} \,|\, P, \text{psy})$$

$$E(Y_{1} \,|\, \text{S, med}) = E(Y_{1} \,|\, P, \text{med})$$ {#eq-2}

signifying that the effect found in the population would be expected to be found in the same population for each treatment.

This assumption may be hard to meet. Clinical experience and empirical evidence (ref: (<https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4156137/>, <https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7669695/#:~:text=Adolescents%20also%20may%20be%20more,medications%2C%20and%20fewer%20anxiety%20symptoms.> , <https://pubmed.ncbi.nlm.nih.gov/16502131/>) indicate that patients and parents often have preferences between psychotherapy and medication, meaning that there is likely to be a self-selection bias with respect to who participates in psychotherapy versus medication trials. Moreover, these treatment preferences correlate with clinically-relevant characteristics of the participants, including severity, gender and comorbidity. Some of these characteristics, particularly baseline severity, may moderate treatment response (ref) and therefore, if they differ across psychotherapy and medication trials, they may confound comparisons.

Here we will compare between four key sample characteristics, namely baseline severity in the outcome, number of sites, gender and age in each sample.

In terms of the second point, differences in trial design may impact outcomes in a differential way between antidepressant and psychotherapy trials. Perhaps the most obvious way in which this happens is the fact that participants in psychotherapy trials are unblinded in terms of treatment allocation; by contrast, in new antidepressant trials, patients (and raters) were found to be unlikely to be able to judge treatment allocation (cite: Assessment of blinding in randomized controlled trials of antidepressants for depressive disorders 2000−2020: A systematic review and meta-analysis, Lancet). This creates differential expectations between medication and psychotherapy trials. Importantly, the placebo control has been developed in order to match as closely as possible the intervention condition, so as to minimise differences in expectancy between conditions. By contrast, psychotherapy control arms vary across waitlist controls, treatment as usual and active controls. These differences in expectation essentially favour the psychotherapy active condition and disadvantage the psychotherapy control (thus potentially the difference between them), because participants in the active condition could be content for receiving the “cutting edge treatment” whilst those in the control will be dissatisfied for having  missed out on “proper treatment”. And this is important because expectancy is associated with treatment outcomes. This also poses problems when trying to infer indirect comparisons in network metanalyses where, for example, CBT is compared to placebo control and medication — given the likely differential expectation effects, comparisons to the unblinded psychotherapy controls may be biased.

Another difference in design that has potential implications for which patients are selected into the trial is the number of sites in a trial. Previous research has shown that in medication trials the number of sites is positively related to the magnitude of the placebo response. This phenomenon is hard to explain fully with the data available, but may be due to lower quality of phenotyping in such studies, with higher measurement (and therefore diagnostic classification) errors, leading to issues such as spontaneous remission or regression to the mean. In contrast to medication trials, psychotherapy trials typically involve one or fewer sites.

An inter-related issue concerns the nature of control interventions being used. Control conditions in trials are meant to generate counterfactual conditions (CITE RUBIN et al) to the intervention: what would have been the outcome in an individual had they not received the intervention. This implies that there is a latent distribution of values that represent that counterfactual condition which control arms in trials are meant to approximate. Those latent values of the counterfactual condition should, on average, not be different between psychotherapy and medication trials, provided that the samples are representative of the same underlying population (see Eqs 1& 2). If the average response to placebo and psychotherapy controls differs systematically from each other, then this is either due to the fact that the trials sample from different populations or because the control conditions differ in their effects (e.g. because of how much they do or do not protect against expectancy effects).

Understanding the nature of controls for each treatment modality is important for a similar reason. Often, in the public domain, psychotherapy and medication are compared to each other on the basis of their respective effect sizes. However, as evident from Eqs 1 and 2, these effect sizes represent differences between the active intervention and the control condition (placebo or psychotherapy control). For these effect sizes to be comparable, placebo and psychotherapy controls ought to be the equal. Otherwise, misleading conclusions could be drawn, e.g. two effect sizes of 40% would be considered equal, even if one arose from a difference of 100% - 60% and another arose from a difference of 40% - 0%.

In this paper, we examine psychotherapy and medication trials in the following ways. First, we compare key baseline characteristics of medication vs psychotherapy trials, specifically the extent to which they are comparable in a) baseline severity; b) number of sites involved; c) gender composition and d) age. Second, we assess the mean efficacy of psychotherapy controls vs placebo. We recognise that both these sets of analyses pull data out of the randomised comparisons between treatment arms. However, we believe that this is justified for two reasons: a) because it is important to make transparent to all, particularly decision makers in healthcare, the raw data before these are entered into more complex modelling such as network metanalysis, as this allows them to form opinions about the quality and limitations of the input data; b) our aim here is not to make claims about the relative efficacy of each treatment arm, but rather establish whether the conditions for the possibility of comparisons are fulfilled. Thus, the comparisons between control conditions are not meant to lead to inferences about placebo being “more efficacious” than psychotherapy controls, but rather to make transparent the problem that they seem to derive from different distributions. Third, we examine the quality of psychotherapy controls as such, by scrutinising the extent to which they are matched to the active intervention in ways such as number or frequency of sessions, and therefore, whether they represent fair pairings from which to draw valid efficacy inferences.

@stringarisDevelopmentalPathwaysChildhood2014

# Method

## Studies included

For this review, we primarily drew upon studies included in two recent meta-analyses. Please refer to these original meta-analyses for a detailed description of study selection criteria. Psychotherapy studies were drawn from Cuijpers' (2023) systematic review and meta-analysis of randomised trials comparing psychotherapy for youth depression against control conditions. Cuijpers made available a full dataset of psychotherapy trials (via https://www.metapsy.org/), which we used for the current study. Whilst Cuijpers (2023) excluded those studies for which the primary outcome variable could not be calculated due to missing data, we included these studies and performed the imputations outlined below; hence we have more psychotherapy studies included in this review compared to Cuijpers' (2023) meta-analysis. Whilst the online database is regularly updated, we chose to exclude studies published after the final date of Cuipers' (2023) literature search.

Medication studies were drawn from Cipriani's (2016) network meta-analysis examining the efficacy and tolerability of a range of antidepressants and placebo for major depressive disorder in young people. The study dataset was made available online though did not include means or standard deviations at baseline or post-test. We emailed the study authors requesting a full dataset with this data, though did not receive a reply, and hence conducted data extraction for medication studies. We excluded three studies because they did not include a control arm. We were unable to locate and therefore complete extraction for two papers. Many studies did not report complete data, and so we emailed all corresponding authors to request missing data, though did not receive any responses. We conducted a systematic search for medication studies published after the final search date of Cipriani's (2016) review up to the final search date of Cuijpers' (2023) review to ensure we analysed an equivalently up-to-date database of medication trials. We used the same search terms outlined in Cipriani (2016).

## Statistical Analysis

### Trial and sample characteristics

We conducted a series of random effects meta-analyses and tested for subgroup differences between psychotherapy versus medication trials in sample characteristics including gender, age, and severity of depressive symptoms at baseline. Meta-analyses were implemented using R's meta package.

In order to compare depression severity across different instruments, we performed a min-max normalisation to turn each study arm mean score at baseline into a percentage using the following formalism:

$$
\text{outcome}_{percent} = \frac{\bar{X} - \text{scale}_{min}}{\text{scale}_{max} - \text{scale}_{min}}
$$ {#eq-3}

where, $$\bar{X}$$ is the mean score for each study arm on the primary outcome questionnaire, and ${scale}_{min}$ and ${scale}_{max}$ are the minimum and maximum possible values of the scale in question, respectively. The standard deviation is calculated thus:

$$
\text{SD}_{Xpercent} = \frac{\text{SD}_{X}}{\text{scale}_{max} - \text{scale}_{min}}
$$ {#eq-4}

where $\text{SD}_{X}$ is the original standard deviation of the mean at baseline.

We also conducted a t-test to compare mean number of trial sites between psychotherapy and medication trials.

### Measures of Effect

As the measure of effect of each individual study, we used the within-group Standardised Mean Difference (SMD), which we defined following NOTE: Charlotte cite Lakens 2013 and in there Cummings 2012, as:

$$SMD_{change} = \frac{Mean_{t_{2}} - Mean_{t_{1}}}{\frac{SD_{t{2}} + SD_{t{1}}}2}\ $$ {#eq-3}

where, $Mean_{t_{2}}$ and $Mean_{t_{1}}$ refer to the means of the main outcome score at the end and beginning of the intervention respectively and $SD_{t_{2}}$ and $SD_{t_{1}}$ to the respective standard deviations. Where individual studies did not report all data required to calculate the SMD, we imputed missing data according to the methods summarised by X (citation), in the following order. If a study reported the standard error of the mean, the SD was obtained simply by multiplying the SE by the square root of the sample size. For conditions where the SD was missing at one time point, the baseline SD was substituted by the post-test SD, and vice versa. If the SD was not available at either time point, missing values were replaced by the mean of the SDs available for comparable cases (defined as same trial type (psy or med), same instrument, same timepoint (pre or post), and same arm (control or active)). Where there were missing means at either baseline or post-test, missing values were calculated using mean change scores, preferring the change scores reported in the paper itself, though where this was unavailable, using the change scores reported in the dataset from Cipriani's meta-analysis (citation).

For the purposes of metanalysis, it is necessary to estimate a standard error of the SMD. This is calculated according to:

$$ SE_{SMD} = \sqrt{\frac{2(1 -r_{t_{1}t_{2}})}{n} + \frac{SMD^2}{2n}}$$ {#eq-5}

where $n$ refers to the study sample size and $r_{t_{1}t_{2}}$ refers to the correlation between the outcome score obtained at baseline and at the end point. This correlation is typically not reported in studies and is often imputed using previously reported correlations for the instruments used. However, this practice has given rise to concerns about misestimation. Whilst such misestimation is possible, there is no reason to expect that it would be systematic, i.e. bias estimation of the effects for the control group of medication compared to those of psychotherapy. Still, to alleviate such concerns we have used a simulations.

In particular, we simulated one thousand truncated distribution of standard errors with the following general characteristics:

$$r_{t_{1}t_{2}} \sim \mathcal{TN}(\mu, \sigma, a, b)$$ {#eq-6}

for which we chose the mean to be $\mu = 0.65$, the standard deviation to be $\ sigma = 0.2$, and the upper and lower bounds to be $a = 0.45$ and $b = 0.9$, respectively. We then used these simulated datasets in the subsequent metanalyses.

### Random Effects Metaregression

We estimated the pooled standardised mean difference for each arm by using a random effects metanalysis implemented in R's metafor package. The main underlying assumption of random effects metanalysis is that each study's true effect size $\theta_{k}$ is affected not only by sampling error $\epsilon_{k}$, but also by $\zeta$ which represents heterogeneity between studies, allowing each study's estimate to vary along a distribution of effects, and the distribution of true effect sizes termed $\tau^2$. Therefore, we can estimate a two stage model with:

$$Y_i \sim \mathcal{N}(\theta_i, \sigma_i^2)
$$ {#eq-7}

$$\theta_i \sim \mathcal{N}(x_i\beta, \tau_i^2)
$$ {#eq-8}

where $Y_i$ is the estimated effect size for study i, has a normal distribution with $\theta_i$ as its true mean effect and sampling error $\sigma^2$. Whereas $\theta_i$ is a study-specific instantiation of the distribution of effect sizes, with $\tau^2$ representing heterogeneity.

This then gives rise to:

$$Y_i = x_i\beta_i + u_i + \epsilon_i$$ {#eq-9}

where,

$$u_i \sim N(0, \tau^2) $$ {#eq-10}

describes the deviation of each study from the mean of the distribution, and,

$$\epsilon_i \sim N(0, \sigma^2)$$ {#eq-11}

describes the sampling error.

We can then specify the following model to obtain the means of each arm of the trials as follows:

$$ \begin{aligned}
    Υ_i &=
    \begin{cases}
        0 & MedControl:  b_0 + u_i + \epsilon_i\\
        1 & MedActive:   b_0 + b_{1_{i}} +  u_i + \epsilon_i \\
        2 & PsyActive:   b_0 + b_{2_{i}}  +  u_i + \epsilon_i \\
        3 & PsyControl:  b_0 + b_{3_{i}}  +  u_i + \epsilon_i \\
    \end{cases}
\end{aligned}$$ {#eq-12}

where to obtain the mean of each level is the sum of $b_0$, the intercept for the reference category of medication control, with the coefficient of each level, e.g. for level 3, $b_{3_{i}}$ the psychotherapy controls. The confidence intervals of the means are constructed in the standard way using the standard errors of the mean. Similarly, each coefficient represents the contrast between the reference category and each level, for an example and of main interest to us $b_{3_{i}}$ represents the contrast between psychotherapy and medication control arms. Inference on the contrasts is done as follows:

$$
z = \frac{\hat{\beta}}{\text{SE}(\hat{\beta})}
$$ {#eq-13}

We used maximum likelihood (ML) to estimate model and applied Hartung-Knapp adjustment to reduce the chance of false positives (NOTE: **Charlotte**cite Ioannides on this).

We present our main results as means of the estimates across simulated datasets, for example, the SMDs of each level of the dummy variable above are means across the simulations.

### Sensitivity Analyses

We conducted a number of additional analyses to test for the robustness of our results.

First, for each of the the analyses comparing trial and sample characteristics at baseline, and control efficacy between study arms, we performed sensitivity analyses where we restricted the included studies in the following ways. First, we excluded studies which recruited participants with subclinical levels of depression. Second, we restricted the analyses to studies using CBT and the FDA-approved antidepressants fluoxetine and escitalopram. Third, we excluded studies that used waitlist as their control condition.

Further, we tested whether the simulated values for the standard error had a substantial influence on the estimation of the differences between the medication vs psychotherapy control condition. To inspect whether this is the case, we plotted the z-value of the difference between the two coefficients against the number of simulations. We make inference on the stability of the difference, by counting the proportion of times that the z-value is above the

```{r, warning=FALSE, message = FALSE, echo = FALSE}
# my alpha for the z 
target_probability <- 1 - 0.05

#start here 
x <- 1

# two decimals should be fine
tolerance <- 1e-3

#find the z value using the pnorm function
while (abs(pnorm(x)) < target_probability ) {
  x <- x + tolerance  
}

# Adjust the step size based on your problem  
cat("critical value of z =", x, "corresponding to an alpha = 0.05." , "\n")

```

### Comparing the control versus active arms of psychotherapy trials

We ran t-tests to compare the active versus control arms of psychotherapy trials on key variables of interest regarding the nature and intensity of the interventions. We extracted data pertaining to the number, duration and intensity of sessions, and the total cumulative hours and period of the intervention. Where a range was provided, the maximum was encoded (e.g. if a paper reported that an intervention involved 8-10 sessions lasting 50-60 minutes, we encoded the number and duration of sessions as 10 and 60, respectively). If sessions varied in frequency across an intervention, we calculated an average by dividing total number of sessions by length of intervention period. Similarly, if the length of sessions varied across the course of the intervention, we calculated a weighted average. Phone call, web-chat and online sessions were encoded as sessions, however guided self-help components were not.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
###########Libraries needed
library(tidyverse)
library(truncnorm)
library(metafor)
library(meta)
library(kableExtra)
```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#############Some custom made functions#############

###### 1. Simulation Function for SEs
simulate_dataframes_for_st_errors <- function(df, num_repetitions, seed, n, a, b, mean, sd) { # n refers to the number
                                                                                              # of unique ids to which a correlation coef
                                                                                              # is allocated.
  set.seed(seed)

  # Empty list to store simulated dfs
  list_df_simulated <- list()

  for (i in 1:num_repetitions) {
    # Simulate the vector
    simulated_correlations_vector <- truncnorm::rtruncnorm(n, a, b, mean, sd)  #using the truncnorm to create correlations

    # Create a copy of the original dataframe
    df_simulated_copy <- df

    # Add/update the sims column
    df_simulated_copy$correlation_sim_values <- simulated_correlations_vector[match(df_simulated_copy$new_study_id, unique(df_simulated_copy$new_study_id))]

    # Remove duplicated columns
    df_simulated_copy <- df_simulated_copy[, !duplicated(colnames(df_simulated_copy))]

    # Calculate the ses
    df_simulated_copy <- df_simulated_copy %>% 
      group_by(new_study_id, baseline_n) %>% 
      mutate(simulated_se = sqrt(((2*(1-correlation_sim_values))/baseline_n) + 
                                   (cohens_d^2/(2*baseline_n))))

    # Add the simulated dataframe to the list
    list_df_simulated[[i]] <- df_simulated_copy
  }

  return(list_df_simulated)
}

###############2. Simulation function for metaregression
run_metaregression <- function(list_of_datasets, model_formula) {
  list_model_meta_reg <- list()

  for (i in seq_along(list_of_datasets)) {
    list_model_meta_reg[[i]] <- metafor::rma(yi = cohens_d,
                                             sei = simulated_se,
                                             data = list_of_datasets[[i]],
                                             method = "ML",
                                             mods = model_formula,
                                             test = "knha")
  }

  return(list_model_meta_reg)
}

#########2. Function to count studies with missingness etc.

count_studies <- function(df) {
  df_count_studies <- df %>%
    group_by(psy_or_med, arm_effect_size) %>%
    count(is.na(cohens_d)) %>%
    rename(is_missing = `is.na(cohens_d)`) %>%
    mutate(
      condition = case_when(
        psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
        psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
        psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
        psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control"
      )
    )

  df_count_studies$condition <- factor(df_count_studies$condition)  # turn to factor

  # relevel so that medication control becomes the reference category for the regression
  df_count_studies$condition <- relevel(df_count_studies$condition, ref = "medication_control")

  return(df_count_studies)
}

##########3. Function to extract mean coefficients and to extract model characteristics

aggregate_model_results <- function(list_model_1_meta_reg, condition) {
  df_coefs <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  df_se <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  df_z_value <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  df_lower_ci <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  df_upper_ci <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  tau_sq <- rep(0, length(list_model_1_meta_reg))
  i_sq <- rep(0, length(list_model_1_meta_reg))
  k <- rep(0, length(list_model_1_meta_reg))
  r_sq <- rep(0, length(list_model_1_meta_reg))

  for (i in 1:length(list_model_1_meta_reg)) {
    for (j in 1:length(condition)) {
      df_coefs[i, j] <- list_model_1_meta_reg[[i]]$beta[[j]]
      df_se[i, j] <- list_model_1_meta_reg[[i]]$se[[j]]
      df_z_value[i, j] <- list_model_1_meta_reg[[i]]$zval[[j]]
      df_lower_ci[i, j] <- list_model_1_meta_reg[[i]]$ci.lb[[j]]
      df_upper_ci[i, j] <- list_model_1_meta_reg[[i]]$ci.ub[[j]]
      tau_sq[i] <- list_model_1_meta_reg[[i]]$tau2
      i_sq[i] <- list_model_1_meta_reg[[i]]$I2
      k[i] <- list_model_1_meta_reg[[i]]$k
      r_sq[i] <- list_model_1_meta_reg[[i]]$R2
    }
  }

  df_coefficients_model <- data.frame(
    coefficients = colMeans(df_coefs),
    se = colMeans(df_se),
    z_value = colMeans(df_z_value),
    lower_ci = colMeans(df_lower_ci),
    upper_ci = colMeans(df_upper_ci),
    tau_sq = mean(tau_sq),
    i_sq = mean(i_sq),
    k = mean(k),
    r_sq = mean(r_sq)
  )

  return(df_coefficients_model)
}

###########4. Function to extract the SMDs  (SMD, ses, CI) for each level of the dummy variable. The output is used below in 5.
# it is a slightly awkward one, because I couldn't' come up with a better way to add the intercept to each coefficient
# given the named output of the built in coef function. here I extract the coefficients

extract_coefficients_func <- function(df_with_coefs){
  list_coefs <- list()
  coefs <- 0
  coefficient_output <- 0

  for(i in 1: length(coefficients(df_with_coefs))){

    temp_vec <- c(0, rep(coefficients(df_with_coefs)[[1]], 
                         length(coefficients(df_with_coefs)) - 1 ))
    coefs[i] <- coef(df_with_coefs)[[i]]

    coefficient_output <- temp_vec + coefs

    st_error_output <- df_with_coefs$se

    df_coefficients <- data.frame(cbind(coefficients = coefficient_output, se = st_error_output ))

    df_coefficients$lower_bound <- df_coefficients$coefficients - 1.96*df_coefficients$se
    df_coefficients$upper_bound <- df_coefficients$coefficients + 1.96*df_coefficients$se
  }
  return (df_coefficients )
}

############5. Function to get the means out of the SMDs (this is similar to Func 3 and I should at some point create one more general one)
# here the argument list_of_dfs is the list of extracted coefficients that you get out of the extract_coefficients_func
# which itself you get from the run_metaregression function. The argument conditions refers to the levels of the variable that describes each study arm

calculate_mean_coefs_ses <- function(list_of_dfs, conditions) {
  df_coefs <- matrix(NA, nrow = length(list_of_dfs), ncol = length(conditions))
  df_se <- matrix(NA, nrow = length(list_of_dfs), ncol = length(conditions))

  for (i in 1:length(list_of_dfs)) {
    for (j in 1:length(conditions)) {
      df_coefs[i, j] <- list_of_dfs[[i]]$coefficients[j]
      df_se[i, j] <- list_of_dfs[[i]]$se[j]
    }
  }

  df_mean_coefs_ses <- data.frame(
    coef_means = colMeans(df_coefs),
    se_means = colMeans(df_se)
  )

  df_mean_coefs_ses$lower_ci <- df_mean_coefs_ses$coef_means - 1.96 * df_mean_coefs_ses$se_means
  df_mean_coefs_ses$upper_ci <- df_mean_coefs_ses$coef_means + 1.96 * df_mean_coefs_ses$se_means

  return(df_mean_coefs_ses)
}

###############6. Function for plotting the SMDs
plot_means_function <- function(dataframe, subtitle) {
  dataframe$text_label <- 0

  for (i in 1:nrow(dataframe)) {
    dataframe$text_label[i] <- paste(
      dataframe$condition[i], "\n",
      round(dataframe$coef_means[i], 2),
      "[" ,
      round(dataframe$lower_ci[i], 2), 
      round(dataframe$upper_ci[i], 2),
      "]"
    ) 
  }

  # Recode condition for ease of plotting
  dataframe$condition <- str_replace_all(dataframe$condition, "_", " ")

  # Encode colours
  redish_palette <- c("medication control" = "deeppink", "medication active" = "deeppink3")
  blueish_palette <- c("psychotherapy active" = "steelblue1", "psychotherapy control" = "steelblue3")

  # Plotting
  plot_means <- ggplot(dataframe, aes(x = coef_means, y = 1:nrow(dataframe),
                                      colour = condition, label = text_label)) +
    geom_point(label=dataframe$k) +
    geom_errorbar(aes(xmin = lower_ci, xmax = upper_ci), width = 0.2, position = position_dodge(0.5)) +
    geom_text(vjust = +1.5, size = 4) +
    scale_size_continuous(guide = "none") +
    guides(colour = FALSE) + 
    scale_color_manual(values = c(redish_palette, blueish_palette)) +
    theme_minimal() +
    labs(x = "TE-random", y = NULL, title = "Adolescent Depression Trial Efficacy by Treatment Arm",
         subtitle = subtitle) +
    xlab("Standardized Mean Difference (SMD) with 95% CIs") +
    ylab("") +
    ylim(0, nrow(dataframe) + 1) +
    xlim(-3.0, 0.5) +
    geom_vline(xintercept = 0, linetype = "dashed", size = 1.5, colour = "grey") +
    geom_segment(x = -1.25, xend = -1.75, y = 4.7, yend = 4.7, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
    geom_text(x = -1.55, y = 4.9, label = "More Effective", color = "grey", vjust = 0.5, hjust = 1) +
    geom_segment(x = -1.35, xend = -0.85, y = 4.9, yend = 4.9, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
    geom_text(x = -1.05, y = 5.05, label = "Less Effective", color = "grey", vjust = 0.5, hjust = 0)+
    theme(axis.title.x = element_text(size = 12), 
          axis.text.x  = element_text(size = 12),
          axis.text.y = (element_blank()),
          plot.title = element_text(size = 12),
          plot.subtitle = element_text(size = 12)) +
    geom_curve(aes(x = 0.18, y = 4.73, xend = 0.02, yend = 4.5), color = "grey", curvature = -0.2, arrow = arrow(length = unit(0.25, "cm"), type = "closed")) +
    geom_text(x = 0.2, y = 4.75, label = "No Effect", color = "grey", vjust = 0.5, hjust = 0)

  return(plot_means)
}

###########7. Function for baseline SMD results
library(broom)
run_metareg_and_extract <- function(dataframe, model_1, condition) {
  # Run metareg
  inst_baseline_metareg <- rma(yi = baseline_mean,
                               sei = baseline_stand_error,
                               data = dataframe,
                               method = "ML",
                               mods = model_1,
                               test = "knha")

  # Extract coefficients from the list
  baseline_metareg_results <- tidy(inst_baseline_metareg)
  baseline_metareg_results$condition <- condition
  coefs_for_inst_baseline <- tidy(inst_baseline_metareg)$estimate

  # Calculate SMDs
  smds_for_inst_baseline <- c(coefs_for_inst_baseline[1], coefs_for_inst_baseline[-1] + coefs_for_inst_baseline[1])

  # Calculate upper and lower CIs
  upper_cis_for_inst_baseline <- smds_for_inst_baseline + 1.96 * tidy(inst_baseline_metareg)$std.error
  lower_cis_for_inst_baseline <- smds_for_inst_baseline - 1.96 * tidy(inst_baseline_metareg)$std.error

  # Create dataframe with results
  df_baseline_means_with_cis <- data.frame(
    baseline_smds = smds_for_inst_baseline,
    upper_cis_smds = upper_cis_for_inst_baseline,
    lower_cis_smds = lower_cis_for_inst_baseline,
    condition = condition
  )

  # Return the two dataframes
  return(list(baseline_metareg_results = baseline_metareg_results, df_baseline_means_with_cis = df_baseline_means_with_cis))
}

```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
library(tidyverse)
# Transforming data ------------------------------------------------------
# import master dataset

df_appl_v_orange <-readxl:: read_excel("df_appl_v_orange.xlsx")

# Here we add in instrument end point values (mins / maxs).
# We will use these to transform the scale into % as follows: (observed score - min of the scale)/(max-min)
# We have done this for both means and standard deviations.

df_scale_ranges <- read.csv("scale_ranges.csv")
df_appl_v_orange <- full_join(df_appl_v_orange, df_scale_ranges, by = "instrument_name")

df_appl_v_orange <- df_appl_v_orange %>% 
  mutate(perc_baseline_mean_active = (baseline_mean_active - scale_min) / (scale_max - scale_min)) %>% 
  mutate(perc_baseline_sd_active = baseline_sd_active / (scale_max - scale_min)) %>%  # this is the best way to estimate the SD of the linear transformation of the score 
  mutate(perc_baseline_mean_control = (baseline_mean_control - scale_min) / (scale_max - scale_min)) %>% 
  mutate(perc_baseline_sd_control = baseline_sd_control  / (scale_max - scale_min)) %>% 
  mutate(perc_post_mean_active = (post_mean_active - scale_min) / (scale_max - scale_min)) %>% 
  mutate(perc_post_sd_active = post_sd_active  / (scale_max - scale_min)) %>% 
  mutate(perc_post_mean_control = (post_mean_control - scale_min) / (scale_max - scale_min)) %>% 
  mutate(perc_post_sd_control = post_sd_control  / (scale_max - scale_min))

# Need to use SMDs (ie our Cohen's d) and then use standard error of SMD, to achieve this I need test-retest reliabilities of instruments.

# CDRS reliability taken from here https://www.liebertpub.com/doi/epdf/10.1089/104454601317261546 
# Using a   2-week interval, and different psychiatrists from the first to the second assessment, 
# Poz-nanski et    al. (1984) demonstrated high reliability (r=   0.86) 
# for the CDRS-R total score in 53 clinic-referred 6- to 12-year-olds.

### A few more tidying things from Argyris before doing metanalyses
# Discovered some errors in the percentage women variable. Have gone back and checked. 
df_appl_v_orange[df_appl_v_orange$study_ID=="Fristad, 2019_cbt + placebo_placebo",]$percent_women <-43.1
df_appl_v_orange[df_appl_v_orange$study=="Clarke, 1995",]$percent_women <-47.3
df_appl_v_orange[df_appl_v_orange$study=="Moeini, 2019",]$percent_women <-100
df_appl_v_orange[df_appl_v_orange$study=="Santomauro, 2016",]$percent_women <-40

# Calculate SE for baseline severity --------------------------------------

df_appl_v_orange$baseline_st_error_active <- 
  df_appl_v_orange$baseline_sd_active/sqrt(df_appl_v_orange$baseline_n_active)

df_appl_v_orange$baseline_st_error_control <-
  df_appl_v_orange$baseline_sd_control/sqrt(df_appl_v_orange$baseline_n_control)

# Turn into a long database with unique rows ------------------------------

### Important: create a dataset that will have unique control studies (see problem that we identified, namely common control conditions)
# create new id to help with better identification and work with duplicates (see below) 
df_appl_v_orange  <- df_appl_v_orange  %>%
  mutate(new_study_id = case_when(psy_or_med == 0 ~ paste(study,year, sep = ", "),
                                  .default = study ))

# We are going to use the descr_active and descr_control columns rather than active_type and control_type columns, as these have unique values for each study. However the descr columns are empty for med studies, so I'm going to paste values across to fix this error.

df_appl_v_orange <- df_appl_v_orange %>% 
  mutate(descr_active = if_else(is.na(descr_active ), active_type , descr_active ))

df_appl_v_orange <- df_appl_v_orange %>% 
  mutate(descr_control = if_else(is.na(descr_control), control_type, descr_control ))

# Step 1: keep only the rows with the top instrument in our hierarchy
df_with_distinct_instruments <-  df_appl_v_orange %>%          
  group_by(new_study_id) %>% 
  filter(instrument_value == min(instrument_value)) # coded for the smallest number to be best.

write.csv(df_with_distinct_instruments, "df_with_distinct_instruments.csv")

# Step 2: turn dataframes to long

# A small bit of cleaning so the code below will work

df_with_distinct_instruments <- df_with_distinct_instruments %>% 
  rename (percent_women_active = active_percent_women, 
          percent_women_control = control_percent_women,
          percent_women_overall = percent_women)

# A: turn long the rows with active
active_df <- df_with_distinct_instruments %>% 
  dplyr:: select(new_study_id, active_type, descr_active, psy_or_med, baseline_n_active, cohens_d_active,
                 baseline_mean_active, baseline_sd_active, post_mean_active, post_sd_active,
                 perc_baseline_mean_active, perc_baseline_sd_active, perc_post_mean_active, perc_post_sd_active, percent_women_overall, percent_women_active, age_m_active, age_sd_active, age_m_overall, age_sd_overall, no_sites) %>% 
  rename(treatment = descr_active) %>% 
  rename(type = active_type) %>% 
  rename_with(~ str_remove(., "_active")) %>% 
  mutate(arm_effect_size = "cohens_d_active") %>% # chose this clunky name because I had used this variable name in previous code and don't want to bread the rest
  group_by(new_study_id) %>% #go through each study id and remove duplicates
  distinct(treatment, .keep_all = TRUE)

control_df <- df_with_distinct_instruments %>% 
  dplyr:: select(new_study_id, control_type, descr_control, psy_or_med, baseline_n_control, cohens_d_control,
                 baseline_mean_control, baseline_sd_control,post_mean_control, post_sd_control,
                 perc_baseline_mean_control, perc_baseline_sd_control, perc_post_mean_control, perc_post_sd_control, percent_women_overall, percent_women_control, age_m_control, age_sd_control, age_m_overall, age_sd_overall, no_sites) %>% 
  rename(treatment = descr_control) %>% 
  rename(type = control_type) %>% 
  rename_with(~ str_remove(., "_control")) %>% 
  mutate(arm_effect_size = "cohens_d_control") %>% #chose this clunky name because I had used this variable name in previous code and don't want to bread the rest
  group_by(new_study_id) %>% #go through each study id and remove duplicates
  distinct(treatment, .keep_all = TRUE)

df_long_for_metan <- rbind(active_df, control_df)

# I'm going to remove the March / TADS psy control row. It is currently filled with NAs which implies missing data, however it is actually the case that this condition doesn't exist as such (the control is placebo, which is not a psy control).

df_long_for_metan <- df_long_for_metan %>%
  filter(!(new_study_id == "March, 2004" & is.na(treatment)))

# I'm going to perform  the operations to calculate overall age and percent women here now. We just want this per study rather than per arm
# as we are going to compare psy vs med (2 subgroups) rather than the 4 arms

df_long_for_metan <- df_long_for_metan %>% 
  group_by(new_study_id) %>%
  mutate(n_overall = sum(baseline_n)) %>% 
  mutate(age_overall = sum(age_m * baseline_n) / sum(baseline_n)) %>% 
  mutate(mean_age = coalesce(age_overall, age_m_overall)) %>% 
  mutate(pooled_sd_age = sqrt(sum((age_sd^2) * (baseline_n - 1)) / sum(baseline_n - 1))) %>% 
  mutate(se_age = pooled_sd_age / sqrt(n_overall)) %>% 
  mutate(gender_overall = sum(percent_women * baseline_n) / sum(baseline_n)) %>% 
  mutate(percent_women = coalesce(percent_women_overall, gender_overall))

# Now just get rid of the unnecessary columns as these are causing clutter

df_long_for_metan <- df_long_for_metan %>% 
  select(-c(percent_women_overall, age_m, age_sd, age_m_overall, age_sd_overall, age_overall, gender_overall))

# Remember that these are now per study, not per arm.

# Create SEs proportions for percentage women --------------------------------------------

# We also need to calculate SE for proportion women for the baseline calculations
# for proportions, this is calculated as sqrt(p(1-p)/n), which I implement stepwise below

df_long_for_metan <- df_long_for_metan %>% 
  mutate(product_perc_women = (percent_women/100)* (1-(percent_women/100)) )

df_long_for_metan <- df_long_for_metan %>% 
  mutate(percent_women_std_error = sqrt(product_perc_women/ n_overall ))

#dim(df_long_for_metan ) #check dimension
#make sure no study lost
#length(unique(df_long_for_metan$new_study_id ) ) == # #length(unique(df_appl_v_orange$new_study_id ) )

# Now check again the studies with muliple arms
# Now checking if this worked with the studies that I used to illustrate the problem above. 
# df_long_for_metan[df_long_for_metan$new_study_id == "Stallard, 2012",]
# df_long_for_metan[df_long_for_metan$new_study_id == "Atkinson, 2014",]
# # also check one which is single to make sure it is kept
# df_long_for_metan[df_long_for_metan$new_study_id == "Ackerson, 1998",]

# save this as the master dataframe  --------------------------------------
# use this dataframe to perform all operations below

write.csv(df_long_for_metan, "df_long_for_metan.csv", row.names = F)

########### simulate so that each study has a value from a distribution of 
######## of correlations.

# I will generate random numbers per study id from a distribution with these parameters. 
# it is reasonable to generate one random correlation value per study as there is no reason why the correlation should
# systematically vary within studies

# Use the function I have created for this

df_long_for_metan <- read.csv("df_long_for_metan.csv")

list_df_simulated <-  simulate_dataframes_for_st_errors (df = df_long_for_metan, 
                                              num_repetitions = 1000, 
                                              seed  = 1974,
                                              n = length(unique(df_long_for_metan$new_study_id)), 
                                              a = 0.45, 
                                              b = .9, 
                                              mean = 0.65, 
                                              sd = 0.2)

#check this worked

#list_df_simulated[[10]][,c("new_study_id", "correlation_sim_values", "simulated_se")] # looks right
#summary(list_df_simulated[[1000]]$correlation_sim_values, na.rm = T) # as expected # the mean is around the parameter I gave

```

# Results

## Included studies

The data for the studies included in this metanalysis are summarised in Supplementary Table 1 and are also available as a csv dataframe on \[<https://github.com/transatlantic-comppsych/apples_oranges>\].

In total, there were `r length(unique(df_long_for_metan$new_study_id))` studies which included `r df_count_studies[df_count_studies$condition == "medication_active"& df_count_studies$is_missing == FALSE,]$n` active arms and `r df_count_studies[df_count_studies$condition == "medication_control"& df_count_studies$is_missing == FALSE,]$n` control arms of antidepressant trials; and `r df_count_studies[df_count_studies$condition == "psychotherapy_active"& df_count_studies$is_missing == FALSE,]$n` active arms and `r df_count_studies[df_count_studies$condition == "psychotherapy_control"& df_count_studies$is_missing == FALSE,]$n` control arms from psychotherapy trials. Note that the number of active and control arms does not exactly match because some studies feature more than one control or active arm. There were also missing data for `r df_count_studies[df_count_studies$condition == "medication_active"& df_count_studies$is_missing == TRUE,]$n`, `r df_count_studies[df_count_studies$condition == "medication_control"& df_count_studies$is_missing == TRUE,]$n`, `r df_count_studies[df_count_studies$condition == "psychotherapy_active"& df_count_studies$is_missing == TRUE,]$n`, and `r df_count_studies[df_count_studies$condition == "psychotherapy_control"& df_count_studies$is_missing == TRUE,]$n` trial arms for medication active, medication control, psychotherapy active, and psychotherapy control conditions respectively, as the data needed to calculate the SMD was missing and could not be imputed by any of the methods outlined above.

<!--# Charlotte, for the paragraph below, please provide the numbers for each of the numbers wherever I have put an X. CB - done. I removed psychodynamic because there weren't any active conditions of this type. -->

Placebo was the control condition for all medication trials; the active arm ranged from serotonin reuptake inhibitors, such as 6 fluoxetine and 2 escitalopram, to tricylics, such as 2 nortriptyline. In psychological trials, the control arm included 14 WL controls, 25 care as usual and several other conditions such as 4 attention control conditions; the active arm included 43 CBT and 8 IPT amongst others. All included trials and the types of treatment controls can be found in Supplementary Table 1.

## Sample characteristics at baseline in medication versus psychotherapy trials

Table 1 summarises the results from each of the meta-analyses examining sample characteristics at baseline. The summary statistics are provided for each subgroup (i.e. for medication and psychotherapy trials) and the p-value derives from the test for subgroup differences. Full results for each of the sensitivity analyses are included in the Supplementary Materials (Table X - Y).

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# Start by looking at demographics at baseline
# Currently we have both control and active conditions, we just need each study once

df_baseline_demographics <- df_long_for_metan %>% 
  distinct(new_study_id, .keep_all = TRUE)

# This is the function to extract statistics

extract_stats <- function(df) {

  # get the summary table with results
  statistics <- summary(df, byvar = "psy_or_med")

  df_stats <- data.frame(
  subgroup = statistics$subgroup.levels,
  k_study = statistics$k.study.w,
  TE_random = statistics$TE.random.w,
  seTE_random = statistics$seTE.random.w,
  lower_random = statistics$lower.random.w,
  upper_random = statistics$upper.random.w,
  tau2 = statistics$tau2.w)

  # Add an empty row at the top and column on right (this is for the test for subgroup differences p value)
  df_stats <- rbind(data_frame(subgroup = NA, k_study = NA, TE_random = NA, seTE_random = NA, 
                             lower_random = NA, upper_random = NA, tau2 = NA), df_stats)
  df_stats <- cbind(df_stats, pval_between = NA)

  # Set the value in the last row and last column as the between groups p val
  df_stats[1, "pval_between"] <- statistics$pval.Q.b.random

  return(df_stats)

}

# Now look at baseline severity for overall psy vs med
# Need to calculate the variables first

df_for_baseline_severity <- df_long_for_metan %>% 
  group_by(new_study_id) %>%
  mutate(perc_baseline_overall = sum(perc_baseline_mean * baseline_n) / sum(baseline_n)) %>% 
  mutate(pooled_perc_sd_baseline = sqrt(sum((perc_baseline_sd^2) * (baseline_n - 1)) / sum(baseline_n - 1))) %>% 
  mutate(perc_se_baseline = pooled_perc_sd_baseline / sqrt(n_overall))

df_for_baseline_severity <- df_for_baseline_severity %>% 
  distinct(new_study_id, .keep_all = TRUE)

met_sev <- metagen(TE = perc_baseline_overall,
                           seTE = perc_se_baseline,
                           studlab = new_study_id,
                           data = df_for_baseline_severity,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "baseline severity",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_sev <- extract_stats(met_sev)

# clinical sample

clin_study_ids <-
  df_appl_v_orange[df_appl_v_orange$diagnosis != "sub",]$new_study_id

df_for_baseline_severity_clin <- df_for_baseline_severity %>%
  filter(new_study_id %in% clin_study_ids)

met_sev_clin <- metagen(TE = perc_baseline_overall,
                           seTE = perc_se_baseline,
                           studlab = new_study_id,
                           data = df_for_baseline_severity_clin,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "baseline severity clinical",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_sev_clin <- extract_stats(met_sev_clin)

# cbt + med

cbt_meds_study_ids <-
  df_appl_v_orange[df_appl_v_orange$active_type == "cbt" | df_appl_v_orange$active_type == "Fluoxetine" | df_appl_v_orange$active_type == "Escitalopram",]$new_study_id

df_for_baseline_severity_cbt <- df_for_baseline_severity %>%
  filter(new_study_id %in% cbt_meds_study_ids)

met_sev_cbt <- metagen(TE = perc_baseline_overall,
                           seTE = perc_se_baseline,
                           studlab = new_study_id,
                           data = df_for_baseline_severity_cbt,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "baseline severity cbt + med studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_sev_cbt <- extract_stats(met_sev_cbt)

# no wl

no_wl_ids <- df_appl_v_orange[df_appl_v_orange$control_type != "wl",]$new_study_id

df_for_baseline_severity_no_wl <- df_for_baseline_severity %>%
  filter(new_study_id %in% no_wl_ids)

met_sev_no_wl <- metagen(TE = perc_baseline_overall,
                           seTE = perc_se_baseline,
                           studlab = new_study_id,
                           data = df_for_baseline_severity_no_wl,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "baseline severity no wl studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_sev_no_wl <- extract_stats(met_sev_no_wl)

```
### Baseline severity

On average, severity of depressive symptoms at baseline was significantly higher in medication trials when compared to psychotherapy trials for the overall sample (p = `results_met_sev$pval_between`).The statistics provided in Table 1 represent the transformed mean scores at baseline on the

and when taking only those studies where the active arm was either CBT, fluoxetine or escitalopram (p = `results_met_sev_cbt$pval_between`). When excluding psychotherapy studies that recruited samples with sub-clinical depression, this difference did not reach statistical significance (p = `results_met_sev_clin$pval_between`).

```{r, warning=FALSE, message = FALSE, echo=FALSE}

# Now start looking at perc_women variable 
# The metagen function doesn't like when NAs result in the TE vector being a different length to subgroup vector. Need to start by removing NAs.

df_for_gender_metan <- df_baseline_demographics %>% 
  filter(!is.na(percent_women))

met_perc_women_overall <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_gender_metan,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across studies")

met_perc_women_overall <- update(met_perc_women_overall,
                              subgroup = df_for_gender_metan$psy_or_med, subgroup.name = "treatment modality",
                              tau.common = FALSE)

# Apply the function to the meta-analysis result

results_met_perc_women_overall <- extract_stats(met_perc_women_overall)

# looking at clinical sample only

df_for_gender_metan_clin <- df_for_gender_metan %>%
  filter(new_study_id %in% clin_study_ids)

met_perc_women_clin <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_gender_metan_clin,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across clinical studies")

met_perc_women_clin <- update(met_perc_women_clin,
                           subgroup = df_for_gender_metan_clin$psy_or_med, subgroup.name = "treatment modality",
                           tau.common = FALSE)

# Apply the function to the meta-analysis result
results_met_perc_women_clin <- extract_stats(met_perc_women_clin)

# Now looking at the CBT + med studies only

df_for_gender_metan_cbt <- df_for_gender_metan %>%
  filter(new_study_id %in% cbt_meds_study_ids)

met_perc_women_cbt <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_gender_metan_cbt,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across cbt and med studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

# Apply the function to the meta-analysis result
results_met_perc_women_cbt <- extract_stats(met_perc_women_cbt)

# Now the no waitlist group

df_for_gender_metan_no_wl <- df_for_gender_metan %>%
  filter(new_study_id %in% no_wl_ids)

met_perc_women_no_wl <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_gender_metan_no_wl,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across studies with controls that are not wl",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

# Apply the function to the meta-analysis result
results_met_perc_women_no_wl <- extract_stats(met_perc_women_no_wl)

```
### Gender

Psychotherapy trials featured a significantly higher percentage of women when compared to medication trials (`results_met_perc_women_overall$pval_between`). On average, samples were comprised of `r results_met_perc_women_overall$TE_random[results_met_perc_women_overall$subgroup == 1]`% (*SE* = `r results_met_perc_women_overall$seTE_random[results_met_perc_women_overall$subgroup == 1]` women across psychotherapy trials and `r results_met_perc_women_overall$TE_random[results_met_perc_women_overall$subgroup == 0]`% (*SE* = `r results_met_perc_women_overall$seTE_random[results_met_perc_women_overall$subgroup == 0]`women across medication trials. Excluding sub-clinical studies (p = `results_met_perc_women_clin$pval_between`) and studies that used waitlist as their control condition yielded very similar results (p = `results_met_perc_women_no_wl$pval_between`). This difference in gender composition however did not reach statistical significance when only CBT, fluoxetine and escitalopram trials were included (p = `results_met_perc_women_cbt$pval_between`).

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# Now start looking at age variable

# We have several studies for which we are missing SE for age. I am going to perform an imputation by taking an average
# for the studies we do have, per modality, and substituting this in for missing values.

df_baseline_demographics$se_age <- as.numeric(df_baseline_demographics$se_age)

mean_se_age_psy <- df_baseline_demographics %>%
  filter(psy_or_med == 1) %>%
  summarise(mean_se_age_psy = mean(se_age, na.rm = TRUE))

mean_se_age_med <- df_baseline_demographics %>%
  filter(psy_or_med == 0) %>%
  summarise(mean_se_age_med = mean(se_age, na.rm = TRUE))

df_baseline_demographics <- df_baseline_demographics %>% 
  mutate(se_age = ifelse(!is.na(se_age), se_age,
                         ifelse(psy_or_med == 1, mean_se_age_psy, mean_se_age_med)))

df_for_age_metan <- df_baseline_demographics %>% 
  filter(!is.na(mean_age))

# Age meta analysis

df_for_age_metan$se_age <- as.numeric(df_for_age_metan$se_age)

met_age <- metagen(TE = mean_age,
                           seTE = se_age,
                           studlab = new_study_id,
                           data = df_for_age_metan,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "age across studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_age <- extract_stats(met_age)

# Looking at clinical sample only

clin_study_ids <-
  df_appl_v_orange[df_appl_v_orange$diagnosis != "sub",]$new_study_id

df_for_age_metan_clin <- df_for_age_metan %>%
  filter(new_study_id %in% clin_study_ids)

met_age_clin <- metagen(TE = mean_age,
                           seTE = se_age,
                           studlab = new_study_id,
                           data = df_for_age_metan_clin,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "age across clinical studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_age_clin <- extract_stats(met_age_clin)

# Now cbt + med studies only

df_for_age_metan_cbt <- df_for_age_metan %>%
  filter(new_study_id %in% cbt_meds_study_ids)

met_age_cbt <- metagen(TE = mean_age,
                           seTE = se_age,
                           studlab = new_study_id,
                           data = df_for_age_metan_cbt,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "age across cbt + med studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_age_cbt <- extract_stats(met_age_cbt)

# Now excluding wl

df_for_age_metan_no_wl <- df_for_age_metan %>%
  filter(new_study_id %in% no_wl_ids)

met_age_no_wl <- metagen(TE = mean_age,
                           seTE = se_age,
                           studlab = new_study_id,
                           data = df_for_age_metan_no_wl,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "age across no wl studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_age_no_wl <- extract_stats(met_age_no_wl)

```
### Age

Mean age was `r results_met_age$TE_random[results_met_perc_women_overall$subgroup == 1]` (*SE* = `r results_met_age$seTE_random[results_met_perc_women_overall$subgroup == 1]`) across psychotherapy trials and `r results_met_age$TE_random[results_met_perc_women_overall$subgroup == 0]` (*SE* = `r results_met_age$seTE_random[results_met_perc_women_overall$subgroup == 0]`) across medication trials, with no significant between group differences (p = `r results_met_age$pval_between`). There were no significant differences in mean age between modalities on further sensitivity analyses.

```{r}

# Lets make a table that combines across these baseline characteristics variables

df_overall_baseline_results <- rbind(results_met_perc_women_overall, results_met_age, results_met_sev)

# For row names
df_overall_baseline_results <- df_overall_baseline_results %>%
  mutate(subgroup = case_when(
    row_number() == 1 ~ "Percent Women",
    row_number() == 4 ~ "Age",
    row_number() == 7 ~ "Severity",
    TRUE ~ if_else(subgroup == 0, "Medication", "Psychotherapy")
  ))

# Define new column names
new_column_names <- c("Subgroup", "K", "Mean", "SE", "Lower CI", "Upper CI", "T2", "p-value")

# Assign new column names to the data frame
colnames(df_overall_baseline_results) <- new_column_names

# Rounding
df_overall_baseline_results <- df_overall_baseline_results %>%
  mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
  mutate_at(vars(ncol(.)), ~ round(., 3))

colnames(df_overall_baseline_results)[8] <- paste0(colnames(df_overall_baseline_results)[8], footnote_marker_number(1))

df_overall_baseline_results %>%
  kbl(align = "l", caption = "Table X: Sample characteristics at baseline across medication and psychotherapy studies", escape = F) %>%
  kable_material(bootstrap_options = c("hover", "condensed", "bordered"), full_width = FALSE) %>%
  row_spec(row = 0, background = "#EBEBEB", extra_css = "border-bottom: 1px black;" ) %>% 
  row_spec(row = 1, bold = T) %>% 
  row_spec(c(4, 7), bold = T, extra_css = "border-top: 1px solid black;") %>%
  column_spec(1, border_right = T) %>% 
  add_indent(c(2, 3, 5, 6, 8, 9)) %>% 
  footnote(number = c("Statistic derived from the test for subgroup differences in the relevant meta-analysis"))

df_overall_baseline_results <- options(knitr.kable.NA = "")

```

```{r, warning=FALSE, message = FALSE, echo=FALSE}

df_sites_summary_stats$mean_sites[df_sites_summary_stats$psy_or_med == 0]

### Number of sites
# Write results, tabulate.

## This data lives in this spreadsheet for psy trials, we need to merge it in

df_no_sites <- readxl:: read_excel("for_despina_and_giannis_200224.xlsx")

df_no_sites <- df_no_sites %>% 
  select(c(new_study_id, number_of_sites)) %>% 
  filter(!is.na(number_of_sites)) %>% 
  distinct(new_study_id, .keep_all = TRUE) %>% 
  mutate(number_of_sites = as.numeric(number_of_sites))

df_baseline_demographics <- full_join(df_baseline_demographics, df_no_sites, by = "new_study_id")

# we have different vectors for psy and med trials, lets coalesce

df_baseline_demographics <- df_baseline_demographics %>%
  mutate(no_sites = as.numeric(no_sites)) %>% 
  mutate(sites = coalesce(no_sites, number_of_sites)) %>%
  select(-no_sites, -number_of_sites)

# we have a few rogue studies (Kitchen, McCarty and Wright). Kitchen was too recent (after 2021), McCarty and Wright are not in Cuijpers' master
# dataset for some reason, though they are in the list of included studies in the table in his MA. For now I'm going to remove them because 
# we don't have data for these trials.

df_baseline_demographics <- df_baseline_demographics %>% 
  filter(!(new_study_id %in% c("Kitchen, 2021", "McCarty, 2013", "Wright, 2020")))

# Lets look at results

# summary stats
df_sites_summary_stats <- df_baseline_demographics %>%
  group_by(psy_or_med) %>%
  summarise(
    n = n(),
    mean_sites = mean(sites, na.rm = TRUE),
    sd_sites = sd(sites, na.rm = TRUE),
  )

# Tabulate
# Rounding
df_sites_summary_stats <- df_sites_summary_stats %>%
  mutate(across(where(is.numeric), ~round(., 2)))

# Row names
df_sites_summary_stats <- df_sites_summary_stats %>%
  mutate(psy_or_med = case_when(
    psy_or_med == 0 ~ "Medication",
    psy_or_med == 1 ~ "Psychotherapy",
    TRUE ~ as.character(psy_or_med)  # Keep other values unchanged
  ))

df_sites_summary_stats <- df_sites_summary_stats %>%
  rename(
    Modality = psy_or_med,
    N = n,
    Mean = mean_sites,
    SD = sd_sites
  )

df_sites_summary_stats %>%
  kbl(caption = "Table X: Number of sites across medication and psychotherapy studies") %>% 
  kable_material(bootstrap_options = c("hover", "condensed", "bordered"), full_width = FALSE) %>% 
  row_spec(row = 0, background = "#EBEBEB") %>% 
  column_spec(column = 1, border_right = T)

# t-test
sites_t_test <- t.test(sites ~ psy_or_med, data = df_baseline_demographics)

sites_t_test_stats <- data.frame(
  statistic = sites_t_test$statistic,
  df = sites_t_test$parameter,
  p_value = sites_t_test$p.value,
  ci_lower = sites_t_test$conf.int[1],
  ci_upper = sites_t_test$conf.int[2]
)

sites_t_test_stats <- sites_t_test_stats %>%
  dplyr:: mutate(across(.cols = c(statistic, df, ci_lower, ci_upper), ~ round(., 2)))

sites_t_test_stats %>%
  kbl(caption = "Table X: Results for t-test comparing number of sites between medication and psychotherapy studies") %>% 
  kable_material(bootstrap_options = c("hover", "condensed", "bordered"), full_width = FALSE) %>% 
  row_spec(row = 0, background = "#EBEBEB") %>% 
  column_spec(column = 1, border_right = T)

# plot

ggplot(df_baseline_demographics, aes(x = factor(psy_or_med), y = sites, color = factor(psy_or_med))) +
  geom_point(position = position_jitter(width = 0.2), size = 2) +
  labs(title = "Number of Sites for Psychotherapy and Medication Trials",
       x = "Treatment Modality",
       y = "Number of Sites") +
  scale_color_discrete(name = "Treatment Type", labels = c("Medication", "Psychotherapy"))

```
### Number of trial sites

Average number of trial sites was significantly higher across medication studies ( *M* = `r `df_sites_summary_stats$mean_sites[df_sites_summary_stats$psy_or_med == 0]`, *SD* = `r `df_sites_summary_stats$sd_sites[df_sites_summary_stats$psy_or_med == 0]` compared to psychotherapy studies ( *M* = `r `df_sites_summary_stats$mean_sites[df_sites_summary_stats$psy_or_med == 1]`, *SD* = `r `df_sites_summary_stats$sd_sites[df_sites_summary_stats$psy_or_med == 1]`(*t* (`r sites_t_test$parameter`) = `r sites_t_test$statistic`, *p* = `r sites_t_test$p.value`).

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# Get Ns, run metareg, prepare graphing ------------------------------------------------------
library(tidyverse)
library(metafor)

# add a new multilevel variable to the simulated data for the regression

for(i in 1: length(list_df_simulated)){

list_df_simulated[[i]]$arm_effect_size <- factor(list_df_simulated[[i]]$arm_effect_size)
  list_df_simulated[[i]] <- list_df_simulated[[i]] %>% 
    mutate(four_level_var = case_when(psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
                                      psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control")
    )

  list_df_simulated[[i]]$four_level_var <- factor(list_df_simulated[[i]]$four_level_var) # turn to factor

  # relevel so that medication control becomes the reference category for the regression
#  list_df_simulated[[i]]$four_level_var <- relevel(list_df_simulated[[1]]$four_level_var, ref = "medication_control")
list_df_simulated[[i]]$four_level_var <-   fct_relevel(list_df_simulated[[i]]$four_level_var,
                                                       "medication_control",
                                                       "medication_active", 
                                                       "psychotherapy_control",
                                                       "psychotherapy_active")
}

# check it worked
# list_df_simulated[[2]] %>% 
# dplyr:: select(psy_or_med, arm_effect_size, four_level_var)

######## NOW RUN METAREGRESSIONS AND GET Coefficients for 
# A) the overall sample
# B) the CBT vs Fluox and Escitalopram sample
# C) the no WL sample
# E) the CDRS only sample (**Argyris** do we still want this?)
# F) Clinical Levels Depression

# A. Estimate Means and CIs for the overall ----------------------------------

# count studies
# count the number of studies
n_unique_studies <- length(unique(df_long_for_metan$new_study_id)) # CHARLOTTE please check
df_count_studies <- count_studies(df_long_for_metan ) # CHARLOTTE please check
# CB checked both, both correct.

# specify model
model_1 <- as.formula(~ four_level_var)

# run metareg function 
# here you can use the simulated results directly
list_model_1_meta_reg <- run_metaregression(list_df_simulated, model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_overall <-  aggregate_model_results(list_model_1_meta_reg, condition)
aggregate_results_overall <- cbind(aggregate_results_overall, condition)

# extract SMDs from the  list
list_dummy_var_means <-  lapply(list_model_1_meta_reg, extract_coefficients_func)

# calculate mean SMDs and ses
df_mean_coefs_from_sim <- calculate_mean_coefs_ses(list_dummy_var_means, condition)

df_mean_coefs_from_sim <- data.frame(cbind(condition,df_mean_coefs_from_sim))

# add the ns
df_mean_coefs_from_sim <- data.frame(cbind(df_mean_coefs_from_sim, n = df_count_studies[df_count_studies$is_missing == FALSE, ]$n))

# B. Estimate Means and CIs for CBT Fluox Esc --------------------------------
#  grab ids for those studies
cbt_fluox_esc_study_ids <-
  df_appl_v_orange[df_appl_v_orange$active_type == "cbt" | df_appl_v_orange$active_type == "Fluoxetine" | df_appl_v_orange$active_type == "Escitalopram",]$new_study_id

# Use those ids to subset all dataframes in the list
# By searching for IDs we have mistakenly included multiple active arms. E.g. for Atkinson which has dulox and fluox. 
# So we need to further subset the simulated dfs. For the active arms, I am filtering out conditions that do not meet our criteria. 
# This leaves all the control conditions from trials where the active is cbt, fluox or esc. 
# We may have multiple controls per study id for this reason, though this isn't a problem 
# conceptually.

list_cbt_fluox_esc_study <- lapply(list_df_simulated, function(df) {
  df %>%
    filter(new_study_id %in% cbt_fluox_esc_study_ids) %>%
    filter(!(arm_effect_size == "cohens_d_active" & !(type %in% c("Fluoxetine", "Escitalopram", "cbt"))))
})

# count the number of studies
cbt_fluox_esc_study <- df_long_for_metan %>%
  filter(new_study_id %in% cbt_fluox_esc_study_ids,
         !(arm_effect_size == "cohens_d_active" & !(type %in% c("Fluoxetine", "Escitalopram", "cbt")))) 
study_count_cbt_fluox_esc_study <- count_studies(cbt_fluox_esc_study)
# CB checked, looks good now.

# apply the metareg function to the list of cbt, fluox, esc
model_1 <- as.formula(~ four_level_var)
list_model_1_cbt_fluox_esc_study <- run_metaregression(list_cbt_fluox_esc_study , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_cbt_fluox_esc_study <-  aggregate_model_results(list_model_1_cbt_fluox_esc_study, condition)
aggregate_results_cbt_fluox_esc_study <- cbind(aggregate_results_cbt_fluox_esc_study, condition)

# extract SMDs from the list for cbt fluox escit
means_cbt_fluox_esc_study <-  lapply(list_model_1_cbt_fluox_esc_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_cbt_fluox_esc_study <- calculate_mean_coefs_ses(means_cbt_fluox_esc_study, condition)
 coef_and_se_means_cbt_fluox_esc_study <- data.frame(cbind(condition, coef_and_se_means_cbt_fluox_esc_study))

# add the ns to the dataframe
 coef_and_se_means_cbt_fluox_esc_study$n  <- study_count_cbt_fluox_esc_study[study_count_cbt_fluox_esc_study$is_missing ==F,]$n

# C. Estimate Means and CIs without WL ---------------------------------------

 # grab the ids for no waitlist
no_wl_ids <-df_appl_v_orange[df_appl_v_orange$control_type != "wl",]$new_study_id

# use them to loop over the simulated list to subset all the dataframes.

list_no_wl <- lapply(list_df_simulated, function(df) {
  df %>%
    filter(new_study_id %in% no_wl_ids) %>%
    filter(!(arm_effect_size == "cohens_d_control" & type == "wl")) 
})

# count studies 
no_wl_studies <- df_long_for_metan %>% 
  filter(new_study_id %in% no_wl_ids) %>%
  filter(!(arm_effect_size == "cohens_d_control" & type == "wl"))
study_count_no_wl_study <- count_studies(no_wl_studies)
# CB checked, looks good!

# use this model as above
model_1 <- as.formula(~ four_level_var)

# apply the metareg function to the list of no wl
list_model_1_no_wl <- run_metaregression(list_no_wl , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_no_wl <-  aggregate_model_results(list_model_1_no_wl, condition)
aggregate_results_no_wl <- cbind(aggregate_results_no_wl, condition)

# now use the function to extract the coefficients
means_no_wl <-  lapply(list_model_1_no_wl, extract_coefficients_func)

# run the mean function for no wl
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_no_wl <- calculate_mean_coefs_ses(means_no_wl, condition)
coef_and_se_means_no_wl <- data.frame(cbind(condition,coef_and_se_means_no_wl))

coef_and_se_means_no_wl$n <- study_count_no_wl_study[study_count_no_wl_study$is_missing 
                                                     ==FALSE,]$n

# E. CDRS Sensitivity Analysis ---------------------------------------

cdrs_study_ids <-
  df_appl_v_orange[df_appl_v_orange$instrument_name == "cdrs",]$new_study_id

# use those ids to subset all dataframes in the list
list_cdrs_study <- list()

for(i in 1: length(list_df_simulated)){

  list_cdrs_study[[i]] <- list_df_simulated[[i]][list_df_simulated[[i]]$new_study_id
                                                          %in% cdrs_study_ids,]

}

# count the number of studies
study_count_cdrs_study <- count_studies(df_long_for_metan[df_long_for_metan$new_study_id %in% cdrs_study_ids,])
# CB checked, correct.

# apply the metareg function to the list of cdrs studies
model_1 <- as.formula(~ four_level_var)
list_model_1_cdrs_study <- run_metaregression(list_cdrs_study , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_cdrs_study <-  aggregate_model_results(list_model_1_cdrs_study, condition)

# extract SMDs from the list for cdrs
means_cdrs_study <-  lapply(list_model_1_cdrs_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_cdrs_study <- calculate_mean_coefs_ses(means_cdrs_study, condition)
coef_and_se_means_cdrs_study <- data.frame(cbind(condition, coef_and_se_means_cdrs_study))

# add the ns to the dataframe
coef_and_se_means_cdrs_study$n  <- study_count_cdrs_study[study_count_cdrs_study$is_missing ==F,]$n

# F Clinical Levels Depression Sensitivity Analysis ---------------------------------------

# We need to decide how we want to do this. We have mdd, mood, sub, and cut categories. For now I'll just exclude sub.

#take study ids where sample has clinical levels of depression
clin_study_ids <-
  df_appl_v_orange[df_appl_v_orange$diagnosis != "sub",]$new_study_id

# use those ids to subset all dataframes in the list
list_clin_study <- list()

for(i in 1: length(list_df_simulated)){

  list_clin_study[[i]] <- list_df_simulated[[i]][list_df_simulated[[i]]$new_study_id
                                                          %in% clin_study_ids,]

}

# count the number of studies
study_count_clin_study <- count_studies(df_long_for_metan[df_long_for_metan$new_study_id %in% clin_study_ids,])
# CB checked, correct.

# apply the metareg function to the list of clinical studies
model_1 <- as.formula(~ four_level_var)
list_model_1_clin_study <- run_metaregression(list_clin_study , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_clin_study <-  aggregate_model_results(list_model_1_clin_study, condition)

# extract SMDs from the list for clinical studies
means_clin_study <-  lapply(list_model_1_clin_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_clin_study <- calculate_mean_coefs_ses(means_clin_study, condition)
coef_and_se_means_clin_study <- data.frame(cbind(condition, coef_and_se_means_clin_study))

# add the ns to the dataframe
coef_and_se_means_clin_study$n  <- study_count_clin_study[study_count_clin_study$is_missing ==F,]$n

```

#### Mean efficacy of psychotherapy versus medication control conditions

We applied metaregression to obtain the SMDs and confidence intervals of each of the four arms.

As can be seen in Figure 1, there were substantial differences between the four arms of the metanalysis with striking difference being the medication and the psychotherapy control arms. In particular, placebo had an SMD = `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$coef_means, 2)`, 95% CI: `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$lower_ci, 2)` to `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$upper_ci, 2)`, whereas psychotherapy controls had an SMD = `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$coef_means, 2)` , 95% CI: `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$lower_ci, 2)` to `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$upper_ci, 2)`.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
subtitle_text_all <- "Metanalytic estimates of within-group changes: all studies"
plot_means_all <- plot_means_function(df_mean_coefs_from_sim, subtitle_text_all)
#print(plot_means_all)
pdf(file = "plot_means_all.pdf")
plot_means_all
```

![Figure 1.](plot_means_all.pdf)

Table X summarises the statistics of the estimated SMDs. <!--# Charlotte, we need to turn these into nice tables with proper columns, rounding etc. -->

```{r, warning=FALSE, message = FALSE, echo=FALSE}

# Prepare to tabulate

df_mean_coefs_from_sim <- df_mean_coefs_from_sim %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition))) %>% 
  relocate("n", .after = "condition")

df_mean_coefs_from_sim %>%
  kbl(caption = "Table X: Summary statistics of the estimated SMDs") %>% 
  kable_material(bootstrap_options = c("hover", "condensed", "bordered"), full_width = FALSE) %>% 
  row_spec(row = 0, background = "#EBEBEB") %>% 
  column_spec(column = 1, border_right = T)

```

In Table 2, we present the regression that tests our hypothesis about differences between medication and psychotherapy control. In particular, in this metaregression model, medication control is the reference category (termed intercept) to which all other categories of the dummy variable, including psychotherapy control, are compared. The strongest difference betwen arms, as judged by the z-value, is between the psychotherapy and medication control with a z-value `r round(aggregate_results_overall[aggregate_results_overall$condition == "psychotherapy_control", ]$z_value, 2)` , which yields a very low p-value ( `r sprintf("%.3e",(1-(pnorm(8.949921))))` ).

```{r, warning=FALSE, message = FALSE, echo=FALSE}

# Prepare to tabulate

aggregate_results_overall <- aggregate_results_overall %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition))) %>% 
  relocate("condition", .before = "coefficients")

aggregate_results_overall %>%
  kbl(caption = "Table X: Meta-regression results") %>% 
  kable_material(bootstrap_options = c("hover", "condensed", "bordered"), full_width = FALSE) %>% 
  row_spec(row = 0, background = "#EBEBEB") %>% 
  column_spec(column = 1, border_right = T)

```

### Sensitivity Analyses

We then conducted a series of sensitivity analyses of our results.

#### Effect of standard errors of the SMDs .

It could be argued that the choice of standard errors of the changes for the calculation of the confidence intervals could have affect the results in one or the other direction. To address such concerns we have simulated 1000 different datasets with SMDs coming from a broad distribution. If standard error distributions were influential, this should show up as substantial variability across simulations. We test this idea in the Figure 1 which displays across the 1000 simulations the z-value of the contrast between medication and psychotherapy control arms (the mean of which we presented in Table 2). As can be seen, the variability in the z-score is minimal and consistently far away from the threshold for significance, i.e. the value of z = 1.645.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# examine the stability of the simulations

z_psy_con_vs_med_con <- 0
for(i in 1: length(list_model_1_meta_reg)){
z_psy_con_vs_med_con[i] <- list_model_1_meta_reg[[i]]$zval[condition=="psychotherapy_control"]

}

stability_sims <- data.frame(n_sims = 1:1000, z_value = z_psy_con_vs_med_con)

stab_sims <- stability_sims %>%
  ggplot(aes(x = n_sims, y = z_value)) +
  geom_point() +
  geom_hline(yintercept = 1.65, linetype = "dotted", color = "red") +
  geom_segment(aes(x = 300, xend = 300, y = 2, yend = 1.65),
               arrow = arrow(length = unit(0.3, "cm")),
               color = "black") +
  annotate("text", x = 300, y = 2.0, label = "p < 0.05 threshold (values above line significant)", hjust = -0.1, vjust = 0) +
  geom_segment(aes(x = 300, xend = 300, y = 7.5, yend = min(stability_sims$z_value-0.3)),
               arrow = arrow(length = unit(0.3, "cm")),
               color = "black") +
  annotate("text", x = 300, y = 7.5, label = "simulated z-values",
           hjust = -0.1, vjust = 0) +
  theme_minimal() +
  labs(
    x = "Number of Simulations",
    y = "z-value",
    title = "Stability of the Statistic of the Difference between \nMedication and Psychotherapy Control Arms",
    subtitle = "Results from 1000 simulations with within-group standard errors"
  )

pdf(file = "stab_sims.pdf")
stab_sims
dev.off()
```

![](stab_sims.pdf)

#### Comparison of CBT with fluoxetine and escitalopram

We next compared the control and active arms of CBT studies to those of fluoxetine and escitalopram studies. As can be seen, the pattern of results is very similar to that of the ovarall analyses.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
subtitle_text_cbt <- "Metanalytic estimates of within-group changes: cbt, fluox, escit"
plot_means_cbt <- plot_means_function(coef_and_se_means_cbt_fluox_esc_study , subtitle_text_cbt)
pdf(file = "plot_means_cbt.pdf")
plot_means_cbt
dev.off()
```

![](plot_means_cbt.pdf)

#### Excluding studies with waitlist control arms.

We next analysed the data after excluding waitlist control studies. As can be seen, the pattern of results is very similar to that of the overall analyses.

```{r, warning=FALSE, message = FALSE, echo=FALSE}

subtitle_text_no_wl <- "Metanalytic estimates of within-group changes: wl excluded"
plot_means_wl <- plot_means_function(coef_and_se_means_no_wl, subtitle_text_no_wl)
#print(plot_means_wl)

pdf(file = "plot_means_wl.pdf")
plot_means_wl
dev.off()
```

![](plot_means_wl.pdf)

```{r, warning=FALSE, message = FALSE, echo=FALSE}

# We need to read in the appropriate dataset which describes the active and control conditions used in
# psychotherapy trials.

df_descr_psy_conditions <- readxl:: read_excel("for_despina_and_giannis_200224.xlsx")

# checking id matchup
# discrepancies <- setdiff(df_descr_psy_conditions$new_study_id, df_long_for_metan$new_study_id)

# Cleaning the dataset
# For now lets look at variables pertaining to the nature / intensity of interventions

colnames(df_descr_psy_conditions)

df_descr_psy_conditions <- df_descr_psy_conditions %>% 
  select(c(new_study_id, `Psychotherapy/control`, levels_var, number_of_sites, Number_of_sessions,
           frequency_weeks, Length_of_sessions_mins, Total_hours_of_intervention_hours, total_hours_gsh,
           Total_period_of_intervention_weeks, Format_group_individual, Delivery, Involvement_of_others))

# rename for clarity and to match master dataframe
df_descr_psy_conditions <- df_descr_psy_conditions %>% 
  rename(treatment = 'Psychotherapy/control', no_sites = number_of_sites, no_sessions =
           Number_of_sessions, freq_weeks = frequency_weeks, length_sessions_mins =
           Length_of_sessions_mins, total_hours = Total_hours_of_intervention_hours,
         total_period_weeks = Total_period_of_intervention_weeks, format = Format_group_individual, delivery =
           Delivery, family_involv = Involvement_of_others)

check <- df_descr_psy_conditions %>% 
  filter(treatment == 'wlc')

# Visual inspection

# There are two studies where the waitlist condition has some intervention (i.e. values should not be set to 0)
# Otherwise, I want values to be set to 0, not NA.

df_descr_psy_conditions <- df_descr_psy_conditions %>%
  mutate(
    no_sessions = ifelse(treatment == "wlc" & !(new_study_id %in% c("Ackerson, 1998", "Diamond, 2002")), 0, no_sessions),
    freq_weeks = ifelse(treatment == "wlc" & !(new_study_id %in% c("Ackerson, 1998", "Diamond, 2002")), 0, freq_weeks),
    length_sessions_mins = ifelse(treatment == "wlc" & !(new_study_id %in% c("Ackerson, 1998", "Diamond, 2002")), 0, length_sessions_mins),
    total_hours = ifelse(treatment == "wlc" & !(new_study_id %in% c("Ackerson, 1998", "Diamond, 2002")), 0, total_hours)
  )

# Now cau

check <- df_descr_psy_conditions %>% 
  filter(treatment == 'cau')

# All studies except for Martinovic, Srivastava, Stikkelbroek and Weisz 2009 should have variables set to NA. 
# This is because we cannot say that they did not feature 0 hours of intervention, but that for the most part
# it is not reported or it is not possible to quantify.

df_descr_psy_conditions <- df_descr_psy_conditions %>%
  mutate(
    no_sessions = ifelse(treatment == "cau" & !(new_study_id %in% c("Martinovic, 2006", "Srivastava, 2020", "Stikkelbroek, 2020", "Weisz, 2009")), NA, no_sessions),
    freq_weeks = ifelse(treatment == "cau" & !(new_study_id %in% c("Martinovic, 2006", "Srivastava, 2020", "Stikkelbroek, 2020", "Weisz, 2009")), NA, freq_weeks),
    length_sessions_mins = ifelse(treatment == "cau" & !(new_study_id %in% c("Martinovic, 2006", "Srivastava, 2020", "Stikkelbroek, 2020", "Weisz, 2009")), NA, length_sessions_mins),
    total_hours = ifelse(treatment == "cau" & !(new_study_id %in% c("Martinovic, 2006", "Srivastava, 2020", "Stikkelbroek, 2020", "Weisz, 2009")), NA, total_hours)
  )

# Now other controls

check <- df_descr_psy_conditions %>%
  filter(!(treatment %in% c('cau', 'wlc'))) %>% 
  filter(levels_var == "psychotherapy_control")

# Made a few manual adjustments in the original excel sheet.

# List of columns to clean
columns_to_clean <- c(
  "no_sessions",
  "freq_weeks",
  "length_sessions_mins",
  "total_hours",
  "total_period_weeks"
)

# Convert non-numeric values to NA in specified columns
df_descr_psy_conditions <- df_descr_psy_conditions %>%
  mutate_at(vars(columns_to_clean), ~as.numeric(as.character(.)))

# Giannis' function for summary stats

group_function <- function(dataframe, my_groups, columns) {
  result_dataframe <- data.frame(
    Group = character(),
    Column = character(),
    N = numeric(),
    Mean = numeric(),
    SD = numeric(),
    Median = numeric(),
    IQR = numeric(),
    stringsAsFactors = FALSE
  )
  unique_groups <- unique(dataframe[[my_groups]])
  for (group in unique_groups) {
    group_data <- dataframe[dataframe[[my_groups]] == group, ]
    for (col in columns) {
      col_data <- group_data[[col]]
      n <- sum(!is.na(col_data))
      mean_val <- mean(col_data, na.rm = TRUE)
      SD <- sd(col_data, na.rm = TRUE)
      median_val <- median(col_data, na.rm = TRUE)
      iqr <- IQR(col_data, na.rm = TRUE)
      result_dataframe <- rbind(result_dataframe,
                                data.frame(Group = group,
                                           Column = col, N = n, Mean = mean_val,
                                           SD = SD, Median = median_val,
                                           IQR = iqr, stringsAsFactors = FALSE))
    }
  }
  return(result_dataframe)
}

sessions <- group_function(df_descr_psy_conditions,"levels_var","no_sessions")
frequency <- group_function(df_descr_psy_conditions,"levels_var","freq_weeks")
length_sessions_mins <- group_function(df_descr_psy_conditions,"levels_var","length_sessions_mins")
total_hours <- group_function(df_descr_psy_conditions,"levels_var","total_hours")
total_period_weeks <- group_function(df_descr_psy_conditions,"levels_var","total_period_weeks")

control_summary <- rbind(sessions, frequency, length_sessions_mins, total_hours, total_period_weeks)

control_summary <- control_summary %>% 
  select(-c(Column, Median, IQR))

control_summary <- control_summary %>%
  mutate(Group = case_when(
    Group == "psychotherapy_active" ~ "Active",
    Group == "psychotherapy_control" ~ "Control",
    TRUE ~ as.character(Group)  # Keep other values unchanged
  ))

control_summary <- control_summary %>% 
  mutate(across(where(is.numeric), ~round(., 2)))

control_summary %>%
  kbl(align = "l", caption = "Table X: Comparing the intensity of the intervention between active and control arms of psychotherapy studies", escape = F) %>%
  kable_material(bootstrap_options = c("hover", "condensed", "bordered"), full_width = FALSE) %>%
  row_spec(row = 0, background = "#EBEBEB", extra_css = "border-bottom: 1px black;" ) %>% 
  column_spec(1, border_right = T) %>% 
  pack_rows("Number of sessions", 1, 2) %>% 
  pack_rows("Frequency (weeks)", 3, 4) %>% 
  pack_rows("Session length (mins)", 5, 6) %>%
  pack_rows("Total intervention hours", 7, 8) %>%
  pack_rows("Total intervention period (weeks)", 9, 10)

# t tests

session_no_t_test <- t.test(no_sessions ~ levels_var, data = df_descr_psy_conditions)
freq_t_test <- t.test(freq_weeks ~ levels_var, data = df_descr_psy_conditions)
session_length_t_test <- t.test(length_sessions_mins ~ levels_var, data = df_descr_psy_conditions)
total_hours_t_test <- t.test(total_hours ~ levels_var, data = df_descr_psy_conditions)
total_period_t_test <- t.test(total_period_weeks ~ levels_var, data = df_descr_psy_conditions)

extract_t_test_stats <- function(df) {

  t_test_stats <- data.frame(
  outcome = df$data.name,
  statistic = df$statistic,
  df = df$parameter,
  p_value = df$p.value,
  ci_lower = df$conf.int[1],
  ci_upper = df$conf.int[2])

  return(t_test_stats)

}

results_session_no_t_test <- extract_t_test_stats(session_no_t_test)
results_freq_t_test <- extract_t_test_stats(freq_t_test)
results_session_length_t_test <- extract_t_test_stats(session_length_t_test)
results_total_hours_t_test <- extract_t_test_stats(total_hours_t_test)
results_total_period_t_test <- extract_t_test_stats(total_period_t_test)

control_summary_t_tests <- rbind(results_session_no_t_test, results_freq_t_test, results_session_length_t_test, results_total_hours_t_test, results_total_period_t_test)

```
### Comparing the nature and intensity of control conditions in psychotherapy trials

Write results here.

#### Addressing Regression to the Mean

We first conducted a meta-analysis to test for differences in means at baseline in the two instruments, CDRS-R and HAM-D on which there was a sufficient number of studies to metanalyse.

As can be seen in Table 3a and 3b, there baseline scores in medication studies were on average substantially higher for the CDRS (10.9 points for the control arms) and the HAM-D (7.3 points for the control arms), respectively. As can be seen in Supplementary Tables XX <!--# Charlotte, please create these out of results_cdrs_metareg_baseline[[1]]  and the same for HAM-D--> the differences were significant at p\<0.05 for CDRS, thought they did not reach statistical significance for the HAM-D (p = 0.0573). <!--# Charlotte, please add the n for each row. Easiest to modify the function so that it appends the baseline n together with the condition. -->

```{r, warning=FALSE, message = FALSE, echo=FALSE}
knitr:: kable(results_cdrs_metareg_baseline[[2]], caption="Table 3a CDRS baselilne scores")
```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
knitr:: kable(results_hamd_metareg_baseline[[2]], caption="Table 3b HAM-D baselilne scores")
```

As can be seen in Table 4, the baseline mean has a very strong association with the SMD.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# knitr:: kable(results_adjusted_cohens_d 
#               
#               , caption = "Table 4. Effect of baseline values on SMDs")
```

Table 5 shows the adjusted SMDs with their confidence intervals that we estimated in metaregression. As can be inferred from the non-overlapping confidence intervals, the difference between the two control conditions was significant (details of the regression model can be found in Supplementary Table XX <!--# Charlotte, please add there the aggregate_results_overall_adjusted_cohens_d -->). Differences were significant between active medication and psychotherapy control, but not between any of the other groups.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
knitr:: kable (df_mean_coefs_from_sim_adjusted_cohens_d, caption = "Table 5. SMDs Adjusted by baseline score" )
```

Finally, we tested whether the simulated correlation values we used for simulation of the standard errors had an appreciable effect on the values of the adjusted SMDs. Supplemental Figure X shows that the variability of the estimated z-value was minimal<!--# Charlotte, the figure below should be supplemental-->.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# 
# # plot the mean z for the difference between controls
# z_psy_con_vs_med_con <- 0
# for(i in 1: length(list_model_1_meta_reg_adj_cohens_d )){
#   z_psy_con_vs_med_con[i] <- list_model_1_meta_reg_adj_cohens_d [[i]]$zval[condition=="psychotherapy_control"]
#   
# }
# 
# 
# 
# stability_sims_adjusted_cohens_d <- data.frame(n_sims = 1:1000, z_value = z_psy_con_vs_med_con)
# 
# 
# stab_sims_adjusted_cohens_d <- stability_sims_adjusted_cohens_d %>%
#   ggplot(aes(x = n_sims, y = z_value)) +
#   geom_point() +
#   geom_hline(yintercept = 1.65, linetype = "dotted", color = "red") +
#   geom_segment(aes(x = 300, xend = 300, y = 2, yend = 1.65),
#                arrow = arrow(length = unit(0.3, "cm")),
#                color = "black") +
#   annotate("text", x = 300, y = 2.0, label = "p < 0.05 threshold (values above line significant)", hjust = -0.1, vjust = 0) +
#   annotate("text", x = 300, y = stability_sims_adjusted_cohens_d$z_value[1] - 0.1, label = "simulated z-values",
#            hjust = -0.1, vjust = 0) +
#   theme_minimal() +
#   labs(
#     x = "Number of Simulations",
#     y = "z-value",
#     title = "Stability of the Statistic of the Difference between \nMedication and Psychotherapy Control Arms",
#     subtitle = "Results from 1000 simulations with within-group standard errors"
#   )
# stab_sims_adjusted_cohens_d
# pdf(file = "stab_sims_adjusted_cohens_d.pdf")
# stab_sims_adjusted_cohens_d
# dev.off()

```

Tables 6a and 6b show the baeline-mean adjusted SMDs for the CDRS at outcome when the baseline measurement has been taken into account. <!--# Here Intercept refers to teh psychotherapy control. Sorry about the labelling of the rows, will improve. -->

```{r}
knitr:: kable(cdrs_smd_with_cis, caption = "Table 6a. CDRS baseline mean adjsuted SMDs")
```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
knitr:: kable(hamd_smd_with_cis, caption = "Table 6. HAM-D baseline-mean adjusted SMDs")
```

### Supplement

```{r}

# Tables for baseline comparisons

# Lets make a table that combines across the sensitivity analyses for severity

df_severity_sens <- rbind(results_met_sev_cbt, results_met_sev_clin, results_met_sev_no_wl)

# For row names
df_severity_sens <- df_severity_sens %>%
  mutate(subgroup = case_when(
    row_number() == 1 ~ "CBT, fluoxetine and escitalopram",
    row_number() == 4 ~ "Excluding subclinical",
    row_number() == 7 ~ "Excluding waitlist",
    TRUE ~ if_else(subgroup == 0, "Medication", "Psychotherapy")
  ))

# Define new column names
new_column_names <- c(" ", "K", "Mean", "SE", "Lower CI", "Upper CI", "T2", "p-value")

# Assign new column names to the data frame
colnames(df_severity_sens) <- new_column_names

# Rounding
df_severity_sens <- df_severity_sens %>%
  mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
  mutate_at(vars(ncol(.)), ~ round(., 3))

colnames(df_severity_sens)[8] <- paste0(colnames(df_severity_sens)[8], footnote_marker_number(1))

df_severity_sens %>%
  kbl(align = "l", caption = "Table X: Sensitivity analyses for comparison of baseline severity levels between psychotherapy and medication trials", escape = F) %>%
  kable_material(bootstrap_options = c("hover", "condensed", "bordered"), full_width = FALSE) %>%
  row_spec(row = 0, background = "#EBEBEB", extra_css = "border-bottom: 1px black;" ) %>% 
  row_spec(row = 1, bold = T) %>% 
  row_spec(c(4, 7), bold = T, extra_css = "border-top: 1px solid black;") %>%
  column_spec(1, border_right = T) %>% 
  add_indent(c(2, 3, 5, 6, 8, 9)) %>% 
  footnote(number = c("Statistic derived from the test for subgroup differences in the relevant meta-analysis"))

# Now gender

df_gender_sens <- rbind(results_met_perc_women_cbt, results_met_perc_women_clin, results_met_perc_women_no_wl)

# For row names
df_gender_sens <- df_gender_sens %>%
  mutate(subgroup = case_when(
    row_number() == 1 ~ "CBT, fluoxetine and escitalopram",
    row_number() == 4 ~ "Excluding subclinical",
    row_number() == 7 ~ "Excluding waitlist",
    TRUE ~ if_else(subgroup == 0, "Medication", "Psychotherapy")
  ))

# Define new column names
new_column_names <- c(" ", "K", "Mean", "SE", "Lower CI", "Upper CI", "T2", "p-value")

# Assign new column names to the data frame
colnames(df_gender_sens) <- new_column_names

# Rounding
df_gender_sens <- df_gender_sens %>%
  mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
  mutate_at(vars(ncol(.)), ~ round(., 3))

colnames(df_gender_sens)[8] <- paste0(colnames(df_gender_sens)[8], footnote_marker_number(1))

df_gender_sens %>%
  kbl(align = "l", caption = "Table X: Sensitivity analyses for comparison of gender composition between psychotherapy and medication trials", escape = F) %>%
  kable_material(bootstrap_options = c("hover", "condensed", "bordered"), full_width = FALSE) %>%
  row_spec(row = 0, background = "#EBEBEB", extra_css = "border-bottom: 1px black;" ) %>% 
  row_spec(row = 1, bold = T) %>% 
  row_spec(c(4, 7), bold = T, extra_css = "border-top: 1px solid black;") %>%
  column_spec(1, border_right = T) %>% 
  add_indent(c(2, 3, 5, 6, 8, 9)) %>% 
  footnote(number = c("Statistic derived from the test for subgroup differences in the relevant meta-analysis"))

# Now age

df_age_sens <- rbind(results_met_age_cbt, results_met_age_clin, results_met_age_clin)

# For row names
df_age_sens <- df_age_sens %>%
  mutate(subgroup = case_when(
    row_number() == 1 ~ "CBT, fluoxetine and escitalopram",
    row_number() == 4 ~ "Excluding subclinical",
    row_number() == 7 ~ "Excluding waitlist",
    TRUE ~ if_else(subgroup == 0, "Medication", "Psychotherapy")
  ))

# Define new column names
new_column_names <- c(" ", "K", "Mean", "SE", "Lower CI", "Upper CI", "T2", "p-value")

# Assign new column names to the data frame
colnames(df_age_sens) <- new_column_names

# Rounding
df_age_sens <- df_age_sens %>%
  mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
  mutate_at(vars(ncol(.)), ~ round(., 3))

colnames(df_age_sens)[8] <- paste0(colnames(df_age_sens)[8], footnote_marker_number(1))

df_age_sens %>%
  kbl(align = "l", caption = "Table X: Sensitivity analyses for comparison of age between psychotherapy and medication trials", escape = F) %>%
  kable_material(bootstrap_options = c("hover", "condensed", "bordered"), full_width = FALSE) %>%
  row_spec(row = 0, background = "#EBEBEB", extra_css = "border-bottom: 1px black;" ) %>% 
  row_spec(row = 1, bold = T) %>% 
  row_spec(c(4, 7), bold = T, extra_css = "border-top: 1px solid black;") %>%
  column_spec(1, border_right = T) %>% 
  add_indent(c(2, 3, 5, 6, 8, 9)) %>% 
  footnote(number = c("Statistic derived from the test for subgroup differences in the relevant meta-analysis"))

# Plots  -----------------------

#define groups
df_long_for_metan <- df_long_for_metan %>%
  mutate(
    group = case_when(
      psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "Medication Active",
      psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "Medication Control",
      psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "Psychotherapy Active",
      psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "Psychotherapy Control",
      TRUE ~ NA_character_
    )
  )

group_colors <- c("Medication Active" = "deeppink1", "Medication Control" = "deeppink4", "Psychotherapy Active" = "steelblue1", "Psychotherapy Control" = "steelblue3")

# Plot cohens d per arm

ggplot(df_long_for_metan, aes(x = group, y = cohens_d, color = group)) +
  geom_point(position = position_jitter(width = 0.2), size = 3) +
  geom_segment(x = 4.5, xend = 4.5, y = -3.5, yend = -3,
               arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = "Psychotherapy Control", y = -3, label = "Less Effective", color = "grey", vjust = 0, hjust = .2) +
  geom_segment(x = 4.5, xend = 4.5, y = -4, yend = -4.5,
               arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = "Psychotherapy Control", y = -4.5, label = "More Effective", color = "grey", vjust = 0, hjust = .2) +
  scale_color_manual(values = group_colors) +
  labs(title = "Scatter plot of Cohen's d by group", x = NULL, y = "Cohen's d") +
  theme_minimal() +
  theme(legend.position = "none")

# Plot percentage baseline means

ggplot(df_long_for_metan, aes(x = group, y = perc_baseline_mean, color = group)) +
  geom_point(position = position_jitter(width = 0.2), size = 3) +
  scale_color_manual(values = group_colors) +
  labs(title = "Scatter plot of % baseline mean by group", x = NULL, y = "Percentage Baseline Mean") +
  theme_minimal() +
  theme(legend.position = "none")

# Box plot Cohens d by group

ggplot(df_long_for_metan, aes(x = group, y = cohens_d, fill = group)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(0.8), color = "black", alpha = 0.5) +
  scale_y_continuous(limits = c(-3.5, 1), breaks = seq(-4, 1, by = 1)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_segment(x = 4.5, xend = 4.5, y = -2.5, yend = -2,
               arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = "Psychotherapy Control", y = -2.2, label = "Less Effective", color = "grey", vjust = 0, hjust = .2) +
  geom_segment(x = 4.5, xend = 4.5, y = -3, yend = -3.5,
               arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = "Psychotherapy Control", y = -3.2, label = "More Effective", color = "grey", vjust = 0, hjust = .2) +
  scale_fill_manual(values = group_colors) +
  labs(title = "Box plot of cohens_d by group", x = "Group", y = "Cohen's d") +
  theme_minimal() +
  theme(legend.position = "none")

```