### 2024-04-19 Manuscript_for_rendering_in_word.qmd
Lots of formatting things, success in fixing up tables Next to address are issues with the figures

---
title: "Fundamental Problems with Child and Adolescent Depression Treatments"
subtitle: "A Quantitative Reappraisal of the Evidence Base"
author:
  - name: Argyris Stringaris
    email: a.stringaris@ucl.ac.uk
    affiliations: 
        - id: ucl_1
          name: University College London
          department: Divisions of Psychiatry and Psychology and Language Science
          address: 1-19 Torrington Place
          city: London, UK
          state: 
          postal-code: WC1E 7HB
          country: United Kingdom
    attributes:
        corresponding: true
    note: 
  - name: Charlotte Burman
    email: charlotte.burman@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
  - name: Dayna Bhudia
    email: dayna.bhudia@ucl.ac.uk
    affiliations:
        - ref: ucl_1
  - name: Despina Miliou
    email: dgmiliou@outlook.com.gr
    affiliations:
        - ref: ucl_1
  - name: Giannis Rokas
    email: <rokasgiannis93@gmail.com
    affiliations:
        - ref: ucl_1
  - name: Lucy Foulkes
    email: lucy.foulkes@psych.ox.ac.uk
    affiliations:
        - id: OU
          name: University of Oxford
          department: Experimental Psychology
          address: Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Road
          city: Oxford
          postal-code: OX2 6GG
          country: United Kingdom
    note:
  - name: Carmen Moreno
    email: cmoreno@hggm.es
    affiliations:
        - id: HGM
          name: Hospital Gregorio Marañón
          department: Department of Psychiatry
          address: 46 C. del Dr. Esquerdo
          city: Madrid
          state:
          postal-code: 28007
          country: Spain
  - name: Samuele Cortese
    email: samuele.cortese@soton.ac.uk
    affiliations:
        - id: soton
          name: University of Southampton
          department: Centre for Innovation in Mental Health (CIMH), School of Psychology
          address: Highfield Campus, Building 44
          city: Southampton
          state:
          postal-code: SO17 1BJ
          country: United Kingdom
  - name: Georgina Krebs
    email: g.krebs@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
keywords: 
  - depression
  - adolescents
  - antidepressants
  - psychotherapy
date: last-modified
format:
  docx:
    reference-doc: custom-reference-doc.docx
editor: visual
bibliography: "emotion_concepts_paper.bib"
font:
  family: "Calibri"
---

##

------------------------------------------------------------------------

------------------------------------------------------------------------

```{r, warning=FALSE, message = FALSE, echo=FALSE}

###########Libraries needed
library(tidyverse)
library(truncnorm)
library(metafor)
library(meta)
library(kableExtra)
library(flextable)
library(magrittr)
library(officer)
library(RefManageR)
library(ftExtra)
library(dplyr)

# For referencing
bib_data <- ReadBib("emotion_concepts_paper.bib")

```

# Introduction

There are two principal treatment modalities for adolescent depression: psychotherapy and medication. Where does one turn to find the evidence that will inform their treatment options? This question is relevant for patients and their carers, for clinicians and for policy makers. However, the question is particularly difficult to answer for adolescent depression, where there are limited data from head-to-head trials of medication and psychotherapy, and where recommendations must therefore be derived from indirect comparisons of treatment efficacy.

Guidelines, such as the ones that the internationally influential UK National Institute of Health and Care Excellence (NICE) have produced for adolescent depression, are a principal source of such information. NICE recommend that youth are first offered psychological therapy (specifically cognitive behaviour therapy and interpersonal therapy) over medication for most presentations of depression [@niceDepressionChildrenYoung2023]. This conclusion is in keeping with two sources of evidence: medication meta-analyses that cast doubt on the efficacy of most antidepressants, with the exception of fluoxetine, for child and adolescent depression [@ciprianiComparativeEfficacyTolerability2016]; and psychotherapy meta-analyses that conclude psychotherapy to be efficacious for child and adolescent depression [@cuijpersEffectsPsychologicalTreatments2021]. However, such conclusions seem at odds with those of a recent network meta-analysis [@zhouComparativeEfficacyAcceptability2020], an established method of comparing treatments with each other using both direct (head-to-head) and indirect (indirectly treatment A with treatment C, via studies that directly compare treatments A with B and B with C) evidence. Indeed, the network meta-analysis concluded that only fluoxetine alone and fluoxetine administered together with CBT were significantly more effective than medication control (i.e. pill placebo) or psychotherapy controls. Given this confusing evidence base, how should patients, carers, clinicians and policy makers make treatment decisions for adolescent depression?

In this paper, we examine whether the existing evidence base for adolescent depression treatment can offer valid answers to such questions. Below, we provide a conceptual framework for answering such questions and test a series of hypotheses using data from existing trials. Two points are crucial to indirect comparisons of treatment modalities with each other.

First, whether the participants of trials in one modality are comparable to those in another modality. Second, whether key conditions of the trial, such as the effects of control conditions or the number of sites involved in a trial, are comparable.

Starting with the first point, to be able to compare between different trials one must assume that these trials sample from the same population. If they do not, then the validity of any comparisons, including those conducted through network metanalysis (which rests on the principal of transitivity, i.e. the requirement that the different sets of randomized trials are similar on average [@chaimaniChapter11Undertaking2023]) are questionable (we provide a more formal account of this in the Supplement).

Indeed, comparing outcomes (Y) between psychotherapy and medication trials requires us to contrast what is called the sample average treatment effect $\tau$ of each, defined in the following way:

$$
\tau_{psy} =  E(Y_{1} - Y_{0} \,|\, S, psy)
$$ $$
\tau_{med} =  E(Y_{1} - Y_{0} \,|\, S, med)\ 
$$

where the operator E denotes the expectation over the differences in outcomes between those who received the intervention (T = 1) and those who received the control condition (T = 0), for each sample, $S$, where psy and med stand for psychotherapy and medication respectively.

Obviously, this comparison rests on the assumption that trials in both modalities sample from the same population, $P$, of patients. Formally, this can be expressed as follows:

$$E(Y_{1} \,|\, \text{S, psy}) = E(Y_{1} \,|\, P, \text{psy})$$

$$E(Y_{1} \,|\, \text{S, med}) = E(Y_{1} \,|\, P, \text{med})$$

signifying that the effect found in the population would be expected to be found in the same population for each treatment.

This assumption, that medication and psychotherapy trials sample from the same population, may be hard to meet. Clinical experience and empirical evidence indicate that patients and parents often have preferences between psychotherapy and medication [@jaycoxAdolescentPrimaryCare2006; @langerParentYouthPreferences2021; @mchughPatientPreferencePsychological2013], meaning that there is likely to be a self-selection bias with respect to who participates in psychotherapy versus medication trials. Moreover, these treatment preferences correlate with clinically-relevant characteristics of the participants, including severity, sex and comorbidity. Some of these characteristics, particularly baseline severity, may moderate treatment response [@courtneyPredictorsModeratorsMediators2022; @lorenzo-luacesDoubleTroubleSymptom2020] and therefore, if they differ across psychotherapy and medication trials, they may confound comparisons.

In terms of the second point, differences in trial design may impact outcomes in a differential way between antidepressant and psychotherapy trials [@delgiovaneCombiningPharmacologicalNonpharmacological2019]. Perhaps the most obvious way in which this happens is the fact that participants in psychotherapy trials are unblinded in terms of treatment allocation, with the exception perhaps of trials that compare two equally plausible treatment arms [@calvoInterventionAdolescentsEarlyOnset2014]. By contrast, in new antidepressant trials, patients (and raters) were found to be unlikely to be able to judge treatment allocation [@linAssessmentBlindingRandomized2022]. This creates differential expectations between medication and psychotherapy trials. Importantly, pill-placebo has been developed as the standard medication control in order to match as closely as possible the intervention condition, so as to minimise differences in expectancy between conditions. By contrast, psychotherapy control arms vary across waitlist controls, treatment as usual and active controls. These differences in expectation favour the psychotherapy active condition and disadvantage the psychotherapy control (thus potentially magnifying the difference between them). This is because of what has been aptly described as "disappointment bias" [@thespininvestigatorsEthicsTrialsCohorts2017], where participants in the active condition are content for receiving the "cutting edge treatment" whilst those in the control are dissatisfied for having missed out on "proper treatment". Expectancy is substantially associated with treatment outcomes [@constantinoExpectations2011]; i.e., it is an effect modifier. If expectancy effects are different for psychotherapy controls and pill placebo, comparisons between them, including in network metanalysis, become questionable.

Another difference in design that has potential implications for which patients are selected into the trial is the number of sites in a trial. Previous research has shown that in medication trials the number of sites is positively related to the magnitude of the placebo response [@bridgePlaceboResponseRandomized2009; @dechartresSingleCenterTrialsShow2011]. This phenomenon may be due to lower quality of phenotyping in such studies, with higher rates of classification errors, entailing higher rates of spontaneous remission or regression to the mean. It is therefore important to examine whether this basic trial design feature differs between medication and psychotherapy trials.

An inter-related issue concerns the nature of control interventions being used. Control conditions in trials are meant to generate counterfactual conditions [@guoChapterCounterfactualFramework2014] to the intervention: what would have been the outcome in an individual had they not received the intervention, with all else being equal. A control condition such as pill placebo, where the appearance of the drug is faithfully emulated, is an effort for all else to be equal. The question is whether psychotherapy trials make a similar effort: are for example the number of hours of contact with a human comparable between active and control arms in psychotherapy trials?

Understanding the nature of controls for each treatment modality is important for a similar reason. Often, in the public domain, psychotherapy and medication are compared to each other on the basis of their respective effect sizes. However, these effect sizes represent differences between the active intervention and the control condition (placebo or psychotherapy control). For these effect sizes to be comparable, placebo and psychotherapy controls ought to be the equal. Otherwise, misleading conclusions could be drawn, e.g. two effect sizes of 40% would be considered equal, even if one arose from a difference of 100% - 60% and another arose from a difference of 40% - 0%.

In this paper, we examine randomised controlled trials (RCTs) of psychotherapy and medication for child and adolescent depression in the following ways. First, we compare key baseline characteristics of medication vs psychotherapy trials, specifically the extent to which they are comparable in a) baseline severity of depression; b) percentage females vs males; c) mean age; and d) number of sites involved. Second, we determine the standardised mean differences of psychotherapy controls vs medication controls using random effects metanalysis. Third, we examine the quality of psychotherapy control conditions by scrutinising the extent to which they are matched to the active intervention in ways such as number or frequency of sessions, and therefore, whether they represent fair pairings from which to draw valid efficacy inferences.

# Method

This study was registered on the Open Science Framework (OSF); the protocol can be found [here](https://osf.io/bfmc6). In Supplement X we outline deviations from this pre-registered protocol.

## Studies included

We primarily drew upon RCTs included in two recent comprehensive meta-analyses with open data available for each medication and psychotherapy, and supplemented them with an updated systematic review, described below and in extensive detail in the supplement. Please refer to these original meta-analyses for a detailed description of their search strategy and study selection criteria. Psychotherapy studies were drawn from a systematic review and meta-analysis of randomised trials comparing psychotherapy for youth depression against control conditions [@cuijpersEffectsPsychologicalTreatments2021]. Cuijpers et al. made available a full dataset of psychotherapy trials (via https://www.metapsy.org/), which we used for the current study. Whilst Cuijpers et al. (2023) excluded those studies for which the primary outcome variable could not be calculated due to missing data, we included these studies and performed the imputations outlined below; hence we have more psychotherapy studies included in this review compared to the original meta-analysis. Whilst the online database is regularly updated, we chose to exclude studies published after the final date of Cuijpers et al.'s (2023) literature search.

Medication studies were drawn from a network meta-analysis examining the efficacy and tolerability of a range of antidepressants and placebo for major depressive disorder in children and adolescents [@ciprianiComparativeEfficacyTolerability2016]. A dataset was made available online though did not include means or standard deviations at baseline or post-test. We were unable to access the full dataset used in this meta-analysis, and hence completed extraction from the included studies ourselves. We excluded three studies because they did not include a control arm. We were unable to locate and therefore complete extraction for two papers. Many studies did not report complete data; we contacted all corresponding authors to request missing data, though did not receive any responses.

We conducted a systematic search for medication studies published after the final search date of Cipriani's (2016) review up to the final search date of Cuijpers et al's (2023) review (May 31 2015 up to Jan 1 2021) to ensure we analysed an equivalently up-to-date database of medication trials. We searched PubMed, the Cochrane Central Register of Controlled Trials, Embase, Web of Science, CINAHL, PsycINFO and LiLACS for randomised controlled trials (RCTs) comparing any antidepressant with placebo in the treatment of children and adolescents with a primary diagnosis of major depressive disorder. We also searched trial registers for published and unpublished studies. We used the same search terms as Cipriani (2016) with one additional search term to include only placebo-controlled trials (see Supplement X for further details). We used Covidence, an online software tool, to manage our systematic review. In total, our search produced 538 studies, 88 of which were duplicates and subsequently removed. Two study authors independently completed title and abstract screenings for 450 studies, and full text screening for 38 studies. 7 studies met inclusion criteria and data extraction was completed for these papers. Clinical trial registers were also systematically searched however all RCTs meeting inclusion criteria had already been identified from the database search outlined above.

## Statistical Analysis

### Trial and sample characteristics

We conducted a series of random effects meta-analyses and tested for subgroup differences between psychotherapy versus medication trials in sample characteristics including sex, age, and severity of depressive symptoms at baseline. Meta-analyses were implemented using R's Meta package.

In order to compare depression severity across the variety of instruments the studies used, we performed a min-max normalisation to turn each study arm mean score at baseline into a percentage using the following formalism:

$$
\text{outcome}_{percent} = \frac{\bar{X} - \text{scale}_{min}}{\text{scale}_{max} - \text{scale}_{min}}
$$

where, $$\bar{X}$$ is the mean score for each study arm on the primary outcome questionnaire, and ${scale}_{min}$ and ${scale}_{max}$ are the minimum and maximum possible values of the scale in question, respectively. The standard deviation is calculated thus:

$$
\text{SD}_{Xpercent} = \frac{\text{SD}_{X}}{\text{scale}_{max} - \text{scale}_{min}}
$$

where $\text{SD}_{X}$ is the original standard deviation of the mean at baseline.

We also conducted a t-test to compare mean number of trial sites between psychotherapy and medication trials.

### Measures of Effect

As the measure of effect of each individual study, we used the within-group Standardised Mean Difference (SMD), which we defined following [@lakensCalculatingReportingEffect2013, @cummingUnderstandingNewStatistics2013] as:

$$SMD_{change} = \frac{Mean_{t_{2}} - Mean_{t_{1}}}{\frac{SD_{t{2}} + SD_{t{1}}}2}\ $$

where, $Mean_{t_{2}}$ and $Mean_{t_{1}}$ refer to the means of the main outcome score at the end and beginning of the intervention respectively and $SD_{t_{2}}$ and $SD_{t_{1}}$ to the respective standard deviations. Where individual studies did not report all data required to calculate the SMD, we imputed missing data according to the methods summarised in this Cochrane Handbook [@higginsChapterChoosingEffect2023], in the following order. If a study reported the standard error of the mean, the SD was obtained simply by multiplying the SE by the square root of the sample size. For conditions where the SD was missing at one time point, the baseline SD was substituted by the post-test SD, and vice versa. If the SD was not available at either time point, missing values were replaced by the mean of the SDs available for comparable cases (defined as same trial type (psy or med), same instrument, same timepoint (pre or post), and same arm (control or active)). Where there were missing means at either baseline or post-test, missing values were calculated using mean change scores, preferring the change scores reported in the paper itself, though where this was unavailable, using the change scores reported in the dataset from Cipriani's meta-analysis.

For the purposes of metanalysis, it is necessary to estimate a standard error of the SMD. This is calculated according to:

$$ SE_{SMD} = \sqrt{\frac{2(1 -r_{t_{1}t_{2}})}{n} + \frac{SMD^2}{2n}}$$

where $n$ refers to the study sample size and $r_{t_{1}t_{2}}$ refers to the correlation between the outcome score obtained at baseline and at the end point. This correlation is typically not reported in studies and is often imputed using previously reported correlations for the instruments used. However, this practice has given rise to concerns about misestimation. Whilst such misestimation is possible, there is no reason to expect that it would be systematic, i.e. bias estimation of the effects for the control group of medication compared to those of psychotherapy. Still, to alleviate such concerns we have used a simulations.

In particular, we simulated one thousand truncated distribution of standard errors with the following general characteristics:

$$r_{t_{1}t_{2}} \sim \mathcal{TN}(\mu, \sigma, a, b)$$

for which we chose the mean to be $\mu = 0.65$, the standard deviation to be $\ sigma = 0.2$, and the upper and lower bounds to be $a = 0.45$ and $b = 0.9$, respectively. We then used these simulated datasets in the subsequent metanalyses.

### Random Effects Metaregression

We estimated the pooled standardised mean difference for each arm by using a random effects metanalysis implemented in R's metafor package. The main underlying assumption of random effects metanalysis is that each study's true effect size $\theta_{k}$ is affected not only by sampling error $\epsilon_{k}$, but also by $\zeta$ which represents heterogeneity between studies, allowing each study's estimate to vary along a distribution of effects, and the distribution of true effect sizes termed $\tau^2$. Therefore, we can estimate a two stage model with:

$$Y_i \sim \mathcal{N}(\theta_i, \sigma_i^2)
$$

$$\theta_i \sim \mathcal{N}(x_i\beta, \tau_i^2)
$$

where $Y_i$ is the estimated effect size for study i, has a normal distribution with $\theta_i$ as its true mean effect and sampling error $\sigma^2$. Whereas $\theta_i$ is a study-specific instantiation of the distribution of effect sizes, with $\tau^2$ representing heterogeneity.

This then gives rise to:

$$Y_i = x_i\beta_i + u_i + \epsilon_i$$

where,

$$u_i \sim N(0, \tau^2) $$

describes the deviation of each study from the mean of the distribution, and,

$$\epsilon_i \sim N(0, \sigma^2)$$

describes the sampling error.

We can then specify the following model to obtain the means of each arm of the trials as follows:

$$ \begin{aligned}
    Υ_i &=
    \begin{cases}
        0 & MedControl:  b_0 + u_i + \epsilon_i\\
        1 & MedActive:   b_0 + b_{1_{i}} +  u_i + \epsilon_i \\
        2 & PsyActive:   b_0 + b_{2_{i}}  +  u_i + \epsilon_i \\
        3 & PsyControl:  b_0 + b_{3_{i}}  +  u_i + \epsilon_i \\
    \end{cases}
\end{aligned}$$

where to obtain the mean of each level is the sum of $b_0$, the intercept for the reference category of medication control, with the coefficient of each level, e.g. for level 3, $b_{3_{i}}$ the psychotherapy controls. The confidence intervals of the means are constructed in the standard way using the standard errors of the mean. Similarly, each coefficient represents the contrast between the reference category and each level, for an example and of main interest to us $b_{3_{i}}$ represents the contrast between psychotherapy and medication control arms. Inference on the contrasts is done as follows:

$$
z = \frac{\hat{\beta}}{\text{SE}(\hat{\beta})}
$$

We used maximum likelihood (ML) to estimate model and applied Hartung-Knapp adjustment to reduce the chance of false positives [@inthoutHartungKnappSidikJonkmanMethodRandom2014].

We present the SMDs of each level of the dummy variable of the four treatment arms (medication control, medication active, psychotherapy control, psychotherapy active) under investigation. The SMDs are the means across the 1000 simulated datasets.

### Sensitivity Analyses

```{r, warning=FALSE, message = FALSE, echo = FALSE, output = FALSE }
# my alpha for the z 
target_probability <- 1 - 0.05

#start here 
x <- 1

# two decimals should be fine
tolerance <- 1e-3

#find the z value using the pnorm function
while (abs(pnorm(x)) < target_probability ) {
  x <- x + tolerance  
}

# Adjust the step size based on your problem  
cat("critical value of z =", x, "corresponding to an alpha = 0.05." , "\n")

x

```

We then conducted a series of sensitivity analyses of our results. For each of the metanalyses we excluded 1) studies which recruited participants with subclinical levels of depression and 2) studies that used waitlist as their control condition.

Next, we compared the control and active arms of studies where either the CDRS or the HAM-D where used as outcome instruments.

Further, we tested whether the simulated values for the standard error had a substantial influence on the estimation of the differences between the medication vs psychotherapy control condition. To inspect whether this is the case, we plotted the z-value of the difference between the two coefficients against the number of simulations. We make inference on the stability of the difference, by counting the proportion of times that the z-value is above the critical value of z = `r x` corresponding to an alpha = 0.05.

### Comparing the control versus active arms of psychotherapy trials

We ran t-tests to compare the active versus control arms of psychotherapy trials on key variables of interest regarding the nature and intensity of the interventions. We extracted data pertaining to the number, duration and intensity of sessions, and the total cumulative hours and period of the intervention. Where a range was provided, the maximum was encoded (e.g. if a paper reported that an intervention involved 8-10 sessions lasting 50-60 minutes, we encoded the number and duration of sessions as 10 and 60, respectively). If sessions varied in frequency across an intervention, we calculated an average by dividing total number of sessions by length of intervention period. Similarly, if the length of sessions varied across the course of the intervention, we calculated a weighted average. Phone call, web-chat and online sessions were encoded as sessions, however guided self-help components were not.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#############Some custom made functions#############

###### 1. Simulation Function for SEs
simulate_dataframes_for_st_errors <- function(df, num_repetitions, seed, n, a, b, mean, sd) { # n refers to the number
                                                                                              # of unique ids to which a correlation coef
                                                                                              # is allocated.
  set.seed(seed)

  # Empty list to store simulated dfs
  list_df_simulated <- list()

  for (i in 1:num_repetitions) {
    # Simulate the vector
    simulated_correlations_vector <- truncnorm::rtruncnorm(n, a, b, mean, sd)  #using the truncnorm to create correlations

    # Create a copy of the original dataframe
    df_simulated_copy <- df

    # Add/update the sims column
    df_simulated_copy$correlation_sim_values <- simulated_correlations_vector[match(df_simulated_copy$new_study_id, unique(df_simulated_copy$new_study_id))]

    # Remove duplicated columns
    df_simulated_copy <- df_simulated_copy[, !duplicated(colnames(df_simulated_copy))]

    # Calculate the ses
    df_simulated_copy <- df_simulated_copy %>% 
      group_by(new_study_id, baseline_n) %>% 
      mutate(simulated_se = sqrt(((2*(1-correlation_sim_values))/baseline_n) + 
                                   (cohens_d^2/(2*baseline_n))))

    # Add the simulated dataframe to the list
    list_df_simulated[[i]] <- df_simulated_copy
  }

  return(list_df_simulated)
}

###############2. Simulation function for metaregression
run_metaregression <- function(list_of_datasets, model_formula) {
  list_model_meta_reg <- list()

  for (i in seq_along(list_of_datasets)) {
    list_model_meta_reg[[i]] <- metafor::rma(yi = cohens_d,
                                             sei = simulated_se,
                                             data = list_of_datasets[[i]],
                                             method = "ML",
                                             mods = model_formula,
                                             test = "knha")
  }

  return(list_model_meta_reg)
}

#########2. Function to count studies with missingness etc.

count_studies <- function(df) {
  df_count_studies <- df %>%
    group_by(psy_or_med, arm_effect_size) %>%
    count(is.na(cohens_d)) %>%
    rename(is_missing = `is.na(cohens_d)`) %>%
    mutate(
      condition = case_when(
        psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
        psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
        psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
        psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control"
      )
    )

  df_count_studies$condition <- factor(df_count_studies$condition)  # turn to factor

  # relevel so that medication control becomes the reference category for the regression
  df_count_studies$condition <- relevel(df_count_studies$condition, ref = "medication_control")

  return(df_count_studies)
}

##########3. Function to extract mean coefficients and to extract model characteristics

aggregate_model_results <- function(list_model_1_meta_reg, condition) {
  df_coefs <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  df_se <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  df_z_value <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  df_lower_ci <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  df_upper_ci <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  tau_sq <- rep(0, length(list_model_1_meta_reg))
  i_sq <- rep(0, length(list_model_1_meta_reg))
  k <- rep(0, length(list_model_1_meta_reg))
  r_sq <- rep(0, length(list_model_1_meta_reg))

  for (i in 1:length(list_model_1_meta_reg)) {
    for (j in 1:length(condition)) {
      df_coefs[i, j] <- list_model_1_meta_reg[[i]]$beta[[j]]
      df_se[i, j] <- list_model_1_meta_reg[[i]]$se[[j]]
      df_z_value[i, j] <- list_model_1_meta_reg[[i]]$zval[[j]]
      df_lower_ci[i, j] <- list_model_1_meta_reg[[i]]$ci.lb[[j]]
      df_upper_ci[i, j] <- list_model_1_meta_reg[[i]]$ci.ub[[j]]
      tau_sq[i] <- list_model_1_meta_reg[[i]]$tau2
      i_sq[i] <- list_model_1_meta_reg[[i]]$I2
      k[i] <- list_model_1_meta_reg[[i]]$k
      r_sq[i] <- list_model_1_meta_reg[[i]]$R2
    }
  }

  df_coefficients_model <- data.frame(
    coefficients = colMeans(df_coefs),
    se = colMeans(df_se),
    z_value = colMeans(df_z_value),
    lower_ci = colMeans(df_lower_ci),
    upper_ci = colMeans(df_upper_ci),
    tau_sq = mean(tau_sq),
    i_sq = mean(i_sq),
    k = mean(k),
    r_sq = mean(r_sq)
  )

  return(df_coefficients_model)
}

###########4. Function to extract the SMDs  (SMD, ses, CI) for each level of the dummy variable. The output is used below in 5.
# it is a slightly awkward one, because I couldn't' come up with a better way to add the intercept to each coefficient
# given the named output of the built in coef function. here I extract the coefficients

extract_coefficients_func <- function(df_with_coefs){
  list_coefs <- list()
  coefs <- 0
  coefficient_output <- 0

  for(i in 1: length(coefficients(df_with_coefs))){

    temp_vec <- c(0, rep(coefficients(df_with_coefs)[[1]], 
                         length(coefficients(df_with_coefs)) - 1 ))
    coefs[i] <- coef(df_with_coefs)[[i]]

    coefficient_output <- temp_vec + coefs

    st_error_output <- df_with_coefs$se

    df_coefficients <- data.frame(cbind(coefficients = coefficient_output, se = st_error_output ))

    df_coefficients$lower_bound <- df_coefficients$coefficients - 1.96*df_coefficients$se
    df_coefficients$upper_bound <- df_coefficients$coefficients + 1.96*df_coefficients$se
  }
  return (df_coefficients )
}

############5. Function to get the means out of the SMDs (this is similar to Func 3 and I should at some point create one more general one)
# here the argument list_of_dfs is the list of extracted coefficients that you get out of the extract_coefficients_func
# which itself you get from the run_metaregression function. The argument conditions refers to the levels of the variable that describes each study arm

calculate_mean_coefs_ses <- function(list_of_dfs, conditions) {
  df_coefs <- matrix(NA, nrow = length(list_of_dfs), ncol = length(conditions))
  df_se <- matrix(NA, nrow = length(list_of_dfs), ncol = length(conditions))

  for (i in 1:length(list_of_dfs)) {
    for (j in 1:length(conditions)) {
      df_coefs[i, j] <- list_of_dfs[[i]]$coefficients[j]
      df_se[i, j] <- list_of_dfs[[i]]$se[j]
    }
  }

  df_mean_coefs_ses <- data.frame(
    coef_means = colMeans(df_coefs),
    se_means = colMeans(df_se)
  )

  df_mean_coefs_ses$lower_ci <- df_mean_coefs_ses$coef_means - 1.96 * df_mean_coefs_ses$se_means
  df_mean_coefs_ses$upper_ci <- df_mean_coefs_ses$coef_means + 1.96 * df_mean_coefs_ses$se_means

  return(df_mean_coefs_ses)
}

###############6. Function for plotting the SMDs
plot_means_function <- function(dataframe, subtitle) {
  dataframe$text_label <- 0

  for (i in 1:nrow(dataframe)) {
    dataframe$text_label[i] <- paste(
      dataframe$condition[i], "\n",
      round(dataframe$coef_means[i], 2),
      "[" ,
      round(dataframe$lower_ci[i], 2), 
      round(dataframe$upper_ci[i], 2),
      "]"
    ) 
  }

  # Recode condition for ease of plotting
  dataframe$condition <- str_replace_all(dataframe$condition, "_", " ")

  # Encode colours
  redish_palette <- c("medication control" = "deeppink", "medication active" = "deeppink3")
  blueish_palette <- c("psychotherapy active" = "steelblue1", "psychotherapy control" = "steelblue3")

  # Plotting
  plot_means <- ggplot(dataframe, aes(x = coef_means, y = 1:nrow(dataframe),
                                      colour = condition, label = text_label)) +
    geom_errorbar(aes(xmin = lower_ci, xmax = upper_ci), width = 0.2, position = position_dodge(0.5)) +
    geom_text(vjust = +1.5, size = 4) +
    scale_size_continuous(guide = "none") +
    guides(colour = FALSE) + 
    scale_color_manual(values = c(redish_palette, blueish_palette)) +
    theme_minimal() +
    labs(x = "TE-random", y = NULL, title = "Adolescent Depression Trial Efficacy by Treatment Arm",
         subtitle = subtitle) +
    xlab("Standardized Mean Difference (SMD) with 95% CIs") +
    ylab("") +
    ylim(0, nrow(dataframe) + 1) +
    xlim(-3.0, 0.5) +
    geom_vline(xintercept = 0, linetype = "dashed", size = 1.5, colour = "grey") +
    geom_segment(x = -1.25, xend = -1.75, y = 4.7, yend = 4.7, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
    geom_text(x = -1.55, y = 4.9, label = "More Effective", color = "grey", vjust = 0.5, hjust = 1) +
    geom_segment(x = -1.35, xend = -0.85, y = 4.9, yend = 4.9, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
    geom_text(x = -1.05, y = 5.05, label = "Less Effective", color = "grey", vjust = 0.5, hjust = 0)+
    theme(axis.title.x = element_text(size = 12), 
          axis.text.x  = element_text(size = 12),
          axis.text.y = (element_blank()),
          plot.title = element_text(size = 12),
          plot.subtitle = element_text(size = 12)) +
    geom_curve(aes(x = 0.18, y = 4.73, xend = 0.02, yend = 4.5), color = "grey", curvature = -0.2, arrow = arrow(length = unit(0.25, "cm"), type = "closed")) +
    geom_text(x = 0.2, y = 4.75, label = "No Effect", color = "grey", vjust = 0.5, hjust = 0)

  return(plot_means)
}

###########7. Function for baseline SMD results
library(broom)
run_metareg_and_extract <- function(dataframe, model_1, condition) {
  # Run metareg
  inst_baseline_metareg <- rma(yi = baseline_mean,
                               sei = baseline_stand_error,
                               data = dataframe,
                               method = "ML",
                               mods = model_1,
                               test = "knha")

  # Extract coefficients from the list
  baseline_metareg_results <- tidy(inst_baseline_metareg)
  baseline_metareg_results$condition <- condition
  coefs_for_inst_baseline <- tidy(inst_baseline_metareg)$estimate

  # Calculate SMDs
  smds_for_inst_baseline <- c(coefs_for_inst_baseline[1], coefs_for_inst_baseline[-1] + coefs_for_inst_baseline[1])

  # Calculate upper and lower CIs
  upper_cis_for_inst_baseline <- smds_for_inst_baseline + 1.96 * tidy(inst_baseline_metareg)$std.error
  lower_cis_for_inst_baseline <- smds_for_inst_baseline - 1.96 * tidy(inst_baseline_metareg)$std.error

  # Create dataframe with results
  df_baseline_means_with_cis <- data.frame(
    baseline_smds = smds_for_inst_baseline,
    upper_cis_smds = upper_cis_for_inst_baseline,
    lower_cis_smds = lower_cis_for_inst_baseline,
    condition = condition
  )

  # Return the two dataframes
  return(list(baseline_metareg_results = baseline_metareg_results, df_baseline_means_with_cis = df_baseline_means_with_cis))
}

```

```{r, warning=FALSE, message = FALSE, echo=FALSE}

# Transforming data ------------------------------------------------------
# import master dataset

df_appl_v_orange <-readxl:: read_excel("df_appl_v_orange.xlsx")

# Here we add in instrument end point values (mins / maxs).
# We will use these to transform the scale into % as follows: (observed score - min of the scale)/(max-min)
# We have done this for both means and standard deviations.

df_scale_ranges <- read.csv("scale_ranges.csv")
df_appl_v_orange <- full_join(df_appl_v_orange, df_scale_ranges, by = "instrument_name")

df_appl_v_orange <- df_appl_v_orange %>% 
  mutate(perc_baseline_mean_active = (baseline_mean_active - scale_min) / (scale_max - scale_min)) %>% 
  mutate(perc_baseline_sd_active = baseline_sd_active / (scale_max - scale_min)) %>%  # this is the best way to estimate the SD of the linear transformation of the score 
  mutate(perc_baseline_mean_control = (baseline_mean_control - scale_min) / (scale_max - scale_min)) %>% 
  mutate(perc_baseline_sd_control = baseline_sd_control  / (scale_max - scale_min)) %>% 
  mutate(perc_post_mean_active = (post_mean_active - scale_min) / (scale_max - scale_min)) %>% 
  mutate(perc_post_sd_active = post_sd_active  / (scale_max - scale_min)) %>% 
  mutate(perc_post_mean_control = (post_mean_control - scale_min) / (scale_max - scale_min)) %>% 
  mutate(perc_post_sd_control = post_sd_control  / (scale_max - scale_min))

# Need to use SMDs (ie our Cohen's d) and then use standard error of SMD, to achieve this I need test-retest reliabilities of instruments.

# CDRS reliability taken from here https://www.liebertpub.com/doi/epdf/10.1089/104454601317261546 
# Using a   2-week interval, and different psychiatrists from the first to the second assessment, 
# Poz-nanski et    al. (1984) demonstrated high reliability (r=   0.86) 
# for the CDRS-R total score in 53 clinic-referred 6- to 12-year-olds.

### A few more tidying things from Argyris before doing metanalyses
# Discovered some errors in the percentage women variable. Have gone back and checked. 
df_appl_v_orange[df_appl_v_orange$study_ID=="Fristad, 2019_cbt + placebo_placebo",]$percent_women <-43.1
df_appl_v_orange[df_appl_v_orange$study=="Clarke, 1995",]$percent_women <-47.3
df_appl_v_orange[df_appl_v_orange$study=="Moeini, 2019",]$percent_women <-100
df_appl_v_orange[df_appl_v_orange$study=="Santomauro, 2016",]$percent_women <-40

# Calculate SE for baseline severity --------------------------------------

df_appl_v_orange$baseline_st_error_active <- 
  df_appl_v_orange$baseline_sd_active/sqrt(df_appl_v_orange$baseline_n_active)

df_appl_v_orange$baseline_st_error_control <-
  df_appl_v_orange$baseline_sd_control/sqrt(df_appl_v_orange$baseline_n_control)

# Turn into a long database with unique rows ------------------------------

### Important: create a dataset that will have unique control studies (see problem that we identified, namely common control conditions)
# create new id to help with better identification and work with duplicates (see below) 
df_appl_v_orange  <- df_appl_v_orange  %>%
  mutate(new_study_id = case_when(psy_or_med == 0 ~ paste(study,year, sep = ", "),
                                  .default = study ))

# We are going to use the descr_active and descr_control columns rather than active_type and control_type columns, as these have unique values for each study. However the descr columns are empty for med studies, so I'm going to paste values across to fix this error.

df_appl_v_orange <- df_appl_v_orange %>% 
  mutate(descr_active = if_else(is.na(descr_active ), active_type , descr_active ))

df_appl_v_orange <- df_appl_v_orange %>% 
  mutate(descr_control = if_else(is.na(descr_control), control_type, descr_control ))

# Step 1: keep only the rows with the top instrument in our hierarchy
df_with_distinct_instruments <-  df_appl_v_orange %>%          
  group_by(new_study_id) %>% 
  filter(instrument_value == min(instrument_value)) # coded for the smallest number to be best.

write.csv(df_with_distinct_instruments, "df_with_distinct_instruments.csv")

# Step 2: turn dataframes to long

# A small bit of cleaning so the code below will work

df_with_distinct_instruments <- df_with_distinct_instruments %>% 
  rename (percent_women_active = active_percent_women, 
          percent_women_control = control_percent_women,
          percent_women_overall = percent_women)

# A: turn long the rows with active
active_df <- df_with_distinct_instruments %>% 
  dplyr:: select(new_study_id, active_type, descr_active, psy_or_med, baseline_n_active, cohens_d_active,
                 baseline_mean_active, baseline_sd_active, post_mean_active, post_sd_active,
                 perc_baseline_mean_active, perc_baseline_sd_active, perc_post_mean_active, perc_post_sd_active, percent_women_overall, percent_women_active, age_m_active, age_sd_active, age_m_overall, age_sd_overall, no_sites) %>% 
  rename(treatment = descr_active) %>% 
  rename(type = active_type) %>% 
  rename_with(~ str_remove(., "_active")) %>% 
  mutate(arm_effect_size = "cohens_d_active") %>% # chose this clunky name because I had used this variable name in previous code and don't want to bread the rest
  group_by(new_study_id) %>% #go through each study id and remove duplicates
  distinct(treatment, .keep_all = TRUE)

control_df <- df_with_distinct_instruments %>% 
  dplyr:: select(new_study_id, control_type, descr_control, psy_or_med, baseline_n_control, cohens_d_control,
                 baseline_mean_control, baseline_sd_control,post_mean_control, post_sd_control,
                 perc_baseline_mean_control, perc_baseline_sd_control, perc_post_mean_control, perc_post_sd_control, percent_women_overall, percent_women_control, age_m_control, age_sd_control, age_m_overall, age_sd_overall, no_sites) %>% 
  rename(treatment = descr_control) %>% 
  rename(type = control_type) %>% 
  rename_with(~ str_remove(., "_control")) %>% 
  mutate(arm_effect_size = "cohens_d_control") %>% #chose this clunky name because I had used this variable name in previous code and don't want to bread the rest
  group_by(new_study_id) %>% #go through each study id and remove duplicates
  distinct(treatment, .keep_all = TRUE)

df_long_for_metan <- rbind(active_df, control_df)

# I'm going to remove the March / TADS psy control row. It is currently filled with NAs which implies missing data, however it is actually the case that this condition doesn't exist as such (the control is placebo, which is not a psy control).

df_long_for_metan <- df_long_for_metan %>%
  filter(!(new_study_id == "March, 2004" & is.na(treatment)))

# I'm going to perform  the operations to calculate overall age and percent women here now. We just want this per study rather than per arm
# as we are going to compare psy vs med (2 subgroups) rather than the 4 arms

df_long_for_metan <- df_long_for_metan %>% 
  group_by(new_study_id) %>%
  mutate(n_overall = sum(baseline_n)) %>% 
  mutate(age_overall = sum(age_m * baseline_n) / sum(baseline_n)) %>% 
  mutate(mean_age = coalesce(age_overall, age_m_overall)) %>% 
  mutate(pooled_sd_age = sqrt(sum((age_sd^2) * (baseline_n - 1)) / sum(baseline_n - 1))) %>% 
  mutate(se_age = pooled_sd_age / sqrt(n_overall)) %>% 
  mutate(sex_overall = sum(percent_women * baseline_n) / sum(baseline_n)) %>% 
  mutate(percent_women = coalesce(percent_women_overall, sex_overall))

# Now just get rid of the unnecessary columns as these are causing clutter

df_long_for_metan <- df_long_for_metan %>% 
  select(-c(percent_women_overall, age_m, age_sd, age_m_overall, age_sd_overall, age_overall, sex_overall))

# Remember that these are now per study, not per arm.

# Create SEs proportions for percentage women --------------------------------------------

# We also need to calculate SE for proportion women for the baseline calculations
# for proportions, this is calculated as sqrt(p(1-p)/n), which I implement stepwise below

df_long_for_metan <- df_long_for_metan %>% 
  mutate(product_perc_women = (percent_women/100)* (1-(percent_women/100)) )

df_long_for_metan <- df_long_for_metan %>% 
  mutate(percent_women_std_error = sqrt(product_perc_women/ n_overall ))

#dim(df_long_for_metan ) #check dimension
#make sure no study lost
#length(unique(df_long_for_metan$new_study_id ) ) == # #length(unique(df_appl_v_orange$new_study_id ) )

# Now check again the studies with muliple arms
# Now checking if this worked with the studies that I used to illustrate the problem above. 
# df_long_for_metan[df_long_for_metan$new_study_id == "Stallard, 2012",]
# df_long_for_metan[df_long_for_metan$new_study_id == "Atkinson, 2014",]
# # also check one which is single to make sure it is kept
# df_long_for_metan[df_long_for_metan$new_study_id == "Ackerson, 1998",]

# For response rate, recalculating after turning long and doing all the imputations

# df_long_for_metan <- df_long_for_metan %>%
#   rowwise() %>%
#   mutate(responders = sum(rnorm(baseline_n, post_mean, 
#                                 post_sd) < (baseline_mean/ 2)))
# 
# df_long_for_metan <- df_long_for_metan %>%
#   rowwise() %>%
#   mutate(resp_rate = responders/baseline_n)
# 
# summary_resp_rate <- df_long_for_metan %>%
#   group_by(group) %>%
#   summarise(mean_resp_rate = mean(resp_rate, na.rm = TRUE),
#             sd_resp_rate = sd(resp_rate, na.rm = TRUE), 
#             n = n())

# save this as the master dataframe  --------------------------------------
# use this dataframe to perform all operations below

write.csv(df_long_for_metan, "df_long_for_metan.csv", row.names = F)

########### simulate so that each study has a value from a distribution of 
######## of correlations.

# I will generate random numbers per study id from a distribution with these parameters. 
# it is reasonable to generate one random correlation value per study as there is no reason why the correlation should
# systematically vary within studies

# Use the function I have created for this

df_long_for_metan <- read.csv("df_long_for_metan.csv")

list_df_simulated <-  simulate_dataframes_for_st_errors (df = df_long_for_metan, 
                                              num_repetitions = 1000, 
                                              seed  = 1974,
                                              n = length(unique(df_long_for_metan$new_study_id)), 
                                              a = 0.45, 
                                              b = .9, 
                                              mean = 0.65, 
                                              sd = 0.2)

#check this worked

#list_df_simulated[[10]][,c("new_study_id", "correlation_sim_values", "simulated_se")] # looks right
#summary(list_df_simulated[[1000]]$correlation_sim_values, na.rm = T) # as expected # the mean is around the parameter I gave

```
# Results

```{r, warning=FALSE, message = FALSE, echo=FALSE}

# This chunk of code needs to be here because some of the data needed for the included studies summary lives here

# Get Ns, run metareg, prepare graphing ------------------------------------------------------

# add a new multilevel variable to the simulated data for the regression

for(i in 1: length(list_df_simulated)){

list_df_simulated[[i]]$arm_effect_size <- factor(list_df_simulated[[i]]$arm_effect_size)
  list_df_simulated[[i]] <- list_df_simulated[[i]] %>% 
    mutate(four_level_var = case_when(psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
                                      psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control")
    )

  list_df_simulated[[i]]$four_level_var <- factor(list_df_simulated[[i]]$four_level_var) # turn to factor

  # relevel so that medication control becomes the reference category for the regression
#  list_df_simulated[[i]]$four_level_var <- relevel(list_df_simulated[[1]]$four_level_var, ref = "medication_control")
list_df_simulated[[i]]$four_level_var <-   fct_relevel(list_df_simulated[[i]]$four_level_var,
                                                       "medication_control",
                                                       "medication_active", 
                                                       "psychotherapy_control",
                                                       "psychotherapy_active")
}

# check it worked
# list_df_simulated[[2]] %>% 
# dplyr:: select(psy_or_med, arm_effect_size, four_level_var)

######## NOW RUN METAREGRESSIONS AND GET Coefficients for 
# A) the overall sample
# B) the CBT vs Fluox and Escitalopram sample
# C) the no WL sample
# E) the CDRS only sample (**Argyris** do we still want this?)
# F) Clinical Levels Depression

# A. Estimate Means and CIs for the overall ----------------------------------

# count studies
# count the number of studies
n_unique_studies <- length(unique(df_long_for_metan$new_study_id)) # CHARLOTTE please check
df_count_studies <- count_studies(df_long_for_metan ) # CHARLOTTE please check
# CB checked both, both correct.

# specify model
model_1 <- as.formula(~ four_level_var)

# run metareg function 
# here you can use the simulated results directly
list_model_1_meta_reg <- run_metaregression(list_df_simulated, model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_overall <-  aggregate_model_results(list_model_1_meta_reg, condition)
aggregate_results_overall <- cbind(aggregate_results_overall, condition)

# extract SMDs from the  list
list_dummy_var_means <-  lapply(list_model_1_meta_reg, extract_coefficients_func)

# calculate mean SMDs and ses
df_mean_coefs_from_sim <- calculate_mean_coefs_ses(list_dummy_var_means, condition)

df_mean_coefs_from_sim <- data.frame(cbind(condition,df_mean_coefs_from_sim))

# add the ns
df_mean_coefs_from_sim <- left_join(df_mean_coefs_from_sim, df_count_studies[df_count_studies$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

# B. Estimate Means and CIs for CBT Fluox Esc --------------------------------
#  grab ids for those studies
cbt_fluox_esc_study_ids <-
  df_appl_v_orange[df_appl_v_orange$active_type == "cbt" | df_appl_v_orange$active_type == "Fluoxetine" | df_appl_v_orange$active_type == "Escitalopram",]$new_study_id

# Use those ids to subset all dataframes in the list
# By searching for IDs we have mistakenly included multiple active arms. E.g. for Atkinson which has dulox and fluox. 
# So we need to further subset the simulated dfs. For the active arms, I am filtering out conditions that do not meet our criteria. 
# This leaves all the control conditions from trials where the active is cbt, fluox or esc. 
# We may have multiple controls per study id for this reason, though this isn't a problem 
# conceptually.

list_cbt_fluox_esc_study <- lapply(list_df_simulated, function(df) {
  df %>%
    filter(new_study_id %in% cbt_fluox_esc_study_ids) %>%
    filter(!(arm_effect_size == "cohens_d_active" & !(type %in% c("Fluoxetine", "Escitalopram", "cbt"))))
})

# count the number of studies
cbt_fluox_esc_study <- df_long_for_metan %>%
  filter(new_study_id %in% cbt_fluox_esc_study_ids,
         !(arm_effect_size == "cohens_d_active" & !(type %in% c("Fluoxetine", "Escitalopram", "cbt")))) 
study_count_cbt_fluox_esc_study <- count_studies(cbt_fluox_esc_study)
# CB checked, looks good now.

# apply the metareg function to the list of cbt, fluox, esc
model_1 <- as.formula(~ four_level_var)
list_model_1_cbt_fluox_esc_study <- run_metaregression(list_cbt_fluox_esc_study , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_cbt_fluox_esc_study <-  aggregate_model_results(list_model_1_cbt_fluox_esc_study, condition)
aggregate_results_cbt_fluox_esc_study <- cbind(aggregate_results_cbt_fluox_esc_study, condition)

# extract SMDs from the list for cbt fluox escit
means_cbt_fluox_esc_study <-  lapply(list_model_1_cbt_fluox_esc_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_cbt_fluox_esc_study <- calculate_mean_coefs_ses(means_cbt_fluox_esc_study, condition)
 coef_and_se_means_cbt_fluox_esc_study <- data.frame(cbind(condition, coef_and_se_means_cbt_fluox_esc_study))

# add the ns to the dataframe
coef_and_se_means_cbt_fluox_esc_study <- left_join(coef_and_se_means_cbt_fluox_esc_study, study_count_cbt_fluox_esc_study[study_count_cbt_fluox_esc_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

# C. Estimate Means and CIs without WL ---------------------------------------

 # grab the ids for no waitlist
no_wl_ids <-df_appl_v_orange[df_appl_v_orange$control_type != "wl",]$new_study_id

# use them to loop over the simulated list to subset all the dataframes.

list_no_wl <- lapply(list_df_simulated, function(df) {
  df %>%
    filter(new_study_id %in% no_wl_ids) %>%
    filter(!(arm_effect_size == "cohens_d_control" & type == "wl")) 
})

# count studies 
no_wl_studies <- df_long_for_metan %>% 
  filter(new_study_id %in% no_wl_ids) %>%
  filter(!(arm_effect_size == "cohens_d_control" & type == "wl"))
study_count_no_wl_study <- count_studies(no_wl_studies)
# CB checked, looks good!

# use this model as above
model_1 <- as.formula(~ four_level_var)

# apply the metareg function to the list of no wl
list_model_1_no_wl <- run_metaregression(list_no_wl , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_no_wl <-  aggregate_model_results(list_model_1_no_wl, condition)
aggregate_results_no_wl <- cbind(aggregate_results_no_wl, condition)

# now use the function to extract the coefficients
means_no_wl <-  lapply(list_model_1_no_wl, extract_coefficients_func)

# run the mean function for no wl
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_no_wl <- calculate_mean_coefs_ses(means_no_wl, condition)
coef_and_se_means_no_wl <- data.frame(cbind(condition,coef_and_se_means_no_wl))

coef_and_se_means_no_wl <- left_join(coef_and_se_means_no_wl, study_count_no_wl_study[study_count_no_wl_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

# E. CDRS Sensitivity Analysis ---------------------------------------

cdrs_study_ids <-
  df_appl_v_orange[df_appl_v_orange$instrument_name == "cdrs",]$new_study_id

# use those ids to subset all dataframes in the list
list_cdrs_study <- list()

for(i in 1: length(list_df_simulated)){

  list_cdrs_study[[i]] <- list_df_simulated[[i]][list_df_simulated[[i]]$new_study_id
                                                          %in% cdrs_study_ids,]
}

# count the number of studies
study_count_cdrs_study <- count_studies(df_long_for_metan[df_long_for_metan$new_study_id %in% cdrs_study_ids,])
# CB checked, correct.

# apply the metareg function to the list of cdrs studies
model_1 <- as.formula(~ four_level_var)
list_model_1_cdrs_study <- run_metaregression(list_cdrs_study , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_cdrs_study <-  aggregate_model_results(list_model_1_cdrs_study, condition)

# extract SMDs from the list for cdrs
means_cdrs_study <-  lapply(list_model_1_cdrs_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_cdrs_study <- calculate_mean_coefs_ses(means_cdrs_study, condition)
coef_and_se_means_cdrs_study <- data.frame(cbind(condition, coef_and_se_means_cdrs_study))

# add the ns to the dataframe
coef_and_se_means_cdrs_study <- left_join(coef_and_se_means_cdrs_study, study_count_cdrs_study[study_count_cdrs_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

# F Clinical Levels Depression Sensitivity Analysis ---------------------------------------

# We need to decide how we want to do this. We have mdd, mood, sub, and cut categories. For now I'll just exclude sub.

#take study ids where sample has clinical levels of depression
clin_study_ids <-
  df_appl_v_orange[df_appl_v_orange$diagnosis != "sub",]$new_study_id

# use those ids to subset all dataframes in the list
list_clin_study <- list()

for(i in 1: length(list_df_simulated)){

  list_clin_study[[i]] <- list_df_simulated[[i]][list_df_simulated[[i]]$new_study_id
                                                          %in% clin_study_ids,]
}

# count the number of studies
study_count_clin_study <- count_studies(df_long_for_metan[df_long_for_metan$new_study_id %in% clin_study_ids,])
# CB checked, correct.

# apply the metareg function to the list of clinical studies
model_1 <- as.formula(~ four_level_var)
list_model_1_clin_study <- run_metaregression(list_clin_study , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_clin_study <-  aggregate_model_results(list_model_1_clin_study, condition)

# extract SMDs from the list for clinical studies
means_clin_study <-  lapply(list_model_1_clin_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_clin_study <- calculate_mean_coefs_ses(means_clin_study, condition)
coef_and_se_means_clin_study <- data.frame(cbind(condition, coef_and_se_means_clin_study))

# add the ns to the dataframe
coef_and_se_means_clin_study <- left_join(coef_and_se_means_clin_study, study_count_clin_study[study_count_clin_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

```

## Included studies

The data for the studies included in this metanalysis are summarised in Supplementary Table 1 and are also available as a csv dataframe on \[<https://github.com/transatlantic-comppsych/apples_oranges>\].

In total, there were `r length(unique(df_long_for_metan$new_study_id))` RCTs which included `r df_count_studies[df_count_studies$condition == "medication_active"& df_count_studies$is_missing == FALSE,]$n` active arms and `r df_count_studies[df_count_studies$condition == "medication_control"& df_count_studies$is_missing == FALSE,]$n` control arms of antidepressant trials; and `r df_count_studies[df_count_studies$condition == "psychotherapy_active"& df_count_studies$is_missing == FALSE,]$n` active arms and `r df_count_studies[df_count_studies$condition == "psychotherapy_control"& df_count_studies$is_missing == FALSE,]$n` control arms from psychotherapy RCTs Note that the number of active and control arms does not exactly match because some studies feature more than one control or active arm. There were also missing data for `r df_count_studies[df_count_studies$condition == "medication_active"& df_count_studies$is_missing == TRUE,]$n`, `r df_count_studies[df_count_studies$condition == "medication_control"& df_count_studies$is_missing == TRUE,]$n`, `r df_count_studies[df_count_studies$condition == "psychotherapy_active"& df_count_studies$is_missing == TRUE,]$n`, and `r df_count_studies[df_count_studies$condition == "psychotherapy_control"& df_count_studies$is_missing == TRUE,]$n` trial arms for medication active, medication control, psychotherapy active, and psychotherapy control conditions respectively, as the data needed to calculate the SMD were missing and could not be imputed by any of the methods outlined above.

Placebo pill was the control condition for all medication trials. In psychotherapy trials, the control arm included 14 WL controls, 25 care as usual and several other conditions such as 4 attention control conditions. All included trials and the types of treatment controls can be found in Supplementary Table 1.

## Sample characteristics at baseline in medication versus psychotherapy trials

Table 1 summarises the results from each of the meta-analyses examining sample characteristics at baseline. The summary statistics are provided for each subgroup (i.e. for medication and psychotherapy trials) and the p-value derives from the test for subgroup differences. Full results for each of the sensitivity analyses are included in the Supplementary Materials (Tables S1 - S3).

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# Start by looking at demographics at baseline
# Currently we have both control and active conditions, we just need each study once

df_baseline_demographics <- df_long_for_metan %>% 
  distinct(new_study_id, .keep_all = TRUE)

# This is the function to extract statistics

extract_stats <- function(df) {

  # get the summary table with results
  statistics <- summary(df, byvar = "psy_or_med")

  df_stats <- tibble(
  subgroup = statistics$subgroup.levels,
  k_study = statistics$k.study.w,
  TE_random = statistics$TE.random.w,
  seTE_random = statistics$seTE.random.w,
  lower_random = statistics$lower.random.w,
  upper_random = statistics$upper.random.w,
  tau2 = statistics$tau2.w)

  # Add an empty row at the top and column on right (this is for the test for subgroup differences p value)
  df_stats <- rbind(data.frame(subgroup = NA, k_study = NA, TE_random = NA, seTE_random = NA, 
                             lower_random = NA, upper_random = NA, tau2 = NA), df_stats)
  df_stats <- cbind(df_stats, pval_between = NA)

  # Set the value in the last row and last column as the between groups p val
  df_stats[1, "pval_between"] <- statistics$pval.Q.b.random

  return(df_stats)

}

# Now look at baseline severity for overall psy vs med
# Need to calculate the variables first

df_for_baseline_severity <- df_long_for_metan %>% 
  group_by(new_study_id) %>%
  mutate(perc_baseline_overall = sum(perc_baseline_mean * baseline_n) / sum(baseline_n)) %>% 
  mutate(pooled_perc_sd_baseline = sqrt(sum((perc_baseline_sd^2) * (baseline_n - 1)) / sum(baseline_n - 1))) %>% 
  mutate(perc_se_baseline = pooled_perc_sd_baseline / sqrt(n_overall))

df_for_baseline_severity <- df_for_baseline_severity %>% 
  distinct(new_study_id, .keep_all = TRUE)

met_sev <- metagen(TE = perc_baseline_overall,
                           seTE = perc_se_baseline,
                           studlab = new_study_id,
                           data = df_for_baseline_severity,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "baseline severity",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_sev <- extract_stats(met_sev)

# clinical sample

clin_study_ids <-
  df_appl_v_orange[df_appl_v_orange$diagnosis != "sub",]$new_study_id

df_for_baseline_severity_clin <- df_for_baseline_severity %>%
  filter(new_study_id %in% clin_study_ids)

met_sev_clin <- metagen(TE = perc_baseline_overall,
                           seTE = perc_se_baseline,
                           studlab = new_study_id,
                           data = df_for_baseline_severity_clin,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "baseline severity clinical",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_sev_clin <- extract_stats(met_sev_clin)

# cbt + med

# cbt_meds_study_ids <-
#   df_appl_v_orange[df_appl_v_orange$active_type == "cbt" | df_appl_v_orange$active_type == "Fluoxetine" | df_appl_v_orange$active_type == "Escitalopram",]$new_study_id
# 
# df_for_baseline_severity_cbt <- df_for_baseline_severity %>%
#   filter(new_study_id %in% cbt_meds_study_ids) 
# 
# met_sev_cbt <- metagen(TE = perc_baseline_overall,
#                            seTE = perc_se_baseline,
#                            studlab = new_study_id,
#                            data = df_for_baseline_severity_cbt,
#                            sm = "",
#                            fixed = FALSE,
#                            random = TRUE,
#                            method.tau = "REML",
#                            hakn = TRUE,
#                            title = "baseline severity cbt + med studies",
#                            subgroup = psy_or_med,
#                            tau.common = FALSE)
# 
# results_met_sev_cbt <- extract_stats(met_sev_cbt)

# no wl

no_wl_ids <- df_appl_v_orange[df_appl_v_orange$control_type != "wl",]$new_study_id

df_for_baseline_severity_no_wl <- df_for_baseline_severity %>%
  filter(new_study_id %in% no_wl_ids)

met_sev_no_wl <- metagen(TE = perc_baseline_overall,
                           seTE = perc_se_baseline,
                           studlab = new_study_id,
                           data = df_for_baseline_severity_no_wl,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "baseline severity no wl studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_sev_no_wl <- extract_stats(met_sev_no_wl)

# Now start looking at perc_women variable 
# The metagen function doesn't like when NAs result in the TE vector being a different length to subgroup vector. Need to start by removing NAs.

df_for_sex_metan <- df_baseline_demographics %>% 
  filter(!is.na(percent_women))

df_for_sex_metan <- df_for_sex_metan %>% 
  filter(percent_women != 100)

met_perc_women_overall <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_sex_metan,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across studies")

met_perc_women_overall <- update(met_perc_women_overall,
                              subgroup = df_for_sex_metan$psy_or_med, subgroup.name = "treatment modality",
                              tau.common = FALSE)

# Apply the function to the meta-analysis result

results_met_perc_women_overall <- extract_stats(met_perc_women_overall)

# looking at clinical sample only

df_for_sex_metan_clin <- df_for_sex_metan %>%
  filter(new_study_id %in% clin_study_ids)

met_perc_women_clin <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_sex_metan_clin,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across clinical studies")

met_perc_women_clin <- update(met_perc_women_clin,
                           subgroup = df_for_sex_metan_clin$psy_or_med, subgroup.name = "treatment modality",
                           tau.common = FALSE)

# Apply the function to the meta-analysis result
results_met_perc_women_clin <- extract_stats(met_perc_women_clin)

# # Now looking at the CBT + med studies only
# 
# df_for_sex_metan_cbt <- df_for_sex_metan %>%
#   filter(new_study_id %in% cbt_meds_study_ids) 
# 
# met_perc_women_cbt <- metagen(TE = percent_women,
#                            seTE = percent_women_std_error,
#                            studlab = new_study_id,
#                            data = df_for_sex_metan_cbt,
#                            sm = "",
#                            fixed = FALSE,
#                            random = TRUE,
#                            method.tau = "REML",
#                            hakn = TRUE,
#                            title = "percentage women across cbt and med studies",
#                            subgroup = psy_or_med,
#                            tau.common = FALSE)
# 
# # Apply the function to the meta-analysis result
# results_met_perc_women_cbt <- extract_stats(met_perc_women_cbt)

# Now the no waitlist group

df_for_sex_metan_no_wl <- df_for_sex_metan %>%
  filter(new_study_id %in% no_wl_ids)

met_perc_women_no_wl <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_sex_metan_no_wl,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across studies with controls that are not wl",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

# Apply the function to the meta-analysis result
results_met_perc_women_no_wl <- extract_stats(met_perc_women_no_wl)

# Now start looking at age variable

# We have several studies for which we are missing SE for age. I am going to perform an imputation by taking an average
# for the studies we do have, per modality, and substituting this in for missing values.

df_baseline_demographics$se_age <- as.numeric(df_baseline_demographics$se_age)

mean_se_age_psy <- df_baseline_demographics %>%
  filter(psy_or_med == 1) %>%
  summarise(mean_se_age_psy = mean(se_age, na.rm = TRUE))

mean_se_age_med <- df_baseline_demographics %>%
  filter(psy_or_med == 0) %>%
  summarise(mean_se_age_med = mean(se_age, na.rm = TRUE))

df_baseline_demographics <- df_baseline_demographics %>% 
  mutate(se_age = ifelse(!is.na(se_age), se_age,
                         ifelse(psy_or_med == 1, mean_se_age_psy, mean_se_age_med)))

df_for_age_metan <- df_baseline_demographics %>% 
  filter(!is.na(mean_age))

weighted_average_age <- sum(df_baseline_demographics$mean_age * df_baseline_demographics$baseline_n, na.rm = TRUE) / sum(df_baseline_demographics$baseline_n, na.rm = TRUE)
tst <- mean(df_baseline_demographics$mean_age, na.rm = TRUE)

# Age meta analysis

df_for_age_metan$se_age <- as.numeric(df_for_age_metan$se_age)

met_age <- metagen(TE = mean_age,
                           seTE = se_age,
                           studlab = new_study_id,
                           data = df_for_age_metan,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "age across studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_age <- extract_stats(met_age)

# Looking at clinical sample only

clin_study_ids <-
  df_appl_v_orange[df_appl_v_orange$diagnosis != "sub",]$new_study_id

df_for_age_metan_clin <- df_for_age_metan %>%
  filter(new_study_id %in% clin_study_ids)

met_age_clin <- metagen(TE = mean_age,
                           seTE = se_age,
                           studlab = new_study_id,
                           data = df_for_age_metan_clin,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "age across clinical studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_age_clin <- extract_stats(met_age_clin)

# # Now cbt + med studies only
# 
# df_for_age_metan_cbt <- df_for_age_metan %>%
#   filter(new_study_id %in% cbt_meds_study_ids) 
# 
# met_age_cbt <- metagen(TE = mean_age,
#                            seTE = se_age,
#                            studlab = new_study_id,
#                            data = df_for_age_metan_cbt,
#                            sm = "",
#                            fixed = FALSE,
#                            random = TRUE,
#                            method.tau = "REML",
#                            hakn = TRUE,
#                            title = "age across cbt + med studies",
#                            subgroup = psy_or_med,
#                            tau.common = FALSE)
# 
# results_met_age_cbt <- extract_stats(met_age_cbt)

# Now excluding wl

df_for_age_metan_no_wl <- df_for_age_metan %>%
  filter(new_study_id %in% no_wl_ids)

met_age_no_wl <- metagen(TE = mean_age,
                           seTE = se_age,
                           studlab = new_study_id,
                           data = df_for_age_metan_no_wl,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "age across no wl studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_age_no_wl <- extract_stats(met_age_no_wl)

# Lets make a table that combines across these baseline characteristics variables
# This is for overall sample only, with separate sensitivity analyses in the supplement

# df_overall_baseline_results <- rbind(results_met_sev, results_met_perc_women_overall, results_met_age)
# 
# # For row names
# df_overall_baseline_results <- df_overall_baseline_results %>%
#   mutate(subgroup = case_when(
#     row_number() == 1 ~ "Severity",
#     row_number() == 4 ~ "Percent Female",
#     row_number() == 7 ~ "Age",
#     TRUE ~ if_else(subgroup == 0, "Medication", "Psychotherapy")
#   )) 
# 
# # Define new column names
# new_column_names <- c("Subgroup", "K", "Mean", "SE", "Lower CI", "Upper CI", "T2", "p-value")
# 
# # Assign new column names to the data frame
# colnames(df_overall_baseline_results) <- new_column_names
# 
# # Rounding
# df_overall_baseline_results <- df_overall_baseline_results %>%
#   mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
#   mutate_at(vars(ncol(.)), ~ round(., 3)) %>%
#   mutate_all(~if_else(is.na(.), "", as.character(.))) 
# 
# big_border = fp_border(color="black", width = 1)
# small_border = fp_border(color="lightgray", width = 1)

# Too basic, limited formatting options
# 
# knitr:: kable(df_overall_baseline_results, caption="Sample characteristics at baseline across medication and psychotherapy studies")
# 
# gt(df_overall_baseline_results) %>%
#   tab_header(
#     title = md("Test this"),
#     subtitle = md("And this"))
# 
# This doesn't work when rendering to word doc
# 
#   kbl(align = "l", caption = "Sample characteristics at baseline across medication and psychotherapy studies", escape = F) %>%
#   kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
#   row_spec(row = 0, background = "#EBEBEB", extra_css = "border-bottom: 1px black;" ) %>%
#   row_spec(row = 1, bold = T) %>%
#   row_spec(c(4, 7), bold = T, extra_css = "border-top: 1px solid black;") %>%
#   column_spec(1, border_right = T) %>%
#   add_indent(c(2, 3, 5, 6, 8, 9))
# 
# this doesn't have a caption when rendering to word. workaround below, adding caption as header row.
# 
# flextable(df_overall_baseline_results) %>%
#   set_caption("Sample characteristics at baseline across medication and psychotherapy studies") %>%
#   bold(part = "header") %>%  # Bold the header row
#   bold(i = c(1, 4, 7), j = NULL, bold = TRUE, part = "body") %>%  # Bold specific rows
#   border_outer(border = small_border) %>%  # Add outer borders
#   hline(i = c(1, 2, 4, 5, 7, 8), border = small_border) %>%  # Add light grey horizontal lines all  rows
#   hline(i = c(3, 6), border = big_border) %>%  # Add darker horizontal lines after specific rows %>%
#   hline(part = "header", border = big_border) %>%
#   vline(j = 1, border = big_border) %>%
#   set_table_properties(width = .75) %>%  # Set table width
#   bg(bg = "#F2F2F2", part = "header") %>%  # Add light grey background to header row
#   autofit() %>%  # Autofit column widths
#   font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
#   padding(i = c(2, 3, 5, 6, 8, 9), j = 1, padding.left = 10) %>%
#   align(j = 2:ncol(df_overall_baseline_results), align = "center", part = "all")

# This is if we want to have all sensitivity analyses in table one

df_overall_baseline_results <- rbind(results_met_sev, results_met_sev_clin, results_met_sev_no_wl, results_met_perc_women_overall, results_met_perc_women_clin, results_met_perc_women_no_wl, results_met_age, results_met_age_clin, results_met_age_no_wl)

# For row names
df_overall_baseline_results <- df_overall_baseline_results %>%
  mutate(subgroup = case_when(
    row_number() %in% c(1, 10, 19) ~ "Overall",
    row_number() %in% c(4, 13, 22) ~ "Excluding subclinical",
    row_number() %in% c(7, 16, 25) ~ "Excluding waitlist",
    TRUE ~ if_else(subgroup == 0, "Medication", "Psychotherapy")
  ))

# Define new column names
new_column_names <- c("Subgroup", "K", "Mean", "SE", "Lower CI", "Upper CI", "T2", "p-value")

# Assign new column names to the data frame
colnames(df_overall_baseline_results) <- new_column_names

# Rounding
df_overall_baseline_results <- df_overall_baseline_results %>%
  mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
  mutate_at(vars(ncol(.)), ~ round(., 3)) %>%
  mutate_all(~if_else(is.na(.), "", as.character(.)))

big_border = fp_border(color="black", width = 1)
small_border = fp_border(color="lightgray", width = 1)

df_overall_baseline_results %>% 
  mutate(category = case_when(
    row_number() %in% 1:9 ~ "Baseline Severity",
    row_number() %in% 10:18 ~ "Percent Female",
    row_number() %in% 19:27 ~ "Age",
    TRUE ~ ""
  )) %>% # grouping rows
  group_by(category) %>%
  as_flextable(hide_grouplabel = TRUE) %>% 
  paginate(group = "category", group_def = "rle") %>% 
  bold(i = c(1, 2, 5, 8, 11, 12, 15, 18, 21, 22, 25, 28), j = NULL, bold = TRUE, part = "body") %>%  # Bold specific rows
  set_table_properties(width = 1, layout = "autofit") %>%  # Set table width
  bg(i = c(1, 11, 21), bg  = "#F2F2F2", part = "body") %>% # identify category sections
  autofit() %>%  # Autofit column widths
  padding(i = c(3, 4, 6, 7, 9, 10, 13, 14, 16, 17, 19, 20, 23, 24, 26, 27, 29, 30), j = 1, padding.left = 10) %>% # indent some rows
  align(j = 2:ncol(df_overall_baseline_results), align = "center", part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  add_header_row(values = "Sample characteristics at baseline across medication and psychotherapy studies", colwidths = 8) %>% # This is a work around for the problems with rendering captions
  add_header_row(values = "Table 1", colwidths = 8) %>% 
  bold(i = c(1, 3), part = "header") %>%   # Bold the relevant header rows
  bg(i = 3, bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(i = 3, part = "header", border = big_border) %>% 
  padding(i = 2, padding.bottom = 15, part = "header") %>% 
  line_spacing(space = .7, part = "all")

```

### Baseline severity

On average, severity of depressive symptoms at baseline was significantly higher in medication trials when compared to psychotherapy trials. The statistics provided in Table 1 are baseline depression scores transformed to reflect percentage of a scale range. To take an example, the CDRS gives a possible total score from 17 to 113 (i.e. range of 96). From Table 1, mean severity was `r round(results_met_sev$TE_random[results_met_sev$subgroup == 1][[2]], 2)` for psychotherapy studies and `r round(results_met_sev$TE_random[results_met_sev$subgroup == 0][[2]], 2)` for medication studies, which would translate to 51.56 (17 + 0.36 x 96) and 57.32 (17 + 0.42 x 96), respectively, as equivalent scores on the CDRS.

When excluding RCTs that used waitlist as their control, baseline severity remained significantly higher in medication trials compared to psychotherapy trials. This difference did not reach statistical significance when excluding studies that recruited samples with sub-clinical depression.

### Sex

For this analysis, we excluded the two psychotherapy trials which included entirely female samples. As can be seen in Table 1, psychotherapy trials featured a significantly higher percentage of females when compared to medication trials. On average, samples were `r round(results_met_perc_women_overall$TE_random[results_met_perc_women_overall$subgroup == 1][[2]],2)`% (*SE* = `r round(results_met_perc_women_overall$seTE_random[results_met_perc_women_overall$subgroup == 1][[2]],2)`) female across psychotherapy trials and `r round(results_met_perc_women_overall$TE_random[results_met_perc_women_overall$subgroup == 0][[2]],2)`% (*SE* = `r round(results_met_perc_women_overall$seTE_random[results_met_perc_women_overall$subgroup == 0][[2]],2)`) female across medication trials. Excluding sub-clinical and waitlist control studies yielded similar results.

### Age

As can be seen in Table 1, mean age was `r round(results_met_age$TE_random[results_met_age$subgroup == 1][[2]],2)` (*SE* = `r round(results_met_age$seTE_random[results_met_age$subgroup == 1][[2]],2)`) across psychotherapy trials and `r round(results_met_age$TE_random[results_met_age$subgroup == 0][[2]],2)` (*SE* = `r round(results_met_age$seTE_random[results_met_age$subgroup == 0][[2]],2)`) across medication trials, with no significant between group differences. There were no significant differences in mean age between modalities on further sensitivity analyses.

## Number of trial sites

```{r, warning=FALSE, message = FALSE, echo=FALSE}

### Number of sites
# Write results, tabulate.

## This data lives in this spreadsheet for psy trials, we need to merge it in

df_no_sites <- readxl:: read_excel("for_despina_and_giannis_200224.xlsx")

df_no_sites <- df_no_sites %>% 
  select(c(new_study_id, number_of_sites)) %>% 
  filter(!is.na(number_of_sites)) %>% 
  distinct(new_study_id, .keep_all = TRUE) %>% 
  mutate(number_of_sites = as.numeric(number_of_sites))

df_baseline_demographics <- full_join(df_baseline_demographics, df_no_sites, by = "new_study_id")

# we have different vectors for psy and med trials, lets coalesce

df_baseline_demographics <- df_baseline_demographics %>%
  mutate(no_sites = as.numeric(no_sites)) %>% 
  mutate(sites = coalesce(no_sites, number_of_sites)) %>%
  select(-no_sites, -number_of_sites)

# we have a few rogue studies (Kitchen, McCarty and Wright). Kitchen was too recent (after 2021), McCarty and Wright are not in Cuijpers' master
# dataset for some reason, though they are in the list of included studies in the table in his MA. For now I'm going to remove them because 
# we don't have data for these trials.

df_baseline_demographics <- df_baseline_demographics %>% 
  filter(!(new_study_id %in% c("Kitchen, 2021", "McCarty, 2013", "Wright, 2020")))

# Now create binary variable, single or multi site trial

df_baseline_demographics <- df_baseline_demographics %>% 
  mutate(site_binary = case_when(sites == 1 ~ "single", 
                                 sites > 1 ~ "multi",
                                 TRUE ~ NA_character_))

# df_baseline_demographics %>%
#   group_by(psy_or_med, site_binary) %>%
#   summarize(num_trials = n())

# Lets look at results

# summary stats
df_sites_summary_stats <- df_baseline_demographics %>%
  group_by(psy_or_med) %>%
  summarise(
    n = n(),
    mean_sites = mean(sites, na.rm = TRUE),
    sd_sites = sd(sites, na.rm = TRUE),
  )

# Tabulate
# Rounding
df_sites_summary_stats <- df_sites_summary_stats %>%
  mutate(across(where(is.numeric), ~round(., 2)))

# Row names
df_sites_summary_stats <- df_sites_summary_stats %>%
  mutate(psy_or_med = case_when(
    psy_or_med == 0 ~ "Medication",
    psy_or_med == 1 ~ "Psychotherapy",
    TRUE ~ as.character(psy_or_med)  # Keep other values unchanged
  ))

df_sites_summary_stats <- df_sites_summary_stats %>%
  rename(
    Modality = psy_or_med,
    N = n,
    Mean = mean_sites,
    SD = sd_sites
  )

# t-test
sites_t_test <- t.test(sites ~ psy_or_med, data = df_baseline_demographics)

sites_t_test_stats <- data.frame(
  statistic = sites_t_test$statistic,
  df = sites_t_test$parameter,
  p_value = sites_t_test$p.value,
  ci_lower = sites_t_test$conf.int[1],
  ci_upper = sites_t_test$conf.int[2]
)

sites_t_test_stats <- sites_t_test_stats %>%
  mutate(across(.cols = c(statistic, df, ci_lower, ci_upper), ~ round(., 2))) %>% 
  mutate(p_value = ifelse(p_value < 0.001, "< 0.001", as.character(round(p_value, 3))))

flextable(df_sites_summary_stats) %>%
  set_table_properties(width = 1, layout = "autofit") %>%
  add_footer_row(values = as_paragraph(as_i("t"), "(", sites_t_test_stats$df, ") = ", sites_t_test_stats$statistic,
                      "; ", as_i("p "), sites_t_test_stats$p_value, 
                      "; 95% CI: ", sites_t_test_stats$ci_lower,
                      "-", sites_t_test_stats$ci_upper, "."), colwidths = 4) %>% 
  add_header_row(values = "Number of sites across medication and psychotherapy studies", colwidths = 4) %>%
  add_header_row(values = "Table 2", colwidths = 4) %>% 
  bold(i = c(1, 3), part = "header") %>%   # Bold the relevant header rows
  bg(i = 3, bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(i = 3, part = "header", border = big_border) %>% 
  padding(i = 2, padding.bottom = 15, part = "header") %>%
  padding(i = c(1, 2), j = 1, padding.left = 10) %>% 
  align(j = 2:ncol(df_sites_summary_stats), align = "center", part = "all") %>% 
  align(j = 1, align = "left", part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>% 
  autofit() %>% 
  paginate(init = TRUE, hdr_ftr = TRUE) %>% 
  line_spacing(space = .7, part = "all")

# This is for if we want to render to pdf

# df_sites_summary_stats <- df_sites_summary_stats %>%
#   kbl(caption = "Number of sites across medication and psychotherapy studies") %>%
#   kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
#   row_spec(row = 0, background = "#EBEBEB") %>%
#   column_spec(column = 1, border_right = T)

# sites_t_test_stats %>%
#   kbl(align = "l", caption = "Results for t-test comparing number of sites between medication and psychotherapy studies") %>%
#   kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
#   row_spec(row = 0, background = "#EBEBEB") %>%
#   column_spec(column = 1, border_right = T)

```

There was a significant difference between the number of sites in medication versus psychotherapy trials, as can be seen in Table 2. Average number of trial sites was significantly higher across medication studies (*M* = `r df_sites_summary_stats$Mean[df_sites_summary_stats$Modality == "Medication"]`, *SD* =`r df_sites_summary_stats$SD[df_sites_summary_stats$Modality == "Medication"]`) compared to psychotherapy studies (*M* =`r df_sites_summary_stats$Mean[df_sites_summary_stats$Modality == "Psychotherapy"]`, *SD* =`r df_sites_summary_stats$SD[df_sites_summary_stats$Modality == "Psychotherapy"]`)(*t* (`r round (sites_t_test$parameter, 2)`) = `r round(sites_t_test$statistic,2)`, *p* =`r sites_t_test_stats$p_value`). Of those studies for which we had data on number of sites, 26 of 28 (93%) medication trials were multisite, compared to 24 of 45 (54%) psychotherapy studies.

## Standardised mean differences of control conditions in psychotherapy and medication studies

We applied metaregression to obtain the SMDs and confidence intervals of each of the four arms.

As can be seen in Figure 1, there were substantial differences between the four arms of the metanalysis with striking differences between the medication and the psychotherapy control arms. In particular, pill placebo had an SMD = `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$coef_means, 2)` (95% CI: `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$lower_ci, 2)` to `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$upper_ci, 2)`) whereas psychotherapy controls had an SMD = `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$coef_means, 2)` (95% CI: `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$lower_ci, 2)` to `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$upper_ci, 2)` ) (see Table 3 below).

```{r, warning=FALSE, message = FALSE, echo=FALSE}

subtitle_text_all <- "Metanalytic estimates of within-group changes: all studies"
plot_means_all <- plot_means_function(df_mean_coefs_from_sim, subtitle_text_all)
pdf(file = "plot_means_all.pdf")

# Prepare to tabulate

table_mean_coefs_from_sim <- df_mean_coefs_from_sim %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition))) %>% 
  relocate("n", .after = "condition")

new_column_names <- c("Condition", "N", "Coefficient", "SE", "Lower CI", "Upper CI")

# Assign new column names to the data frame
colnames(table_mean_coefs_from_sim) <- new_column_names

# Regression results

aggregate_results_overall <- aggregate_results_overall %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition))) %>% 
  relocate("condition", .before = "coefficients")

new_column_names <- c("Condition", "Coefficient", "SE", "z value", "Lower CI", "Upper CI", "T2", "I2", "k", "R2")

# Assign new column names to the data frame
colnames(aggregate_results_overall) <- new_column_names

flextable(table_mean_coefs_from_sim) %>%
  set_table_properties(width = 1, layout = "autofit") %>%
  autofit() %>%
  font(fontname = "Calibri", part = "all") %>% 
  add_header_row(values = "Summary statistics of the estimated SMDs", colwidths = 6) %>%
  add_header_row(values = "Table 3", colwidths = 6) %>% 
  bold(i = c(1, 3), part = "header") %>%   # Bold the relevant header rows
  bg(i = 3, bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(i = 3, part = "header", border = big_border) %>% 
  padding(i = 2, padding.bottom = 15, part = "header") %>% 
  align(j = 2:ncol(df_mean_coefs_from_sim), align = "center", part = "all") %>% 
  align(j = 1, align = "left", part = "all") %>% 
  paginate(init = TRUE, hdr_ftr = TRUE) %>% 
  line_spacing(space = .7, part = "all")

```

![Figure 1.](plot_means_all.pdf)

![](plot_means_all.pdf)

In Table 4, we present the regression that tests our hypothesis about differences between medication and psychotherapy controls. In particular, in this metaregression model, medication control is the reference category (termed intercept) to which all other categories of the dummy variable are compared. The strongest difference between arms, as judged by the z-value, is between the psychotherapy and medication controls with a z-value `r aggregate_results_overall[aggregate_results_overall$condition == "psychotherapy_control", ]$z_value` , which yields a very low p-value ( `r sprintf("%.3e",(1-(pnorm(8.949921))))` ).

```{r, warning=FALSE, message = FALSE, echo = FALSE}

table_aggregate_results_overall <- aggregate_results_overall[, 1:6]

flextable(table_aggregate_results_overall) %>%
  set_table_properties(width = 1, layout = "autofit") %>%
  autofit() %>%
  add_footer_row(values = as_paragraph("T", as_sup("2"), " = ", aggregate_results_overall[1, "T2"],
                      "; ", as_i("I"), as_sup("2"), " = ", aggregate_results_overall[1, "I2"], 
                      "; ", as_i("K"), " = ", aggregate_results_overall[1, "k"],
                      "; ", as_i("R"), as_sup("2"), " = ", aggregate_results_overall[1, "R2"], "."), colwidths = 6) %>%
  add_header_row(values = "Results from metaregression with overall sample", colwidths = 6) %>%
  add_header_row(values = "Table 4", colwidths = 6) %>% 
  bold(i = c(1, 3), part = "header") %>%   # Bold the relevant header rows
  bg(i = 3, bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(i = 3, part = "header", border = big_border) %>% 
  padding(i = 2, padding.bottom = 15, part = "header") %>% 
  align(j = 2:ncol(table_aggregate_results_overall), align = "center", part = "all") %>% 
  align(j = 1, align = "left", part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>% 
  paginate(init = TRUE, hdr_ftr = TRUE) %>% 
  line_spacing(space = .7, part = "all")

```

##### Sensitivity analyses

We then conducted a series of sensitivity analyses of our results. First, we analysed the data after excluding waitlist control studies. As can be seen, the pattern of results is very similar to that of the overall analyses.

```{r, warning=FALSE, message = FALSE, echo=FALSE}

subtitle_text_no_wl <- "Metanalytic estimates of within-group changes: wl excluded"
plot_means_wl <- plot_means_function(coef_and_se_means_no_wl, subtitle_text_no_wl)
#print(plot_means_wl)
pdf(file = "plot_means_wl.pdf")
plot_means_wl
dev.off()

```

![](plot_means_wl.pdf)

We next analysed the data after excluding sub-clinical studies. Again, the pattern of results was very similar to that of the overall analyses.

```{r, warning=FALSE, message = FALSE, echo=FALSE}

subtitle_text_clin <- "Metanalytic estimates of within-group changes: subclinical studies excluded"
plot_means_clin <- plot_means_function(coef_and_se_means_clin_study, subtitle_text_clin)
pdf(file = "plot_means_clin.pdf")
plot_means_clin
dev.off()

```

![](plot_means_clin.pdf)

Finally, we examined the data including only those studies that used the CDRS, which yielded a very similar pattern of results.

```{r, warning=FALSE, message = FALSE, echo=FALSE}

subtitle_text_cdrs <- "Metanalytic estimates of within-group changes: CDRS studies only"
plot_means_cdrs <- plot_means_function(coef_and_se_means_cdrs_study, subtitle_text_cdrs)
pdf(file = "plot_means_cdrs.pdf")
plot_means_cdrs
dev.off()

```

![](plot_means_cdrs.pdf)

#### Effect of standard errors of the SMDs .

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# examine the stability of the simulations

z_psy_con_vs_med_con <- 0
for(i in 1: length(list_model_1_meta_reg)){
z_psy_con_vs_med_con[i] <- list_model_1_meta_reg[[i]]$zval[condition=="psychotherapy_control"]

}

stability_sims <- data.frame(n_sims = 1:1000, z_value = z_psy_con_vs_med_con)

stab_sims <- stability_sims %>%
  ggplot(aes(x = n_sims, y = z_value)) +
  geom_point() +
  geom_hline(yintercept = 1.65, linetype = "dotted", color = "red") +
  geom_segment(aes(x = 300, xend = 300, y = 2, yend = 1.65),
               arrow = arrow(length = unit(0.3, "cm")),
               color = "black") +
  annotate("text", x = 300, y = 2.0, label = "p < 0.05 threshold (values above line significant)", hjust = -0.1, vjust = 0) +
  geom_segment(aes(x = 300, xend = 300, y = 7.5, yend = min(stability_sims$z_value-0.3)),
               arrow = arrow(length = unit(0.3, "cm")),
               color = "black") +
  annotate("text", x = 300, y = 7.5, label = "simulated z-values",
           hjust = -0.1, vjust = 0) +
  theme_minimal() +
  labs(
    x = "Number of Simulations",
    y = "z-value",
    title = "Stability of the Statistic of the Difference between Medication and Psychotherapy Control Arms",
    subtitle = "Results from 1000 simulations with within-group standard errors"
  )

pdf(file = "stab_sims.pdf")
stab_sims
dev.off()
```

It could be argued that the choice of standard errors of the changes for the calculation of the confidence intervals could have affect the results in one or the other direction. To address such concerns we have simulated 1000 different datasets with SMDs coming from a broad distribution. If standard error distributions were influential, this should show up as substantial variability across simulations. We test this idea in the Figure X which displays across the 1000 simulations the z-value of the contrast between medication and psychotherapy control arms (the mean of which we presented in Table X). As can be seen, the variability in the z-score is minimal and consistently far away from the threshold for significance, i.e. the value of z = 1.645.

![](stab_sims.pdf)

#### Addressing Regression to the Mean

We first conducted a meta-analysis to test for differences in means at baseline in the two instruments, CDRS-R and HAM-D, on which there was a sufficient number of studies to metanalyse.

```{r, warning=FALSE, message = FALSE, echo=FALSE}

### get CDRS studies
cdrs_studies_for_baseline_metareg <- list_cdrs_study[[1]] # created this above, you can take any of these, as the means and sds are repeated.

#turn sds to sec
cdrs_studies_for_baseline_metareg$baseline_stand_error <- cdrs_studies_for_baseline_metareg$baseline_sd/
                                        sqrt(cdrs_studies_for_baseline_metareg$baseline_n)

model_1 <- as.formula(~ four_level_var)
condition <-  levels(list_df_simulated[[1]]$four_level_var) # this is the four level var 
                                                              #throughout

results_cdrs_metareg_baseline <- run_metareg_and_extract(cdrs_studies_for_baseline_metareg,  
                                                         model_1, condition)

### get HAM-D studies
hamd_study_ids <-
  df_appl_v_orange[df_appl_v_orange$instrument_name == "hamd",]$new_study_id

list_hamd_study <- list_df_simulated[[1]][list_df_simulated[[1]]$new_study_id
                                                          %in% hamd_study_ids,]

hamd_studies_for_baseline_metareg <- list_hamd_study #

#turn sds to sec
hamd_studies_for_baseline_metareg$baseline_stand_error <- hamd_studies_for_baseline_metareg$baseline_sd/
                                        sqrt(hamd_studies_for_baseline_metareg$baseline_n)

model_1 <- as.formula(~ four_level_var)
condition <-  levels(list_df_simulated[[1]]$four_level_var) # this is the four level var 
                                                              #throughout
results_hamd_metareg_baseline <- run_metareg_and_extract(hamd_studies_for_baseline_metareg, 
                                                         model_1, condition)

results_hamd_metareg_baseline$condition <-  levels(list_df_simulated[[1]]$four_level_var)

```

As can be seen in Tables X and Y, the baseline scores in medication studies were on average substantially higher for the CDRS (10.9 points for the control arms) and the HAM-D (7.3 points for the control arms), respectively. The differences were significant at p\<0.05 for CDRS, thought they did not reach statistical significance for the HAM-D (p = 0.0573).

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# knitr:: kable(results_cdrs_metareg_baseline[[2]], caption="Table 3a CDRS baselilne scores")
```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# knitr:: kable(results_hamd_metareg_baseline[[2]], caption="Table 3b HAM-D baselilne scores")
```

We next addressed potential regression to the mean by residualising SMDs by the baseline scores in one regression model.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#### Let's start from the beginning and regress cohens on the baseline dataframe.

# create a new list with dfs for the sims
df_long_for_metan_adj_cohens_d <- df_long_for_metan
# rename the cohens_d value to make space for the new cohen's d
df_long_for_metan_adj_cohens_d <-
  df_long_for_metan_adj_cohens_d %>%
rename(unadjusted_cohens_d = cohens_d,
)
# estimate the cohen's d from the residuals
model_adjust_cohens_d <-  lm(df_long_for_metan_adj_cohens_d$unadjusted_cohens_d ~ df_long_for_metan_adj_cohens_d$perc_baseline_mean, na.action=na.exclude)

# there is a great effect of the baseline means, as shown here
results_adjusted_cohens_d <- broom:: tidy(model_adjust_cohens_d )

# extract the residuals which are the adjusted cohen's d, to get the simulated standard errors
df_long_for_metan_adj_cohens_d$cohens_d <- residuals(model_adjust_cohens_d )

# apply function to simulate standard errors
list_df_simulated_adj_cohens_d <- simulate_dataframes_for_st_errors(df = df_long_for_metan_adj_cohens_d,
                                   num_repetitions = 1000,
                                   seed  = 1974,
                                   n = length(unique(df_long_for_metan_adj_cohens_d$new_study_id)),
                                   a = 0.45,
                                   b = .9,
                                   mean = 0.65,
                                   sd = 0.2)

# add the four level variable and re-level
for(i in 1: length(list_df_simulated_adj_cohens_d)){

  list_df_simulated_adj_cohens_d[[i]]$arm_effect_size <- factor(list_df_simulated_adj_cohens_d[[i]]$arm_effect_size)
  list_df_simulated_adj_cohens_d[[i]] <- list_df_simulated_adj_cohens_d[[i]] %>%
    mutate(four_level_var = case_when(psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
                                      psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control")
    )

  list_df_simulated_adj_cohens_d[[i]]$four_level_var <- factor(list_df_simulated_adj_cohens_d[[i]]$four_level_var) # turn to factor

  # relevel so that medication control becomes the reference category for the regression
  #  list_df_simulated[[i]]$four_level_var <- relevel(list_df_simulated[[1]]$four_level_var, ref = "medication_control")
  list_df_simulated_adj_cohens_d[[i]]$four_level_var <-   fct_relevel(list_df_simulated_adj_cohens_d[[i]]$four_level_var,
                                                         "medication_control",
                                                         "medication_active",
                                                         "psychotherapy_control",
                                                         "psychotherapy_active")
}

# Run metaregression model (assuming that the four level variable is present given that I create this dataframe from the list_)
# specify model
model_x <- as.formula(~ four_level_var)

# run metareg function
# here you can use the simulated results directly
list_model_1_meta_reg_adj_cohens_d <- run_metaregression(list_df_simulated_adj_cohens_d, model_x)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated_adj_cohens_d[[1]]$four_level_var)
aggregate_results_overall_adjusted_cohens_d <-  aggregate_model_results(list_model_1_meta_reg_adj_cohens_d, condition)
aggregate_results_overall_adjusted_cohens_d <- cbind(aggregate_results_overall_adjusted_cohens_d, condition)

# extract SMDs from the  list
list_dummy_var_means_adjusted_cohens_d <-  lapply(list_model_1_meta_reg_adj_cohens_d, extract_coefficients_func)

# calculate mean SMDs and ses
df_mean_coefs_from_sim_adjusted_cohens_d <- calculate_mean_coefs_ses(list_dummy_var_means_adjusted_cohens_d, condition)

df_mean_coefs_from_sim_adjusted_cohens_d <- data.frame(cbind(condition, df_mean_coefs_from_sim_adjusted_cohens_d))

```

Table X shows the adjusted SMDs with their confidence intervals that we estimated in metaregression. As can be inferred from the non-overlapping confidence intervals, there were significant differences between the medication control and psychotherapy control arms.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# knitr:: kable (df_mean_coefs_from_sim_adjusted_cohens_d, caption = "Table 5. SMDs Adjusted by baseline score" )
```

Finally, we tested whether the simulated correlation values we used for simulation of the standard errors had an appreciable effect on the values of the adjusted SMDs. Supplemental Figure X shows that the variability of the estimated z-value was minimal<!--# Charlotte, the figure below should be supplemental-->.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# 
# # plot the mean z for the difference between controls
# z_psy_con_vs_med_con <- 0
# for(i in 1: length(list_model_1_meta_reg_adj_cohens_d )){
#   z_psy_con_vs_med_con[i] <- list_model_1_meta_reg_adj_cohens_d [[i]]$zval[condition=="psychotherapy_control"]
#   
# }
# 
# 
# 
# stability_sims_adjusted_cohens_d <- data.frame(n_sims = 1:1000, z_value = z_psy_con_vs_med_con)
# 
# 
# stab_sims_adjusted_cohens_d <- stability_sims_adjusted_cohens_d %>%
#   ggplot(aes(x = n_sims, y = z_value)) +
#   geom_point() +
#   geom_hline(yintercept = 1.65, linetype = "dotted", color = "red") +
#   geom_segment(aes(x = 300, xend = 300, y = 2, yend = 1.65),
#                arrow = arrow(length = unit(0.3, "cm")),
#                color = "black") +
#   annotate("text", x = 300, y = 2.0, label = "p < 0.05 threshold (values above line significant)", hjust = -0.1, vjust = 0) +
#   annotate("text", x = 300, y = stability_sims_adjusted_cohens_d$z_value[1] - 0.1, label = "simulated z-values",
#            hjust = -0.1, vjust = 0) +
#   theme_minimal() +
#   labs(
#     x = "Number of Simulations",
#     y = "z-value",
#     title = "Stability of the Statistic of the Difference between Medication and Psychotherapy Control Arms",
#     subtitle = "Results from 1000 simulations with within-group standard errors"
#   )
# stab_sims_adjusted_cohens_d
# pdf(file = "stab_sims_adjusted_cohens_d.pdf")
# stab_sims_adjusted_cohens_d
# dev.off()

```

## Comparing the nature and intensity of control conditions in psychotherapy trials

```{r, warning=FALSE, message = FALSE, echo=FALSE}

# We need to read in the appropriate dataset which describes the active and control conditions used in
# psychotherapy trials.

df_descr_psy_conditions <- readxl:: read_excel("for_despina_and_giannis_200224.xlsx")

# checking id matchup
# discrepancies <- setdiff(df_descr_psy_conditions$new_study_id, df_long_for_metan$new_study_id)

# Cleaning the dataset
# For now lets look at variables pertaining to the nature / intensity of interventions

df_descr_psy_conditions <- df_descr_psy_conditions %>% 
  select(c(new_study_id, `Psychotherapy/control`, levels_var, number_of_sites, Number_of_sessions,
           frequency_weeks, Length_of_sessions_mins, Total_hours_of_intervention_hours, total_hours_gsh,
           Total_period_of_intervention_weeks, Format_group_individual, Delivery, Involvement_of_others))

# rename for clarity and to match master dataframe
df_descr_psy_conditions <- df_descr_psy_conditions %>% 
  rename(treatment = 'Psychotherapy/control', no_sites = number_of_sites, no_sessions =
           Number_of_sessions, freq_weeks = frequency_weeks, length_sessions_mins =
           Length_of_sessions_mins, total_hours = Total_hours_of_intervention_hours,
         total_period_weeks = Total_period_of_intervention_weeks, format = Format_group_individual, delivery =
           Delivery, family_involv = Involvement_of_others)

check <- df_descr_psy_conditions %>% 
  filter(treatment == 'wlc')

# Visual inspection

# There are two studies where the waitlist condition has some intervention (i.e. values should not be set to 0)
# Otherwise, I want values to be set to 0, not NA.

df_descr_psy_conditions <- df_descr_psy_conditions %>%
  mutate(
    no_sessions = ifelse(treatment == "wlc" & !(new_study_id %in% c("Ackerson, 1998", "Diamond, 2002")), 0, no_sessions),
    freq_weeks = ifelse(treatment == "wlc" & !(new_study_id %in% c("Ackerson, 1998", "Diamond, 2002")), 0, freq_weeks),
    length_sessions_mins = ifelse(treatment == "wlc" & !(new_study_id %in% c("Ackerson, 1998", "Diamond, 2002")), 0, length_sessions_mins),
    total_hours = ifelse(treatment == "wlc" & !(new_study_id %in% c("Ackerson, 1998", "Diamond, 2002")), 0, total_hours)
  )

# Now cau

check <- df_descr_psy_conditions %>% 
  filter(treatment == 'cau')

# All studies except for Martinovic, Srivastava, Stikkelbroek and Weisz 2009 should have variables set to NA. 
# This is because we cannot say that they did not feature 0 hours of intervention, but that for the most part
# it is not reported or it is not possible to quantify.

df_descr_psy_conditions <- df_descr_psy_conditions %>%
  mutate(
    no_sessions = ifelse(treatment == "cau" & !(new_study_id %in% c("Martinovic, 2006", "Srivastava, 2020", "Stikkelbroek, 2020", "Weisz, 2009")), NA, no_sessions),
    freq_weeks = ifelse(treatment == "cau" & !(new_study_id %in% c("Martinovic, 2006", "Srivastava, 2020", "Stikkelbroek, 2020", "Weisz, 2009")), NA, freq_weeks),
    length_sessions_mins = ifelse(treatment == "cau" & !(new_study_id %in% c("Martinovic, 2006", "Srivastava, 2020", "Stikkelbroek, 2020", "Weisz, 2009")), NA, length_sessions_mins),
    total_hours = ifelse(treatment == "cau" & !(new_study_id %in% c("Martinovic, 2006", "Srivastava, 2020", "Stikkelbroek, 2020", "Weisz, 2009")), NA, total_hours)
  )

# Now other controls

check <- df_descr_psy_conditions %>%
  filter(!(treatment %in% c('cau', 'wlc'))) %>% 
  filter(levels_var == "psychotherapy_control")

# Made a few manual adjustments in the original excel sheet.

# List of columns to clean
columns_to_clean <- c(
  "no_sessions",
  "freq_weeks",
  "length_sessions_mins",
  "total_hours",
  "total_period_weeks"
)

# Convert non-numeric values to NA in specified columns
df_descr_psy_conditions <- df_descr_psy_conditions %>%
  mutate_at(vars(columns_to_clean), ~as.numeric(as.character(.)))

# Giannis' function for summary stats

group_function <- function(dataframe, my_groups, columns) {
  result_dataframe <- data.frame(
    Group = character(),
    Column = character(),
    N = numeric(),
    Mean = numeric(),
    SD = numeric(),
    Median = numeric(),
    IQR = numeric(),
    stringsAsFactors = FALSE
  )
  unique_groups <- unique(dataframe[[my_groups]])
  for (group in unique_groups) {
    group_data <- dataframe[dataframe[[my_groups]] == group, ]
    for (col in columns) {
      col_data <- group_data[[col]]
      n <- sum(!is.na(col_data))
      mean_val <- mean(col_data, na.rm = TRUE)
      SD <- sd(col_data, na.rm = TRUE)
      median_val <- median(col_data, na.rm = TRUE)
      iqr <- IQR(col_data, na.rm = TRUE)
      result_dataframe <- rbind(result_dataframe,
                                data.frame(Column = col, Group = group, N = n, Mean = mean_val,
                                           SD = SD, Median = median_val,
                                           IQR = iqr, stringsAsFactors = FALSE))
    }
  }
  return(result_dataframe)
}

# Filter out wl conditions
df_descr_psy_conditions <- df_descr_psy_conditions %>% 
  filter(new_study_id %in% no_wl_ids)

# count_wl <- df_long_for_metan %>% 
#   filter(treatment == "wlc") %>% 
#   distinct(new_study_id) %>% 
#   nrow()
# 
# count_cau <- df_long_for_metan %>% 
#   filter(treatment == "cau") %>% 
#   distinct(new_study_id) %>% 
#   nrow()

sessions <- group_function(df_descr_psy_conditions,"levels_var","no_sessions")
frequency <- group_function(df_descr_psy_conditions,"levels_var","freq_weeks")
length_sessions_mins <- group_function(df_descr_psy_conditions,"levels_var","length_sessions_mins")
total_hours <- group_function(df_descr_psy_conditions,"levels_var","total_hours")
total_period_weeks <- group_function(df_descr_psy_conditions,"levels_var","total_period_weeks")

control_summary <- rbind(sessions, frequency, length_sessions_mins, total_hours, total_period_weeks)

control_summary <- control_summary %>% 
  select(-c(Median, IQR))

control_summary <- control_summary %>%
  mutate(Group = case_when(
    Group == "psychotherapy_active" ~ "Active",
    Group == "psychotherapy_control" ~ "Control",
    TRUE ~ as.character(Group)  # Keep other values unchanged
  ))

control_summary <- control_summary %>%
  mutate(Column = case_when(
    Column == "no_sessions" ~ "Number of sessions",
    Column == "freq_weeks" ~ "Intensity (sessions per week)",
    Column == "length_sessions_mins" ~ "Session length (mins)",
    Column == "total_hours" ~ "Total intervention hours",
    Column == "total_period_weeks" ~ "Intervention duration (weeks)",
    TRUE ~ as.character(Column)  # Keep original value if no match
  ))

control_summary <- control_summary %>% 
  mutate(across(where(is.numeric), ~round(., 2)))

as_grouped_data(control_summary, groups = "Column") %>% 
  as_flextable(hide_grouplabel = TRUE ) %>% 
  bold(i = c(1, 4, 7, 10, 13), bold = TRUE, part = "body") %>%  
  set_table_properties(width = 1, layout = "autofit") %>%  # Set table width 
  autofit() %>%  # Autofit column widths 
  align(j = 2:4, align = "center", part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  add_header_row(values = "Comparing the intensity of the intervention between active and control arms of psychotherapy studies", colwidths = 4) %>% # This is a work around for the problems with rendering captions
  add_header_row(values = "Table 5", colwidths = 4) %>% 
  bold(i = c(1, 3), part = "header") %>%   # Bold the relevant header rows
  bg(i = 3, bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(i = 3, part = "header", border = big_border) %>% 
  padding(i = 2, padding.bottom = 15, part = "header") %>% 
  paginate(init = TRUE) %>% 
  line_spacing(space = .7, part = "all")

# t tests

session_no_t_test <- t.test(no_sessions ~ levels_var, data = df_descr_psy_conditions)
freq_t_test <- t.test(freq_weeks ~ levels_var, data = df_descr_psy_conditions)
session_length_t_test <- t.test(length_sessions_mins ~ levels_var, data = df_descr_psy_conditions)
total_hours_t_test <- t.test(total_hours ~ levels_var, data = df_descr_psy_conditions)
total_period_t_test <- t.test(total_period_weeks ~ levels_var, data = df_descr_psy_conditions)

extract_t_test_stats <- function(df) {

  t_test_stats <- data.frame(
  outcome = df$data.name,
  statistic = df$statistic,
  df = df$parameter,
  p_value = df$p.value,
  ci_lower = df$conf.int[1],
  ci_upper = df$conf.int[2])

  return(t_test_stats)

}

results_session_no_t_test <- extract_t_test_stats(session_no_t_test)
results_freq_t_test <- extract_t_test_stats(freq_t_test)
results_session_length_t_test <- extract_t_test_stats(session_length_t_test)
results_total_hours_t_test <- extract_t_test_stats(total_hours_t_test)
results_total_period_t_test <- extract_t_test_stats(total_period_t_test)

control_summary_t_tests <- rbind(results_session_no_t_test, results_freq_t_test, results_session_length_t_test, results_total_hours_t_test, results_total_period_t_test)

# For row names
control_summary_t_tests <- control_summary_t_tests %>%
  mutate(outcome = case_when(
    row_number() == 1 ~ "Number of sessions",
    row_number() == 2 ~ "Intensity (sessions per week)",
    row_number() == 3 ~ "Session length (mins)",
    row_number() == 4 ~ "Total intervention hours",
    row_number() == 5 ~ "Intervention duration (weeks)"
  ))

# Define new column names
new_column_names <- c("Outcome", "t statistic", "df", "p-value", "Lower CI", "Upper CI")

# Assign new column names to the data frame
colnames(control_summary_t_tests) <- new_column_names

#Rounding
control_summary_t_tests <- control_summary_t_tests %>%
  mutate(across(where(is.numeric) & !matches("p-value"), ~ round(., 2))) %>%
  mutate(`p-value` = ifelse(`p-value` < 0.001, "< 0.001", sprintf("%.3f", `p-value`)))

flextable(control_summary_t_tests) %>%
  set_table_properties(width = 1, layout = "autofit") %>%  # Set table width %>% 
  autofit() %>%  # Autofit column widths 
  align(j = 2:ncol(control_summary_t_tests), align = "center", part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  add_header_row(values = "Results for t-tests comparing intervention intensity between active and control arms of psychotherapy trials", colwidths = 6) %>% # This is a work around for the problems with rendering captions
  add_header_row(values = "Table 6", colwidths = 6) %>% 
  bold(i = c(1, 3), part = "header") %>%   # Bold the relevant header rows
  bg(i = 3, bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(i = 3, part = "header", border = big_border) %>% 
  padding(i = 2, padding.bottom = 15, part = "header") %>% 
  paginate(init = TRUE) %>% 
  line_spacing(space = .7, part = "all")

```

Overall, the active arms of psychotherapy studies were considerably more intensive than the control arms they were compared against (see Table 6 for summary statistics). Active intervention arms featured significantly more sessions than control arms (*t* (`r control_summary_t_tests$df[control_summary_t_tests$Outcome == "Number of sessions"]`) = `r control_summary_t_tests$"t statistic"[control_summary_t_tests$Outcome == "Number of sessions"]`, *p* = `r control_summary_t_tests$"p-value"[control_summary_t_tests$Outcome == "Number of sessions"]`). These sessions were also significantly longer (*t* (`r control_summary_t_tests$df[control_summary_t_tests$Outcome == "Session length (mins)"]`) = `r control_summary_t_tests$"t statistic"[control_summary_t_tests$Outcome == "Session length (mins)"]`, *p* = `r control_summary_t_tests$"p-value"[control_summary_t_tests$Outcome == "Session length (mins)"]`) and more frequent (*t* (`r control_summary_t_tests$df[control_summary_t_tests$Outcome == "Intensity (sessions per week)"]`) = `r control_summary_t_tests$"t statistic"[control_summary_t_tests$Outcome == "Intensity (sessions per week)"]`, *p* = `r control_summary_t_tests$"p-value"[control_summary_t_tests$Outcome == "Intensity (sessions per week)"]`). The total hours involved in an intervention were higher in active vs control arms (*t* (`r control_summary_t_tests$df[control_summary_t_tests$Outcome == "Total intervention hours"]`) = `r control_summary_t_tests$"t statistic"[control_summary_t_tests$Outcome == "Total intervention hours"]`, *p* = `r control_summary_t_tests$"p-value"[control_summary_t_tests$Outcome == "Total intervention hours"]`), though the total period of the intervention was similar (*t* (`r control_summary_t_tests$df[control_summary_t_tests$Outcome == "Total intervention period (weeks)"]`) = `r control_summary_t_tests$"t statistic"[control_summary_t_tests$Outcome == "Intervention duration (weeks)"]`, *p* = `r control_summary_t_tests$"p-value"[control_summary_t_tests$Outcome == "Intervention duration (weeks)"]`).

# Discussion

This paper took as its starting point the question of how anyone, be it a patient, parent, or clinician, should decide whether to opt for medication or psychotherapy for the treatment of adolescent depression. In order to address this question, we asked whether the two treatment modalities --- medication and psychotherapy --- are comparable on the basis of the existing evidence. Specifically, our paper answered two questions that are fundamental to any attempts at comparing the two treatment modalities.

First, whether the participants of trials in one modality are comparable to those in another modality. Second, whether key conditions of the trial, such as the effects of control conditions or the number of sites involved in a trial, are comparable.

Starting with the first question, we find evidence that participants in the two treatment modalities differ in the following key aspects: those enrolled in medication trials are more likely to be male and have more severe depression compared to those in psychotherapy trials. By contrast, there were no significant differences in age.

Severity is important as it may moderate treatment response, with some evidence suggesting that those with higher baseline scores respond more to antidepressant therapy [@stoneResponseAcuteMonotherapy2022] or that their response to pill placebo is lower [@bridgePlaceboResponseRandomized2009]. Some studies argue against severity as a treatment moderator [@trogerBaselineDepressionSeverity2024; @weitzBaselineDepressionSeverity2015], however, these are within modality (e.g. within CBT studies), i.e. within those people who have chosen to be in the particular trial and modality. Moreover, there is evidence that severity may represent different subtypes in terms of course of depression and its outcomes in real life settings [@lamersSixyearLongitudinalCourse2016; @simmonds-buckleyDepressionSubtypesTheir2021].

Similarly, we found that on average there were about 7% more females in psychotherapy trials. Whilst there is little evidence that sex or sex moderate treatment response within modality in adult trials of depression [@cuijpersGENDERPREDICTORMODERATOR2014], this is yet another indicator that different people enter trials of each modality type, which would violate basic assumptions of comparability between trials.

We then turned to our second basic question, about whether key conditions of the trial design were comparable between modalities. We find a series of critical differences between the two modalities.

First, medication trials are vastly more likely to be multi-site than their psychotherapy counterparts; in the current study, 93% of medication RCTs were multisite compared to 54% of psychotherapy RCTs Multisite trials are associated with higher pill placebo response [@bridgePlaceboResponseRandomized2009], and are less common in NIH-funded (i.e. non industry) trials which are also more likely to show higher pill placebo efficacy. Also, in single-site trials, principal investigators are often intellectually invested in the treatment (in psychotherapy these are often treatments developed or modified by the PI); this is in stark contrast to the incentive structure in multi-site trials where primacy is given to the number of recruited participants which is the primary unit of reimbursement.

Second, psychotherapy controls have moderate effect size (-0.5) whereas pill-placebo has very large effect size (-1.9). Our analysis could be critiqued as it relies on comparing the within arm symptom change of each trial taken out of randomization, which is generally advised against (CITE). This criticism would apply if our aim were to draw inferences about the efficacy of each arm --- in which case preserving the randomization (in order to balance confounders) is critical. However, we note that, our findings are largely in keeping with those of the network metanalysis, which is designed to preserve the randomization structure. More importantly, we do not make the claim with our analyses that these differences are genuinely due to efficacy differences; they may well be because of the fact that the people who attend psychotherapy and medication trials are different and therefore respond differently. In either case (difference in efficacy vs difference in trial participant profile), the vast disparity in the response to control conditions is reason for major concern about our ability to draw inferences concerning the comparison between the two modalities. This is all the more so as clinicians as well as policy makers often resort to effect sizes to summarise findings, with SMD being the one used as per standard. Our findings make obvious that comparing treatment modalities based on the effect size of each modality is misleading.

Moving beyond the comparison of the controls between the two modalities, we examined whether psychotherapy controls are reasonable counterfactuals to receiving the treatment. An optimal control condition is one where the treatment differs (e.g. treatment vs no treatment), but everything else is equal (in technical terms, the ceteris paribus assumption). An obvious disadvantage of psychotherapy trials is that they are typically unblinded (and hard to blind). Yet, our results show that psychotherapy trials are unlikely to fulfill some other very basic conditions of the "all else is equal" assumption. Most (XXX %) of psychotherapy RCT control conditions are either waitlist control or treatment as usual, both of which are very likely to create negative expectations for participants not randomized to treatment; thus, the comparison is not between treatment and no treatment, but rather treatment and the poor luck of being randomized to waiting. But we also find that even psychological controls were unreasonable counterfactuals. In order to test that a psychological treatment is effective per se (e.g. because of the cognitive techniques the therapist deploys) rather than because of generic effects (e.g. pleasant human contact), aspects such as human contact time should be matched. We find that there are vast differences between treatment and control arms in psychotherapy trials (CITE NUMBER) which may lead to an inflation of the true efficacy of psychological treatments.

Given all of the above, the certainty with which guidelines, including UK's NICE recommend psychotherapy over medication for adolescent depression is surprising. Indeed, we believe that our findings have a number of profound implications for patients, their families, clinicians and policy makers and we list these below.

First, the grounds for a comparison between medication and psychotherapy should be seen as shaky, rather than offering confidence, and there is an urgent need to revisit guidelines and public information in light of the limitations.

Second, the low quality of control conditions of psychotherapy trials for adolescent depression should prompt a rethinking of how these studies are appraised and what needs to be done to create fair and realistic comparators. Indeed, investment of effort and funds should be directed into providing rigorous evidence that established depression psychotherapies are more efficacious than fair controls. There are examples of such psychotherapy RCTS, e.g. in social anxiety [@clarkCognitiveTherapyExposure2006a] where such rigor has been applied.

Third, our findings make clear the inherent difficulties of comparing psychotherapy with medication trials [@delgiovaneCombiningPharmacologicalNonpharmacological2019]. The first obstacle is the comparability of the populations taking part. Even if a trial were designed to conduct a head-to-head comparison of psychotherapy with medication (as has been done in [@FluoxetineCognitiveBehavioralTherapy2004]), it might sample the population of those who are indifferent to which one they would receive. And even in such a design, difficulties with blinding of the psychotherapy control would have to be overcome to draw valid inferences.

In summary, our data give cause for consternation about the state of the evidence of youth depression, one of the most common and debilitating disorders in young people. Our data question the state of knowledge about the efficacy of youth psychotherapies and, in particular, the extent to which giving them primacy in the treatment of depression is justified and beneficial for young people. Returning to our motivating question, the stakeholders, including patients and clinicians, deserve better evidence on which to base their choice than currently exists.

# Supplement

PRISMA chart summarising the screening process for the systematic review.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# # Tables for baseline comparisons
# 
# # Lets make a table that combines across the sensitivity analyses for severity
# 
# df_severity_sens <- rbind(results_met_sev_cbt, results_met_sev_clin, results_met_sev_no_wl)
# 
# # For row names
# df_severity_sens <- df_severity_sens %>%
#   mutate(subgroup = case_when(
#     row_number() == 1 ~ "CBT, fluoxetine and escitalopram",
#     row_number() == 4 ~ "Excluding subclinical",
#     row_number() == 7 ~ "Excluding waitlist",
#     TRUE ~ if_else(subgroup == 0, "Medication", "Psychotherapy")
#   )) 
# 
# # Define new column names
# new_column_names <- c(" ", "K", "Mean", "SE", "Lower CI", "Upper CI", "T2", "p-value")
# 
# # Assign new column names to the data frame
# colnames(df_severity_sens) <- new_column_names
# 
# # Rounding
# df_severity_sens <- df_severity_sens %>%
#   mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
#   mutate_at(vars(ncol(.)), ~ round(., 3))
# 
# df_severity_sens %>%
#   mutate_all(~if_else(is.na(.), "", as.character(.))) 
# 
# flextable(df_severity_sens) %>%
#   set_caption(caption = "Table S1: Sensitivity analyses for comparison of baseline severity levels between psychotherapy and medication trials") %>%
#   bold(part = "header") %>%  # Bold the header row
#   bold(i = c(1, 4, 7), j = NULL, bold = TRUE, part = "body") %>%  # Bold specific rows
#   border_outer(border = small_border) %>%  # Add outer borders
#   hline(i = c(1, 2, 4, 5, 7, 8), border = small_border) %>%  # Add light grey horizontal lines all  rows
#   hline(i = c(3, 6), border = big_border) %>%  # Add darker horizontal lines after specific rows %>% 
#   hline(part = "header", border = big_border) %>% 
#   vline(j = 1, border = big_border) %>% 
#   set_table_properties(width = .75) %>%  # Set table width
#   bg(bg = "#F2F2F2", part = "header") %>%  # Add light grey background to header row
#   autofit() %>%  # Autofit column widths
#   font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri 
#   padding(i = c(2, 3, 5, 6, 8, 9), j = 1, padding.left = 10) %>% 
#   align(j = 2:ncol(df_overall_baseline_results), align = "center", part = "all")
# 
# # Now sex
# 
# df_sex_sens <- rbind(results_met_perc_women_cbt, results_met_perc_women_clin, results_met_perc_women_no_wl) 
# 
# # For row names
# df_sex_sens <- df_sex_sens %>%
#   mutate(subgroup = case_when(
#     row_number() == 1 ~ "CBT, fluoxetine and escitalopram",
#     row_number() == 4 ~ "Excluding subclinical",
#     row_number() == 7 ~ "Excluding waitlist",
#     TRUE ~ if_else(subgroup == 0, "Medication", "Psychotherapy")
#   )) 
# 
# # Define new column names
# new_column_names <- c(" ", "K", "Mean", "SE", "Lower CI", "Upper CI", "T2", "p-value")
# 
# # Assign new column names to the data frame
# colnames(df_sex_sens) <- new_column_names
# 
# # Rounding
# df_sex_sens <- df_sex_sens %>%
#   mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
#   mutate_at(vars(ncol(.)), ~ round(., 3))
# 
# df_sex_sens %>%
#   mutate_all(~if_else(is.na(.), "", as.character(.))) 
# 
# flextable(df_sex_sens) %>%
#   set_caption(caption = "Table S2: Sensitivity analyses for comparison of sex composition between psychotherapy and medication trials") %>%
#   bold(part = "header") %>%  # Bold the header row
#   bold(i = c(1, 4, 7), j = NULL, bold = TRUE, part = "body") %>%  # Bold specific rows
#   border_outer(border = small_border) %>%  # Add outer borders
#   hline(i = c(1, 2, 4, 5, 7, 8), border = small_border) %>%  # Add light grey horizontal lines all  rows
#   hline(i = c(3, 6), border = big_border) %>%  # Add darker horizontal lines after specific rows %>% 
#   hline(part = "header", border = big_border) %>% 
#   vline(j = 1, border = big_border) %>% 
#   set_table_properties(width = .75) %>%  # Set table width
#   bg(bg = "#F2F2F2", part = "header") %>%  # Add light grey background to header row
#   autofit() %>%  # Autofit column widths
#   font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri 
#   padding(i = c(2, 3, 5, 6, 8, 9), j = 1, padding.left = 10) %>% 
#   align(j = 2:ncol(df_overall_baseline_results), align = "center", part = "all")
# 
# # Now age
# 
# df_age_sens <- rbind(results_met_age_cbt, results_met_age_clin, results_met_age_clin) 
# 
# # For row names
# df_age_sens <- df_age_sens %>%
#   mutate(subgroup = case_when(
#     row_number() == 1 ~ "CBT, fluoxetine and escitalopram",
#     row_number() == 4 ~ "Excluding subclinical",
#     row_number() == 7 ~ "Excluding waitlist",
#     TRUE ~ if_else(subgroup == 0, "Medication", "Psychotherapy")
#   )) 
# 
# # Define new column names
# new_column_names <- c(" ", "K", "Mean", "SE", "Lower CI", "Upper CI", "T2", "p-value")
# 
# # Assign new column names to the data frame
# colnames(df_age_sens) <- new_column_names
# 
# # Rounding
# df_age_sens <- df_age_sens %>%
#   mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
#   mutate_at(vars(ncol(.)), ~ round(., 3))
# 
# df_age_sens %>%
#   mutate_all(~if_else(is.na(.), "", as.character(.))) 
# 
# flextable(df_age_sens) %>%
#   set_caption(caption = "Table S3: Sensitivity analyses for comparison of age between psychotherapy and medication trials") %>%
#   bold(part = "header") %>%  # Bold the header row
#   bold(i = c(1, 4, 7), j = NULL, bold = TRUE, part = "body") %>%  # Bold specific rows
#   border_outer(border = small_border) %>%  # Add outer borders
#   hline(i = c(1, 2, 4, 5, 7, 8), border = small_border) %>%  # Add light grey horizontal lines all  rows
#   hline(i = c(3, 6), border = big_border) %>%  # Add darker horizontal lines after specific rows %>% 
#   hline(part = "header", border = big_border) %>% 
#   vline(j = 1, border = big_border) %>% 
#   set_table_properties(width = .75) %>%  # Set table width
#   bg(bg = "#F2F2F2", part = "header") %>%  # Add light grey background to header row
#   autofit() %>%  # Autofit column widths
#   font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri 
#   padding(i = c(2, 3, 5, 6, 8, 9), j = 1, padding.left = 10) %>% 
#   align(j = 2:ncol(df_overall_baseline_results), align = "center", part = "all")

```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# Plots  -----------------------

#define groups
df_long_for_metan <- df_long_for_metan %>%
  mutate(
    group = case_when(
      psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "Medication Active",
      psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "Medication Control",
      psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "Psychotherapy Active",
      psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "Psychotherapy Control",
      TRUE ~ NA_character_
    )
  )

group_colors <- c("Medication Active" = "deeppink1", "Medication Control" = "deeppink4", "Psychotherapy Active" = "steelblue1", "Psychotherapy Control" = "steelblue3")

# # Scatter plot cohens d per arm
# 
# ggplot(df_long_for_metan, aes(x = group, y = cohens_d, color = group)) +
#   geom_point(position = position_jitter(width = 0.2), size = 3) +
#   geom_segment(x = 4.5, xend = 4.5, y = -3.5, yend = -3,
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -3, label = "Less Effective", color = "grey", vjust = 0, hjust = .2) +
#   geom_segment(x = 4.5, xend = 4.5, y = -4, yend = -4.5,
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -4.5, label = "More Effective", color = "grey", vjust = 0, hjust = .2) +
#   scale_color_manual(values = group_colors) +
#   labs(title = "Scatter plot of Cohen's d by group", x = NULL, y = "Cohen's d") +
#   theme_minimal() +
#   theme(legend.position = "none")

# Box plot Cohens d by group

ggplot(df_long_for_metan, aes(x = group, y = cohens_d, fill = group)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(0.8), color = "black", alpha = 0.5) +
  scale_y_continuous(limits = c(-3.5, 1), breaks = seq(-4, 1, by = 1)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_segment(x = 4.5, xend = 4.5, y = -2.5, yend = -2,
               arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = "Psychotherapy Control", y = -2.2, label = "Less Effective", color = "grey", vjust = 0, hjust = .2) +
  geom_segment(x = 4.5, xend = 4.5, y = -3, yend = -3.5,
               arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = "Psychotherapy Control", y = -3.2, label = "More Effective", color = "grey", vjust = 0, hjust = .2) +
  scale_fill_manual(values = group_colors) +
  labs(title = "Box plot of cohens_d by group", x = "Group", y = "Cohen's d") +
  theme_minimal() +
  theme(legend.position = "none")

# # Plot percentage baseline means
# 
# ggplot(df_long_for_metan, aes(x = group, y = perc_baseline_mean, color = group)) +
#   geom_point(position = position_jitter(width = 0.2), size = 3) +
#   scale_color_manual(values = group_colors) +
#   labs(title = "Scatter plot of % baseline mean by group", x = NULL, y = "Percentage Baseline Mean") +
#   theme_minimal() +
#   theme(legend.position = "none")

```

# References