### 2024-05-16 Manuscript_for_rendering_in_word.qmd
Vancouver formatting style and word cuts

---
title: "Comparing Apples and Oranges in Youth Depression Treatments?"
subtitle: "A Quantitative Critique of the Evidence Base and Guidelines"
author:
  - name: Argyris Stringaris
    email: a.stringaris@ucl.ac.uk
    affiliations: 
        - id: ucl_1
          name: University College London
          department: Divisions of Psychiatry and Psychology and Language Science
          address: 1-19 Torrington Place
          city: London, UK
          state: 
          postal-code: WC1E 7HB
          country: United Kingdom
    attributes:
        corresponding: true
    note: 
  - name: Charlotte Burman
    email: charlotte.burman@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
  - name: Rudolf Uher
    email: uher@Dal.Ca
    affiliations:
        - id: dal
          name: Dalhousie University
          department: Department of Psychiatry
          address: 
          city: Nova Scotia
          state: 
          postal-code: 
          country: Canada
  - name: Dayna Bhudia
    email: dayna.bhudia@ucl.ac.uk
    affiliations:
        - ref: ucl_1
  - name: Despoina Miliou
    email: dgmiliou@outlook.com.gr
    affiliations:
        - ref: ucl_1
  - name: Ioannis-Marios Rokas
    email: <rokasgiannis93@gmail.com
    affiliations:
        - ref: ucl_1
  - name: Marinos Kyriakopoulos 
  - name: Lucy Foulkes
    email: lucy.foulkes@psych.ox.ac.uk
    affiliations:
        - id: OU
          name: University of Oxford
          department: Experimental Psychology
          address: Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Road
          city: Oxford
          postal-code: OX2 6GG
          country: United Kingdom
    note:
  - name: Carmen Moreno
    email: cmoreno@hggm.es
    affiliations:
        - id: HGM
          name: Hospital Gregorio Marañón
          department: Department of Psychiatry
          address: 46 C. del Dr. Esquerdo
          city: Madrid
          state:
          postal-code: 28007
          country: Spain
  - name: Samuele Cortese
    email: samuele.cortese@soton.ac.uk
    affiliations:
        - id: soton
          name: University of Southampton
          department: Centre for Innovation in Mental Health (CIMH), School of Psychology
          address: Highfield Campus, Building 44
          city: Southampton
          state:
          postal-code: SO17 1BJ
          country: United Kingdom
  - name: Georgina Krebs
    email: g.krebs@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
keywords: 
  - depression
  - adolescents
  - antidepressants
  - psychotherapy
date: last-modified
format:
  docx:
    reference-doc: custom-reference-doc.docx
    fig-width: 9
    fig-height: 7
    fig_caption: yes
    tbl_caption: yes
    numer-sections: true
editor: visual
bibliography: "emotion_concepts_paper.bib"
csl: "vancouver.csl"
font:
  family: "Calibri"
prefer-html: true
fig-format: png
fig-cap-location: top

---

##

------------------------------------------------------------------------

------------------------------------------------------------------------

```{r, warning=FALSE, message = FALSE, echo=FALSE}

###########Libraries needed
library(tidyverse)
library(truncnorm)
library(metafor)
library(meta)
library(kableExtra)
library(flextable)
library(magrittr)
library(officer)
library(RefManageR)
library(ftExtra)
library(dplyr)
library(broom)
library(effsize)

# For referencing
bib_data <- ReadBib("emotion_concepts_paper.bib")

```
```{r global_options, include=FALSE}
knitr::opts_chunk$set(tab.cap.fp_text = officer::fp_text_lite(italic = FALSE, bold = TRUE))
```

# Abstract
**Question:** Should a young person receive psychotherapy or medication for their depression, and on what evidence do we base this decision? In this paper, we test whether the basic conditions required to draw valid inferences to answer this question are currently met.
**Study selection and analysis:** We included 88 RCTs of psychotherapy and medication for child and adolescent depression (mean age 4-18 years). Using meta-analyses, we compared a) participant characteristics and b) trial characteristics in medication and psychotherapy trials. Lastly, we examined whether psychotherapy controls are well-matched to active conditions. 
**Findings:** Participants in medication RCTs had higher depression severity and were more frequently male compared to psychotherapy RCTs. There was a dramatic difference in the within-subject improvement due to placebo (SMD=-1.9 (95% CI: -2.10 to -1.70)) vs psychotherapy controls (SMD=-0.5 (95% CI: -0.75 to -0.25)). Within psychotherapy RCTs, control conditions were less intensive on average than active conditions.
**Conclusions:** Medication and psychotherapy RCTs differ on fundamental participant and methodological characteristics, thereby violating key conditions for valid comparison between them. Psychotherapy controls often involve little therapist contact and are easy-to-beat comparators. These findings cast doubt on the confidence with which psychotherapy is recommended for youth depression, and highlight the pressing need to improve the evidence base.

# Key messages
Psychotherapy is recommended before medication for most cases of depression in children and adolescents, a recommendation that is based on indirect comparisons of outcomes from randomised controlled trials (RCTs) within each treatment modality. We examine the validity of these inferences by scrutinising the comparability of psychotherapy and medication RCTs. We find significant differences in sample characteristics (namely depression severity and sex composition) and trial design features, such that the effect sizes of medication controls (i.e. pill placebo) are much larger than those for psychotherapy controls, and that medication RCTs feature significantly more trial sites. We also examine the quality of controls used in psychotherapy RCTs and find that they are poorly matched to active intervention arms in ways such as human contact hours, and hence represent poor and easy to beat comparators. Our findings underscore the need for a higher quality evidence base upon which to base treatment guidelines and clinical decision making.

# Background

Should a child or an adolescent receive psychotherapy or medication for their depression, and what information should be used to guide decision-making?

For adolescent depression, there are limited head-to-head trials of medication and psychotherapy, and hence recommendations are derived from indirect comparisons of treatment efficacy. The National Institute of Health and Care Excellence (NICE) guidelines for adolescent depression recommend psychotherapy (specifically cognitive behaviour therapy (CBT) or interpersonal therapy (IPT)) over medication in most cases [@niceDepressionChildrenYoung2019]. This is in keeping with two sources of evidence relating to child and adolescent depression: meta-analyses of medication randomised controlled trials (RCTs) that cast doubt on the efficacy of antidepressants, with the exception of fluoxetine [@ciprianiComparativeEfficacyTolerability2016]; and meta-analyses of psychotherapy RCTs that conclude psychotherapy to be efficacious [@cuijpersEffectsPsychologicalTreatments2021]. However, a recent network meta-analysis (NMA) [@zhouComparativeEfficacyAcceptability2020], an established method of comparing treatments using both direct and indirect (i.e. treatment A with treatment C, via studies that directly compare A with B and B with C) evidence, concluded that only fluoxetine alone and fluoxetine administered together with CBT were significantly more effective than medication control (i.e. pill placebo) or psychotherapy controls. The only head-to-head RCT comparing modalities found that fluoxetine, alone and in combination with CBT, was superior to pill placebo, though CBT alone was not [@marchFluoxetineCognitiveBehavioralTherapy2004]. Given this confusing evidence base, how should patients, carers, clinicians and policy makers make treatment decisions?

In this paper, we examine whether the existing evidence for adolescent depression treatments can offer valid answers to this question. We provide a conceptual framework and test a series of hypotheses using data from existing trials. Two points are crucial to indirect comparisons of treatment modalities. First, whether the participants in trials are comparable across modalities. Second, whether key conditions of the trial, such as the effects of control conditions or the number of sites involved, are comparable.

Starting with the first point, comparison between different trials assumes that they sample from the same population. If not, the validity of any comparisons, including those conducted through NMA (which rests on the principal of transitivity, i.e. the requirement that the different sets of randomized trials are similar on average [@chaimaniChapter11Undertaking2023]) are questionable.

<!-- Indeed, comparing outcomes (Y) between psychotherapy and medication trials requires us to contrast what is called the sample average treatment effect $\tau$ of each, defined in the following way: -->

<!-- $$ -->

<!-- \tau_{psy} =  E(Y_{1} - Y_{0} \,|\, S, psy) -->

<!-- $$ $$ -->

<!-- \tau_{med} =  E(Y_{1} - Y_{0} \,|\, S, med)\  -->

<!-- $$ -->

<!-- where the operator E denotes the expectation over the differences in outcomes between those who received the intervention (T = 1) and those who received the control condition (T = 0), for each sample, $S$, where psy and med stand for psychotherapy and medication respectively. -->

<!-- Obviously, this comparison rests on the assumption that trials in both modalities sample from the same population, $P$, of patients. Formally, this can be expressed as follows: -->

<!-- $$E(Y_{1} \,|\, \text{S, psy}) = E(Y_{1} \,|\, P, \text{psy})$$ -->

<!-- $$E(Y_{1} \,|\, \text{S, med}) = E(Y_{1} \,|\, P, \text{med})$$ -->

<!-- signifying that the effect found in the population would be expected to be found in the same population for each treatment. -->

The assumption that medication and psychotherapy trials sample from the same population may not be valid as patients and parents often have treatment preferences [@jaycoxAdolescentPrimaryCare2006; @langerParentYouthPreferences2021; @mchughPatientPreferencePsychological2013], meaning that there is likely to be a self-selection bias in who participates in psychotherapy and medication trials. Moreover, treatment preferences correlate with clinically-relevant participant characteristics, including severity and sex. Some of these characteristics, such as severity, may moderate treatment response [@courtneyPredictorsModeratorsMediators2022; @lorenzo-luacesDoubleTroubleSymptom2020] and if they differ across psychotherapy and medication trials they may confound comparisons.

Regarding the second point, differences in trial design may impact outcomes in a differential way between medication and psychotherapy trials [@delgiovaneCombiningPharmacologicalNonpharmacological2019]. Most obviously, participants in psychotherapy trials are generally unblinded to treatment allocation, with the exception perhaps of trials that compare two equally plausible treatment arms [@calvoInterventionAdolescentsEarlyOnset2014]. This creates differential expectations which may favour the psychotherapy active condition, as participants are content to be receiving the “cutting edge” treatment, whilst those in the control are dissatisfied for having missed out (i.e. “disappointment bias” [@reltonEthicsTrialsCohorts2017]. By contrast, in new antidepressant trials, patients (and raters) were largely unable to judge treatment allocation [@linAssessmentBlindingRandomized2022], suggesting that expectancy effects are well-matched across conditions. Since expectancy is substantially associated with treatment outcomes [@constantinoExpectations2011], if expectancy differs between medication and psychotherapy trials, comparisons between them, including in NMA, become questionable.

Another difference in design is the number of trial sites. The number of sites in medication trials is positively related to the magnitude of placebo response [@bridgePlaceboResponseRandomized2009; @dechartresSingleCenterTrialsShow2011]. This phenomenon may be due to lower quality of assessments in multi-site trials, with higher rates of classification errors and therefore higher apparent spontaneous remission or regression to the mean.

An inter-related issue concerns the effect of control conditions. Often psychotherapy and medication are compared on the basis of their respective effect sizes (i.e. differences between the active and control conditions for each modality). For these to be comparable, medication and psychotherapy controls ought to be equal in their effects. Otherwise, misleading conclusions could be drawn, e.g. two effect sizes of 40% would be considered equal, even if one arose from a difference of 100%-60% and another from a difference of 40%-0%.

Additionally, control conditions in RCTs should generate counterfactual conditions to the intervention [@guoChapterCounterfactualFramework2014]: what would have been the outcome had an individual not received the intervention, with all else being equal. Pill placebo, where the appearance of the drug is faithfully emulated, is an effort for all else to be equal. In psychotherapy trials, control conditions may not be so well matched to the intervention (e.g. in number of hours of therapist contact).

# Objective

In this study we examine RCTs of psychotherapy and medication for child and adolescent depression (mean age 4-18 years). We hypothesise that there are substantial differences between psychotherapy and medication RCTs, making their comparison problematic, which we examine in the following ways. First, we conduct meta-analyses to compare sample characteristics of medication and psychotherapy trials including: a) baseline depression severity; b) percentage females; and c) mean age. Second, we examine trial characteristics including the efficacy of the control arms, using random-effects meta-regression, and the number of trial sites. Third, we examine the quality of psychotherapy control conditions by scrutinising the extent to which they are matched to the active intervention in ways such as number and frequency of sessions, and hence whether they represent fair pairings from which to draw valid efficacy inferences.

# Study selection and analysis

The study protocol was registered on the Open Science Framework (OSF) and can be found [here](https://osf.io/bfmc6).

A detailed description of our methods, including formalisms on which analyses are based, can be found in the Supplement.

## Included studies

We included RCTs identified in a recent meta-analysis of psychotherapy versus control [@cuijpersEffectsPsychologicalTreatments2021] and an NMA examining the efficacy of antidepressants [@ciprianiComparativeEfficacyTolerability2016] for depression in children and adolescents. For the psychotherapy trials, we utilised open data from the previous meta-analysis (see [here](https://docs.metapsy.org/databases/depression-childadol-psyctr/)). For medication trials, we were unable to access the full dataset used in the NMA and hence extracted data from the included studies ourselves.

For medication trials, we also conducted a systematic search for studies published after the final search date of Cipriani et al.'s [@ciprianiComparativeEfficacyTolerability2016] review up to the final search date of Cuijpers et al's [@cuijpersEffectsPsychologicalTreatments2021] review to ensure we analysed an equivalently up-to-date database of medication trials. Two authors screened 450 titles and abstracts, and 38 full text records. Seven studies met inclusion criteria and one author completed data extraction for these papers (see @fig-prisma for PRISMA diagram).

## Statistical Analysis

### Sample characteristics

We conducted random-effects meta-analyses and tested subgroup differences (psychotherapy vs medication trials) in severity of depressive symptoms, sex and age. Meta-analyses were implemented using R's meta package.

<!-- In order to compare depression severity across the variety of instruments the studies used, we performed a min-max normalisation to turn each study arm mean score at baseline into a percentage using the following formalism: -->

<!-- $$ -->
<!-- \text{outcome}_{percent} = \frac{\bar{X} - \text{scale}_{min}}{\text{scale}_{max} - \text{scale}_{min}} -->
<!-- $$ -->

<!-- where, $$\bar{X}$$ is the mean score for each study arm on the primary outcome questionnaire, and ${scale}_{min}$ and ${scale}_{max}$ are the minimum and maximum possible values of the scale in question, respectively. The standard deviation is calculated thus: -->

<!-- $$ -->
<!-- \text{SD}_{Xpercent} = \frac{\text{SD}_{X}}{\text{scale}_{max} - \text{scale}_{min}} -->
<!-- $$ -->

<!-- where $\text{SD}_{X}$ is the original standard deviation of the mean at baseline. -->

### Trial design

#### Measures of Effect

As the measure of effect of each individual study, we used the within-group Standardised Mean Difference (SMD) for the primary depression scale used (selected according to the hierarchy defined in the Supplement).

<!-- which we defined following [@lakensCalculatingReportingEffect2013, @cummingUnderstandingNewStatistics2013] as: -->

<!-- $$SMD_{change} = \frac{Mean_{t_{2}} - Mean_{t_{1}}}{\frac{SD_{t{2}} + SD_{t{1}}}2}\ $$ -->

<!-- where, $Mean_{t_{2}}$ and $Mean_{t_{1}}$ refer to the means of the main outcome score at the end and beginning of the intervention respectively and $SD_{t_{2}}$ and $SD_{t_{1}}$ to the respective standard deviations.  -->

Where individual studies did not report all data required to calculate the SMD, we imputed missing data according to the methods summarised in this Cochrane Handbook [@higginsChapterChoosingEffect2023].

<!-- If a study reported the standard error of the mean, the SD was obtained simply by multiplying the SE by the square root of the sample size. For conditions where the SD was missing at one time point, the baseline SD was substituted by the post-test SD, and vice versa. If the SD was not available at either time point, missing values were replaced by the mean of the SDs available for comparable cases (defined as same trial type (psy or med), same instrument, same timepoint (pre or post), and same arm (control or active)). Where there were missing means at either baseline or post-test, missing values were calculated using mean change scores, preferring the change scores reported in the paper itself, though where this was unavailable, using the change scores reported in the dataset from Cipriani et al.'s meta-analysis. -->

For meta-analysis it is necessary to estimate a standard error of the SMD. This requires a correlation between the pre- and post-measures, a statistic typically not reported. To ensure that our results are not biased by misestimation, we simulated n=1000 datasets for different values (0.45 to 0.9) of this correlation and used these datasets in subsequent analyses (please refer to Supplement for full details).

<!-- This is calculated according to: -->

<!-- $$ SE_{SMD} = \sqrt{\frac{2(1 -r_{t_{1}t_{2}})}{n} + \frac{SMD^2}{2n}}$$ -->

<!-- where $n$ refers to the study sample size and $r_{t_{1}t_{2}}$ refers to the correlation between the outcome score obtained at baseline and at the end point. This correlation is typically not reported in studies and is often imputed using previously reported correlations for the instruments used. However, this practice has given rise to concerns about misestimation. Whilst such misestimation is possible, there is no reason to expect that it would be systematic, i.e. bias estimation of the effects for the control group of medication compared to those of psychotherapy. Still, to alleviate such concerns we have used a simulations. -->

<!-- In particular, we simulated one thousand truncated distribution of standard errors with the following general characteristics: -->

<!-- $$r_{t_{1}t_{2}} \sim \mathcal{TN}(\mu, \sigma, a, b)$$ -->

<!-- for which we chose the mean to be $\mu = 0.65$, the standard deviation to be $\ sigma = 0.2$, and the upper and lower bounds to be $a = 0.45$ and $b = 0.9$, respectively. We then used these simulated datasets in the subsequent meta-analyses. -->

#### Random-effects Metaregression

We estimated the pooled SMD for each arm by using a random-effects meta-analysis implemented in R's metafor package.

<!-- The main underlying assumption of random-effects meta-analysis is that each study's true effect size $\theta_{k}$ is affected not only by sampling error $\epsilon_{k}$, but also by $\zeta$ which represents heterogeneity between studies, allowing each study's estimate to vary along a distribution of effects, and the distribution of true effect sizes termed $\tau^2$. Therefore, we can estimate a two stage model with: -->

<!-- $$Y_i \sim \mathcal{N}(\theta_i, \sigma_i^2) -->
<!-- $$ -->

<!-- $$\theta_i \sim \mathcal{N}(x_i\beta, \tau_i^2) -->
<!-- $$ -->

<!-- where $Y_i$ is the estimated effect size for study i, has a normal distribution with $\theta_i$ as its true mean effect and sampling error $\sigma^2$. Whereas $\theta_i$ is a study-specific instantiation of the distribution of effect sizes, with $\tau^2$ representing heterogeneity. -->

<!-- This then gives rise to: -->

<!-- $$Y_i = x_i\beta_i + u_i + \epsilon_i$$ -->

<!-- where, -->

<!-- $$u_i \sim N(0, \tau^2) $$ -->

<!-- describes the deviation of each study from the mean of the distribution, and, -->

<!-- $$\epsilon_i \sim N(0, \sigma^2)$$ -->

<!-- describes the sampling error. -->

<!-- We can then specify the following model to obtain the means of each arm of the trials as follows: -->

<!-- $$ \begin{aligned} -->
<!--     Υ_i &= -->
<!--     \begin{cases} -->
<!--         0 & MedControl:  b_0 + u_i + \epsilon_i\\ -->
<!--         1 & MedActive:   b_0 + b_{1_{i}} +  u_i + \epsilon_i \\ -->
<!--         2 & PsyActive:   b_0 + b_{2_{i}}  +  u_i + \epsilon_i \\ -->
<!--         3 & PsyControl:  b_0 + b_{3_{i}}  +  u_i + \epsilon_i \\ -->
<!--     \end{cases} -->
<!-- \end{aligned}$$ -->

<!-- where to obtain the mean of each level is the sum of $b_0$, the intercept for the reference category of medication control, with the coefficient of each level, e.g. for level 3, $b_{3_{i}}$ the psychotherapy controls. The confidence intervals of the means are constructed in the standard way using the standard errors of the mean. Similarly, each coefficient represents the contrast between the reference category and each level, for an example and of main interest to us $b_{3_{i}}$ represents the contrast between psychotherapy and medication control arms. Inference on the contrasts is done as follows: -->

<!-- $$ -->
<!-- z = \frac{\hat{\beta}}{\text{SE}(\hat{\beta})} -->
<!-- $$ -->

<!-- We used maximum likelihood (ML) to estimate model and applied Hartung-Knapp adjustment to reduce the chance of false positives [@inthoutHartungKnappSidikJonkmanMethodRandom2014]. -->

We present the SMDs of each of the four treatment arms (medication control, medication active, psychotherapy control, psychotherapy active) under investigation. The SMDs are the means across the 1000 simulated datasets.

#### Number of sites

We also conducted a t-test to compare mean number of trial sites between psychotherapy and medication trials.

### Sensitivity Analyses

```{r, warning=FALSE, message = FALSE, echo = FALSE, output = FALSE }
# my alpha for the z 
target_probability <- 1 - 0.05

#start here 
x <- 1

# two decimals should be fine
tolerance <- 1e-3

#find the z value using the pnorm function
while (abs(pnorm(x)) < target_probability ) {
  x <- x + tolerance  
}

# Adjust the step size based on your problem  
cat("critical value of z =", x, "corresponding to an alpha = 0.05." , "\n")

x

```
We conducted a series of sensitivity analyses where we excluded studies that 1) used waitlist as their control and 2) recruited participants with subclinical levels of depression. Next, we conducted two analyses where we included only trials that used the Children's Depression Rating Scale, Revised (CDRS-R) or the Hamilton Depression Rating Scale (HAM-D) as outcome instruments.

Further, we tested whether the simulated values for the standard error had a substantial influence on the estimation of the differences between the medication and psychotherapy control conditions. We plotted the z-value of the difference between the two coefficients against the number of simulations. We make inference on the stability of the difference, by counting the proportion of times that the z-value is above the critical value of z = `r x` corresponding to an alpha = 0.05.

Finally, we examined whether differential regression to the mean may account for differences in effect for psychotherapy and medication trials.

### Comparing the control and active arms of psychotherapy trials

We ran t-tests to compare the active and control arms of psychotherapy trials on key variables of interest regarding the intensity of the interventions: the number, duration and intensity of sessions, and the total cumulative hours and duration of the intervention.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#############Some custom made functions#############

###### 1. Simulation Function for SEs
simulate_dataframes_for_st_errors <- function(df, num_repetitions, seed, n, a, b, mean, sd) { # n refers to the number
                                                                                              # of unique ids to which a correlation coef
                                                                                              # is allocated.
  set.seed(seed)

  # Empty list to store simulated dfs
  list_df_simulated <- list()

  for (i in 1:num_repetitions) {
    # Simulate the vector
    simulated_correlations_vector <- truncnorm::rtruncnorm(n, a, b, mean, sd)  #using the truncnorm to create correlations

    # Create a copy of the original dataframe
    df_simulated_copy <- df

    # Add/update the sims column
    df_simulated_copy$correlation_sim_values <- simulated_correlations_vector[match(df_simulated_copy$new_study_id, unique(df_simulated_copy$new_study_id))]

    # Remove duplicated columns
    df_simulated_copy <- df_simulated_copy[, !duplicated(colnames(df_simulated_copy))]

    # Calculate the ses
    df_simulated_copy <- df_simulated_copy %>% 
      group_by(new_study_id, baseline_n) %>% 
      mutate(simulated_se = sqrt(((2*(1-correlation_sim_values))/baseline_n) + 
                                   (cohens_d^2/(2*baseline_n))))

    # Add the simulated dataframe to the list
    list_df_simulated[[i]] <- df_simulated_copy
  }

  return(list_df_simulated)
}

###############2. Simulation function for metaregression
run_metaregression <- function(list_of_datasets, model_formula) {
  list_model_meta_reg <- list()

  for (i in seq_along(list_of_datasets)) {
    list_model_meta_reg[[i]] <- metafor::rma(yi = cohens_d,
                                             sei = simulated_se,
                                             data = list_of_datasets[[i]],
                                             method = "ML",
                                             mods = model_formula,
                                             test = "knha")
  }

  return(list_model_meta_reg)
}

#########2. Function to count studies with missingness etc.

count_studies <- function(df) {
  df_count_studies <- df %>%
    group_by(psy_or_med, arm_effect_size) %>%
    count(is.na(cohens_d)) %>%
    rename(is_missing = `is.na(cohens_d)`) %>%
    mutate(
      condition = case_when(
        psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
        psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
        psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
        psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control"
      )
    )

  df_count_studies$condition <- factor(df_count_studies$condition)  # turn to factor

  # relevel so that medication control becomes the reference category for the regression
  df_count_studies$condition <- relevel(df_count_studies$condition, ref = "medication_control")

  return(df_count_studies)
}

##########3. Function to extract mean coefficients and to extract model characteristics

aggregate_model_results <- function(list_model_1_meta_reg, condition) {
  df_coefs <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  df_se <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  df_z_value <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  df_lower_ci <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  df_upper_ci <- matrix(NA, nrow = length(list_model_1_meta_reg), ncol = length(condition))
  tau_sq <- rep(0, length(list_model_1_meta_reg))
  i_sq <- rep(0, length(list_model_1_meta_reg))
  k <- rep(0, length(list_model_1_meta_reg))
  r_sq <- rep(0, length(list_model_1_meta_reg))

  for (i in 1:length(list_model_1_meta_reg)) {
    for (j in 1:length(condition)) {
      df_coefs[i, j] <- list_model_1_meta_reg[[i]]$beta[[j]]
      df_se[i, j] <- list_model_1_meta_reg[[i]]$se[[j]]
      df_z_value[i, j] <- list_model_1_meta_reg[[i]]$zval[[j]]
      df_lower_ci[i, j] <- list_model_1_meta_reg[[i]]$ci.lb[[j]]
      df_upper_ci[i, j] <- list_model_1_meta_reg[[i]]$ci.ub[[j]]
      tau_sq[i] <- list_model_1_meta_reg[[i]]$tau2
      i_sq[i] <- list_model_1_meta_reg[[i]]$I2
      k[i] <- list_model_1_meta_reg[[i]]$k
      r_sq[i] <- list_model_1_meta_reg[[i]]$R2
    }
  }

  df_coefficients_model <- data.frame(
    coefficients = colMeans(df_coefs),
    se = colMeans(df_se),
    z_value = colMeans(df_z_value),
    lower_ci = colMeans(df_lower_ci),
    upper_ci = colMeans(df_upper_ci),
    tau_sq = mean(tau_sq),
    i_sq = mean(i_sq),
    k = mean(k),
    r_sq = mean(r_sq)
  )

  return(df_coefficients_model)
}

###########4. Function to extract the SMDs  (SMD, ses, CI) for each level of the dummy variable. The output is used below in 5.
# it is a slightly awkward one, because I couldn't' come up with a better way to add the intercept to each coefficient
# given the named output of the built in coef function. here I extract the coefficients

extract_coefficients_func <- function(df_with_coefs){
  list_coefs <- list()
  coefs <- 0
  coefficient_output <- 0

  for(i in 1: length(coefficients(df_with_coefs))){

    temp_vec <- c(0, rep(coefficients(df_with_coefs)[[1]], 
                         length(coefficients(df_with_coefs)) - 1 ))
    coefs[i] <- coef(df_with_coefs)[[i]]

    coefficient_output <- temp_vec + coefs

    st_error_output <- df_with_coefs$se

    df_coefficients <- data.frame(cbind(coefficients = coefficient_output, se = st_error_output ))

    df_coefficients$lower_bound <- df_coefficients$coefficients - 1.96*df_coefficients$se
    df_coefficients$upper_bound <- df_coefficients$coefficients + 1.96*df_coefficients$se
  }
  return (df_coefficients )
}

############5. Function to get the means out of the SMDs (this is similar to Func 3 and I should at some point create one more general one)
# here the argument list_of_dfs is the list of extracted coefficients that you get out of the extract_coefficients_func
# which itself you get from the run_metaregression function. The argument conditions refers to the levels of the variable that describes each study arm

calculate_mean_coefs_ses <- function(list_of_dfs, conditions) {
  df_coefs <- matrix(NA, nrow = length(list_of_dfs), ncol = length(conditions))
  df_se <- matrix(NA, nrow = length(list_of_dfs), ncol = length(conditions))

  for (i in 1:length(list_of_dfs)) {
    for (j in 1:length(conditions)) {
      df_coefs[i, j] <- list_of_dfs[[i]]$coefficients[j]
      df_se[i, j] <- list_of_dfs[[i]]$se[j]
    }
  }

  df_mean_coefs_ses <- data.frame(
    coef_means = colMeans(df_coefs),
    se_means = colMeans(df_se)
  )

  df_mean_coefs_ses$lower_ci <- df_mean_coefs_ses$coef_means - 1.96 * df_mean_coefs_ses$se_means
  df_mean_coefs_ses$upper_ci <- df_mean_coefs_ses$coef_means + 1.96 * df_mean_coefs_ses$se_means

  return(df_mean_coefs_ses)
}

###############6. Function for plotting the SMDs
plot_means_function <- function(dataframe) {
  dataframe$text_label <- 0

  for (i in 1:nrow(dataframe)) {
    dataframe$text_label[i] <- paste(
      dataframe$condition[i], "\n",
      round(dataframe$coef_means[i], 2),
      "[" ,
      round(dataframe$lower_ci[i], 2), 
      round(dataframe$upper_ci[i], 2),
      "]"
    ) 
  }

  # Recode condition for ease of plotting
  dataframe$condition <- str_replace_all(dataframe$condition, "_", " ")

  # Encode colours
  redish_palette <- c("Medication Control" = "deeppink4", "Medication Active" = "deeppink")
  blueish_palette <- c("Psychotherapy Active" = "steelblue1", "Psychotherapy Control" = "steelblue3")

  # Plotting
  plot_means <- ggplot(dataframe, aes(x = coef_means, y = 1:nrow(dataframe),
                                      colour = condition, label = text_label)) +
    geom_point(label=dataframe$k) +
    geom_errorbar(aes(xmin = lower_ci, xmax = upper_ci), width = 0.2, position = position_dodge(0.5)) +
    geom_text(vjust = +1.5, size = 4) +
    scale_size_continuous(guide = "none") +
    guides(colour = FALSE) + 
    scale_color_manual(values = c(redish_palette, blueish_palette)) +
    theme_minimal() +
    xlab("Standardized Mean Difference (SMD) with 95% CIs") +
    ylab("") +
    ylim(0, nrow(dataframe) + 1) +
    xlim(-3.0, 0.5) +
    geom_vline(xintercept = 0, linetype = "dashed", size = 1.5, colour = "grey") +
    geom_segment(x = -1.25, xend = -1.75, y = 4.7, yend = 4.7, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
    geom_text(x = -1.55, y = 4.9, label = "More Effective", color = "grey", vjust = 0.5, hjust = 1) +
    geom_segment(x = -1.35, xend = -0.85, y = 4.9, yend = 4.9, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
    geom_text(x = -1.05, y = 5.05, label = "Less Effective", color = "grey", vjust = 0.5, hjust = 0)+
    theme(axis.title.x = element_text(size = 12, family  = "Calibri"), 
          axis.text.x  = element_text(size = 12, family = "Calibri"),
          axis.text.y = (element_blank()),
          plot.title = element_text(size = 12, family =  "Calibri", face= "italic"),
          plot.subtitle = element_text(size = 12, family =  "Calibri", face= "italic")) +
    geom_curve(aes(x = 0.18, y = 4.73, xend = 0.02, yend = 4.5), color = "grey", curvature = -0.2, arrow = arrow(length = unit(0.25, "cm"), type = "closed")) +
    geom_text(x = 0.2, y = 4.75, label = "No Effect", color = "grey", vjust = 0.5, hjust = 0)

  return(plot_means)
}

###########7. Function for baseline SMD results
library(broom)
run_metareg_and_extract <- function(dataframe, model_1, condition) {
  # Run metareg
  inst_baseline_metareg <- rma(yi = baseline_mean,
                               sei = baseline_stand_error,
                               data = dataframe,
                               method = "ML",
                               mods = model_1,
                               test = "knha")

  # Extract coefficients from the list
  baseline_metareg_results <- tidy(inst_baseline_metareg)
  baseline_metareg_results$condition <- condition
  coefs_for_inst_baseline <- tidy(inst_baseline_metareg)$estimate

  # Calculate SMDs
  smds_for_inst_baseline <- c(coefs_for_inst_baseline[1], coefs_for_inst_baseline[-1] + coefs_for_inst_baseline[1])

  # Calculate upper and lower CIs
  upper_cis_for_inst_baseline <- smds_for_inst_baseline + 1.96 * tidy(inst_baseline_metareg)$std.error
  lower_cis_for_inst_baseline <- smds_for_inst_baseline - 1.96 * tidy(inst_baseline_metareg)$std.error

  # Create dataframe with results
  df_baseline_means_with_cis <- data.frame(
    baseline_smds = smds_for_inst_baseline,
    upper_cis_smds = upper_cis_for_inst_baseline,
    lower_cis_smds = lower_cis_for_inst_baseline,
    condition = condition
  )

  # Return the two dataframes
  return(list(baseline_metareg_results = baseline_metareg_results, df_baseline_means_with_cis = df_baseline_means_with_cis))
}

```

```{r, warning=FALSE, message = FALSE, echo=FALSE}

# Transforming data ------------------------------------------------------
# import master dataset

df_appl_v_orange <-readxl:: read_excel("df_appl_v_orange.xlsx")

# Here we add in instrument end point values (mins / maxs).
# We will use these to transform the scale into % as follows: (observed score - min of the scale)/(max-min)
# We have done this for both means and standard deviations.

df_scale_ranges <- read.csv("scale_ranges.csv")
df_appl_v_orange <- full_join(df_appl_v_orange, df_scale_ranges, by = "instrument_name")

df_appl_v_orange <- df_appl_v_orange %>% 
  mutate(perc_baseline_mean_active = (baseline_mean_active - scale_min) / (scale_max - scale_min)) %>% 
  mutate(perc_baseline_sd_active = baseline_sd_active / (scale_max - scale_min)) %>%  # this is the best way to estimate the SD of the linear transformation of the score 
  mutate(perc_baseline_mean_control = (baseline_mean_control - scale_min) / (scale_max - scale_min)) %>% 
  mutate(perc_baseline_sd_control = baseline_sd_control  / (scale_max - scale_min)) %>% 
  mutate(perc_post_mean_active = (post_mean_active - scale_min) / (scale_max - scale_min)) %>% 
  mutate(perc_post_sd_active = post_sd_active  / (scale_max - scale_min)) %>% 
  mutate(perc_post_mean_control = (post_mean_control - scale_min) / (scale_max - scale_min)) %>% 
  mutate(perc_post_sd_control = post_sd_control  / (scale_max - scale_min))

# Need to use SMDs (ie our Cohen's d) and then use standard error of SMD, to achieve this I need test-retest reliabilities of instruments.

# CDRS reliability taken from here https://www.liebertpub.com/doi/epdf/10.1089/104454601317261546 
# Using a   2-week interval, and different psychiatrists from the first to the second assessment, 
# Poz-nanski et    al. (1984) demonstrated high reliability (r=   0.86) 
# for the CDRS-R total score in 53 clinic-referred 6- to 12-year-olds.

### A few more tidying things from Argyris before doing meta-analyses
# Discovered some errors in the percentage women variable. Have gone back and checked. 
df_appl_v_orange[df_appl_v_orange$study_ID=="Fristad, 2019_cbt + placebo_placebo",]$percent_women <-43.1
df_appl_v_orange[df_appl_v_orange$study=="Clarke, 1995",]$percent_women <-47.3
df_appl_v_orange[df_appl_v_orange$study=="Moeini, 2019",]$percent_women <-100
df_appl_v_orange[df_appl_v_orange$study=="Santomauro, 2016",]$percent_women <-40

# Calculate SE for baseline severity --------------------------------------

df_appl_v_orange$baseline_st_error_active <- 
  df_appl_v_orange$baseline_sd_active/sqrt(df_appl_v_orange$baseline_n_active)

df_appl_v_orange$baseline_st_error_control <-
  df_appl_v_orange$baseline_sd_control/sqrt(df_appl_v_orange$baseline_n_control)

# Turn into a long database with unique rows ------------------------------

### Important: create a dataset that will have unique control studies (see problem that we identified, namely common control conditions)
# create new id to help with better identification and work with duplicates (see below) 
df_appl_v_orange  <- df_appl_v_orange  %>%
  mutate(new_study_id = case_when(psy_or_med == 0 ~ paste(study,year, sep = ", "),
                                  .default = study ))

# We are going to use the descr_active and descr_control columns rather than active_type and control_type columns, as these have unique values for each study. However the descr columns are empty for med studies, so I'm going to paste values across to fix this error.

df_appl_v_orange <- df_appl_v_orange %>% 
  mutate(descr_active = if_else(is.na(descr_active ), active_type , descr_active ))

df_appl_v_orange <- df_appl_v_orange %>% 
  mutate(descr_control = if_else(is.na(descr_control), control_type, descr_control ))

# Step 1: keep only the rows with the top instrument in our hierarchy
df_with_distinct_instruments <-  df_appl_v_orange %>%          
  group_by(new_study_id) %>% 
  filter(instrument_value == min(instrument_value)) # coded for the smallest number to be best.

write.csv(df_with_distinct_instruments, "df_with_distinct_instruments.csv")

# Step 2: turn dataframes to long

# A small bit of cleaning so the code below will work

df_with_distinct_instruments <- df_with_distinct_instruments %>% 
  rename (percent_women_active = active_percent_women, 
          percent_women_control = control_percent_women,
          percent_women_overall = percent_women)

# A: turn long the rows with active
active_df <- df_with_distinct_instruments %>% 
  dplyr:: select(new_study_id, active_type, descr_active, psy_or_med, baseline_n_active, cohens_d_active,
                 baseline_mean_active, baseline_sd_active, post_mean_active, post_sd_active,
                 perc_baseline_mean_active, perc_baseline_sd_active, perc_post_mean_active, perc_post_sd_active, percent_women_overall, percent_women_active, age_m_active, age_sd_active, age_m_overall, age_sd_overall, no_sites, instrument_name) %>% 
  rename(treatment = descr_active) %>% 
  rename(type = active_type) %>% 
  rename_with(~ str_remove(., "_active")) %>% 
  mutate(arm_effect_size = "cohens_d_active") %>% # chose this clunky name because I had used this variable name in previous code and don't want to bread the rest
  group_by(new_study_id) %>% #go through each study id and remove duplicates
  distinct(treatment, .keep_all = TRUE)

control_df <- df_with_distinct_instruments %>% 
  dplyr:: select(new_study_id, control_type, descr_control, psy_or_med, baseline_n_control, cohens_d_control,
                 baseline_mean_control, baseline_sd_control,post_mean_control, post_sd_control,
                 perc_baseline_mean_control, perc_baseline_sd_control, perc_post_mean_control, perc_post_sd_control, percent_women_overall, percent_women_control, age_m_control, age_sd_control, age_m_overall, age_sd_overall, no_sites, instrument_name) %>% 
  rename(treatment = descr_control) %>% 
  rename(type = control_type) %>% 
  rename_with(~ str_remove(., "_control")) %>% 
  mutate(arm_effect_size = "cohens_d_control") %>% #chose this clunky name because I had used this variable name in previous code and don't want to bread the rest
  group_by(new_study_id) %>% #go through each study id and remove duplicates
  distinct(treatment, .keep_all = TRUE)

df_long_for_metan <- rbind(active_df, control_df)

# I'm going to remove the March / TADS psy control row. It is currently filled with NAs which implies missing data, however it is actually the case that this condition doesn't exist as such (the control is placebo, which is not a psy control).

df_long_for_metan <- df_long_for_metan %>%
  filter(!(new_study_id == "March, 2004" & is.na(treatment)))

# I'm going to perform  the operations to calculate overall age and percent women here now. We just want this per study rather than per arm for the first set of meta-analyses as we are going to compare psy vs med (2 subgroups) rather than the 4 arms

df_long_for_metan <- df_long_for_metan %>% 
  group_by(new_study_id) %>%
  mutate(n_overall = sum(baseline_n)) %>% 
  mutate(age_overall = sum(age_m * baseline_n) / sum(baseline_n)) %>% 
  mutate(mean_age = coalesce(age_overall, age_m_overall)) %>% 
  mutate(pooled_sd_age = sqrt(sum((age_sd^2) * (baseline_n - 1)) / sum(baseline_n - 1))) %>% 
  mutate(se_age = pooled_sd_age / sqrt(n_overall)) %>% 
  mutate(sex_overall = sum(percent_women * baseline_n) / sum(baseline_n)) %>% 
  mutate(percent_women = coalesce(percent_women_overall, sex_overall))

# Now just get rid of the unnecessary columns as these are causing clutter

df_long_for_metan <- df_long_for_metan %>% 
  select(-c(percent_women_overall, age_m, age_sd, age_m_overall, age_sd_overall, age_overall, sex_overall))

# Remember that these are now per study, not per arm.

# Create SEs proportions for percentage women --------------------------------------------

# We also need to calculate SE for proportion women for the baseline calculations
# for proportions, this is calculated as sqrt(p(1-p)/n), which I implement stepwise below

df_long_for_metan <- df_long_for_metan %>% 
  mutate(product_perc_women = (percent_women/100)* (1-(percent_women/100)) )

df_long_for_metan <- df_long_for_metan %>% 
  mutate(percent_women_std_error = sqrt(product_perc_women/ n_overall ))

#dim(df_long_for_metan ) #check dimension
#make sure no study lost
#length(unique(df_long_for_metan$new_study_id ) ) == # #length(unique(df_appl_v_orange$new_study_id ) )

# Now check again the studies with muliple arms
# Now checking if this worked with the studies that I used to illustrate the problem above. 
# df_long_for_metan[df_long_for_metan$new_study_id == "Stallard, 2012",]
# df_long_for_metan[df_long_for_metan$new_study_id == "Atkinson, 2014",]
# # also check one which is single to make sure it is kept
# df_long_for_metan[df_long_for_metan$new_study_id == "Ackerson, 1998",]

# For response rate, recalculating after turning long and doing all the imputations

# df_long_for_metan <- df_long_for_metan %>%
#   rowwise() %>%
#   mutate(responders = sum(rnorm(baseline_n, post_mean, 
#                                 post_sd) < (baseline_mean/ 2)))
# 
# df_long_for_metan <- df_long_for_metan %>%
#   rowwise() %>%
#   mutate(resp_rate = responders/baseline_n)
# 
# summary_resp_rate <- df_long_for_metan %>%
#   group_by(group) %>%
#   summarise(mean_resp_rate = mean(resp_rate, na.rm = TRUE),
#             sd_resp_rate = sd(resp_rate, na.rm = TRUE), 
#             n = n())

# save this as the master dataframe  --------------------------------------
# use this dataframe to perform all operations below

write.csv(df_long_for_metan, "df_long_for_metan.csv", row.names = F)

########### simulate so that each study has a value from a distribution of 
######## of correlations.

# I will generate random numbers per study id from a distribution with these parameters. 
# it is reasonable to generate one random correlation value per study as there is no reason why the correlation should
# systematically vary within studies

# Use the function I have created for this

df_long_for_metan <- read.csv("df_long_for_metan.csv")

list_df_simulated <-  simulate_dataframes_for_st_errors (df = df_long_for_metan, 
                                              num_repetitions = 1000, 
                                              seed  = 1974,
                                              n = length(unique(df_long_for_metan$new_study_id)), 
                                              a = 0.45, 
                                              b = .9, 
                                              mean = 0.65, 
                                              sd = 0.2)

#check this worked

#list_df_simulated[[10]][,c("new_study_id", "correlation_sim_values", "simulated_se")] # looks right
#summary(list_df_simulated[[1000]]$correlation_sim_values, na.rm = T) # as expected # the mean is around the parameter I gave

```

# Findings

```{r, warning=FALSE, message = FALSE, echo=FALSE}

# This chunk of code needs to be here because some of the data needed for the included studies summary lives here

# Get Ns, run metareg, prepare graphing ------------------------------------------------------

# add a new multilevel variable to the simulated data for the regression

for(i in 1: length(list_df_simulated)){

list_df_simulated[[i]]$arm_effect_size <- factor(list_df_simulated[[i]]$arm_effect_size)
  list_df_simulated[[i]] <- list_df_simulated[[i]] %>% 
    mutate(four_level_var = case_when(psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
                                      psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control")
    )

  list_df_simulated[[i]]$four_level_var <- factor(list_df_simulated[[i]]$four_level_var) # turn to factor

  # relevel so that medication control becomes the reference category for the regression
#  list_df_simulated[[i]]$four_level_var <- relevel(list_df_simulated[[1]]$four_level_var, ref = "medication_control")
list_df_simulated[[i]]$four_level_var <-   fct_relevel(list_df_simulated[[i]]$four_level_var,
                                                       "medication_control",
                                                       "medication_active", 
                                                       "psychotherapy_control",
                                                       "psychotherapy_active")
}

# check it worked
# list_df_simulated[[2]] %>% 
# dplyr:: select(psy_or_med, arm_effect_size, four_level_var)

######## NOW RUN METAREGRESSIONS AND GET Coefficients for 
# A) the overall sample
# B) the CBT vs Fluox and Escitalopram sample
# C) the no WL sample
# E) the CDRS only sample (**Argyris** do we still want this?)
# F) Clinical Levels Depression

# A. Estimate Means and CIs for the overall ----------------------------------

# count studies
# count the number of studies
n_unique_studies <- length(unique(df_long_for_metan$new_study_id)) # CHARLOTTE please check
df_count_studies <- count_studies(df_long_for_metan ) # CHARLOTTE please check
# CB checked both, both correct.

# specify model
model_1 <- as.formula(~ four_level_var)

# run metareg function 
# here you can use the simulated results directly
list_model_1_meta_reg <- run_metaregression(list_df_simulated, model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_overall <-  aggregate_model_results(list_model_1_meta_reg, condition)
aggregate_results_overall <- cbind(aggregate_results_overall, condition)

# extract SMDs from the  list
list_dummy_var_means <-  lapply(list_model_1_meta_reg, extract_coefficients_func)

# calculate mean SMDs and ses
df_mean_coefs_from_sim <- calculate_mean_coefs_ses(list_dummy_var_means, condition)

df_mean_coefs_from_sim <- data.frame(cbind(condition,df_mean_coefs_from_sim))

# add the ns
df_mean_coefs_from_sim <- left_join(df_mean_coefs_from_sim, df_count_studies[df_count_studies$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

# B. Estimate Means and CIs for CBT Fluox Esc --------------------------------
#  grab ids for those studies
cbt_fluox_esc_study_ids <-
  df_appl_v_orange[df_appl_v_orange$active_type == "cbt" | df_appl_v_orange$active_type == "Fluoxetine" | df_appl_v_orange$active_type == "Escitalopram",]$new_study_id

# Use those ids to subset all dataframes in the list
# By searching for IDs we have mistakenly included multiple active arms. E.g. for Atkinson which has dulox and fluox. 
# So we need to further subset the simulated dfs. For the active arms, I am filtering out conditions that do not meet our criteria. 
# This leaves all the control conditions from trials where the active is cbt, fluox or esc. 
# We may have multiple controls per study id for this reason, though this isn't a problem 
# conceptually.

list_cbt_fluox_esc_study <- lapply(list_df_simulated, function(df) {
  df %>%
    filter(new_study_id %in% cbt_fluox_esc_study_ids) %>%
    filter(!(arm_effect_size == "cohens_d_active" & !(type %in% c("Fluoxetine", "Escitalopram", "cbt"))))
})

# count the number of studies
cbt_fluox_esc_study <- df_long_for_metan %>%
  filter(new_study_id %in% cbt_fluox_esc_study_ids,
         !(arm_effect_size == "cohens_d_active" & !(type %in% c("Fluoxetine", "Escitalopram", "cbt")))) 
study_count_cbt_fluox_esc_study <- count_studies(cbt_fluox_esc_study)
# CB checked, looks good now.

# apply the metareg function to the list of cbt, fluox, esc
model_1 <- as.formula(~ four_level_var)
list_model_1_cbt_fluox_esc_study <- run_metaregression(list_cbt_fluox_esc_study , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_cbt_fluox_esc_study <-  aggregate_model_results(list_model_1_cbt_fluox_esc_study, condition)
aggregate_results_cbt_fluox_esc_study <- cbind(aggregate_results_cbt_fluox_esc_study, condition)

# extract SMDs from the list for cbt fluox escit
means_cbt_fluox_esc_study <-  lapply(list_model_1_cbt_fluox_esc_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_cbt_fluox_esc_study <- calculate_mean_coefs_ses(means_cbt_fluox_esc_study, condition)
 coef_and_se_means_cbt_fluox_esc_study <- data.frame(cbind(condition, coef_and_se_means_cbt_fluox_esc_study))

# add the ns to the dataframe
coef_and_se_means_cbt_fluox_esc_study <- left_join(coef_and_se_means_cbt_fluox_esc_study, study_count_cbt_fluox_esc_study[study_count_cbt_fluox_esc_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

# C. Estimate Means and CIs without WL ---------------------------------------

 # grab the ids for no waitlist
no_wl_ids <-df_appl_v_orange[df_appl_v_orange$control_type != "wl",]$new_study_id

# use them to loop over the simulated list to subset all the dataframes.

list_no_wl <- lapply(list_df_simulated, function(df) {
  df %>%
    filter(new_study_id %in% no_wl_ids) %>%
    filter(!(arm_effect_size == "cohens_d_control" & type == "wl")) 
})

# count studies 
no_wl_studies <- df_long_for_metan %>% 
  filter(new_study_id %in% no_wl_ids) %>%
  filter(!(arm_effect_size == "cohens_d_control" & type == "wl"))
study_count_no_wl_study <- count_studies(no_wl_studies)
# CB checked, looks good!

# use this model as above
model_1 <- as.formula(~ four_level_var)

# apply the metareg function to the list of no wl
list_model_1_no_wl <- run_metaregression(list_no_wl , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_no_wl <-  aggregate_model_results(list_model_1_no_wl, condition)
aggregate_results_no_wl <- cbind(aggregate_results_no_wl, condition)

# now use the function to extract the coefficients
means_no_wl <-  lapply(list_model_1_no_wl, extract_coefficients_func)

# run the mean function for no wl
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_no_wl <- calculate_mean_coefs_ses(means_no_wl, condition)
coef_and_se_means_no_wl <- data.frame(cbind(condition,coef_and_se_means_no_wl))

coef_and_se_means_no_wl <- left_join(coef_and_se_means_no_wl, study_count_no_wl_study[study_count_no_wl_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

# D Clinical Levels Depression Sensitivity Analysis ---------------------------------------

# We need to decide how we want to do this. We have mdd, mood, sub, and cut categories. For now I'll just exclude sub.

#take study ids where sample has clinical levels of depression
clin_study_ids <-
  df_appl_v_orange[df_appl_v_orange$diagnosis != "sub",]$new_study_id

# use those ids to subset all dataframes in the list
list_clin_study <- list()

for(i in 1: length(list_df_simulated)){

  list_clin_study[[i]] <- list_df_simulated[[i]][list_df_simulated[[i]]$new_study_id
                                                          %in% clin_study_ids,]
}

# count the number of studies
study_count_clin_study <- count_studies(df_long_for_metan[df_long_for_metan$new_study_id %in% clin_study_ids,])
# CB checked, correct.

# apply the metareg function to the list of clinical studies
model_1 <- as.formula(~ four_level_var)
list_model_1_clin_study <- run_metaregression(list_clin_study , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_clin_study <-  aggregate_model_results(list_model_1_clin_study, condition)

# extract SMDs from the list for clinical studies
means_clin_study <-  lapply(list_model_1_clin_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_clin_study <- calculate_mean_coefs_ses(means_clin_study, condition)
coef_and_se_means_clin_study <- data.frame(cbind(condition, coef_and_se_means_clin_study))

# add the ns to the dataframe
coef_and_se_means_clin_study <- left_join(coef_and_se_means_clin_study, study_count_clin_study[study_count_clin_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

# E. CDRS Sensitivity Analysis ---------------------------------------

cdrs_study_ids <-
  df_appl_v_orange[df_appl_v_orange$instrument_name == "cdrs",]$new_study_id

# use those ids to subset all dataframes in the list
list_cdrs_study <- list()

for(i in 1: length(list_df_simulated)){

  list_cdrs_study[[i]] <- list_df_simulated[[i]][list_df_simulated[[i]]$new_study_id
                                                          %in% cdrs_study_ids,]
}

# count the number of studies
study_count_cdrs_study <- count_studies(df_long_for_metan[df_long_for_metan$new_study_id %in% cdrs_study_ids,])
# CB checked, correct.

# apply the metareg function to the list of cdrs studies
model_1 <- as.formula(~ four_level_var)
list_model_1_cdrs_study <- run_metaregression(list_cdrs_study , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_cdrs_study <-  aggregate_model_results(list_model_1_cdrs_study, condition)

# extract SMDs from the list for cdrs
means_cdrs_study <-  lapply(list_model_1_cdrs_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_cdrs_study <- calculate_mean_coefs_ses(means_cdrs_study, condition)
coef_and_se_means_cdrs_study <- data.frame(cbind(condition, coef_and_se_means_cdrs_study))

# add the ns to the dataframe
coef_and_se_means_cdrs_study <- left_join(coef_and_se_means_cdrs_study, study_count_cdrs_study[study_count_cdrs_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

# E. HAMD Sensitivity Analysis ---------------------------------------

hamd_study_ids <-
  df_appl_v_orange[df_appl_v_orange$instrument_name == "hamd",]$new_study_id

# use those ids to subset all dataframes in the list
list_hamd_study <- list()

for(i in 1: length(list_df_simulated)){

  list_hamd_study[[i]] <- list_df_simulated[[i]][list_df_simulated[[i]]$new_study_id
                                                          %in% hamd_study_ids,]
}

# count the number of studies
study_count_hamd_study <- count_studies(df_long_for_metan[df_long_for_metan$new_study_id %in% hamd_study_ids,])
# CB checked, correct.

# apply the metareg function to the list of cdrs studies
model_1 <- as.formula(~ four_level_var)
list_model_1_hamd_study <- run_metaregression(list_hamd_study , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_hamd_study <-  aggregate_model_results(list_model_1_hamd_study, condition)

# extract SMDs from the list for cdrs
means_hamd_study <-  lapply(list_model_1_hamd_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_hamd_study <- calculate_mean_coefs_ses(means_hamd_study, condition)
coef_and_se_means_hamd_study <- data.frame(cbind(condition, coef_and_se_means_hamd_study))

# add the ns to the dataframe
coef_and_se_means_hamd_study <- left_join(coef_and_se_means_hamd_study, study_count_hamd_study[study_count_hamd_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

```

## Included studies

Data for included are summarised in @tbl-all-trials and are also available as a csv dataframe on \[<https://github.com/transatlantic-comppsych/apples_oranges>\]. Please see @fig-prisma for a summary of the sources of included RCTs.

In total, there were `r length(unique(df_long_for_metan$new_study_id))` RCTs which included `r sum(df_count_studies[df_count_studies$condition == "medication_active", "n"])` active arms and `r sum(df_count_studies[df_count_studies$condition == "medication_control", "n"])` control arms of medication trials; and `r sum(df_count_studies[df_count_studies$condition == "psychotherapy_active", "n"])` active arms and `r sum(df_count_studies[df_count_studies$condition == "psychotherapy_control", "n"])` control arms from psychotherapy RCTs. Note that the number of active and control arms does not exactly match because some studies feature more than one control or active arm.

Placebo pill was the control condition for all medication trials. In psychotherapy trials, the control arm included 14 waitlist, 25 treatment-as-usual (TAU), and 19 other control conditions (e.g. relaxation training, attention control, counselling).

```{r, echo=FALSE}
#| label: fig-prisma
#| fig-cap: "PRISMA chart summarising sources of included studies"
#| fig-cap-location: top
knitr::include_graphics("PRISMA.png")
```

## Sample characteristics at baseline in medication and psychotherapy trials

@tbl-baseline_results summarises the results from each of the meta-analyses examining sample characteristics at baseline. The summary statistics are provided for each subgroup (i.e. medication and psychotherapy) and the p-value derives from the test for subgroup differences.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# Start by looking at demographics at baseline
# Currently we have both control and active conditions, we just need each study once

df_baseline_demographics <- df_long_for_metan %>% 
  distinct(new_study_id, .keep_all = TRUE)

# This is the function to extract statistics

extract_stats <- function(df) {

  # get the summary table with results
  statistics <- summary(df, byvar = "psy_or_med")

  df_stats <- tibble(
  subgroup = statistics$subgroup.levels,
  k_study = statistics$k.study.w,
  TE_random = statistics$TE.random.w,
  seTE_random = statistics$seTE.random.w,
  lower_random = statistics$lower.random.w,
  upper_random = statistics$upper.random.w,
  tau2 = statistics$tau2.w)

  # Add an empty row at the top and column on right (this is for the test for subgroup differences p value)
  df_stats <- rbind(data.frame(subgroup = NA, k_study = NA, TE_random = NA, seTE_random = NA, 
                             lower_random = NA, upper_random = NA, tau2 = NA), df_stats)
  df_stats <- cbind(df_stats, pval_between = NA)

  # Set the value in the last row and last column as the between groups p val
  df_stats[1, "pval_between"] <- statistics$pval.Q.b.random

  return(df_stats)

}

# Now look at baseline severity for overall psy vs med
# Need to calculate the variables first

df_for_baseline_severity <- df_long_for_metan %>% 
  group_by(new_study_id) %>%
  mutate(perc_baseline_overall = sum(perc_baseline_mean * baseline_n) / sum(baseline_n)) %>% 
  mutate(pooled_perc_sd_baseline = sqrt(sum((perc_baseline_sd^2) * (baseline_n - 1)) / sum(baseline_n - 1))) %>% 
  mutate(perc_se_baseline = pooled_perc_sd_baseline / sqrt(n_overall)) %>%  # now also do this for untransformed variable
  mutate(baseline_overall = sum(baseline_mean * baseline_n) / sum(baseline_n)) %>% 
  mutate(pooled_sd_baseline = sqrt(sum((baseline_sd^2) * (baseline_n - 1)) / sum(baseline_n - 1))) %>% 
  mutate(se_baseline = pooled_sd_baseline / sqrt(n_overall))

df_for_baseline_severity <- df_for_baseline_severity %>% 
  distinct(new_study_id, .keep_all = TRUE)

df_for_baseline_severity <- df_for_baseline_severity %>% 
    filter(new_study_id != "March, 2004")  # removing TADS because it does not technically belong to either the psy or med groups

met_sev <- metagen(TE = perc_baseline_overall,
                           seTE = perc_se_baseline,
                           studlab = new_study_id,
                           data = df_for_baseline_severity,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "baseline severity",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_sev <- extract_stats(met_sev)

# clinical sample

clin_study_ids <-
  df_appl_v_orange[df_appl_v_orange$diagnosis != "sub",]$new_study_id

df_for_baseline_severity_clin <- df_for_baseline_severity %>%
  filter(new_study_id %in% clin_study_ids)

met_sev_clin <- metagen(TE = perc_baseline_overall,
                           seTE = perc_se_baseline,
                           studlab = new_study_id,
                           data = df_for_baseline_severity_clin,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "baseline severity clinical",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_sev_clin <- extract_stats(met_sev_clin)

# cbt + med

# cbt_meds_study_ids <-
#   df_appl_v_orange[df_appl_v_orange$active_type == "cbt" | df_appl_v_orange$active_type == "Fluoxetine" | df_appl_v_orange$active_type == "Escitalopram",]$new_study_id
# 
# df_for_baseline_severity_cbt <- df_for_baseline_severity %>%
#   filter(new_study_id %in% cbt_meds_study_ids) 
# 
# met_sev_cbt <- metagen(TE = perc_baseline_overall,
#                            seTE = perc_se_baseline,
#                            studlab = new_study_id,
#                            data = df_for_baseline_severity_cbt,
#                            sm = "",
#                            fixed = FALSE,
#                            random = TRUE,
#                            method.tau = "REML",
#                            hakn = TRUE,
#                            title = "baseline severity cbt + med studies",
#                            subgroup = psy_or_med,
#                            tau.common = FALSE)
# 
# results_met_sev_cbt <- extract_stats(met_sev_cbt)

# no wl

no_wl_ids <- df_appl_v_orange[df_appl_v_orange$control_type != "wl",]$new_study_id

df_for_baseline_severity_no_wl <- df_for_baseline_severity %>%
  filter(new_study_id %in% no_wl_ids)

met_sev_no_wl <- metagen(TE = perc_baseline_overall,
                           seTE = perc_se_baseline,
                           studlab = new_study_id,
                           data = df_for_baseline_severity_no_wl,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "baseline severity no wl studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_sev_no_wl <- extract_stats(met_sev_no_wl)

# Now start looking at perc_women variable 
# The metagen function doesn't like when NAs result in the TE vector being a different length to subgroup vector. Need to start by removing NAs.

df_for_sex_metan <- df_baseline_demographics %>% 
  filter(!is.na(percent_women))

df_for_sex_metan <- df_for_sex_metan %>% 
  filter(percent_women != 100) # filter out studies with only women %>%

df_for_sex_metan <- df_for_sex_metan %>% 
  filter(new_study_id != "March, 2004") # removing TADS because it does not technically belong to either the psy or med groups

met_perc_women_overall <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_sex_metan,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across studies")

met_perc_women_overall <- update(met_perc_women_overall,
                              subgroup = df_for_sex_metan$psy_or_med, subgroup.name = "treatment modality",
                              tau.common = FALSE)

# Apply the function to the meta-analysis result

results_met_perc_women_overall <- extract_stats(met_perc_women_overall)

# looking at clinical sample only

df_for_sex_metan_clin <- df_for_sex_metan %>%
  filter(new_study_id %in% clin_study_ids)

met_perc_women_clin <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_sex_metan_clin,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across clinical studies")

met_perc_women_clin <- update(met_perc_women_clin,
                           subgroup = df_for_sex_metan_clin$psy_or_med, subgroup.name = "treatment modality",
                           tau.common = FALSE)

# Apply the function to the meta-analysis result
results_met_perc_women_clin <- extract_stats(met_perc_women_clin)

# # Now looking at the CBT + med studies only
# 
# df_for_sex_metan_cbt <- df_for_sex_metan %>%
#   filter(new_study_id %in% cbt_meds_study_ids) 
# 
# met_perc_women_cbt <- metagen(TE = percent_women,
#                            seTE = percent_women_std_error,
#                            studlab = new_study_id,
#                            data = df_for_sex_metan_cbt,
#                            sm = "",
#                            fixed = FALSE,
#                            random = TRUE,
#                            method.tau = "REML",
#                            hakn = TRUE,
#                            title = "percentage women across cbt and med studies",
#                            subgroup = psy_or_med,
#                            tau.common = FALSE)
# 
# # Apply the function to the meta-analysis result
# results_met_perc_women_cbt <- extract_stats(met_perc_women_cbt)

# Now the no waitlist group

df_for_sex_metan_no_wl <- df_for_sex_metan %>%
  filter(new_study_id %in% no_wl_ids)

met_perc_women_no_wl <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_sex_metan_no_wl,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across studies with controls that are not wl",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

# Apply the function to the meta-analysis result
results_met_perc_women_no_wl <- extract_stats(met_perc_women_no_wl)

# Now start looking at age variable

# We have several studies for which we are missing SE for age. I am going to perform an imputation by taking an average
# for the studies we do have, per modality, and substituting this in for missing values.

df_baseline_demographics$se_age <- as.numeric(df_baseline_demographics$se_age)

mean_se_age_psy <- df_baseline_demographics %>%
  filter(psy_or_med == 1) %>%
  summarise(mean_se_age_psy = mean(se_age, na.rm = TRUE))

mean_se_age_med <- df_baseline_demographics %>%
  filter(psy_or_med == 0) %>%
  summarise(mean_se_age_med = mean(se_age, na.rm = TRUE))

df_baseline_demographics <- df_baseline_demographics %>% 
  mutate(se_age = ifelse(!is.na(se_age), se_age,
                         ifelse(psy_or_med == 1, mean_se_age_psy, mean_se_age_med)))

df_for_age_metan <- df_baseline_demographics %>% 
  filter(!is.na(mean_age))

df_for_age_metan <- df_for_age_metan %>% 
  filter(new_study_id != "March, 2004") # removing TADS because it does not technically belong to either the psy or med groups

# Age meta analysis

df_for_age_metan$se_age <- as.numeric(df_for_age_metan$se_age)

met_age <- metagen(TE = mean_age,
                           seTE = se_age,
                           studlab = new_study_id,
                           data = df_for_age_metan,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "age across studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_age <- extract_stats(met_age)

# Looking at clinical sample only

clin_study_ids <-
  df_appl_v_orange[df_appl_v_orange$diagnosis != "sub",]$new_study_id

df_for_age_metan_clin <- df_for_age_metan %>%
  filter(new_study_id %in% clin_study_ids)

met_age_clin <- metagen(TE = mean_age,
                           seTE = se_age,
                           studlab = new_study_id,
                           data = df_for_age_metan_clin,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "age across clinical studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_age_clin <- extract_stats(met_age_clin)

# # Now cbt + med studies only
# 
# df_for_age_metan_cbt <- df_for_age_metan %>%
#   filter(new_study_id %in% cbt_meds_study_ids) 
# 
# met_age_cbt <- metagen(TE = mean_age,
#                            seTE = se_age,
#                            studlab = new_study_id,
#                            data = df_for_age_metan_cbt,
#                            sm = "",
#                            fixed = FALSE,
#                            random = TRUE,
#                            method.tau = "REML",
#                            hakn = TRUE,
#                            title = "age across cbt + med studies",
#                            subgroup = psy_or_med,
#                            tau.common = FALSE)
# 
# results_met_age_cbt <- extract_stats(met_age_cbt)

# Now excluding wl

df_for_age_metan_no_wl <- df_for_age_metan %>%
  filter(new_study_id %in% no_wl_ids)

met_age_no_wl <- metagen(TE = mean_age,
                           seTE = se_age,
                           studlab = new_study_id,
                           data = df_for_age_metan_no_wl,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "age across no wl studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_age_no_wl <- extract_stats(met_age_no_wl)

# This is if we want to have all sensitivity analyses in table one

df_overall_baseline_results <- rbind(results_met_sev, results_met_sev_clin, results_met_sev_no_wl, results_met_perc_women_overall, results_met_perc_women_clin, results_met_perc_women_no_wl, results_met_age, results_met_age_clin, results_met_age_no_wl)

# For row names
df_overall_baseline_results <- df_overall_baseline_results %>%
  mutate(subgroup = case_when(
    row_number() %in% c(1, 10, 19) ~ "Overall",
    row_number() %in% c(4, 13, 22) ~ "Excluding subclinical",
    row_number() %in% c(7, 16, 25) ~ "Excluding waitlist",
    TRUE ~ if_else(subgroup == 0, "Medication", "Psychotherapy")
  ))

# Define new column names
new_column_names <- c("Subgroup", "K", "Mean", "SE", "Lower CI", "Upper CI", "T2", "p-value")

# Assign new column names to the data frame
colnames(df_overall_baseline_results) <- new_column_names

# Rounding
df_overall_baseline_results <- df_overall_baseline_results %>%
  mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
  mutate_at(vars(ncol(.)), ~ round(., 3)) %>%
  mutate_all(~if_else(is.na(.), "", as.character(.)))

big_border = fp_border(color="black", width = 1)
small_border = fp_border(color="lightgray", width = 1)

```

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
#| label: tbl-baseline_results
#| tbl-cap: "Sample characteristics at baseline across medication and psychotherapy studies: Results for overall sample and sensitivity analyses"
#| tbl-cap-location: top

df_overall_baseline_results %>% 
  mutate(category = case_when(
    row_number() %in% 1:9 ~ "Baseline Severity of Depressive Symptoms*",
    row_number() %in% 10:18 ~ "Percent Female",
    row_number() %in% 19:27 ~ "Age",
    TRUE ~ ""
  )) %>% # grouping rows
  group_by(category) %>%
  as_flextable(hide_grouplabel = TRUE) %>% 
  paginate(group = "category") %>% 
  bold(i = c(1, 2, 5, 8, 11, 12, 15, 18, 21, 22, 25, 28), j = NULL, bold = TRUE, part = "body") %>%  # Bold specific rows
  # set_table_properties(layout = "fixed") %>%  # Autofit column widths
  bg(i = c(1, 11, 21), bg  = "#F2F2F2", part = "body") %>% # identify category sections
  padding(i = c(3, 4, 6, 7, 9, 10, 13, 14, 16, 17, 19, 20, 23, 24, 26, 27, 29, 30), j = 1, padding.left = 10) %>% # indent some rows
  align(j = 2:ncol(df_overall_baseline_results), align = "center", part = "all") %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row  
  # italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "all" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  # padding(i = 2, padding.bottom = 15, part = "header") %>% 
  line_spacing(space = .7, part = "all") %>% 
  add_footer_row(values = as_paragraph("*These are baseline depression scores transformed to reflect percentage of a scale range (see Supplement for detailed description).To take an example, the CDRS gives a possible total score from 17 to 113 (i.e. range of 96). Mean severity was 0.36 for psychotherapy studies and 0.42 for medication studies, which would translate to 51.56 (17 + 0.36 x 96) and 57.32 (17 + 0.42 x 96), respectively, as equivalent scores on the CDRS."), colwidths = 8) %>%
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  autofit()

```
### Baseline severity

On average, depression severity at baseline was significantly higher in medication trials compared to psychotherapy trials (see @tbl-baseline_results). When excluding RCTs that used waitlist as their control, baseline severity remained significantly higher in medication trials compared to psychotherapy trials. This difference did not reach statistical significance when excluding studies that recruited samples with sub-clinical depression.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# looking at CDRS studies only, this is with only 2 levels

df_for_baseline_severity_cdrs <- df_for_baseline_severity %>%
  filter(new_study_id %in% cdrs_study_ids)

met_sev_cdrs <- metagen(TE = baseline_overall,
                           seTE = se_baseline,
                           studlab = new_study_id,
                           data = df_for_baseline_severity_cdrs,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "baseline severity cdrs",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_sev_cdrs <- extract_stats(met_sev_cdrs)

# looking at HAMD studies only,

df_for_baseline_severity_hamd <- df_for_baseline_severity %>%
  filter(new_study_id %in% hamd_study_ids)

met_sev_hamd <- metagen(TE = baseline_overall,
                           seTE = se_baseline,
                           studlab = new_study_id,
                           data = df_for_baseline_severity_hamd,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "baseline severity cdrs",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

results_met_sev_hamd <- extract_stats(met_sev_hamd)
```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
## this is for four levels
### get CDRS studies
cdrs_studies_for_baseline_metareg <- list_cdrs_study[[1]] # created this above, you can take any of these, as the means and sds are repeated.

#turn sds to sec
cdrs_studies_for_baseline_metareg$baseline_stand_error <- cdrs_studies_for_baseline_metareg$baseline_sd/
                                        sqrt(cdrs_studies_for_baseline_metareg$baseline_n)

model_1 <- as.formula(~ four_level_var)
condition <-  levels(list_df_simulated[[1]]$four_level_var) # this is the four level var
                                                              #throughout

results_cdrs_metareg_baseline <- run_metareg_and_extract(cdrs_studies_for_baseline_metareg,
                                                         model_1, condition)

results_cdrs_metareg_baseline[[2]] <- left_join(results_cdrs_metareg_baseline[[2]], study_count_cdrs_study[study_count_cdrs_study$is_missing == FALSE, ] %>%
                                        ungroup() %>%
                                      select(condition, n),
                                    by = "condition")

### get HAM-D studies
hamd_study_ids <-
  df_appl_v_orange[df_appl_v_orange$instrument_name == "hamd",]$new_study_id

list_hamd_study <- list_df_simulated[[1]][list_df_simulated[[1]]$new_study_id
                                                          %in% hamd_study_ids,]

hamd_studies_for_baseline_metareg <- list_hamd_study #

#turn sds to sec
hamd_studies_for_baseline_metareg$baseline_stand_error <- hamd_studies_for_baseline_metareg$baseline_sd/
                                        sqrt(hamd_studies_for_baseline_metareg$baseline_n)

model_1 <- as.formula(~ four_level_var)
condition <-  levels(list_df_simulated[[1]]$four_level_var) # this is the four level var
                                                              #throughout
results_hamd_metareg_baseline <- run_metareg_and_extract(hamd_studies_for_baseline_metareg,
                                                         model_1, condition)

results_hamd_metareg_baseline$condition <-  levels(list_df_simulated[[1]]$four_level_var)

results_hamd_metareg_baseline[[2]] <- left_join(results_hamd_metareg_baseline[[2]], study_count_hamd_study[study_count_hamd_study$is_missing == FALSE, ] %>%
                                        ungroup() %>%
                                      select(condition, n),
                                    by = "condition")

```

To ensure that this was not an artefact of variable transformation, we also compared means at baseline in the two instruments, CDRS and HAM-D, on which there was a sufficient number of studies to metanalyse. As can be seen in @tbl-hamd-baseline and @tbl-cdrs-baseline (see Supplemental Materials), the number of studies is much smaller, but the pattern of differences is the same for the HAM-D and the CDRS, though it does not reach statistical significance for the latter.

### Sex

For this analysis, we excluded the two psychotherapy trials which included entirely female samples (Moeini, 2019; Shomaker, 2016). As can be seen in @tbl-baseline_results, psychotherapy trials featured a significantly higher percentage of females when compared to medication trials. On average, samples were `r round(results_met_perc_women_overall$TE_random[results_met_perc_women_overall$subgroup == 1][[2]],2)`% (*SE* = `r round(results_met_perc_women_overall$seTE_random[results_met_perc_women_overall$subgroup == 1][[2]],2)`) female across psychotherapy trials and `r round(results_met_perc_women_overall$TE_random[results_met_perc_women_overall$subgroup == 0][[2]],2)`% (*SE* = `r round(results_met_perc_women_overall$seTE_random[results_met_perc_women_overall$subgroup == 0][[2]],2)`) female across medication trials. Excluding sub-clinical and waitlist control studies yielded similar results.

### Age

As can be seen in @tbl-baseline_results, mean age was `r round(results_met_age$TE_random[results_met_age$subgroup == 1][[2]],2)` (*SE* = `r round(results_met_age$seTE_random[results_met_age$subgroup == 1][[2]],2)`) across psychotherapy trials and `r round(results_met_age$TE_random[results_met_age$subgroup == 0][[2]],2)` (*SE* = `r round(results_met_age$seTE_random[results_met_age$subgroup == 0][[2]],2)`) across medication trials, with no significant between group differences. There were no significant differences in mean age between modalities on further sensitivity analyses.

## Trial design

### Standardised mean differences of control conditions in psychotherapy and medication studies

We applied metaregression to obtain the SMDs and confidence intervals of each of the four study arms. As seen in @fig-plot-means-all there were substantial differences between the four arms of the meta-analysis with striking differences between the medication and psychotherapy control arms. In particular, pill placebo had an SMD = `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$coef_means, 2)` (95% CI: `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$lower_ci, 2)` to `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "medication_control", ]$upper_ci, 2)`) whereas psychotherapy controls had an SMD = `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$coef_means, 2)` (95% CI: `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$lower_ci, 2)` to `r round(df_mean_coefs_from_sim[df_mean_coefs_from_sim$condition == "psychotherapy_control", ]$upper_ci, 2)` ).

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#| label: fig-plot-means-all
#| fig-cap: "Meta-analytic estimates of within-group changes: all studies"
#| fig-cap-location: top

df_mean_coefs_from_sim <- df_mean_coefs_from_sim %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition)))

# subtitle_text_all <- "Metanalytic estimates of within-group changes: all studies"
plot_means_all <- plot_means_function(df_mean_coefs_from_sim)
print(plot_means_all)

```

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
# #| label: tbl-coefs-overall
# #| tbl-cap: "Summary statistics of the estimated SMDs"
# #| tbl-cap-location: top
# 
# # Prepare to tabulate
# 
# table_mean_coefs_from_sim <- df_mean_coefs_from_sim %>%
#   mutate(across(where(is.numeric), ~round(., 2))) %>%
#   relocate("n", .after = "condition") 
# 
# new_column_names <- c("Condition", "N", "Coefficient", "SE", "Lower CI", "Upper CI")
# 
# # Assign new column names to the data frame
# colnames(table_mean_coefs_from_sim) <- new_column_names
# 
# flextable(table_mean_coefs_from_sim) %>%
#   set_table_properties(width = 1) %>%
#   autofit() %>%
#   font(fontname = "Calibri", part = "all") %>% 
#   # add_header_row(values = "Summary statistics of the estimated SMDs", colwidths = 6) %>%
#   # add_header_row(values = "Table 3", colwidths = 6) %>% 
#   bold(part = "header") %>%   # Bold the relevant header rows
#   bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
#   # italic(i = 2, italic = TRUE, part = "header") %>% 
#   border_remove() %>% # this is to remove borders around the caption that look strange
#   hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
#   vline(j = 1, part = "body", border = small_border) %>% 
#   border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
#   hline(part = "header", border = big_border) %>% 
#   # padding(i = 2, padding.bottom = 15, part = "header") %>% 
#   align(j = 2:ncol(df_mean_coefs_from_sim), align = "center", part = "all") %>% 
#   align(j = 1, align = "left", part = "all") %>% 
#   paginate(init = TRUE, hdr_ftr = TRUE) %>% 
#   line_spacing(space = .7, part = "all")

```

In @tbl-results-overall, we present the regression that tests our hypothesis about differences between medication and psychotherapy controls. Here, medication control is the reference category to which all others are compared. The strongest difference between arms, as judged by the z-value, is between the psychotherapy and medication controls with a z-value of `r round(aggregate_results_overall[aggregate_results_overall$condition == "psychotherapy_control", ]$z_value, 2)` (p<0.0001).

```{r, warning=FALSE, message = FALSE, echo = FALSE, ft.align="left"}
#| label: tbl-results-overall
#| tbl-cap: "Results from metaregression with overall sample"
#| tbl-cap-location: top

# Regression results

aggregate_results_overall <- aggregate_results_overall %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition))) %>% 
  relocate("condition", .before = "coefficients")

new_column_names <- c("Condition", "Coefficient", "SE", "z value", "Lower CI", "Upper CI", "T2", "I2", "k", "R2")

# Assign new column names to the data frame
colnames(aggregate_results_overall) <- new_column_names

table_aggregate_results_overall <- aggregate_results_overall[, 1:6]
table_aggregate_results_overall[1, 4] <- NA

flextable(table_aggregate_results_overall) %>%
  set_table_properties(width = 1) %>%
  autofit() %>%
  add_footer_row(values = as_paragraph("T", as_sup("2"), " = ", aggregate_results_overall[1, "T2"],
                      "; ", as_i("I"), as_sup("2"), " = ", aggregate_results_overall[1, "I2"], 
                      "; ", as_i("K"), " = ", aggregate_results_overall[1, "k"],
                      "; ", as_i("R"), as_sup("2"), " = ", aggregate_results_overall[1, "R2"], "."), colwidths = 6) %>%
  # add_header_row(values = "Results from metaregression with overall sample", colwidths = 6) %>%
  # add_header_row(values = "Table 4", colwidths = 6) %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  # italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  # padding(i = 2, padding.bottom = 15, part = "header") %>% 
  align(j = 2:ncol(table_aggregate_results_overall), align = "center", part = "all") %>% 
  align(j = 1, align = "left", part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>% 
  paginate(init = TRUE, hdr_ftr = TRUE) %>% 
  line_spacing(space = .7, part = "all")

```

##### Sensitivity analyses

We  conducted a series of sensitivity analyses (all figures in the Supplemental Materials). Excluding waitlist control studies (see @fig-plot-means-no-wl) and sub-clinical studies (see @fig-plot-means-clin) yielded a pattern of results very similar to the overall analyses. Next, we examined the data including only those studies that used the CDRS (see @fig-plot-means-cdrs) or the HAMD (see @fig-plot-means-hamd). Medication control and psychotherapy control conditions remained significantly different, though the small number of studies resulted in less precise estimates of the SMDs. Finally, we showed that different values for the pre-post measure correlation had minimal effect on the estimated outcomes (see @fig-stab-sims).

##### Addressing Regression to the Mean

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#### Let's start from the beginning and regress cohens on the baseline dataframe.

# create a new list with dfs for the sims
df_long_for_metan_adj_cohens_d <- df_long_for_metan
# rename the cohens_d value to make space for the new cohen's d
df_long_for_metan_adj_cohens_d <-
  df_long_for_metan_adj_cohens_d %>%
rename(unadjusted_cohens_d = cohens_d,
)
# estimate the cohen's d from the residuals
model_adjust_cohens_d <-  lm(df_long_for_metan_adj_cohens_d$unadjusted_cohens_d ~ df_long_for_metan_adj_cohens_d$perc_baseline_mean, na.action=na.exclude)

# there is a great effect of the baseline means, as shown here
results_adjusted_cohens_d <- broom:: tidy(model_adjust_cohens_d )

# extract the residuals which are the adjusted cohen's d, to get the simulated standard errors
df_long_for_metan_adj_cohens_d$cohens_d <- residuals(model_adjust_cohens_d )

# apply function to simulate standard errors
list_df_simulated_adj_cohens_d <- simulate_dataframes_for_st_errors(df = df_long_for_metan_adj_cohens_d,
                                   num_repetitions = 1000,
                                   seed  = 1974,
                                   n = length(unique(df_long_for_metan_adj_cohens_d$new_study_id)),
                                   a = 0.45,
                                   b = .9,
                                   mean = 0.65,
                                   sd = 0.2)

# add the four level variable and re-level
for(i in 1: length(list_df_simulated_adj_cohens_d)){

  list_df_simulated_adj_cohens_d[[i]]$arm_effect_size <- factor(list_df_simulated_adj_cohens_d[[i]]$arm_effect_size)
  list_df_simulated_adj_cohens_d[[i]] <- list_df_simulated_adj_cohens_d[[i]] %>%
    mutate(four_level_var = case_when(psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
                                      psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control")
    )

  list_df_simulated_adj_cohens_d[[i]]$four_level_var <- factor(list_df_simulated_adj_cohens_d[[i]]$four_level_var) # turn to factor

  # relevel so that medication control becomes the reference category for the regression
  #  list_df_simulated[[i]]$four_level_var <- relevel(list_df_simulated[[1]]$four_level_var, ref = "medication_control")
  list_df_simulated_adj_cohens_d[[i]]$four_level_var <-   fct_relevel(list_df_simulated_adj_cohens_d[[i]]$four_level_var,
                                                         "medication_control",
                                                         "medication_active",
                                                         "psychotherapy_control",
                                                         "psychotherapy_active")
}

# Run metaregression model (assuming that the four level variable is present given that I create this dataframe from the list_)
# specify model
model_x <- as.formula(~ four_level_var)

# run metareg function
# here you can use the simulated results directly
list_model_1_meta_reg_adj_cohens_d <- run_metaregression(list_df_simulated_adj_cohens_d, model_x)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated_adj_cohens_d[[1]]$four_level_var)
aggregate_results_overall_adjusted_cohens_d <-  aggregate_model_results(list_model_1_meta_reg_adj_cohens_d, condition)
aggregate_results_overall_adjusted_cohens_d <- cbind(aggregate_results_overall_adjusted_cohens_d, condition)

# extract SMDs from the  list
list_dummy_var_means_adjusted_cohens_d <-  lapply(list_model_1_meta_reg_adj_cohens_d, extract_coefficients_func)

# calculate mean SMDs and ses
df_mean_coefs_from_sim_adjusted_cohens_d <- calculate_mean_coefs_ses(list_dummy_var_means_adjusted_cohens_d, condition)

df_mean_coefs_from_sim_adjusted_cohens_d <- data.frame(cbind(condition, df_mean_coefs_from_sim_adjusted_cohens_d))

```

We addressed potential regression to the mean by residualising SMDs by the percentage adjusted baseline scores in one regression model (see @tbl-adj-smds in the Supplement). The difference between the medication control and psychotherapy control arms remained similar to the overall analyses.

### Number of trial sites

```{r, warning=FALSE, message = FALSE, echo=FALSE}
### Number of sites
# Write results, tabulate.

## This data lives in this spreadsheet for psy trials, we need to merge it in

df_no_sites <- readxl:: read_excel("for_despina_and_giannis_200224.xlsx")

df_no_sites <- df_no_sites %>% 
  select(c(new_study_id, number_of_sites)) %>% 
  filter(!is.na(number_of_sites)) %>% 
  distinct(new_study_id, .keep_all = TRUE) %>% 
  mutate(number_of_sites = as.numeric(number_of_sites))

df_baseline_demographics <- full_join(df_baseline_demographics, df_no_sites, by = "new_study_id")

df_baseline_demographics <- df_baseline_demographics %>% 
  filter(new_study_id != "March, 2004") # removing TADS because it does not technically belong to either the psy or med groups

# we have different vectors for psy and med trials, lets coalesce

df_baseline_demographics <- df_baseline_demographics %>%
  mutate(no_sites = as.numeric(no_sites)) %>% 
  mutate(sites = coalesce(no_sites, number_of_sites)) %>%
  select(-no_sites, -number_of_sites)

# we have a few rogue studies (Kitchen, McCarty and Wright). Kitchen was too recent (after 2021), McCarty and Wright are not in Cuijpers' master
# dataset for some reason, though they are in the list of included studies in the table in his MA. For now I'm going to remove them because 
# we don't have data for these trials.

df_baseline_demographics <- df_baseline_demographics %>% 
  filter(!(new_study_id %in% c("Kitchen, 2021", "McCarty, 2013", "Wright, 2020")))

# Now create binary variable, single or multi site trial

df_baseline_demographics <- df_baseline_demographics %>% 
  mutate(site_binary = case_when(sites == 1 ~ "single", 
                                 sites > 1 ~ "multi",
                                 TRUE ~ NA_character_))

# df_baseline_demographics %>%
#   group_by(psy_or_med, site_binary) %>%
#   summarize(num_trials = n())

# Lets look at results

# summary stats
df_sites_summary_stats <- df_baseline_demographics %>%
  group_by(psy_or_med) %>%
  summarise(
    n = n(),
    mean_sites = mean(sites, na.rm = TRUE),
    sd_sites = sd(sites, na.rm = TRUE),
  )

# Tabulate
# Rounding
df_sites_summary_stats <- df_sites_summary_stats %>%
  mutate(across(where(is.numeric), ~round(., 2)))

# Row names
df_sites_summary_stats <- df_sites_summary_stats %>%
  mutate(psy_or_med = case_when(
    psy_or_med == 0 ~ "Medication",
    psy_or_med == 1 ~ "Psychotherapy",
    TRUE ~ as.character(psy_or_med)  # Keep other values unchanged
  ))

df_sites_summary_stats <- df_sites_summary_stats %>%
  rename(
    Modality = psy_or_med,
    N = n,
    Mean = mean_sites,
    SD = sd_sites
  )

# t-test
sites_t_test <- t.test(sites ~ psy_or_med, data = df_baseline_demographics)

sites_t_test_stats <- data.frame(
  statistic = sites_t_test$statistic,
  df = sites_t_test$parameter,
  p_value = sites_t_test$p.value,
  ci_lower = sites_t_test$conf.int[1],
  ci_upper = sites_t_test$conf.int[2]
)

sites_t_test_stats <- sites_t_test_stats %>%
  mutate(across(.cols = c(statistic, df, ci_lower, ci_upper), ~ round(., 2))) %>% 
  mutate(p_value = ifelse(p_value < 0.001, "< 0.001", as.character(round(p_value, 3))))

```

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
## | label: tbl-sites-summary
## | tbl-cap: "Number of sites across medication and psychotherapy studies"
## | tbl-cap-location: top
#
# flextable(df_sites_summary_stats) %>%
#   set_table_properties(width = 1, layout = "fixed") %>%
#   add_footer_row(values = as_paragraph(as_i("t"), "(", sites_t_test_stats$df, ") = ", sites_t_test_stats$statistic,
#                       "; ", as_i("p "), sites_t_test_stats$p_value, 
#                       "; 95% CI: ", sites_t_test_stats$ci_lower,
#                       "-", sites_t_test_stats$ci_upper, "."), colwidths = 4) %>% 
#   # add_header_row(values = "Number of sites across medication and psychotherapy studies", colwidths = 4) %>%
#   # add_header_row(values = "Table 2", colwidths = 4) %>% 
#   bold(part = "header") %>%   # Bold the relevant header rows
#   bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
#   # italic(i = 2, italic = TRUE, part = "header") %>% 
#   border_remove() %>% # this is to remove borders around the caption that look strange
#   hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
#   vline(j = 1, part = "body", border = small_border) %>% 
#   border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
#   hline(part = "header", border = big_border) %>% 
#   # padding(i = 2, padding.bottom = 15, part = "header") %>%
#   padding(i = c(1, 2), j = 1, padding.left = 10) %>% 
#   align(j = 2:ncol(df_sites_summary_stats), align = "center", part = "all") %>% 
#   align(j = 1, align = "left", part = "all") %>% 
#   font(fontname = "Calibri", part = "all") %>% 
#   autofit() %>% 
#   paginate(init = TRUE, hdr_ftr = TRUE) %>% 
#   line_spacing(space = .7, part = "all")

```

Average number of trial sites was significantly higher in medication trials (*M* = `r df_sites_summary_stats$Mean[df_sites_summary_stats$Modality == "Medication"]`, *SD* =`r df_sites_summary_stats$SD[df_sites_summary_stats$Modality == "Medication"]`) compared to psychotherapy studies (*M* =`r df_sites_summary_stats$Mean[df_sites_summary_stats$Modality == "Psychotherapy"]`, *SD* =`r df_sites_summary_stats$SD[df_sites_summary_stats$Modality == "Psychotherapy"]`)(*t* (`r round (sites_t_test$parameter, 2)`) = `r round(sites_t_test$statistic,2)`, *p* =`r sites_t_test_stats$p_value`). Of those studies with data available, 26 of 28 (93%) medication trials were multisite, compared to 24 of 45 (54%) psychotherapy studies.

## Comparing the nature and intensity of control conditions in psychotherapy trials

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# We need to read in the appropriate dataset which describes the active and control conditions used in
# psychotherapy trials.

df_descr_psy_conditions <- readxl:: read_excel("for_despina_and_giannis_200224.xlsx")

# checking id matchup
# discrepancies <- setdiff(df_descr_psy_conditions$new_study_id, df_long_for_metan$new_study_id)

# Cleaning the dataset
# For now lets look at variables pertaining to the nature / intensity of interventions

df_descr_psy_conditions <- df_descr_psy_conditions %>% 
  select(c(new_study_id, `Psychotherapy/control`, levels_var, number_of_sites, Number_of_sessions,
           frequency_weeks, Length_of_sessions_mins, Total_hours_of_intervention_hours, total_hours_gsh,
           Total_period_of_intervention_weeks, Format_group_individual, Delivery, Involvement_of_others))

# rename for clarity and to match master dataframe
df_descr_psy_conditions <- df_descr_psy_conditions %>% 
  rename(treatment = 'Psychotherapy/control', no_sites = number_of_sites, no_sessions =
           Number_of_sessions, freq_weeks = frequency_weeks, length_sessions_mins =
           Length_of_sessions_mins, total_hours = Total_hours_of_intervention_hours,
         total_period_weeks = Total_period_of_intervention_weeks, format = Format_group_individual, delivery =
           Delivery, family_involv = Involvement_of_others)

check <- df_descr_psy_conditions %>% 
  filter(treatment == 'wlc')

# Visual inspection

# There are two studies where the waitlist condition has some intervention (i.e. values should not be set to 0)
# Otherwise, I want values to be set to 0, not NA.

df_descr_psy_conditions <- df_descr_psy_conditions %>%
  mutate(
    no_sessions = ifelse(treatment == "wlc" & !(new_study_id %in% c("Ackerson, 1998", "Diamond, 2002")), 0, no_sessions),
    freq_weeks = ifelse(treatment == "wlc" & !(new_study_id %in% c("Ackerson, 1998", "Diamond, 2002")), 0, freq_weeks),
    length_sessions_mins = ifelse(treatment == "wlc" & !(new_study_id %in% c("Ackerson, 1998", "Diamond, 2002")), 0, length_sessions_mins),
    total_hours = ifelse(treatment == "wlc" & !(new_study_id %in% c("Ackerson, 1998", "Diamond, 2002")), 0, total_hours)
  )

# Now cau

check <- df_descr_psy_conditions %>% 
  filter(treatment == 'cau')

# All studies except for Martinovic, Srivastava, Stikkelbroek and Weisz 2009 should have variables set to NA. 
# This is because we cannot say that they did not feature 0 hours of intervention, but that for the most part
# it is not reported or it is not possible to quantify.

df_descr_psy_conditions <- df_descr_psy_conditions %>%
  mutate(
    no_sessions = ifelse(treatment == "cau" & !(new_study_id %in% c("Martinovic, 2006", "Srivastava, 2020", "Stikkelbroek, 2020", "Weisz, 2009")), NA, no_sessions),
    freq_weeks = ifelse(treatment == "cau" & !(new_study_id %in% c("Martinovic, 2006", "Srivastava, 2020", "Stikkelbroek, 2020", "Weisz, 2009")), NA, freq_weeks),
    length_sessions_mins = ifelse(treatment == "cau" & !(new_study_id %in% c("Martinovic, 2006", "Srivastava, 2020", "Stikkelbroek, 2020", "Weisz, 2009")), NA, length_sessions_mins),
    total_hours = ifelse(treatment == "cau" & !(new_study_id %in% c("Martinovic, 2006", "Srivastava, 2020", "Stikkelbroek, 2020", "Weisz, 2009")), NA, total_hours)
  )

# Now other controls

check <- df_descr_psy_conditions %>%
  filter(!(treatment %in% c('cau', 'wlc'))) %>% 
  filter(levels_var == "psychotherapy_control")

# Made a few manual adjustments in the original excel sheet.

# List of columns to clean
columns_to_clean <- c(
  "no_sessions",
  "freq_weeks",
  "length_sessions_mins",
  "total_hours",
  "total_period_weeks"
)

# Convert non-numeric values to NA in specified columns
df_descr_psy_conditions <- df_descr_psy_conditions %>%
  mutate_at(vars(columns_to_clean), ~as.numeric(as.character(.)))

# Giannis' function for summary stats

group_function <- function(dataframe, my_groups, columns) {
  result_dataframe <- data.frame(
    Group = character(),
    Column = character(),
    N = numeric(),
    Mean = numeric(),
    SD = numeric(),
    Median = numeric(),
    IQR = numeric(),
    stringsAsFactors = FALSE
  )
  unique_groups <- unique(dataframe[[my_groups]])
  for (group in unique_groups) {
    group_data <- dataframe[dataframe[[my_groups]] == group, ]
    for (col in columns) {
      col_data <- group_data[[col]]
      n <- sum(!is.na(col_data))
      mean_val <- mean(col_data, na.rm = TRUE)
      SD <- sd(col_data, na.rm = TRUE)
      median_val <- median(col_data, na.rm = TRUE)
      iqr <- IQR(col_data, na.rm = TRUE)
      result_dataframe <- rbind(result_dataframe,
                                data.frame(Column = col, Group = group, N = n, Mean = mean_val,
                                           SD = SD, Median = median_val,
                                           IQR = iqr, stringsAsFactors = FALSE))
    }
  }
  return(result_dataframe)
}

# Filter out wl conditions
# df_descr_psy_conditions <- df_descr_psy_conditions %>% 
  # filter(new_study_id %in% no_wl_ids)

# count_wl <- df_long_for_metan %>% 
#   filter(treatment == "wlc") %>% 
#   distinct(new_study_id) %>% 
#   nrow()
# 
# count_cau <- df_long_for_metan %>% 
#   filter(treatment == "cau") %>% 
#   distinct(new_study_id) %>% 
#   nrow()

sessions <- group_function(df_descr_psy_conditions,"levels_var","no_sessions")
frequency <- group_function(df_descr_psy_conditions,"levels_var","freq_weeks")
length_sessions_mins <- group_function(df_descr_psy_conditions,"levels_var","length_sessions_mins")
total_hours <- group_function(df_descr_psy_conditions,"levels_var","total_hours")
total_period_weeks <- group_function(df_descr_psy_conditions,"levels_var","total_period_weeks")

control_summary <- rbind(sessions, frequency, length_sessions_mins, total_hours)

control_summary <- control_summary %>% 
  select(-c(Median, IQR))

control_summary <- control_summary %>%
  mutate(Group = case_when(
    Group == "psychotherapy_active" ~ "Active",
    Group == "psychotherapy_control" ~ "Control",
    TRUE ~ as.character(Group)  # Keep other values unchanged
  ))

control_summary <- control_summary %>%
  mutate(Column = case_when(
    Column == "no_sessions" ~ "Number of sessions",
    Column == "freq_weeks" ~ "Intensity (sessions per week)",
    Column == "length_sessions_mins" ~ "Session length (mins)",
    Column == "total_hours" ~ "Total intervention hours",
    TRUE ~ as.character(Column)  # Keep original value if no match
  ))

# other_ctr_ids <- df_long_for_metan[df_long_for_metan$type == "other ctr",]$new_study_id
# testing <- df_descr_psy_conditions %>% 
#   filter(new_study_id %in% other_ctr_ids)
# view(testing)

# t tests

session_no_t_test <- t.test(no_sessions ~ levels_var, data = df_descr_psy_conditions)
freq_t_test <- t.test(freq_weeks ~ levels_var, data = df_descr_psy_conditions)
session_length_t_test <- t.test(length_sessions_mins ~ levels_var, data = df_descr_psy_conditions)
total_hours_t_test <- t.test(total_hours ~ levels_var, data = df_descr_psy_conditions)

extract_t_test_stats <- function(df) {

  t_test_stats <- data.frame(
  outcome = df$data.name,
  statistic = df$statistic,
  df = df$parameter,
  p_value = df$p.value,
  ci_lower = df$conf.int[1],
  ci_upper = df$conf.int[2])

  return(t_test_stats)

}

results_session_no_t_test <- extract_t_test_stats(session_no_t_test)
results_freq_t_test <- extract_t_test_stats(freq_t_test)
results_session_length_t_test <- extract_t_test_stats(session_length_t_test)
results_total_hours_t_test <- extract_t_test_stats(total_hours_t_test)

# for t test table

control_summary_t_tests <- rbind(results_session_no_t_test, results_freq_t_test, results_session_length_t_test, results_total_hours_t_test)

# For row names
control_summary_t_tests <- control_summary_t_tests %>%
  mutate(outcome = case_when(
    row_number() == 1 ~ "Number of sessions",
    row_number() == 2 ~ "Intensity (sessions per week)",
    row_number() == 3 ~ "Session length (mins)",
    row_number() == 4 ~ "Total intervention hours"
  ))

# Define new column names
new_column_names <- c("Outcome", "t statistic", "df", "p-value", "Lower CI", "Upper CI")

# Assign new column names to the data frame
colnames(control_summary_t_tests) <- new_column_names

#Rounding
control_summary_t_tests <- control_summary_t_tests %>%
  mutate(across(where(is.numeric) & !matches("p-value"), ~ round(., 2))) %>%
  mutate(`p-value` = ifelse(`p-value` < 0.001, "< 0.001", sprintf("%.3f", `p-value`)))

# function for between group effect sizes (cohens d)
# 
# effect_btwn <- function(df) {
#   pooled_sd <- sqrt(((df$N[1]-1)*df$SD[1]^2 + (df$N[2]-1)*df$SD[2]^2) / (df$N[1] + df$N[2] - 2))
#   mean_diff <- df$Mean[1] - df$Mean[2]
#   effect_btwn <- mean_diff / pooled_sd
#   return(effect_btwn)
# }
# 
# session_es <- effect_btwn(sessions)
# frequency_es <- effect_btwn(frequency)
# length_sessions_mins_es <- effect_btwn(length_sessions_mins)
# total_hours_es <- effect_btwn(total_hours)
# total_period_weeks_es <- effect_btwn(total_period_weeks)

# using a package to calculate instead

session_es <- cohen.d(df_descr_psy_conditions$no_sessions ~ df_descr_psy_conditions$levels_var, pooled = TRUE, paired = FALSE, within=FALSE, na.rm=FALSE )
frequency_es <- cohen.d(df_descr_psy_conditions$freq_weeks ~ df_descr_psy_conditions$levels_var, pooled = TRUE, paired = FALSE, within=FALSE, na.rm=FALSE )
length_sessions_mins_es <- cohen.d(df_descr_psy_conditions$length_sessions_mins ~ df_descr_psy_conditions$levels_var, pooled = TRUE, paired = FALSE, within=FALSE, na.rm=FALSE )
total_hours_es <- cohen.d(df_descr_psy_conditions$total_hours ~ df_descr_psy_conditions$levels_var, pooled = TRUE, paired = FALSE, within=FALSE, na.rm=FALSE )

# going to try displaying summary statistics and t-tests in one place

control_summary <- cbind(control_summary, 
                          "Cohen's d" = rep(NA, nrow(control_summary)),
                         "Upper CI" = rep(NA, nrow(control_summary)),
                         "Lower CI" = rep(NA, nrow(control_summary)),
                         t = rep(NA, nrow(control_summary)), 
                          df = rep(NA, nrow(control_summary)), 
                          "p-value" = rep(NA, nrow(control_summary)))

# Update the first row with session_es
control_summary[1, c("Cohen's d", "Upper CI", "Lower CI", "t", "df", "p-value")] <- c(
  session_es$estimate, session_es$conf.int[1], session_es$conf.int[2], 
  control_summary_t_tests$`t statistic`[1], control_summary_t_tests$df[1], 
  control_summary_t_tests$`p-value`[1]
)

# Update the third row with frequency_es
control_summary[3, c("Cohen's d", "Upper CI", "Lower CI", "t", "df", "p-value")] <- c(
  frequency_es$estimate, frequency_es$conf.int[1], frequency_es$conf.int[2], 
  control_summary_t_tests$`t statistic`[2], control_summary_t_tests$df[2], 
  control_summary_t_tests$`p-value`[2]
)

# Update the fifth row with length_sessions_mins_es
control_summary[5, c("Cohen's d", "Upper CI", "Lower CI", "t", "df", "p-value")] <- c(
  length_sessions_mins_es$estimate, length_sessions_mins_es$conf.int[1], length_sessions_mins_es$conf.int[2], 
  control_summary_t_tests$`t statistic`[3], control_summary_t_tests$df[3], 
  control_summary_t_tests$`p-value`[3]
)

# Update the seventh row with total_hours_es
control_summary[7, c("Cohen's d", "Upper CI", "Lower CI", "t", "df", "p-value")] <- c(
  total_hours_es$estimate, total_hours_es$conf.int[1], total_hours_es$conf.int[2], 
  control_summary_t_tests$`t statistic`[4], control_summary_t_tests$df[4], 
  control_summary_t_tests$`p-value`[4]
)

control_summary <- control_summary %>%
  mutate(
    `Cohen's d` = as.numeric(`Cohen's d`),
    `Upper CI` = as.numeric(`Upper CI`),
    `Lower CI` = as.numeric(`Lower CI`)
  ) %>%
  mutate(across(where(is.numeric), ~round(., 2)))

# # plot
# 
# ggplot(df_descr_psy_conditions_no_wl, aes(x = total_hours, fill = levels_var)) +
#   geom_histogram(position = "identity", alpha = 0.5, binwidth = 5, color = "black") +
#   facet_wrap(~ levels_var, ncol = 2) +
#   theme_minimal() +
#   labs(title = "Histograms of Number of Total Hours by Group",
#        x = "Number of Hours",
#        y = "Frequency") +
#   scale_fill_manual(values = c("lightpink", "lightblue"))

```

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
#| label: tbl-intensity-of-int
#| tbl-cap: "Comparing the intensity of the intervention between active and control arms of psychotherapy studies"
#| tbl-cap-location: top

as_grouped_data(control_summary, groups = "Column") %>% 
  as_flextable(hide_grouplabel = TRUE ) %>% 
  bold(i = c(1, 4, 7, 10), bold = TRUE, part = "body") %>%  
  set_table_properties(width = 1) %>%  # Set table width 
  autofit() %>%  # Autofit column widths 
  align(j = 2:9, align = "center", part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  # add_header_row(values = "Comparing the intensity of the intervention between active and control arms of psychotherapy studies", colwidths = 4) %>% # This is a work around for the problems with rendering captions
  # add_header_row(values = "Table 5", colwidths = 4) %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  #italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(j = 1:4, part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = c(1,4), part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  #padding(i = 2, padding.bottom = 15, part = "header") %>% 
  paginate(init = TRUE) %>% 
  line_spacing(space = .7, part = "all") 
```

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
# #| label: tbl-t-tests-intensity
# #| tbl-cap: "Results for t-tests Comparing the intensity of the intervention between active and control arms of psychotherapy studies"
# #| tbl-cap-location: top
# 
# flextable(control_summary_t_tests) %>%
#   set_table_properties(width = 1) %>%  # Set table width %>% 
#   autofit() %>%  # Autofit column widths 
#   align(j = 2:ncol(control_summary_t_tests), align = "center", part = "all") %>% 
#   font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
#   #add_header_row(values = "Results for t-tests comparing intervention intensity between active and control arms of psychotherapy trials", colwidths = 6) %>% # This is a work around for the problems with rendering captions
#   #add_header_row(values = "Table 6", colwidths = 6) %>% 
#   bold(part = "header") %>%   # Bold the relevant header rows
#   bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
#   #italic(i = 2, italic = TRUE, part = "header") %>% 
#   border_remove() %>% # this is to remove borders around the caption that look strange
#   hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
#   vline(j = 1, part = "body", border = small_border) %>% 
#   border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
#   hline(part = "header", border = big_border) %>% 
#   #padding(i = 2, padding.bottom = 15, part = "header") %>% 
#   paginate(init = TRUE) %>% 
#   line_spacing(space = .7, part = "all")

```

As seen in @tbl-intensity-of-int, active conditions featured significantly more sessions when compared to control conditions. Sessions in active conditions were longer and more frequent, resulting in significantly more intervention hours overall. Notably, many control conditions were very poorly described and their intensity could not be quantified, resulting in missing data. We performed a sensitivity analysis where we excluded trials using waitlist controls; with the exception of number of sessions, differences between active and control arms were no longer statistically significant though remained substantial, with between-group Cohen's ds ranging from 0.41-0.54 (please see @tbl-intensity-of-int-no-wl in the Supplement).

# Conclusions and Clinical Implications

```{r, echo=FALSE}
# Filter the dataframe
filtered_df <- df_long_for_metan %>%
  filter(psy_or_med == 1, arm_effect_size == "cohens_d_control")

# Count the occurrences of each type
type_counts <- filtered_df %>%
  count(type) %>%
  mutate(total_count = sum(n),
         wl_sum = sum(n[type == "wl"]),
         wl_percentage = wl_sum / total_count * 100)

# how many examples of matching of active and control in number of hours

# instances_count <- df_descr_psy_conditions_no_wl %>%
#   filter(!is.na(total_hours)) %>%
#   group_by(new_study_id, total_hours) %>%
#   summarise(count = n()) %>%
#   filter(count > 1) %>%
#   summarise(total_instances = n())

```

This paper sought to address the question of whether psychotherapy and medication can be meaningfully compared on the basis of the existing evidence, by looking at two key conditions for comparability. First, whether the participants of trials in one modality are comparable to those in another modality. Second, whether conditions of the trial, such as the effects of control conditions or the number of sites involved, are comparable.

Starting with the first question, we found that participants in medication trials are comparable on age but are more likely to be male and have more severe depression compared to those in psychotherapy trials. This indicates that different people enter medication and psychotherapy trials, thereby violating basic assumptions of comparability.

Severity is particularly important as it may moderate treatment response, with some evidence suggesting that those with higher baseline scores respond more to antidepressants [@stoneResponseAcuteMonotherapy2022] or that their response to pill placebo is lower [@bridgePlaceboResponseRandomized2009]. Other studies argue against severity as a treatment moderator [@trogerBaselineDepressionSeverity2024; @weitzBaselineDepressionSeverity2015], however these are within people who have chosen to be in the particular trial and modality. Moreover, severity may represent different subtypes in terms of course of depression and real-life outcomes [@lamersSixyearLongitudinalCourse2016; @simmonds-buckleyDepressionSubtypesTheir2021].

We then turned to our second question about whether trial design conditions are comparable between modalities. We found that medication trials are vastly more likely to be multi-site than their psychotherapy counterparts; in the current study, 93% of medication RCTs were multisite compared to 54% of psychotherapy RCTs. Multisite trials are associated with higher pill placebo response [@bridgePlaceboResponseRandomized2009], and are less common in publicly-funded trials which show lower pill placebo efficacy. Also, in single-site trials, principal investigators are often intellectually invested in the treatment (in psychotherapy these are often treatments developed or modified by the PI); this is in stark contrast to the incentive structure in multi-site trials where the number of recruited participants is the primary unit of reimbursement.

Second, psychotherapy controls have moderate effect sizes (-0.5) whereas medication controls have very large effect sizes (-1.9). Our analysis could be critiqued as it compares within arm symptom change per trial, therefore effectively breaking randomisation. This criticism would apply if our aim were to draw inferences about the efficacy of each arm --- in which case preserving randomisation to balance confounders is critical. However, we note that our findings are largely in keeping with those of the NMA, which is designed to preserve the randomisation structure. More importantly, we do not claim that these differences are genuinely due to efficacy differences; they may well be because people who attend psychotherapy and medication trials are different and therefore respond differently. In either case (difference in efficacy vs difference in trial participant profile), the vast disparity in the response to control conditions is reason for major concern about our ability to draw inferences from comparisons of modalities. This is particularly problematic as clinicians as well as policy makers often resort to between-group effect sizes to summarise findings, which our findings make obvious is misleading.

Moving beyond comparison between modalities, we examined whether psychotherapy controls are reasonable counterfactuals to receiving treatment. An obvious disadvantage of psychotherapy trials is that they are typically unblinded (and hard to blind) yet psychotherapy trials are unlikely to fulfill some other basic conditions of the "all else is equal" assumption. In order to test that a psychological treatment is effective per se (e.g. because of the specific techniques) rather than because of generic effects (e.g. pleasant human contact), aspects such as therapist contact time should be matched. Many (24%) psychotherapy RCTs use waitlist controls, which by definition do not match for hours of therapist contact, and are often associated with disappointment bias. TAU and other psychotherapy control conditions varied drastically; 9 RCTs used controls that exactly matched the active arm in total number of contact hours, though several studies used bibliotherapy or online-only control conditions which did not involve any direct therapist contact. Importantly, controls were often very poorly described, resulting in difficulties quantifying their intensity and hence evaluating their adequacy as counterfactual conditions. Overall, there is a lack of rigor in matching control to active treatment conditions in psychotherapy RCTs, with the latter typically featuring considerably more contact hours, which may artificially inflate estimates of the efficacy of psychological treatments.

Given all of the above, the certainty with which guidelines recommend psychotherapy over medication for adolescent depression is surprising. Indeed, we believe that our findings have several profound implications for patients, their families, clinicians and policy makers.

First, the grounds for comparison between medication and psychotherapy should be seen as shaky, rather than offering confidence, and there is an urgent need to revisit guidelines and public information in light of the limitations.

Second, the low quality of psychotherapy control conditions should prompt consideration of how to create fair comparators. Investment should be directed into providing rigorous evidence that establishes depression psychotherapies as more efficacious than fair controls. There are examples of RCTs where such rigor has been applied (e.g. Bolton, 2007; Liddle, 1990; Rohde, 2004) in matching active and control arms on variables such as therapist time and attention, provision of homework, and small group interaction.

Third, our findings make clear the inherent difficulties of comparing psychotherapy with medication trials [@delgiovaneCombiningPharmacologicalNonpharmacological2019]. The first obstacle is the comparability of the populations taking part. Head-to-head comparisons of psychotherapy with medication (as done in [@marchFluoxetineCognitiveBehavioralTherapy2004]) are more favourable in this regard, yet even so these trials might sample the population of those who are indifferent to which treatment they receive. And even in such a design, difficulties with blinding of the psychotherapy control would have to be overcome to draw valid inferences.

In summary, the current findings give cause for consternation about the state of the evidence for treatments of youth depression. Our data question the state of knowledge about the efficacy of psychotherapies and the extent to which giving them primacy in the treatment of depression is justified and beneficial for young people. Returning to our motivating question, the stakeholders, including patients and clinicians, deserve better evidence on which to base their choices.

# Supplemental Materials

## Systematic review search terms and PRISMA chart

We conducted a systematic search for medication studies published from 31 May 2015 up to 1 Jan 2021 (i.e. after the final search date of Cipriani et al.'s [@ciprianiComparativeEfficacyTolerability2016] review up to the final search date of Cuijpers et al's [@cuijpersEffectsPsychologicalTreatments2021] review). We searched PubMed, the Cochrane Central Register of Controlled Trials, Embase, Web of Science, CINAHL, PsycINFO and LiLACS for randomised controlled trials (RCTs) comparing any antidepressant with placebo in the treatment of children and adolescents with a primary diagnosis of major depressive disorder. We used the same search terms as Cipriani [@ciprianiComparativeEfficacyTolerability2016] with one additional search term to include only placebo-controlled trials (see below). We additionally applied filters to specify our date range, and to exclude reviews and non-human studies. We also searched clinical trial registers for published and unpublished studies however all RCTs meeting inclusion criteria had already been identified from the database search outlined above. Please see @fig-prisma for the PRISMA flow diagram.

We used Covidence, an online software tool, to manage our systematic review. Our search produced 538 studies, 88 of which were duplicates and subsequently removed. Two authors screened 450 titles and abstracts, and 38 full text records. Seven studies met inclusion criteria and data extraction was completed for these papers.

**Search terms**

Explicit search strategy: title/abstract = (depress* or dysthymi\* or "mood disorder\*" or "affective disorder*") AND (adolesc\* or child\* or boy\* or girl\* or juvenil\* or minors or paediatri\* or pediatri\* or pubescen\* or school\* or student\* or teen\* or young or youth\*) AND (selective serotonin reuptake inhibitor or SSRI or citalopram or fluoxetine or paroxetine or sertraline or escitalopram or fluvoxamine or serotonin norepinephrine reuptake inhibitor\* or SNRI or venlafaxine or duloxetine or milnacipran or reboxetine or bupropion or noradrenergic and specific serotonergic antidepressants or NaSSA or mirtazapine or TCA or tricyclic or amersergide or amineptine or amitriptyline or amoxapine or butriptyline or chlorpoxiten or clomipramine or clorimipramine or demexiptiline or desipramine or dibenzipin or dothiepin or doxepin or imipramine or lofepramine or melitracen or metapramine or nortriptyline or noxiptiline or opipramol or protriptyline or quinupramine or tianeptine or trimipramine) AND (placebo)

**References for included trials**

References for trials included in the existing meta-analyses drawn upon for this study can be found [here](https://docs.metapsy.org/databases/depression-childadol-psyctr/) for psychotherapy RCTs and [here](https://ora.ox.ac.uk/objects/uuid:e0b5ae23-d562-4348-94b8-84f70b7812c5) for medication RCTs. Below are references for the seven additional RCTs identified in the original systematic review conducted for the current study.

*Publications*

Atkinson, S., Lubaczewski, S., Ramaker, S., England, R. D., Wajsbrot, D. B., Abbas, R., & Findling, R. L. (2018). Desvenlafaxine versus placebo in the treatment of children and adolescents with major depressive disorder. *Journal of Child and Adolescent Psychopharmacology, 28(1),* 55-65. DOI: 10.1089/cap.2017.0099

Durgam, S., Chen, C. Z., Migliore, R., Prakash, C., Edwards, J., & Findling, R. L. (2018). A phase 3, double-blind, randomized, placebo-controlled study of vilazodone in adolescents with major depressive disorder. *Pediatric Drugs, 20(4),* 353-363. DOI: 10.1007/s40272-018-0290-4

Findling, R. L., McCusker, E., & Strawn, J. R. (2020). A randomized, double-blind, placebo-controlled trial of vilazodone in children and adolescents with major depressive disorder with twenty-six-week open-label follow-up. *Journal of Child and Adolescent Psychopharmacology, 30(6),* 355-365. DOI: 10.1089/cap.2019.0176

Le Noury, J., Nardo, J. M., Healy, D., Jureidini, J., Raven, M., Tufanaru, C., & Abi-Jaoude, E. (2015). Restoring Study 329: efficacy and harms of paroxetine and imipramine in treatment of major depression in adolescence. *BMJ, 351,*, h4320. DOI: 10.1136/bmj.h4320

Weihs, K. L., Murphy, W., Abbas, R., Chiles, D., England, R. D., Ramaker, S., & Wajsbrot, D. B. (2018). Desvenlafaxine versus placebo in a fluoxetine-referenced study of children and adolescents with major depressive disorder. Journal of Child and Adolescent Psychopharmacology, 28(1), 36-46. DOI: 10.1089/cap.2017.0100

*Unpublished clinical trials*

Active Reference (Fluoxetine) Fixed-dose Study of Vortioxetine in Paediatric Patients Aged 12 to 17 Years With Major Depressive Disorder (MDD). ClinicalTrials.gov Identifier: NCT02709746. Accessed 17 Jan 2024.

Safety and Efficacy of Levomilnacipran ER in Adolescent Participants With Major Depressive Disorder. ClinicalTrials.gov Identifier: NCT02431806. Accessed 17 Jan 2024.

**Hierarchy of depression symptom severity measurement scales**

Where multiple depression rating scales were used, we selected the best available measure according to the following hierarchy used in Cipriani et al. [@ciprianiComparativeEfficacyTolerability2016].

1. Children’s Depression Rating Scale (CDRS)
1. Hamilton Depression Rating Scale (HAMD)
1. Montgomery Asberg Depression Rating Scale (MADRS)
1. Beck Depression Inventory (BDI)
1. Children’s Depression Inventory (CDI)
1. Schedule for Affective Disorders and Schizophrenia for School-Aged Children (K-SADS)
1. Mood and Feeling Questionnaire (MFQ)
1. Reynolds Adolescent Depression Scale (RADS)
1. Bellevue Index of Depression (BID)
1. Child Depression Scale (CDS)
1. Centre for Epidemiological Studies Depression Scale (CES-D)
1. Child Assessment Schedule (CAS)
1. Child Behaviour Checklist-Depression (CBCL-D)

## Full description of methods (including formalisms)

## Included studies

We drew upon RCTs included in two recent comprehensive meta-analyses with open data available for each medication and psychotherapy, and supplemented them with an updated systematic review. Please refer to these original meta-analyses for a detailed description of their search strategy and study selection criteria. Psychotherapy studies were drawn from a systematic review and meta-analysis of randomised trials comparing psychotherapy for youth depression against control conditions [@cuijpersEffectsPsychologicalTreatments2021] (dataset available [here](https://docs.metapsy.org/databases/depression-childadol-psyctr/)). Whilst Cuijpers et al. [@cuijpersEffectsPsychologicalTreatments2021] excluded studies for which the primary outcome variable could not be calculated due to missing data, we included these studies and performed the imputations outlined below. We also included studies which had data available for other variables in interest, including number of sites or baseline demographics; hence we have more psychotherapy studies included in this review compared to the original meta-analysis. Whilst the online database is regularly updated, we chose to exclude studies published after the final date of Cuijpers et al.'s [@cuijpersEffectsPsychologicalTreatments2021] literature search.

Medication studies were drawn from a network meta-analysis examining the efficacy and tolerability of antidepressants and placebo for major depressive disorder in children and adolescents [@ciprianiComparativeEfficacyTolerability2016]. A dataset was made available online though did not include means or standard deviations at baseline or post-test. We were unable to access the full dataset used in this meta-analysis, and hence completed extraction from the included studies ourselves. We excluded three studies because they had no control arm. We were unable to locate and therefore complete extraction for two RCTs (Almeida-Montes, 2005; Eli Lilly, 1986). Many studies did not report complete data; we contacted all corresponding authors to request missing data, though did not receive any responses.

We conducted a systematic search for medication studies published after the final search date of Cipriani et al.'s [@ciprianiComparativeEfficacyTolerability2016] review up to the final search date of Cuijpers et al's [@cuijpersEffectsPsychologicalTreatments2021] review to ensure we analysed an equivalently up-to-date database of medication trials. Please see the Supplemental Materials for further details. Our search produced 538 studies, 88 of which were duplicates and subsequently removed. Two authors screened 450 titles and abstracts, and 38 full text records. Seven studies met inclusion criteria and data extraction was completed for these papers.

## Statistical Analysis

### Sample characteristics

We conducted a series of random-effects meta-analyses and tested for subgroup differences between psychotherapy and medication trials in sample characteristics including sex, age, and severity of depressive symptoms at baseline. Meta-analyses were implemented using R's Meta package (version 7.0-0).

In order to compare depression severity across the variety of instruments the studies used, we performed a min-max normalisation to turn each study arm mean score at baseline into a percentage using the following formalism:

$$
\text{outcome}_{percent} = \frac{\bar{X} - \text{scale}_{min}}{\text{scale}_{max} - \text{scale}_{min}}
$$

where, $$\bar{X}$$ is the mean score for each study arm on the primary outcome questionnaire, and ${scale}_{min}$ and ${scale}_{max}$ are the minimum and maximum possible values of the scale in question, respectively. The standard deviation is calculated thus:

$$
\text{SD}_{Xpercent} = \frac{\text{SD}_{X}}{\text{scale}_{max} - \text{scale}_{min}}
$$

where $\text{SD}_{X}$ is the original standard deviation of the mean at baseline.

### Trial design

#### Measures of Effect

As the measure of effect of each individual study, we used the within-group Standardised Mean Difference (SMD), which we defined following [@lakensCalculatingReportingEffect2013, @cummingUnderstandingNewStatistics2013] as:

$$SMD_{change} = \frac{Mean_{t_{2}} - Mean_{t_{1}}}{\frac{SD_{t{2}} + SD_{t{1}}}2}\ $$

where, $Mean_{t_{2}}$ and $Mean_{t_{1}}$ refer to the means of the main outcome score at the end and beginning of the intervention respectively and $SD_{t_{2}}$ and $SD_{t_{1}}$ to the respective standard deviations. Where individual studies did not report all data required to calculate the SMD, we imputed missing data according to the methods summarised in this Cochrane Handbook [@higginsChapterChoosingEffect2023], in the following order. If a study reported the standard error of the mean, the SD was obtained simply by multiplying the SE by the square root of the sample size. For conditions where the SD was missing at one time point, the baseline SD was substituted by the post-test SD, and vice versa. If the SD was not available at either time point, missing values were replaced by the mean of the SDs available for comparable cases (defined as same trial type (psy or med), same instrument, same timepoint (pre or post), and same arm (control or active)). Where there were missing means at either baseline or post-test, missing values were calculated using mean change scores, preferring the change scores reported in the paper itself, though where this was unavailable, using the change scores reported in the dataset from Cipriani et al.'s meta-analysis (for medication studies only).

For the purposes of meta-analysis, it is necessary to estimate a standard error of the SMD. This is calculated according to:

$$ SE_{SMD} = \sqrt{\frac{2(1 -r_{t_{1}t_{2}})}{n} + \frac{SMD^2}{2n}}$$

where $n$ refers to the study sample size and $r_{t_{1}t_{2}}$ refers to the correlation between the outcome score obtained at baseline and at the end point. This correlation is typically not reported in studies and is often imputed using previously reported correlations for the instruments used. However, this practice has given rise to concerns about misestimation. Whilst such misestimation is possible, there is no reason to expect that it would be systematic, i.e. bias estimation of the effects for the control group of medication compared to those of psychotherapy. Still, to alleviate such concerns we have used a simulations.

In particular, we simulated one thousand truncated distribution of standard errors with the following general characteristics:

$$r_{t_{1}t_{2}} \sim \mathcal{TN}(\mu, \sigma, a, b)$$

for which we chose the mean to be $\mu = 0.65$, the standard deviation to be $\ sigma = 0.2$, and the upper and lower bounds to be $a = 0.45$ and $b = 0.9$, respectively. We then used these simulated datasets in the subsequent meta-analyses.

#### Random-effects Metaregression

We estimated the pooled standardised mean difference for each arm by using a random-effects meta-analysis implemented in R's metafor package (version 4.4-0). The main underlying assumption of random-effects meta-analysis is that each study's true effect size $\theta_{k}$ is affected not only by sampling error $\epsilon_{k}$, but also by $\zeta$ which represents heterogeneity between studies, allowing each study's estimate to vary along a distribution of effects, and the distribution of true effect sizes termed $\tau^2$. Therefore, we can estimate a two stage model with:

$$Y_i \sim \mathcal{N}(\theta_i, \sigma_i^2)
$$

$$\theta_i \sim \mathcal{N}(x_i\beta, \tau_i^2)
$$

where $Y_i$ is the estimated effect size for study i, has a normal distribution with $\theta_i$ as its true mean effect and sampling error $\sigma^2$. Whereas $\theta_i$ is a study-specific instantiation of the distribution of effect sizes, with $\tau^2$ representing heterogeneity.

This then gives rise to:

$$Y_i = x_i\beta_i + u_i + \epsilon_i$$

where,

$$u_i \sim N(0, \tau^2) $$

describes the deviation of each study from the mean of the distribution, and,

$$\epsilon_i \sim N(0, \sigma^2)$$

describes the sampling error.

We can then specify the following model to obtain the means of each arm of the trials as follows:

$$ \begin{aligned}
    Υ_i &=
    \begin{cases}
        0 & MedControl:  b_0 + u_i + \epsilon_i\\
        1 & MedActive:   b_0 + b_{1_{i}} +  u_i + \epsilon_i \\
        2 & PsyActive:   b_0 + b_{2_{i}}  +  u_i + \epsilon_i \\
        3 & PsyControl:  b_0 + b_{3_{i}}  +  u_i + \epsilon_i \\
    \end{cases}
\end{aligned}$$

where to obtain the mean of each level is the sum of $b_0$, the intercept for the reference category of medication control, with the coefficient of each level, e.g. for level 3, $b_{3_{i}}$ the psychotherapy controls. The confidence intervals of the means are constructed in the standard way using the standard errors of the mean. Similarly, each coefficient represents the contrast between the reference category and each level, for an example and of main interest to us $b_{3_{i}}$ represents the contrast between psychotherapy and medication control arms. Inference on the contrasts is done as follows:

$$
z = \frac{\hat{\beta}}{\text{SE}(\hat{\beta})}
$$

We used maximum likelihood (ML) to estimate model and applied Hartung-Knapp adjustment to reduce the chance of false positives [@inthoutHartungKnappSidikJonkmanMethodRandom2014].

We present the SMDs of each of the four treatment arms (medication control, medication active, psychotherapy control, psychotherapy active) under investigation. The SMDs are the means across the 1000 simulated datasets.

#### Number of sites

We also conducted a t-test to compare mean number of trial sites between psychotherapy and medication trials.

### Sensitivity Analyses

```{r, warning=FALSE, message = FALSE, echo = FALSE, output = FALSE }
# my alpha for the z 
target_probability <- 1 - 0.05

#start here 
x <- 1

# two decimals should be fine
tolerance <- 1e-3

#find the z value using the pnorm function
while (abs(pnorm(x)) < target_probability ) {
  x <- x + tolerance  
}

# Adjust the step size based on your problem  
cat("critical value of z =", x, "corresponding to an alpha = 0.05." , "\n")

x

```

We conducted a series of sensitivity analyses. For each of the meta-analyses we excluded studies that 1) used waitlist as their control and 2) recruited participants with subclinical levels of depression. Next, we conducted two analyses where we included only trials that used the Children's Depression Rating Scale, Revised (CDRS-R) or the Hamilton Depression Rating Scale (HAM-D) as outcome instruments.

Further, we tested whether the simulated values for the standard error had a substantial influence on the estimation of the differences between the medication and psychotherapy control conditions. To inspect whether this is the case, we plotted the z-value of the difference between the two coefficients against the number of simulations. We make inference on the stability of the difference, by counting the proportion of times that the z-value is above the critical value of z = `r x` corresponding to an alpha = 0.05.

Finally, we examined whether differential regression to the mean may account for differences in effect for psychotherapy and medication trials.

### Comparing the control and active arms of psychotherapy trials

We ran t-tests to compare the active and control arms of psychotherapy trials on key variables of interest regarding the intensity of the interventions. We extracted data pertaining to the number, duration and intensity of sessions, and the total cumulative hours and duration of the intervention. Where a range was provided, the maximum was encoded (e.g. if a paper reported that an intervention involved 8-10 sessions lasting 50-60 minutes, we encoded the number and duration of sessions as 10 and 60, respectively). If sessions varied in frequency across an intervention, we calculated an average by dividing total number of sessions by length of intervention period. Similarly, if the length of sessions varied across the course of the intervention, we calculated a weighted average. Phone call, web-chat and online sessions were encoded as sessions, however guided self-help components were not.

## Summary of all included trials

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#| label: tbl-all-trials
#| tbl-cap: "Summary of included RCTs"
#| tbl-cap-location: top

table_all_trials <- df_long_for_metan %>% 
  select(psy_or_med, new_study_id, arm_effect_size, type, baseline_n, instrument_name, baseline_mean, baseline_sd, post_mean, post_sd, cohens_d)

table_all_trials <- table_all_trials %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>% 
  mutate(psy_or_med = ifelse(psy_or_med == 0, "Medication", "Psychotherapy")) %>% 
  mutate(arm_effect_size = ifelse(arm_effect_size == "cohens_d_active", "Active", "Control"))

new_column_names <- c("psy_or_med", "Study", "Arm", "Description", "N", "Instrument", "Baseline M", "Baseline SD", "Post M", "Post SD", "Cohen's d")

# Assign new column names to the data frame
colnames(table_all_trials) <- new_column_names

table_all_trials %>% 
  arrange(psy_or_med, Study) %>% 
  group_by(psy_or_med) %>%
  as_flextable(hide_grouplabel = TRUE) %>% 
  merge_v(j = ~Study) %>% 
  align(j = 4:10, align = "center", part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row  
  bg(i = c(1, 86), bg = "#F2F2F2", part = "body") %>% 
  bold(i = c(1, 86), bold = TRUE, part = "body") %>% 
  border_remove() %>% # this is to remove borders that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "all" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  line_spacing(space = .7, part = "all") %>% 
  set_table_properties(width = 1, layout = "autofit")

```

## Metaregression sensitivity analyses

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#| label: fig-plot-means-no-wl
#| fig-cap: "Meta-analytic estimates of within-group changes: waitlist studies excluded"
#| fig-cap-location: top
#| fig-title: Figure S

coef_and_se_means_no_wl <- coef_and_se_means_no_wl %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition)))

# subtitle_text_no_wl <- "Metanalytic estimates of within-group changes: wl excluded"
plot_means_no_wl <- plot_means_function(coef_and_se_means_no_wl)
print(plot_means_no_wl)

# png("plot_means_no_wl.png")
# plot_means_no_wl
# dev.off()

```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#| label: fig-plot-means-clin
#| fig-cap: "Meta-analytic estimates of within-group changes: subclinical studies excluded"
#| fig-cap-location: top

coef_and_se_means_clin_study <- coef_and_se_means_clin_study %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition)))

# subtitle_text_clin <- "Metanalytic estimates of within-group changes: subclinical studies excluded"
plot_means_clin <- plot_means_function(coef_and_se_means_clin_study)
print(plot_means_clin)

# png("plot_means_clin.png")
# plot_means_clin
# dev.off()

```
```{r, warning=FALSE, message = FALSE, echo=FALSE}
#| label: fig-plot-means-cdrs
#| fig-cap: "Meta-analytic estimates of within-group changes: CDRS studies only"
#| fig-cap-location: top

coef_and_se_means_cdrs_study <- coef_and_se_means_cdrs_study %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition)))

# subtitle_text_cdrs <- "Metanalytic estimates of within-group changes: CDRS studies only"
plot_means_cdrs <- plot_means_function(coef_and_se_means_cdrs_study)
print(plot_means_cdrs)

# png("plot_means_cdrs.png")
# plot_means_cdrs
# dev.off()

```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#| label: fig-plot-means-hamd
#| fig-cap: "Meta-analytic estimates of within-group changes: HAM-D studies only"
#| fig-cap-location: top

coef_and_se_means_hamd_study <- coef_and_se_means_hamd_study %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition)))

# subtitle_text_hamd <- "Metanalytic estimates of within-group changes: HAM-D studies only"
plot_means_hamd <- plot_means_function(coef_and_se_means_hamd_study)
print(plot_means_hamd)

# png("plot_means_cdrs.png")
# plot_means_cdrs
# dev.off()

```

#### Effect of standard errors of the SMDs

It could be argued that the choice of standard errors of the changes for the calculation of the confidence intervals could have affect the results in one or the other direction. To address such concerns we have simulated 1000 different datasets with SMDs coming from a broad distribution. If standard error distributions were influential, this should show up as substantial variability across simulations. We test this idea in the @fig-stab-sims which displays across the 1000 simulations the z-value of the contrast between medication and psychotherapy control arms (the mean of which we presented in @tbl-results-overall). As can be seen, the variability in the z-score is minimal and consistently far away from the threshold for significance, i.e. the value of z = 1.645.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#| label: fig-stab-sims
#| fig-cap: "Stability of the Statistic of the Difference between Medication and Psychotherapy Control Arms"
#| fig-cap-location: top

# examine the stability of the simulations

z_psy_con_vs_med_con <- 0
for(i in 1: length(list_model_1_meta_reg)){
z_psy_con_vs_med_con[i] <- list_model_1_meta_reg[[i]]$zval[condition=="psychotherapy_control"]

}

stability_sims <- data.frame(n_sims = 1:1000, z_value = z_psy_con_vs_med_con)

stab_sims <- stability_sims %>%
  ggplot(aes(x = n_sims, y = z_value)) +
  geom_point() +
  geom_hline(yintercept = 1.65, linetype = "dotted", color = "red") +
  geom_segment(aes(x = 300, xend = 300, y = 2, yend = 1.65),
               arrow = arrow(length = unit(0.3, "cm")),
               color = "black") +
  annotate("text", x = 300, y = 2.0, label = "p < 0.05 threshold (values above line significant)", hjust = -0.1, vjust = 0) +
  geom_segment(aes(x = 300, xend = 300, y = 7.5, yend = min(stability_sims$z_value-0.3)),
               arrow = arrow(length = unit(0.3, "cm")),
               color = "black") +
  annotate("text", x = 300, y = 7.5, label = "simulated z-values",
           hjust = -0.1, vjust = 0) +
  theme_minimal() +
  labs(
    x = "Number of Simulations",
    y = "z-value",
    title = "Stability of the Statistic of the Difference between Medication and Psychotherapy Control Arms",
    subtitle = "Results from 1000 simulations with within-group standard errors"
  )

print(stab_sims)

```

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
#| label: tbl-hamd-baseline
#| tbl-cap: "HAM-D Baseline Scores across Psychotherapy and Medication RCTs"
#| tbl-cap-location: top

# now for the ham-d
# For row names
results_met_sev_hamd <- results_met_sev_hamd %>%
  mutate(subgroup = if_else(subgroup == 0, "Medication", "Psychotherapy"))

# Define new column names
new_column_names <- c("Subgroup", "K", "Mean", "SE", "Lower CI", "Upper CI", "T2", "p-value")

# Assign new column names to the data frame
colnames(results_met_sev_hamd) <- new_column_names

# Rounding
results_met_sev_hamd <- results_met_sev_hamd %>%
  mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
  mutate_at(vars(ncol(.)), ~ round(., 3)) %>%
  mutate_all(~if_else(is.na(.), "", as.character(.)))

results_met_sev_hamd %>% 
  as_flextable() %>% 
  align(j = 2:ncol(results_met_sev_hamd), align = "center", part = "all") %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row  
  border_remove() %>% # this is to remove borders that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "all" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  line_spacing(space = .7, part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  autofit()

# # All of the below is for the HAMD baseline meta-analysis when we have 4 levels
# # Prepare to tabulate
# 
# table_2_hamd_metareg_baseline <- results_hamd_metareg_baseline[[2]] %>%
#   mutate(across(where(is.numeric), ~round(., 2))) %>%
#   mutate(condition = case_when(
#     condition == "medication_control" ~ "Medication Control",
#     condition == "medication_active" ~ "Medication Active",
#     condition == "psychotherapy_control" ~ "Psychotherapy Control",
#     condition == "psychotherapy_active" ~ "Psychotherapy Active",
#     TRUE ~ as.character(condition))) %>%
#   relocate("condition", .before = "baseline_smds") %>%
#   relocate("n", .before = "baseline_smds") %>%
#   relocate("lower_cis_smds", .before = "upper_cis_smds")
# 
# new_column_names <- c("Condition", "N", "Baseline Scores", "Lower CI", "Upper CI")
# 
# # Assign new column names to the data frame
# colnames(table_2_hamd_metareg_baseline) <- new_column_names
# 
# flextable(table_2_hamd_metareg_baseline) %>%
#   set_table_properties(width = 1) %>%
#   autofit() %>%
#   font(fontname = "Calibri", part = "all") %>%
#   bold(part = "header") %>%   # Bold the relevant header rows
#   bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>%
#   border_remove() %>% # this is to remove borders around the caption that look strange
#   hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
#   vline(j = 1, part = "body", border = small_border) %>%
#   border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
#   hline(part = "header", border = big_border) %>%
#   # padding(i = 2, padding.bottom = 15, part = "header") %>%
#   align(j = 2:ncol(table_2_hamd_metareg_baseline), align = "center", part = "all") %>%
#   align(j = 1, align = "left", part = "all") %>%
#   paginate(init = TRUE, hdr_ftr = TRUE) %>%
#   line_spacing(space = .7, part = "all")
# 
# # Regression results
# 
# table_1_hamd_metareg_baseline <- results_hamd_metareg_baseline[[1]] %>%
#   mutate(across(where(is.numeric), ~round(., 2))) %>%
#   mutate(condition = case_when(
#     condition == "medication_control" ~ "Medication Control",
#     condition == "medication_active" ~ "Medication Active",
#     condition == "psychotherapy_control" ~ "Psychotherapy Control",
#     condition == "psychotherapy_active" ~ "Psychotherapy Active",
#     TRUE ~ as.character(condition))) %>%
#   relocate("condition", .before = "term") %>%
#   select(-c("term", "type"))
# 
# new_column_names <- c("Condition", "Estimate", "SE", "Statistic", "p value")
# 
# # Assign new column names to the data frame
# colnames(table_1_hamd_metareg_baseline) <- new_column_names
# 
# flextable(table_1_hamd_metareg_baseline) %>%
#   set_table_properties(width = 1) %>%
#   autofit() %>%
#   font(fontname = "Calibri", part = "all") %>%
#   bold(part = "header") %>%   # Bold the relevant header rows
#   bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>%
#   border_remove() %>% # this is to remove borders around the caption that look strange
#   hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
#   vline(j = 1, part = "body", border = small_border) %>%
#   border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
#   hline(part = "header", border = big_border) %>%
#   # padding(i = 2, padding.bottom = 15, part = "header") %>%
#   align(j = 2:ncol(table_1_hamd_metareg_baseline), align = "center", part = "all") %>%
#   align(j = 1, align = "left", part = "all") %>%
#   paginate(init = TRUE, hdr_ftr = TRUE) %>%
#   line_spacing(space = .7, part = "all")

```

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
#| label: tbl-cdrs-baseline
#| tbl-cap: "CDRS Baseline Scores across Psychotherapy and Medication RCTs"
#| tbl-cap-location: top

# For row names
results_met_sev_cdrs <- results_met_sev_cdrs %>%
  mutate(subgroup = if_else(subgroup == 0, "Medication", "Psychotherapy"))

# Define new column names
new_column_names <- c("Subgroup", "K", "Mean", "SE", "Lower CI", "Upper CI", "T2", "p-value")

# Assign new column names to the data frame
colnames(results_met_sev_cdrs) <- new_column_names

# Rounding
results_met_sev_cdrs <- results_met_sev_cdrs %>%
  mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
  mutate_at(vars(ncol(.)), ~ round(., 3)) %>%
  mutate_all(~if_else(is.na(.), "", as.character(.))) %>% 
  slice(1, 3, 2)

results_met_sev_cdrs %>% 
  as_flextable() %>% 
  align(j = 2:ncol(results_met_sev_cdrs), align = "center", part = "all") %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row  
  border_remove() %>% # this is to remove borders that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "all" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  line_spacing(space = .7, part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  autofit()

# # All of the below is for the CDRS baseline meta-analysis when we have 4 levels
# # Prepare to tabulate
# 
# table_2_cdrs_metareg_baseline <- results_cdrs_metareg_baseline[[2]] %>%
#   mutate(across(where(is.numeric), ~round(., 2))) %>%
#   mutate(condition = case_when(
#     condition == "medication_control" ~ "Medication Control",
#     condition == "medication_active" ~ "Medication Active",
#     condition == "psychotherapy_control" ~ "Psychotherapy Control",
#     condition == "psychotherapy_active" ~ "Psychotherapy Active",
#     TRUE ~ as.character(condition))) %>%
#   relocate("condition", .before = "baseline_smds") %>%
#   relocate("n", .before = "baseline_smds") %>%
#   relocate("lower_cis_smds", .before = "upper_cis_smds")
# 
# new_column_names <- c("Condition", "N", "Baseline Scores", "Lower CI", "Upper CI")
# 
# # Assign new column names to the data frame
# colnames(table_2_cdrs_metareg_baseline) <- new_column_names
# 
# flextable(table_2_cdrs_metareg_baseline) %>%
#   set_table_properties(width = 1) %>%
#   autofit() %>%
#   font(fontname = "Calibri", part = "all") %>%
#   bold(part = "header") %>%   # Bold the relevant header rows
#   bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>%
#   border_remove() %>% # this is to remove borders around the caption that look strange
#   hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
#   vline(j = 1, part = "body", border = small_border) %>%
#   border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
#   hline(part = "header", border = big_border) %>%
#   # padding(i = 2, padding.bottom = 15, part = "header") %>%
#   align(j = 2:ncol(table_2_cdrs_metareg_baseline), align = "center", part = "all") %>%
#   align(j = 1, align = "left", part = "all") %>%
#   paginate(init = TRUE, hdr_ftr = TRUE) %>%
#   line_spacing(space = .7, part = "all")
# 
# # Regression results
# 
# table_1_cdrs_metareg_baseline <- results_cdrs_metareg_baseline[[1]] %>%
#   mutate(across(where(is.numeric), ~round(., 2))) %>%
#   mutate(condition = case_when(
#     condition == "medication_control" ~ "Medication Control",
#     condition == "medication_active" ~ "Medication Active",
#     condition == "psychotherapy_control" ~ "Psychotherapy Control",
#     condition == "psychotherapy_active" ~ "Psychotherapy Active",
#     TRUE ~ as.character(condition))) %>%
#   relocate("condition", .before = "term") %>%
#   select(-c("term", "type"))
# 
# new_column_names <- c("Condition", "Estimate", "SE", "Statistic", "p value")
# 
# # Assign new column names to the data frame
# colnames(table_1_cdrs_metareg_baseline) <- new_column_names
# 
# flextable(table_1_cdrs_metareg_baseline) %>%
#   set_table_properties(width = 1) %>%
#   autofit() %>%
#   font(fontname = "Calibri", part = "all") %>%
#   bold(part = "header") %>%   # Bold the relevant header rows
#   bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>%
#   border_remove() %>% # this is to remove borders around the caption that look strange
#   hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
#   vline(j = 1, part = "body", border = small_border) %>%
#   border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
#   hline(part = "header", border = big_border) %>%
#   # padding(i = 2, padding.bottom = 15, part = "header") %>%
#   align(j = 2:ncol(table_1_cdrs_metareg_baseline), align = "center", part = "all") %>%
#   align(j = 1, align = "left", part = "all") %>%
#   paginate(init = TRUE, hdr_ftr = TRUE) %>%
#   line_spacing(space = .7, part = "all")

```

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
#| label: tbl-adj-smds
#| tbl-cap: "SMDs adjusted by baseline scores"
#| tbl-cap-location: top

table_mean_coefs_from_sim_adjusted_cohens_d <- df_mean_coefs_from_sim_adjusted_cohens_d %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition)))

new_column_names <- c("Condition", "Coefficient", "SE", "Lower CI", "Upper CI")

# Assign new column names to the data frame
colnames(table_mean_coefs_from_sim_adjusted_cohens_d) <- new_column_names

flextable(table_mean_coefs_from_sim_adjusted_cohens_d) %>%
  set_table_properties(width = 1) %>%
  autofit() %>%
  font(fontname = "Calibri", part = "all") %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  # padding(i = 2, padding.bottom = 15, part = "header") %>% 
  align(j = 2:ncol(table_mean_coefs_from_sim_adjusted_cohens_d), align = "center", part = "all") %>% 
  align(j = 1, align = "left", part = "all") %>% 
  paginate(init = TRUE, hdr_ftr = TRUE) %>% 
  line_spacing(space = .7, part = "all")

aggregate_results_overall_adjusted_cohens_d <- aggregate_results_overall_adjusted_cohens_d %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition))) %>% 
  relocate("condition", .before = "coefficients")

new_column_names <- c("Condition", "Coefficient", "SE", "z value", "Lower CI", "Upper CI", "T2", "I2", "k", "R2")

# Assign new column names to the data frame
colnames(aggregate_results_overall_adjusted_cohens_d) <- new_column_names

table_results_overall_adjusted_cohens_d <- aggregate_results_overall_adjusted_cohens_d[, 1:6]

flextable(table_results_overall_adjusted_cohens_d) %>%
  set_table_properties(width = 1) %>%
  autofit() %>%
  add_footer_row(values = as_paragraph("T", as_sup("2"), " = ", aggregate_results_overall_adjusted_cohens_d[1, "T2"],
                      "; ", as_i("I"), as_sup("2"), " = ", aggregate_results_overall_adjusted_cohens_d[1, "I2"], 
                      "; ", as_i("K"), " = ", aggregate_results_overall_adjusted_cohens_d[1, "k"],
                      "; ", as_i("R"), as_sup("2"), " = ", aggregate_results_overall_adjusted_cohens_d[1, "R2"], "."), colwidths = 6) %>%
  # add_header_row(values = "Results from metaregression with overall sample", colwidths = 6) %>%
  # add_header_row(values = "Table 4", colwidths = 6) %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  # italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  # padding(i = 2, padding.bottom = 15, part = "header") %>% 
  align(j = 2:ncol(table_aggregate_results_overall), align = "center", part = "all") %>% 
  align(j = 1, align = "left", part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>% 
  paginate(init = TRUE, hdr_ftr = TRUE) %>% 
  line_spacing(space = .7, part = "all")

```

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}

# We want to compare active v control arms when we filter out wl conditions

df_descr_psy_conditions_no_wl <- df_descr_psy_conditions %>% 
  filter(new_study_id %in% no_wl_ids)

df_descr_psy_conditions_no_wl <- df_descr_psy_conditions_no_wl %>% 
    filter(!(treatment == "wlc"))

sessions <- group_function(df_descr_psy_conditions_no_wl,"levels_var","no_sessions")
frequency <- group_function(df_descr_psy_conditions_no_wl,"levels_var","freq_weeks")
length_sessions_mins <- group_function(df_descr_psy_conditions_no_wl,"levels_var","length_sessions_mins")
total_hours <- group_function(df_descr_psy_conditions_no_wl,"levels_var","total_hours")

control_summary_no_wl <- rbind(sessions, frequency, length_sessions_mins, total_hours)

control_summary_no_wl <- control_summary_no_wl %>% 
  select(-c(Median, IQR))

control_summary_no_wl <- control_summary_no_wl %>%
  mutate(Group = case_when(
    Group == "psychotherapy_active" ~ "Active",
    Group == "psychotherapy_control" ~ "Control",
    TRUE ~ as.character(Group)  # Keep other values unchanged
  ))

control_summary_no_wl <- control_summary_no_wl %>%
  mutate(Column = case_when(
    Column == "no_sessions" ~ "Number of sessions",
    Column == "freq_weeks" ~ "Intensity (sessions per week)",
    Column == "length_sessions_mins" ~ "Session length (mins)",
    Column == "total_hours" ~ "Total intervention hours",
    TRUE ~ as.character(Column)  # Keep original value if no match
  ))

# t tests

session_no_t_test <- t.test(no_sessions ~ levels_var, data = df_descr_psy_conditions_no_wl)
freq_t_test <- t.test(freq_weeks ~ levels_var, data = df_descr_psy_conditions_no_wl)
session_length_t_test <- t.test(length_sessions_mins ~ levels_var, data = df_descr_psy_conditions_no_wl)
total_hours_t_test <- t.test(total_hours ~ levels_var, data = df_descr_psy_conditions_no_wl)

extract_t_test_stats <- function(df) {

  t_test_stats <- data.frame(
  outcome = df$data.name,
  statistic = df$statistic,
  df = df$parameter,
  p_value = df$p.value,
  ci_lower = df$conf.int[1],
  ci_upper = df$conf.int[2])

  return(t_test_stats)

}

results_session_no_t_test <- extract_t_test_stats(session_no_t_test)
results_freq_t_test <- extract_t_test_stats(freq_t_test)
results_session_length_t_test <- extract_t_test_stats(session_length_t_test)
results_total_hours_t_test <- extract_t_test_stats(total_hours_t_test)

control_summary_t_tests_no_wl <- rbind(results_session_no_t_test, results_freq_t_test, results_session_length_t_test, results_total_hours_t_test)

# For row names
control_summary_t_tests_no_wl <- control_summary_t_tests_no_wl %>%
  mutate(outcome = case_when(
    row_number() == 1 ~ "Number of sessions",
    row_number() == 2 ~ "Intensity (sessions per week)",
    row_number() == 3 ~ "Session length (mins)",
    row_number() == 4 ~ "Total intervention hours"
  ))

# Define new column names
new_column_names <- c("Outcome", "t statistic", "df", "p-value", "Lower CI", "Upper CI")

# Assign new column names to the data frame
colnames(control_summary_t_tests_no_wl) <- new_column_names

#Rounding
control_summary_t_tests_no_wl <- control_summary_t_tests_no_wl %>%
  mutate(across(where(is.numeric) & !matches("p-value"), ~ round(., 2))) %>%
  mutate(`p-value` = ifelse(`p-value` < 0.001, "< 0.001", sprintf("%.3f", `p-value`)))

# calculating with my function

# session_es_no_wl <- effect_btwn(sessions)
# frequency_es_no_wl <- effect_btwn(frequency)
# length_sessions_mins_es_no_wl <- effect_btwn(length_sessions_mins)
# total_hours_es_no_wl <- effect_btwn(total_hours)
# total_period_weeks_es_no_wl <- effect_btwn(total_period_weeks)

# using a package to calculate instead

session_es_no_wl <- cohen.d(df_descr_psy_conditions_no_wl$no_sessions ~ df_descr_psy_conditions_no_wl$levels_var, pooled = TRUE, paired = FALSE, within=FALSE, na.rm=FALSE )
frequency_es_no_wl <- cohen.d(df_descr_psy_conditions_no_wl$freq_weeks ~ df_descr_psy_conditions_no_wl$levels_var, pooled = TRUE, paired = FALSE, within=FALSE, na.rm=FALSE )
length_sessions_mins_es_no_wl <- cohen.d(df_descr_psy_conditions_no_wl$length_sessions_mins ~ df_descr_psy_conditions_no_wl$levels_var, pooled = TRUE, paired = FALSE, within=FALSE, na.rm=FALSE )
total_hours_es_no_wl <- cohen.d(df_descr_psy_conditions_no_wl$total_hours ~ df_descr_psy_conditions_no_wl$levels_var, pooled = TRUE, paired = FALSE, within=FALSE, na.rm=FALSE )

# going to try displaying summary statistics and t-tests in one place

control_summary_no_wl <- cbind(control_summary_no_wl, 
                          "Cohen's d" = rep(NA, nrow(control_summary_no_wl)),
                         "Upper CI" = rep(NA, nrow(control_summary_no_wl)),
                         "Lower CI" = rep(NA, nrow(control_summary_no_wl)),
                         t = rep(NA, nrow(control_summary_no_wl)), 
                          df = rep(NA, nrow(control_summary_no_wl)), 
                          "p-value" = rep(NA, nrow(control_summary_no_wl)))

# Update the first row with session_es_no_wl
control_summary_no_wl[1, c("Cohen's d", "Upper CI", "Lower CI", "t", "df", "p-value")] <- c(
  session_es_no_wl$estimate, session_es_no_wl$conf.int[1], session_es_no_wl$conf.int[2], 
  control_summary_t_tests_no_wl$`t statistic`[1], control_summary_t_tests_no_wl$df[1], 
  control_summary_t_tests_no_wl$`p-value`[1]
)

# Update the third row with frequency_es_no_wl
control_summary_no_wl[3, c("Cohen's d", "Upper CI", "Lower CI", "t", "df", "p-value")] <- c(
  frequency_es_no_wl$estimate, frequency_es_no_wl$conf.int[1], frequency_es_no_wl$conf.int[2], 
  control_summary_t_tests_no_wl$`t statistic`[2], control_summary_t_tests_no_wl$df[2], 
  control_summary_t_tests_no_wl$`p-value`[2]
)

# Update the fifth row with length_sessions_mins_es_no_wl
control_summary_no_wl[5, c("Cohen's d", "Upper CI", "Lower CI", "t", "df", "p-value")] <- c(
  length_sessions_mins_es_no_wl$estimate, length_sessions_mins_es_no_wl$conf.int[1], length_sessions_mins_es_no_wl$conf.int[2], 
  control_summary_t_tests_no_wl$`t statistic`[3], control_summary_t_tests_no_wl$df[3], 
  control_summary_t_tests_no_wl$`p-value`[3]
)

# Update the seventh row with total_hours_es_no_wl
control_summary_no_wl[7, c("Cohen's d", "Upper CI", "Lower CI", "t", "df", "p-value")] <- c(
  total_hours_es_no_wl$estimate, total_hours_es_no_wl$conf.int[1], total_hours_es_no_wl$conf.int[2], 
  control_summary_t_tests_no_wl$`t statistic`[4], control_summary_t_tests_no_wl$df[4], 
  control_summary_t_tests_no_wl$`p-value`[4]
)

control_summary_no_wl <- control_summary_no_wl %>%
  mutate(
    `Cohen's d` = as.numeric(`Cohen's d`),
    `Upper CI` = as.numeric(`Upper CI`),
    `Lower CI` = as.numeric(`Lower CI`)
  ) %>%
  mutate(across(where(is.numeric), ~round(., 2)))

```

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
#| label: tbl-intensity-of-int-no-wl
#| tbl-cap: "Comparing the intensity of the intervention between active and control arms of psychotherapy studies: waitlist studies excluded"
#| tbl-cap-location: top

as_grouped_data(control_summary_no_wl, groups = "Column") %>% 
  as_flextable(hide_grouplabel = TRUE ) %>% 
  bold(i = c(1, 4, 7, 10), bold = TRUE, part = "body") %>%  
  set_table_properties(width = 1) %>%  # Set table width 
  autofit() %>%  # Autofit column widths 
  align(j = 2:9, align = "center", part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  # add_header_row(values = "Comparing the intensity of the intervention between active and control arms of psychotherapy studies", colwidths = 4) %>% # This is a work around for the problems with rendering captions
  # add_header_row(values = "Table 5", colwidths = 4) %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  #italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(j = 1:4, part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = c(1,4), part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  #padding(i = 2, padding.bottom = 15, part = "header") %>% 
  paginate(init = TRUE) %>% 
  line_spacing(space = .7, part = "all")

```

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
# #| label: tbl-t-tests-intensity-no-wl
# #| tbl-cap: "Results for t-tests comparing the intensity of the intervention between active and control arms of psychotherapy studies: waitlist studies excluded"
# #| tbl-cap-location: top
# 
# flextable(control_summary_t_tests_no_wl) %>%
#   set_table_properties(width = 1) %>%  # Set table width %>% 
#   autofit() %>%  # Autofit column widths 
#   align(j = 2:ncol(control_summary_t_tests_no_wl), align = "center", part = "all") %>% 
#   font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
#   #add_header_row(values = "Results for t-tests comparing intervention intensity between active and control arms of psychotherapy trials", colwidths = 6) %>% # This is a work around for the problems with rendering captions
#   #add_header_row(values = "Table 6", colwidths = 6) %>% 
#   bold(part = "header") %>%   # Bold the relevant header rows
#   bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
#   #italic(i = 2, italic = TRUE, part = "header") %>% 
#   border_remove() %>% # this is to remove borders around the caption that look strange
#   hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
#   vline(j = 1, part = "body", border = small_border) %>% 
#   border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
#   hline(part = "header", border = big_border) %>% 
#   #padding(i = 2, padding.bottom = 15, part = "header") %>% 
#   paginate(init = TRUE) %>% 
#   line_spacing(space = .7, part = "all")

```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# Plots  -----------------------

#define groups
# df_long_for_metan <- df_long_for_metan %>%
#   mutate(
#     group = case_when(
#       psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "Medication Active",
#       psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "Medication Control",
#       psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "Psychotherapy Active",
#       psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "Psychotherapy Control",
#       TRUE ~ NA_character_
#     )
#   )
# 
# group_colors <- c("Medication Active" = "deeppink1", "Medication Control" = "deeppink4", "Psychotherapy Active" = "steelblue1", "Psychotherapy Control" = "steelblue3")

# # Scatter plot cohens d per arm
# 
# ggplot(df_long_for_metan, aes(x = group, y = cohens_d, color = group)) +
#   geom_point(position = position_jitter(width = 0.2), size = 3) +
#   geom_segment(x = 4.5, xend = 4.5, y = -3.5, yend = -3,
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -3, label = "Less Effective", color = "grey", vjust = 0, hjust = .2) +
#   geom_segment(x = 4.5, xend = 4.5, y = -4, yend = -4.5,
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -4.5, label = "More Effective", color = "grey", vjust = 0, hjust = .2) +
#   scale_color_manual(values = group_colors) +
#   labs(title = "Scatter plot of Cohen's d by group", x = NULL, y = "Cohen's d") +
#   theme_minimal() +
#   theme(legend.position = "none")

# Box plot Cohens d by group
# 
# ggplot(df_long_for_metan, aes(x = group, y = cohens_d, fill = group)) +
#   geom_boxplot(outlier.shape = NA, position = position_dodge(0.8), color = "black", alpha = 0.5) +
#   scale_y_continuous(limits = c(-3.5, 1), breaks = seq(-4, 1, by = 1)) +
#   geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
#   geom_segment(x = 4.5, xend = 4.5, y = -2.5, yend = -2,
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -2.2, label = "Less Effective", color = "grey", vjust = 0, hjust = .2) +
#   geom_segment(x = 4.5, xend = 4.5, y = -3, yend = -3.5,
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -3.2, label = "More Effective", color = "grey", vjust = 0, hjust = .2) +
#   scale_fill_manual(values = group_colors) +
#   labs(title = "Box plot of cohens_d by group", x = "Group", y = "Cohen's d") +
#   theme_minimal() +
#   theme(legend.position = "none")

```

# References